---
title: "Water Quality Index data layer"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

## Summary

This layer is derived from the [EPA's National Coastal Condition Assessment Water Quality Index (WQI) data](https://www.epa.gov/national-aquatic-resource-surveys/ncca), specifically the measurements for 2005/2006 and 2010 time periods.

The EPA measures the Water Quality Index for multiple sites throughout the Northeast region. The WQI status for each site are either `GOOD`, `FAIR`, `POOR` or `MISSING`. A score for each OHI-Northeast coastal region is calculated using these WQI values and the site-associated weights. `GOOD` sites are assigned a perfect score (100), `FAIR` sites are assigned a value of 50 and `POOR` sites are assigned a 0. Each site's score is multiplied by it's weight (assigned by the EPA) and then averaged to get final region scores.

Sites with MISSING data are removed.

## Background

> "The National Coastal Condition Assessment (NCCA) is a national coastal monitoring program with rigorous quality assurance protocols and standardized sampling procedures designed to produce national and regional estimates of coastal condition."

> "NCCA evaluates four indices of condition—water quality, sediment quality, benthic community condition, and fish tissue contaminants – to evaluate the ecological condition and recreational potential of coastal waters. Consistent sampling and analytical procedures ensure that the results can be compared across the country and over time."

The WQI is a composite indicator that combines measurements of five components (DIN, DIP, chlorophyll a, water clarity, and dissolved oxygen).

Assessments have taken place in 2000/2001, 2005/2006, 2010 and 2015. As of this writing, the 2015 data is not available.

### Setup

```{r setup, message = F, warning = F, reslts = 'hide'}
knitr::opts_chunk$set(fig.width = 10, fig.height = 6, fig.path = 'figs/', message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')

library(tidyverse)
```

## Load Data

Four datasets were downloaded from [this NCCA data portal](https://www.epa.gov/national-aquatic-resource-surveys/data-national-aquatic-resource-surveys):

- site information for both time periods (2 datasets)
- WQI for both time periods (2 datasets)

The data is loaded and filtered for sites in the Northeast region, identified by postal code. The only information we need from the `sites` data is the weighting factor for each site. This will be used later on to properly weight the contribution of each site to the final score.

```{r load_data}
#load site info for 05/06 time period
site_info_1 <- readxl::read_excel(file.path(dir_anx, "_raw_data/EPA/nca_siteinformationdata_narschallenge.xlsx")) %>% 
  filter(PSTL_CODE %in% c("MA", "ME", "RI", "NY", "NH", "CT")) %>%
  select(siteID = SITE_ID, WGT_NCA_56)

#load site info for 2010 time period
site_info_2 <- read_csv(file.path(dir_anx, "_raw_data/EPA/assessed_ncca2010_siteinfo.revised.06212016.csv")) %>%
  filter(STATE %in% c("MA", "ME", "RI", "NY", "NH", "CT"),
         NCA_REGION == "East Coast") %>%
  select(siteID = SITE_ID, WGT_NCCA10)

#load WQI data for 05/06
water_old <- readxl::read_excel(file.path(dir_anx, "_raw_data/EPA/nca_waterchemdata_narschallenge.xlsx")) %>% 
  filter(PSTL_CODE %in% c("MA", "ME", "RI", "NY", "NH", "CT")) %>%
  select(siteID = SITE_ID, status = Cat_WQI, LON_DD, LAT_DD, PSTL_CODE, Year = SAMPYEAR) %>%
  left_join(site_info_1)

#load WQI data for 2010
water_new <- read_csv(file.path(dir_anx, "_raw_data/EPA/ncca2010_water_quality_indicator_status.csv")) %>%
  filter(PSTL_CODE %in% c("MA", "ME", "RI", "NY", "NH", "CT"),
         NCCA_REG == "Northeast") %>%
  select(siteID = SITE_ID, status = CAT_WQI, LON_DD, LAT_DD, PSTL_CODE) %>%
  mutate(Year = 2010) %>%
  left_join(site_info_2)
```

## Data wrangling

Combining both datasets and assigning the time periods for each. If the year is 2000 or 2001, that's time period 1. 2005-2006 is time period 2 and 2010 is time period 3.

```{r combine_data}
wqi <- water_old %>%
  bind_rows(water_new) %>%
  mutate(time_period = 
           case_when(
             Year %in% c(2000, 2001) ~ 1,
             Year %in% c(2005, 2006) ~ 2,
             Year == 2010 ~ 3
           )) %>%
  filter(time_period != 1) #remove earlier time period
```

Now assign OHI region names or ids to each of the sites. For the most part, the State in the WQI data is the region, but for Massachusetts we need to be specific about what region the sites belong to. The best way to do this is manually so I use `mapview::mapview()` to identify sites in MA and assign to either Massachusetts-North or Massachusetts-South.

Look first just at sites identified by the `MA` postal code.

```{r ma_sites}
#filter wqi for sites only in MA
ma <- wqi %>%
  filter(PSTL_CODE == "MA")

#create a simple features (sf) object from the coordinates contained in the dataset
ma_points = st_as_sf(ma, coords = c("LON_DD", "LAT_DD"), 
                 crs = 4326)
#plot
ggplot() +
  geom_sf(data = rgns %>% filter(rgn_id %in% c(7,8)), aes(fill = rgn_id)) +
  geom_sf(data = ma_points, color = 'darkgreen') +
  theme_bw()
```

To assign region ids, try `sf::st_intersect` between the MA sites and the OHI-Northeast regions file already loaded from common.R
```{r ma_rgns_intersect}
int <- ma_points %>%
  st_transform('+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs') %>%
  st_intersection(rgns)

ggplot() +
  geom_sf(data = rgns %>% filter(rgn_id %in% c(7,8)), aes(fill = rgn_id)) +
  geom_sf(data = int, color = 'darkgreen') +
  theme_bw()
```

It looks like most sites were captured but what points are we missing? We probably didn't get all the sites in the intersection.

```{r missing_ma_sites}
#find sites that are not in the intersection
s <- setdiff(ma_points$siteID, int$siteID)

#limit to just the missing sites
miss <- ma_points %>%
  filter(siteID %in% s)

#use mapview to manually explore and then assign region ids
mapview::mapview(miss)

miss_gf <- miss %>%
  data.frame() %>%
  mutate(rgn_id = 
           case_when(
             siteID == "NCCA10-1001" ~ 8,
             siteID == "NCCA10-1008" ~ 8
           )) %>%
  select(-geometry)
```

Take the info from `int` that we want and combine with these missing sites

```{r}
ma_df <- int %>%
  data.frame() %>%
  select(-area_km2, -geometry, -rgn_name) %>%
  bind_rows(miss_gf)
```

Additional data cleaning that adds the Massachusetts sites back in and removes the few sites that are actually in the George's Bank area.

```{r clean_wqi_data}
wqi_clean <- wqi %>%
  mutate(rgn_name = 
           case_when(
             PSTL_CODE == "ME" ~ "Maine",
             PSTL_CODE == "RI" ~ "Rhode Island",
             PSTL_CODE == "NY" ~ "New York",
             PSTL_CODE == "CT" ~ "Connecticut",
             PSTL_CODE == "NH" ~ "New Hampshire",
             PSTL_CODE == "MA" ~ "NA"
           )) %>%
  left_join(rgn_data) %>%
  left_join(ma_df, by = c("siteID", "time_period", "Year", "status", "PSTL_CODE")) %>%
  mutate(rgn_id = ifelse(is.na(rgn_id.x), rgn_id.y, rgn_id.x)) %>%
  select(-rgn_id.x, -rgn_id.y, -rgn_name, -area_km2, -WGT_NCA_56.y, -WGT_NCCA10.y) %>%
  left_join(rgn_data, by = "rgn_id") %>%
  filter(rgn_name != "Georges Bank") %>%
  rename(WGT_NCA_56 = WGT_NCA_56.x,
         WGT_NCCA10 = WGT_NCCA10.x) %>%
  select(-state_abv.x, -state_abv.y, -state_name.x, -state.x, -state_name.y, -state.y, -area_km2) %>%
  filter(Year > 2004) %>% #remove the early years
  as.data.frame()
```

## Run cat.analysis

Use the [`spsurvey`](https://www.rdocumentation.org/packages/spsurvey/versions/3.4) package to run `cat.analysis` and produce a sub-population estimate that incorporates site weights. 

We remove all sites that have `MISSING` as a status. In this data, New York has 5%, New Hampshire has 3.5% and Maine has 2% missing.

In 2010 only New York has missing sites, a total of 3.35% of area.

**For now I will ignore the MISSING sites**

```{r cat_analysis_0506}
library(spsurvey)

#select just 0506 data
wqi_0506 <- wqi_clean %>% 
  filter(Year < 2007, status != "MISSING")

# Create the required data frames
sites <- wqi_0506 %>%
          select(siteID) %>%
          mutate(use = rep(TRUE, nrow(.)))
subpop <- wqi_0506 %>%
          select(siteID, rgn_name)
design <- wqi_0506%>%
          select(siteID,
                 wgt = WGT_NCA_56,
                 xcoord = LON_DD,
                 ycoord = LAT_DD)
data <- wqi_0506 %>%
        select(siteID, status)

## from this output we only want to grab. Estimate.P is the percent of area (e.g. 98%) for the coastal waters of a subpopulation (e.g. Maine) in a particular category of condition (e.g. Good)
Water_Column_Status_Estimates_0506 <- cat.analysis(sites, subpop, design, data) %>%
  select(rgn_name = Subpopulation, Category, perc = Estimate.P) %>%
  mutate(cat_score = 
      case_when(
        Category == "POOR" ~ 0,
        Category == "FAIR" ~ 50,
        Category == "GOOD" ~ 100),
      year = 2006)
```

Do the same `cat.analysis` but for 2010 data.

```{r cat_analysis_2010}
# Create the required data frames
wqi_2010 <- wqi_clean %>% 
  filter(Year == 2010, status != "MISSING")

sites <- wqi_2010 %>%
          select(siteID) %>%
          mutate(use = rep(TRUE, nrow(.)))
subpop <- wqi_2010%>%
          select(siteID, rgn_name)
design <- wqi_2010%>%
          select(siteID,
                 wgt = WGT_NCCA10,
                 xcoord = LON_DD,
                 ycoord = LAT_DD)
data <- wqi_2010 %>%
        select(siteID, status)

## from this output we only want to grab. Estimate.P is the percent of area (e.g. 98%) for the coastal waters of a subpopulation (e.g. Maine) in a particular category of condition (e.g. Good)
Water_Column_Status_Estimates_2010 <- cat.analysis(sites, subpop, design, data) %>%
  select(rgn_name = Subpopulation, Category, perc = Estimate.P) %>%
  mutate(cat_score = 
      case_when(
        Category == "POOR" ~ 0,
        Category == "FAIR" ~ 50,
        Category == "GOOD" ~ 100),
      year = 2010) 
```

## Create layer

Aggregate by region and get WQI scores

```{r wqi_layer_rgn_scores}
wqi_rgns <- Water_Column_Status_Estimates_0506 %>%
  rbind(Water_Column_Status_Estimates_2010) %>%
  mutate(score = cat_score * (perc/100)) %>%
  filter(Category != "Total") %>%
  group_by(rgn_name, year) %>%
  summarize(wqi_score = sum(score))

write.csv(wqi_rgns, file = "data/wqi_score_rgns.csv")

ggplot(wqi_rgns, aes(x = year, y = wqi_score, color = rgn_name)) +
  geom_line() +
  theme_bw() +
  labs(x = "Year",
       y = "Score",
       color = "Region")
```

## Gap-filling

The plot has guessed the values between 2006 and 2010 but we need to actually assign these before saving this layer. We will assume a simple linear interpolation between years.

```{r gapfill}
df_gf <- wqi_rgns %>%
  complete(year = 2006:2010) %>%
  group_by(rgn_name) %>%
  mutate(score = zoo::na.approx(wqi_score),
         gapfilled = ifelse(is.na(wqi_score), 1, 0))

write.csv(df_gf, file = "data/wqi_score_rgns_gf.csv")
```

## Save layer for toolbox

I need to attach region ID's to each row. I also add the offshore regions 1:4 with NA values for the toolbox to run. 

```{r save_layer}
other_rgns <- data.frame(year     = rep(2006:2010, each = 4),
                         rgn_id   = c(1,2,3,4),
                         rgn_name = c("Offshore", "Georges Bank", "Gulf of Maine", "Mid-Atlantic Bight"),
                         score = NA)

df_gf %>%
  left_join(rgn_data) %>%
  select(year, rgn_id, rgn_name, score) %>%
  bind_rows(other_rgns) %>%
  write_csv(file.path(dir_calc, "layers/cw_wqi.csv"))
```

Save as a pressure layer as well (inverse)

```{r}
df_gf %>%
  left_join(rgn_data) %>%
  mutate(score = 1-score/100) %>%
  select(year, rgn_id, rgn_name, score) %>%
  bind_rows(other_rgns) %>%
  write_csv(file.path(dir_calc, "layers/prs_wqi.csv"))
```


## Citation

Citation: U.S. Environmental Protection Agency. 2016. National Aquatic Resource Surveys. National Coastal Condition Assessment 2010 (data and metadata files). Available from U.S. EPA web page:https://www.epa.gov/national-aquatic-resource-surveys/data-national-aquatic-resource-surveys. Data from the National Aquatic Resource Surveys. Date accessed: 2018-07-31.









