---
title: "OHINE: Climate Change Resilience - ACEE Rankings Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the ACEE scores by OHINE region. This data is used to measure climate change resilience regulations.

# Data

**American Council for Energy-Efficient Economy** (EEE) state and local policy rankings

**Downloaded**: August, 2019

**Description**: Rankings of states' energy efficiency policy and program efforts. https://database.aceee.org/state-scorecard-rank

**Time range**: 2006-2018

**Format**: PDF


# About
American Council for Energy Efficiency Rankings: Each state is scored out of 50 (except in 2006 it's out of 44) on their policies in place for energy efficiency etc. 50 would be a perfect score. This script adjusts those scores on a scale of 0-1. There is no data for 2007, and the method for ranking the states changed by adding a new category from 2006 to 2008 (hence the total possible score changes). I used the 2006 score for 2007, and this was done in the spreadsheet on the server.

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(readxl)
```

Load Data
```{r load_data}
eee <- read_excel(file.path(dir_anx, "_raw_data/EEE/aceee_2006_2018.xlsx"))
```


## Calculate scores
```{r calculate}
acee_rankings <- eee %>% 
  mutate(score_eee = score/target)

#save
write.csv(acee_rankings, file = "data/acee_rankings.csv")
```

## Graph scores
```{r visualize}
acee_rankings_graph <-ggplot(acee_rankings)+
  geom_line(aes(x=year, y = score, color=state))+
  ggtitle("American Council for Energy Efficiency Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 
acee_rankings_graph

ggsave("figs/acee_rankings_graph.jpg", width=7, height=5, dpi=300)
```





