---
title: "OHIEC: Resilience - Climate Change Pressure Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates climate change resilience values by OHINE region. 

**4. EPA Carbon Emissions**
**Downloaded**: August, 2019

**Description**: Carbon emissions from each state from fossil fuels. 'https://www.epa.gov/sites/production/files/2017-09/documents/co2ffc_2015.pdf'

**Time range**: 1990-2016

**Format**: Tabular
1. Take the total CO2 emissions each year and compare to the target, which is calculated by the state set goals
a. Carbon Emission Targets:
Targets pulled from https://www.c2es.org/document/greenhouse-gas-emissions-targets/
Rhode Island is missing from this map so pulled from http://www.planning.ri.gov/documents/LU/energy/energy15.pdf


##Carbon Emmissions
Targets:
Maine= 80% reduction below 1990 levels by 2050
New Hampshire= 80% reduction below 1990 levels by 2050
Connecticut= 80% reduction below 2001 levels by 2050
Massachusetts= 80% reduction below 1990 levels by 2050
New York= 100% reduction below 1990 levels by 2050
Rhode Island= 80% reduction below 1990 levels by 2050


``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(Hmisc)
library(stringi)
library(stringr)
library(readxl)
library(tabulizer)
```

Converting table from a pdf to a df
```{r}
location <- file.path(dir_anx, "_raw_data/EPA/co2_emmissions_EPA.pdf")
out <- extract_tables(location)

#page 1
out1 <- out[[1]] %>%  
  as.data.frame() %>% 
  rename(state = V1, blank = V2, "1990" = V3, "1991"= V4, "1992"= V5, "1993"=V6, "1994"=V7, "1995"=V8, "1996"=V9, "1997"=V10, "1998"=V11, "1999"=V12, "2000"=V13, "2001"=V14, "2002"=V15, "2003"=V16, "2004"=V17, "2005"=V18, "2006"=V19, "2007"=V20, "2008"=V21, "2009"=V22, "2010"=V23, "2011"=V24, "2012"=V25, "2013"=V26, "2014"=V27, "2015"=V28, "2016"=V29) %>% 
  filter(state != "State Sector",
         state != "Commercial",
         state != "Industrial", 
         state != "Residential",
         state != "Transportation",
         state != "Electric Power") %>% 
  select(-blank)
  
#page 2  
out2 <- out[[2]] %>% 
  as.data.frame() %>% 
  rename(state = V1, blank = V2, "1990" = V3, "1991"= V4, "1992"= V5, "1993"=V6, "1994"=V7, "1995"=V8, "1996"=V9, "1997"=V10, "1998"=V11, "1999"=V12, "2000"=V13, "2001"=V14, "2002"=V15, "2003"=V16, "2004"=V17, "2005"=V18, "2006"=V19, "2007"=V20, "2008"=V21, "2009"=V22, "2010"=V23, "2011"=V24, "2012"=V25, "2013"=V26, "2014"=V27, "2015"=V28, "2016"=V29) %>% 
  filter(blank != "Commercial",
         blank != "Industrial", 
         blank != "Residential",
         blank != "Transportation",
         blank != "Electric Power")%>% 
  select(-blank)

#page 3
out3 <- out[[3]] %>% 
   as.data.frame() %>% 
  rename(state = V1, "1990" = V2, "1991"= V3, "1992"= V4, "1993"=V5, "1994"=V6, "1995"=V7, "1996"=V8, "1997"=V9, "1998"=V10, "1999"=V11, "2000"=V12, "2001"=V13, "2002"=V14, "2003"=V15, "2004"=V16, "2005"=V17, "2006"=V18, "2007"=V19, "2008"=V20, "2009"=V21, "2010"=V22, "2011"=V23, "2012"=V24, "2013"=V25, "2014"=V26, "2015"=V27, "2016"=V28) %>% 
  filter(state != "Commercial",
         state != "Industrial", 
         state != "Residential",
         state != "Transportation",
         state != "Electric Power")

#page 4
out4 <- out[[4]] %>%  
   as.data.frame() %>% 
  rename(state = V1, "1990" = V2, "1991"= V3, "1992"= V4, "1993"=V5, "1994"=V6, "1995"=V7, "1996"=V8, "1997"=V9, "1998"=V10, "1999"=V11, "2000"=V12, "2001"=V13, "2002"=V14, "2003"=V15, "2004"=V16, "2005"=V17, "2006"=V18, "2007"=V19, "2008"=V20, "2009"=V21, "2010"=V22, "2011"=V23, "2012"=V24, "2013"=V25, "2014"=V26, "2015"=V27, "2016"=V28) %>% 
  filter(state != "Commercial",
         state != "Industrial", 
         state != "Residential",
         state != "Transportation",
         state != "Electric Power")

#page 5
out5 <- out[[5]] %>%  
     as.data.frame() %>% 
  rename(state = V1, "1990" = V2, "1991"= V3, "1992"= V4, "1993"=V5, "1994"=V6, "1995"=V7, "1996"=V8, "1997"=V9, "1998"=V10, "1999"=V11, "2000"=V12, "2001"=V13, "2002"=V14, "2003"=V15, "2004"=V16, "2005"=V17, "2006"=V18, "2007"=V19, "2008"=V20, "2009"=V21, "2010"=V22, "2011"=V23, "2012"=V24, "2013"=V25, "2014"=V26, "2015"=V27, "2016"=V28) %>% 
  filter(state != "Commercial",
         state != "Industrial", 
         state != "Residential",
         state != "Transportation",
         state != "Electric Power")
```


Combining and Tidying
```{r}
emissions <- rbind(out1, out2, out3, out4, out5) %>% 
  gather(year, emission, 2:28) %>% 
   filter(state == "Maine" | state == "New Hampshire" | state == "Massachusetts" | state == "New Hampshire" | state == "Rhode Island" | state == "Connecticut" | state == "New York") %>% 
  mutate(year = as.numeric(year),
         emission = as.numeric(emission)) %>% 
  mutate(end_target = case_when(state == "Maine" ~ (23.64 - 0.8*(23.64)), #calulate state emission targets
                            state == "New Hampshire" ~ (14.59 - 0.8*(14.59)),
                            state == "Connecticut" ~ (41.53 - 0.8*(41.53)),
                            state == "Massachusetts" ~ (83.09 - 0.8*(83.09)),
                            state == "New York" ~ (207.82 - 1*(207.82)),
                            state == "Rhode Island" ~ (8.91 - 0.8*(8.91))))

#save
write.csv(emissions, file = "data/emissions.csv")
```

Visualize emmissions
```{r}

graph_em <-ggplot(emissions)+
  geom_line(aes(x=year, y = emission, color=state))+
  ggtitle("Total Carbon Emissions") +
  ylab("Emissions (million metric tons)") +
  xlab("Year") +
  theme_classic() 

graph_em 
ggsave("figs/graph_em .jpg", width=7, height=5, dpi=300)

```

Regressions for trend targets

New Hampshire
```{r}
#NH 80% reduction below 1990 levels by 2050
nh_em <- emissions %>% 
  filter(state == "New Hampshire") %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  mutate(year_n = seq(1:13))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
nh_em_target <- tribble(
  ~year, ~emission,
  1, 	21.8	,
  46, 2.92)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_nh_em <- lm(emission~ year, data = nh_em_target)
#coe =   -0.42   ; int= 22.22

nh_em_target <- nh_em %>% 
  mutate(target = (  -0.42  *year_n +22.22),
         emission = round(emission, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, emission, target)
```

Massachusetts
```{r}
#MA 80% reduction below 1990 levels by 2050
ma_em <- emissions %>% 
  filter(state == "Massachusetts") %>% 
  filter(year >= 2004 & year <= 2017)%>% 
  mutate(year_n = seq(1:13))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ma_em_target <- tribble(
  ~year, ~emission,
  1, 82.6	,
  46, 16.6)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ma_em <- lm(emission~ year, data = ma_em_target)
#coe=--1.47   ; int =  84.07 

ma_em_target <- ma_em %>% 
  mutate(target = (-1.47*year_n +  84.07  ),
         emission = round(emission, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, emission, target)
```

Rhode Island
```{r}
#RI 80% reduction below 1990 levels by 2050
ri_em <- emissions %>% 
  filter(state == "Rhode Island") %>% 
  filter(year >= 2004 & year <= 2017)%>% 
  mutate(year_n = seq(1:13))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ri_em_target <- tribble(
  ~year, ~emission,
  1, 10.97,
  46, 1.78
)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ri_em <- lm(emission~ year, data = ri_em_target)
#coe= -0.204  ; int = 11.174 

ri_em_target<- ri_em %>% 
    mutate(target = (-0.204 *year_n+11.174 ),
         emission = round(emission, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, emission, target)
```

Connecticut
```{r}
#CT 80% reduction below 2001 levels by 2050
ct_em <- emissions %>% 
  filter(state == "Connecticut") %>% 
  filter(year >= 2004 & year <= 2017)%>% 
  mutate(year_n = seq(1:13))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ct_em_target <- tribble(
  ~year, ~emission,
  1, 44.5	,
  46, 8.31)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ct_em <- lm(emission~ year, data = ct_em_target)
#coef=-0.804 ; int = 45.304 

ct_em_target <- ct_em %>% 
  mutate(target = (-0.804 *year_n+45.304),
         emission = round(emission, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, emission, target)
```

New York
```{r}
#NY 100% reduction below 1990 levels by 2050
ny_em <- emissions %>% 
  filter(state == "New York") %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  mutate(year_n = seq(1:13))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ny_em_target <- tribble(
  ~year, ~emission,
  1, 215,
  46, 0)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ny_em <- lm(emission~ year, data = ny_em_target)
#coe= -4.78  ; int = 215.00  

ny_em_target <- ny_em %>% 
  mutate(target = (-4.78  *year_n+219.78),
         emission = round(emission, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, emission, target)
```

Maine
```{r}
#MA 80% reduction below 1990 levels by 2050
me_em <- emissions %>% 
  filter(state == "Maine") %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  mutate(year_n = seq(1:13))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression
me_em_target <- tribble(
  ~year, ~emission,
  1, 24.0,
  46, 4.73)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_me_em <- lm(emission~ year, data = me_em_target)
#coe= -0.428; int = 24.428   

me_em_target <- me_em %>% 
  mutate(target = (-0.428*year_n+24.428),
         emission = round(emission, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, emission, target)
```

Combine all states df with target
```{r}
emissions_target <- rbind(nh_em_target, ma_em_target, ri_em_target, ct_em_target, ny_em_target, me_em_target)
```

Calculate scores
```{r}
carbon_em <- emissions_target %>% 
  mutate(score = target/emission,
         score = if_else(score >1, 1, score)) 
#save
write.csv(carbon_em, file = "data/carbon_em.csv")
```

Visualize scores
```{r}  
carbon_em_graph <-ggplot(carbon_em)+
  geom_line(aes(x=year, y = score, color=state))+
  ggtitle("Carbon emission scores") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 
carbon_em_graph
ggsave("figs/carbon_em_graph.jpg", width=7, height=5, dpi=300)
```
