---
title: "OHIEC: Resilience - Climate Friendly Actions Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the scope of climate friendly actions by region

# Data
**America's Pledge**

**Downloaded**: August, 2019

**Description**: Got the raw data for Figure 2-3 Climate Friendly Actions by U.S States: Types and Numbers in the America's Pledge Report: Phase 1 (https://www.bbhub.io/dotorg/sites/28/2017/11/AmericasPledgePhaseOneReportWeb.pdf) from Tom Cyrs tom.cyrs@wri.org 

**Time range**: 2006-2018

**Format**: Tabular

# About
There are 33 different categories of environmental friendly actions each state can make. A state will get a perfect score if they have an action in each of these categories. Here I removed actions that were energy efficiency related because that information is more succinctly summarized in the ACEE data.  
 
There are a total of 30 possible actions categories for this data layer.
Set target as 30. A state would need to have taken some sort of action in each of the following categories to get 100%

Categories:

1. Integrating transportation & land-use in comprehensive plans	
2. Dedicated funding streams for public transit
3. Adopting legislation in line with Complete Streets objectives
4. Financial incentives for high efficiency vehicles
5. Conservation easement tax credits
6. California’s vehicle emission standards
7. Rules and incentives to reduce food waste
8. Most recent building energy codes
9. Appliance and equipment energy efficiency standards
10. Carbon pricing	
11. Zero Emission Vehicle mandate
12. Cost-sharing programs to improve forest systems
13. Wildfire protection incentives
14. Coal mine methane standards
15. Low carbon fuel standard
16. Methane standards for existing oil and natural gas facilities
17. Setting methane emission reduction targets
18. Freight-specific energy efficiency performance metrics
19. HFC management program requiring all leaks to be repaired
20. Financial incentives for CCS
21. Zero-emission credits for nuclear
22. GHG Emission Targets
23. Renewable energy portfolio standards or goals
24. Property Assessed Clean Energy
25. Combined heat and power financing and incentives
26. Landfill gas energy project bond, grant, loan, or rebate programs
27. Energy efficiency resource standard or goals
28. Property tax programs to support sustainable forests
29. Efficient vehicle requirement for public fleet procurement
30. Freight plan  with multimodal freight strategies

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(readxl)
```

Load Data
```{r load_data}
raw_clim_act <- read_excel(file.path(dir_anx, "_raw_data/Americas_Pledge/states_climate_regs_summary.xlsx"), skip=2, sheet = 2)
```

## Calculate scores
```{r calculate_climate_actions_score}
climate_actions <- raw_clim_act %>% 
  filter(State == "Maine" | State == "Massachusetts" | State == "New Hampshire" | State == "Rhode Island" | State == "Connecticut" | State == "New York" | State == "Total # of actions considered") %>% 
  select(-"Energy efficiency", -"Total") %>% 
  mutate(target = 23,
         total = rowSums(.[2:9]),
         score = total/target) %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  filter(State != "Total # of actions considered")

write.csv(climate_actions, file = "data/climate_actions.csv")
```

## Graph scores
```{r graph_climate_action_scores}
climate_actions_graph <-ggplot(climate_actions)+
  geom_line(aes(x=year, y = score, color=State))+
  ggtitle("Climate Friendly Actions Score") +
  ylab("Score") +
  xlab("State") +
  theme_classic() 
climate_actions_graph

ggsave("figs/climate_actions_graph.jpg", width=7, height=5, dpi=300)
```



