---
title: "OHIEC: Resilience - Climate Change Pressure Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates climate change resilience values by OHINE region. 

**3. EIA Consumption Data**
**Downloaded**: August, 2019

**Description**: Energy consumption by energy source.  https://www.eia.gov/beta/states/states/ny/data/dashboard/total-energy

**Time range**: 1960-2017

**Format**: Tabular


2. Take the percent of clean power consumption each year and compare to the target, which is calculated by the state set goals
a. Renewable/Clean Energy Targets:
Drawn from http://www.ncsl.org/research/energy/renewable-portfolio-standards.aspx#me
note: all targets are part of "renewable energy portfolio standards" so I won't be including nuclear energy into calculating if states are on track for targets


##Clean Energy Cosumption

1. Maine = 40% by 2017
2. New Hampshire = 25.2% by 2025
3. Connecticut = 44% by 2030
4. Massachusetts = 35% 2030
5. New York = 50% by 2030
6. Rhode Island = 38.5% by 2035


``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(Hmisc)
library(stringi)
library(stringr)
library(readxl)
library(tabulizer)
```
Load in the energy consumption data for each state. All numbers are in Billion Btu
Did not include the energy imported beacuse there is no way to know what source that came from

```{r}
ct_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ct_consumption.csv"), skip = 1)
me_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/me_consumption.csv"), skip = 1)
ma_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ma_consumption.csv"), skip = 1)
ri_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ri_consumption.csv"), skip = 1)
nh_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/nh_consumption.csv"), skip = 1)
ny_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ny_consumption.csv"), skip = 1)
```

Connecticut
```{r}
ct_con <- ct_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Connecticut",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Connecticut",
         nuc = "Nuclear energy consumed for electricity generation, total, Connecticut",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Connecticut",
         renew = "Renewable energy total consumption, Connecticut",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Connecticut",
         net_elec_import = "Net imports of electricity into the United States, Connecticut") %>% 
  mutate(state = "Connecticut") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Maine
```{r}
me_con <- me_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Maine",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Maine",
         nuc = "Nuclear energy consumed for electricity generation, total, Maine",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Maine",
         renew = "Renewable energy total consumption, Maine",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Maine",
         net_elec_import = "Net imports of electricity into the United States, Maine") %>% 
  mutate(state = "Maine") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Massachusetts
```{r}
ma_con <- ma_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Massachusetts",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Massachusetts",
         nuc = "Nuclear energy consumed for electricity generation, total, Massachusetts",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Massachusetts",
         renew = "Renewable energy total consumption, Massachusetts",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Massachusetts",
         net_elec_import = "Net imports of electricity into the United States, Massachusetts") %>% 
  mutate(state = "Massachusetts") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Rhode Island
```{r}
ri_con <- ri_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Rhode Island",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Rhode Island",
         nuc = "Nuclear energy consumed for electricity generation, total, Rhode Island",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Rhode Island",
         renew = "Renewable energy total consumption, Rhode Island",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Rhode Island",
         net_elec_import = "Net imports of electricity into the United States, Rhode Island") %>% 
  mutate(state = "Rhode Island") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

New Hampshire
```{r}
nh_con <- nh_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, New Hampshire",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), New Hampshire",
         nuc = "Nuclear energy consumed for electricity generation, total, New Hampshire",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, New Hampshire",
         renew = "Renewable energy total consumption, New Hampshire",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), New Hampshire",
         net_elec_import = "Net imports of electricity into the United States, New Hampshire") %>% 
  mutate(state = "New Hampshire") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

New York
```{r}
ny_con <- ny_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, New York",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), New York",
         nuc = "Nuclear energy consumed for electricity generation, total, New York",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, New York",
         renew = "Renewable energy total consumption, New York",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), New York",
         net_elec_import = "Net imports of electricity into the United States, New York") %>% 
  mutate(state = "New York") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Combine all states energy consumption data into one and calculate the percent of renewable. Also calculating percent with clean(ie: nuclear) for fun/its interesting
```{r}
consum <- rbind(me_con, ma_con, ct_con, nh_con, ny_con, ri_con) %>% 
  mutate(coal = as.numeric(coal),
         nat_gas = as.numeric(nat_gas),
         nuc = as.numeric(nuc),
         petrol = as.numeric(petrol),
         renew = as.numeric(renew)) %>% 
  rowwise() %>% 
  mutate(total = sum(coal, nat_gas, nuc, petrol, renew),
         non_renew = sum(coal, nat_gas, nuc, petrol),
         per_renew = renew/total,
         year = as.numeric(year),
         clean = sum(nuc, renew),
         per_clean= clean/total) %>% 
  ungroup()
```

Visualize
```{r}
# renewable only
con_renew_graph <- ggplot(consum)+
  geom_line(aes(x=year, y= per_renew, color= state))+
 ggtitle("Percent of renewable energy used (no nuclear)") +
  ylab("Percent") +
  xlab("Year") +
  theme_classic()

con_renew_graph

#clean(nuclear) and renewable
con_clean_graph <- ggplot(consum)+
  geom_line(aes(x=year, y= per_clean, color= state))+
 ggtitle("Percent of clean/renewable energy used") +
  ylab("Percent") +
  xlab("Year") +
  theme_classic()

con_clean_graph
```

Now that we have the actual percentage of energy consumption that comes from renewable, we can do linear regressions using the state set targets as the end points. With this we can calculate if each state is on track to hitting their renewable energy target goal for a given year.

Maine
```{r}
#MA 40% clean by 2017
me <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Maine") %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
me_target <- tribble(
  ~year, ~per_clean,
  1, 0.277,
  14, 0.4
)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_me <- lm(per_clean ~ year, data = me_target)
#intercept = 0.26754 ; coef = 0.00946 
summary(lm_me)

plot(per_clean ~ year, data = me_target)
abline(lm_me)

###trying to use equation no loop
me_clean_target <- me %>% 
  rowwise() %>% 
  mutate(target = (0.00946*year_n + 0.26754),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter( year != "2004") %>% 
  select(year, state, per_renew, target) 
```

Massachusetts
```{r}
#MA 35% clean by 2030
ma <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Massachusetts")  %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ma_target <- tribble(
  ~year, ~per_clean,
  1, 		0.0387	,
  26, .35)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ma <- lm(per_clean ~ year, data = ma_target)
#coef = 0.0125    ; int = 0.0262  

ma_clean_target <- ma %>% 
  rowwise() %>% 
  mutate(target = (0.0125*year_n+ 0.0262),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter( year != "2004") %>% 
  select(year, state, per_renew, target)
```

New Hampshire
```{r}
#NH 25.2% clean by 2025
nh <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "New Hampshire") %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
nh_target <- tribble(
  ~year, ~per_clean,
  1, 0.0770,
  21, .252
)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_nh <- lm(per_clean ~ year, data = nh_target)
#coef=  0.00875 int=  0.06825 

nh_clean_target <- nh %>% 
  mutate(target = (0.00875*year_n + 0.06825),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
   filter( year != "2004") %>% 
  select(year, state, per_renew, target)
```


Rhode Island
```{r}
#RI 38.5% clean by 2035
ri <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Rhode Island") %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ri_target <- tribble(
  ~year, ~per_clean,
  0, 0.0256	,
  31, .385)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ri <- lm(per_clean ~ year, data = ri_target)
#coef=0.0256  , int = 0.0116  

ri_clean_target <- ri %>% 
   mutate(target = (0.0116*year_n+0.0256),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_renew, target)
```

New York
```{r}
#NY 50% clean by 2030
ny <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "New York") %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ny_target <- tribble(
  ~year, ~per_clean,
  0, 	0.0950	,
  26, .50)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ny <- lm(per_clean ~ year, data = ny_target)
#coef=0.0156   , int =   0.0950

ny_clean_target <- ny %>% 
  mutate(target = (0.0156*year_n+0.0950 ),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_renew, target)
```


Connecticut
```{r}
#CT 44% clean by 2030
ct <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Connecticut")  %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ct_target <- tribble(
  ~year, ~per_clean,
  0, 0.0490,
  26, .44
)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ct <- lm(per_clean ~ year, data = ct_target)
#coef= 0.015  , int= 0.049 

ct_clean_target <- ct %>% 
  mutate(target = (0.015*year_n+0.049),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_renew, target)
```

Combine all states df with target
```{r}
energy_target <- rbind(me_clean_target, nh_clean_target, ma_clean_target, ri_clean_target, ct_clean_target, ny_clean_target)
```


Calculate scores
```{r}
renew_consum <- energy_target %>% 
  mutate(score = per_renew/target,
         score = if_else(score >1, 1, score)) 
#save
write.csv(renew_consum, file = "data/renew_consum.csv")
```

Visualize scores
```{r}  
renew_consum_graph <-ggplot(renew_consum)+
  geom_line(aes(x=year, y = score, color=state))+
  ggtitle("Clean Energy Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 
renew_consum_graph

ggsave("figs/renew_consum_graph.jpg", width=7, height=5, dpi=300)
```