---
title: "OHINE: Climate Change Resilience - Clean Energy Consumption Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates how close each region is meeting their annual target necessary to reach their end clean energy consumption goal.This data is used to measure climate change resilience implementation & enforcement and effectiveness & compliance.

# Data

**Source:** US Energy Information Administration (EIA) Consumption Data

**Downloaded:** August, 2019

**Description:** Energy consumption by energy source.  https://www.eia.gov/beta/states/states/ny/data/dashboard/total-energy

**Time range:** 1960-2017

**Format:** Tabular, csv file

# About
Scores are calculated by taking the EIA energy consumption data, calculating the percent of energy consumed that comes from clean sources, and dividing it by the target emissions. Total annual carbon emissions were pulled from the EIA website. Clean energy consumption targets are calculated by performing a linear regression from 2004 to the target year of their consumption goal. While there are a variety of energy plans and portfolios in the US that differ from state to state, every state in our region of interest has a "renewable energy portfolio standards", so we do not include nuclear energy production as part of the total renewable energy consumption target.

Clean Energy Consumption Targets:
http://www.ncsl.org/research/energy/renewable-portfolio-standards.aspx#me

1. Maine = 40% by 2017
2. New Hampshire = 25.2% by 2025
3. Connecticut = 44% by 2030
4. Massachusetts = 35% 2030
5. New York = 50% by 2030
6. Rhode Island = 38.5% by 2035

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(Hmisc)
library(stringi)
library(stringr)
library(readxl)
library(tabulizer)
```

Load in the energy consumption data for each state. Consumption is in billion Btu. We do not include the energy imported because since there we do not know the sources.
```{r load_data}
ct_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ct_consumption.csv"), skip = 1)
me_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/me_consumption.csv"), skip = 1)
ma_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ma_consumption.csv"), skip = 1)
ri_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ri_consumption.csv"), skip = 1)
nh_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/nh_consumption.csv"), skip = 1)
ny_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ny_consumption.csv"), skip = 1)
```



## Calculate consumption
Now we need to calculate the total energy consumption and the total clean energy consumption by each state each year.

Connecticut
```{r ct_calc_total}
ct_con <- ct_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Connecticut",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Connecticut",
         nuc = "Nuclear energy consumed for electricity generation, total, Connecticut",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Connecticut",
         renew = "Renewable energy total consumption, Connecticut",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Connecticut",
         net_elec_import = "Net imports of electricity into the United States, Connecticut") %>% 
  mutate(state = "Connecticut") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Maine
```{r me_calc_total}
me_con <- me_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Maine",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Maine",
         nuc = "Nuclear energy consumed for electricity generation, total, Maine",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Maine",
         renew = "Renewable energy total consumption, Maine",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Maine",
         net_elec_import = "Net imports of electricity into the United States, Maine") %>% 
  mutate(state = "Maine") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Massachusetts
```{r ma_calc_total}
ma_con <- ma_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Massachusetts",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Massachusetts",
         nuc = "Nuclear energy consumed for electricity generation, total, Massachusetts",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Massachusetts",
         renew = "Renewable energy total consumption, Massachusetts",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Massachusetts",
         net_elec_import = "Net imports of electricity into the United States, Massachusetts") %>% 
  mutate(state = "Massachusetts") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Rhode Island
```{r ri_calc_total}
ri_con <- ri_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Rhode Island",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Rhode Island",
         nuc = "Nuclear energy consumed for electricity generation, total, Rhode Island",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Rhode Island",
         renew = "Renewable energy total consumption, Rhode Island",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Rhode Island",
         net_elec_import = "Net imports of electricity into the United States, Rhode Island") %>% 
  mutate(state = "Rhode Island") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

New Hampshire
```{r nh_calc_total}
nh_con <- nh_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, New Hampshire",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), New Hampshire",
         nuc = "Nuclear energy consumed for electricity generation, total, New Hampshire",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, New Hampshire",
         renew = "Renewable energy total consumption, New Hampshire",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), New Hampshire",
         net_elec_import = "Net imports of electricity into the United States, New Hampshire") %>% 
  mutate(state = "New Hampshire") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

New York
```{r ny_calc_total}
ny_con <- ny_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, New York",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), New York",
         nuc = "Nuclear energy consumed for electricity generation, total, New York",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, New York",
         renew = "Renewable energy total consumption, New York",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), New York",
         net_elec_import = "Net imports of electricity into the United States, New York") %>% 
  mutate(state = "New York") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Combine all states energy consumption data and calculate the percent of renewable. 
```{r calc_percent_consum_clean}
consum <- rbind(me_con, ma_con, ct_con, nh_con, ny_con, ri_con) %>% 
  mutate(coal = as.numeric(coal),
         nat_gas = as.numeric(nat_gas),
         nuc = as.numeric(nuc),
         petrol = as.numeric(petrol),
         renew = as.numeric(renew)) %>% 
  rowwise() %>% 
  mutate(total = sum(coal, nat_gas, nuc, petrol, renew),
         non_renew = sum(coal, nat_gas, nuc, petrol),
         per_renew = renew/total,
         year = as.numeric(year),
         clean = sum(nuc, renew),
         per_clean= clean/total) %>% 
  ungroup()
```

Visualize
```{r visualize_consumption_clean}
con_renew_graph <- ggplot(consum)+
  geom_line(aes(x=year, y= per_renew, color= state))+
 ggtitle("Percent of Renewable Energy Used") +
  ylab("Percent") +
  xlab("Year") +
  theme_classic()

con_renew_graph
```

## Regressions for trend targets

Now that we have the actual percentage of energy consumption that comes from renewable, we can conduct regressions for each state to determine the annual target needed to meet end target. This is done by creating a state specific df with the start (observed in 2004) and end (goal year) carbon emissions, and then using the slope and intercept to calculate the annual targets. With this we can calculate if each state is on track to hitting their renewable energy target goal for a given year.

Maine

Target = 40% clean by 2017
```{r me_lin_reg}
me <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Maine") %>% 
  mutate(year_n = seq(1:14))

#create tribble with the start (value in 2004) and end (goal in whichever year) to do a linear regression 
me_target <- tribble(
  ~year, ~per_clean,
  1, 0.277,
  14, 0.4
)

# simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_me <- lm(per_clean ~ year, data = me_target)

b <- coef(lm_me)["(Intercept)"]
m <- coef <- coef(lm_me)["year"]

me_clean_target <- me %>% 
  rowwise() %>% 
  mutate(target = (m*year_n + b),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter( year != "2004") %>% 
  select(year, state, per_renew, target) 
```

Massachusetts

Target = 35% clean by 2030
```{r ma_lin_reg}
ma <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Massachusetts")  %>% 
  mutate(year_n = seq(1:14))

#create tribble with the start (value in 2004) and end (goal in whichever year) to do a linear regression 
ma_target <- tribble(
  ~year, ~per_clean,
  1, 		0.0387	,
  26, .35)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ma <- lm(per_clean ~ year, data = ma_target)

b <- coef(lm_ma)["(Intercept)"]
m <- coef <- coef(lm_ma)["year"]

ma_clean_target <- ma %>% 
  rowwise() %>% 
  mutate(target = (m*year_n+ b),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter( year != "2004") %>% 
  select(year, state, per_renew, target)
```

New Hampshire

Target = 25.2% clean by 2025
```{r nh_lin_reg}
nh <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "New Hampshire") %>% 
  mutate(year_n = seq(1:14))

#create tribble with the start (value in 2004) and end (goal in whichever year) to do a linear regression 
nh_target <- tribble(
  ~year, ~per_clean,
  1, 0.0770,
  21, .252
)

# simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_nh <- lm(per_clean ~ year, data = nh_target)

b <- coef(lm_nh)["(Intercept)"]
m <- coef <- coef(lm_nh)["year"]

nh_clean_target <- nh %>% 
  mutate(target = (m*year_n + b),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
   filter( year != "2004") %>% 
  select(year, state, per_renew, target)
```

Rhode Island

Target = 38.5% clean by 2035
```{r ri_lin_reg}
ri <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Rhode Island") %>% 
  mutate(year_n = seq(1:14))

#create tribble with the start (value in 2004) and end (goal in whichever year) to do a linear regression 
ri_target <- tribble(
  ~year, ~per_clean,
  0, 0.0256	,
  31, .385)

# simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ri <- lm(per_clean ~ year, data = ri_target)

b <- coef(lm_ri)["(Intercept)"]
m <- coef <- coef(lm_ri)["year"]

ri_clean_target <- ri %>% 
   mutate(target = (m*year_n+b),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_renew, target)
```

New York

Target = 50% clean by 2030
```{r ny_lin_reg}
ny <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "New York") %>% 
  mutate(year_n = seq(1:14))

#create tribble with the start (value in 2004) and end (goal in whichever year) to do a linear regression 
ny_target <- tribble(
  ~year, ~per_clean,
  0, 	0.0950	,
  26, .50)

# simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ny <- lm(per_clean ~ year, data = ny_target)

b <- coef(lm_ny)["(Intercept)"]
m <- coef <- coef(lm_ny)["year"]

ny_clean_target <- ny %>% 
  mutate(target = (m*year_n+b ),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_renew, target)
```

Connecticut

Target = 44% clean by 2030
```{r ct_lin_reg}
ct <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Connecticut")  %>% 
  mutate(year_n = seq(1:14))

#create tribble with the start (value in 2004) and end (goal in whichever year) to do a linear regression 
ct_target <- tribble(
  ~year, ~per_clean,
  0, 0.0490,
  26, .44
)

# simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ct <- lm(per_clean ~ year, data = ct_target)

b <- coef(lm_ct)["(Intercept)"]
m <- coef <- coef(lm_ct)["year"]

ct_clean_target <- ct %>% 
  mutate(target = (m*year_n+b),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_renew, target)
```

Combine all states df with target
```{r combine_state_targets}
energy_target <- rbind(me_clean_target, nh_clean_target, ma_clean_target, ri_clean_target, ct_clean_target, ny_clean_target)
```

# Results
## Calculate score
```{r calc_scores}
renew_consum <- energy_target %>% 
  mutate(score = per_renew/target,
         score = if_else(score >1, 1, score)) 
#save
write.csv(renew_consum, file = "data/renew_consum.csv")
```

## Visualize scores
```{r vis_scores}  
renew_consum_graph <-ggplot(renew_consum)+
  geom_line(aes(x=year, y = score, color=state))+
  ggtitle("Clean Energy Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 
renew_consum_graph

ggsave("figs/renew_consum_graph.jpg", width=7, height=5, dpi=300)
```