---
title: "OHIEC: Resilience - Climate Change Pressure Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates climate change resilience values by OHINE region. 

# Data
1. American Council for Energy-Efficient Economy (EEE) state and local policy rankings
2. America's Pledge Phase 1 Report
3. US Energy Information Administration Energy Consumption Data
4. EPA CO2 Emissions from Fossil Fuel Combustion Data

**1. EEE**
**Downloaded**: August, 2019

**Description**: Rankings of states' energy efficiency policy and program efforts. https://database.aceee.org/state-scorecard-rank

**Time range**: 2006-2018

**Format**: PDF


**2. America's Pledge**
**Downloaded**: August, 2019

**Description**: Got the raw data for Figure 2-3 Climate Friendly Actions by U.S States: Types and Numbers in the America's Pledge Report: Phase 1 (https://www.bbhub.io/dotorg/sites/28/2017/11/AmericasPledgePhaseOneReportWeb.pdf) from Tom Cyrs tom.cyrs@wri.org 

**Time range**: 2006-2018

**Format**: Tabular


**3. EIA Consumption Data**
**Downloaded**: August, 2019

**Description**: Energy consumption by energy source.  https://www.eia.gov/beta/states/states/ny/data/dashboard/total-energy

**Time range**: 1960-2017

**Format**: Tabular


**4. EPA Carbon Emissions**
**Downloaded**: August, 2019

**Description**: Carbon emissions from each state from fossil fuels. 'https://www.epa.gov/sites/production/files/2017-09/documents/co2ffc_2015.pdf'

**Time range**: 1990-2016

**Format**: Tabular

##Using the Data
Regulations:
1. America's Pledge Report: There are 33 differnet cateogies of environmental friendly actions each state can make. A state will get a perfect score if they have an action in each of these categories 
2. American Council for Energy Efficiency Rankings: Each state is scored out of 50 (except in 2006 it's out of 44) on their policies in place for energy efficiency etc. 50 would be a perfect score

Effectiveness:
1. Take the total CO2 emissions each year and compare to the target, which is calculated by the state set goals
a. Carbon Emission Targets:
Targets pulled from https://www.c2es.org/document/greenhouse-gas-emissions-targets/
Rhode Island is missing from this map so pulled from http://www.planning.ri.gov/documents/LU/energy/energy15.pdf

2. Take the percent of clean power consumption each year and compare to the target, which is calculated by the state set goals
a. Renewable/Clean Energy Targets:
Drawn from http://www.ncsl.org/research/energy/renewable-portfolio-standards.aspx#me
note: all targets are part of "renewable energy portfolio standards" so I won't be including nuclear energy into calculating if states are on track for targets

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(Hmisc)
library(stringi)
library(stringr)
library(readxl)
library(tabulizer)
```


Calculate Climate Change Resilience Score

Tidy each piece so can combine
```{r}
##eee scores
eee_score_t <- eee_score %>% 
  rename(rgn_name = state) %>% 
  mutate(rgn_id = ifelse(rgn_name == "Connecticut", 5,
                           ifelse(rgn_name == "Maine", 6,
                                  ifelse(rgn_name == "Massachusetts-North", 7 ,
                                         ifelse(rgn_name == "Massachusetts-South", 8,
                                                ifelse(rgn_name == "New Hampshire", 9,
                                                       ifelse(rgn_name == "New York", 10,
                                                              11))))))) %>% 
  select(year, rgn_name, rgn_id, score_eee) %>% 
  mutate(rgn_id= as.character(rgn_id))

##emmissions scores
#create a seperate df to bind to the main one for year 2017 beacuse no data for that
em_2017<- tribble(
  ~year, ~rgn_name, ~rgn_id,
  2017, "Maine", 6,  
  2017, "New Hampshire", 9,  
  2017, "Massachusetts-North", 7, 
  2017, "Rhode Island", 11,  
  2017, "Connecticut", 5,  
  2017, "New York", 10,  
  2017, "Massachusetts-South", 8
  ) %>% 
  as.data.frame() %>% 
  add_column(score= NA)

emissions_scores_t <- emissions_scores %>% 
  mutate(rgn_id1 = case_when(state == "Maine" ~ 6,
                            state == "New Hampshire" ~ 9,
                            state == "Massachusetts" ~ 7,
                            state == "Rhode Island" ~ 11,
                            state == "Connecticut" ~ 5,
                            state == "New York" ~ 10),
         rgn_id8 = case_when(state == "Massachusetts" ~ 8)) %>% 
  rename(rgn_name = state) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id8, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(year, rgn_name, rgn_id, score) %>% 
  mutate(rgn_name = ifelse(rgn_id == 1, "Offshore", 
                       ifelse(rgn_id==2, "Georges Bank",
                               ifelse(rgn_id==3, "Gulf of Maine",
                                      ifelse(rgn_id==4, "Mid-Atlantic Bight",
                                             ifelse(rgn_id==5, "Connecticut",
                                                    ifelse(rgn_id==6, "Maine",
                                                           ifelse(rgn_id==7, "Massachusetts-North",
                                                                  ifelse(rgn_id==8, "Massachusetts-South",
                                                                         ifelse(rgn_id==9, "New Hampshire",
                                                                                ifelse(rgn_id==10, "New York",
                                                                                       "Rhode Island")))))))))))%>% 
  rbind(em_2017) %>% 
  arrange(rgn_name) %>% 
  fill(score, .direction = "down") %>% 
  rename(score_em = score) 

##power consumption scores
energy_consump_scores_t <- energy_consump_scores %>% 
  mutate(rgn_id1 = case_when(
    state == "Maine" ~ "6", 
    state == "New Hampshire" ~ "9", 
    state == "Massachusetts" ~ "7",
    state == "Rhode Island" ~ "11",
    state == "Connecticut" ~ "5",
    state == "New York" ~ "10"),
    rgn_id2 = case_when(
    state == "Massachusetts" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  rename(rgn_name = "state") %>% 
  select(-filler) %>% 
  select(year, rgn_name, rgn_id, score) %>% 
  mutate(rgn_name = ifelse(rgn_name == "Massachusetts" & rgn_id == 7, "Massachusetts-North",
                           ifelse(rgn_name == "Massachusetts" & rgn_id == 8, "Massachusetts-South", rgn_name))) %>% 
  rename(score_power = score) 

##climate actions scores
clim_act_t <- clim_act %>% 
  select(year, State, score) %>% 
  rename(state=State) %>% 
  mutate(rgn_id1 = case_when(state == "Maine" ~ 6,
                            state == "New Hampshire" ~ 9,
                            state == "Massachusetts" ~ 7,
                            state == "Rhode Island" ~ 11,
                            state == "Connecticut" ~ 5,
                            state == "New York" ~ 10),
         rgn_id8 = case_when(state == "Massachusetts" ~ 8)) %>% 
  rename(rgn_name = state) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id8, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(year, rgn_name, rgn_id, score) %>% 
   mutate(rgn_name = ifelse(rgn_name == "Massachusetts" & rgn_id == 7, "Massachusetts-North",
                           ifelse(rgn_name == "Massachusetts" & rgn_id == 8, "Massachusetts-South", rgn_name))) %>% 
  rename(score_act = score) 
```

Creating a df to add on to the water pollution scores to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  3, "Gulf of Maine",
  2, "Georges Bank",
  4, "Mid-Atlantic Bight",
  1, "Offshore") 

miss_rgns <- miss_rgns %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(score = NA) %>% 
  unique()
```

Combine all scores!
```{r}
cc_res_score <- left_join(emissions_scores_t, clim_act_t, by = c("year", "rgn_name", "rgn_id")) %>% 
  left_join(energy_consump_scores_t, by = c("year", "rgn_name", "rgn_id")) %>% 
  left_join(eee_score_t, by = c("year", "rgn_name", "rgn_id"))%>% 
  rowwise() %>% 
  mutate(score = mean(c(score_power, score_act, score_em, score_eee))) %>% 
  select(year, rgn_name, rgn_id, score) %>% 
  rbind(miss_rgns)

write.csv(cc_res_score, file = "data/cc_res_score.csv")
```

Visualize
```{r}
cc_res_graph <- ggplot(cc_res_score)+
  geom_line(aes(x=year, y= score, color= rgn_name))+
 ggtitle("Climate Change Resilience Scores") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

cc_res_graph

##save fig
ggsave("figs/cc_res_graph.jpg", width=7, height=5, dpi=300)
```


Write as a pressure layer to ne-scores
```{r}
write.csv(cc_res_score, file.path(dir_calc, "layers/res_climate_change.csv"))
```















