---
title: "OHINE: Resilience - Climate Change Resilience Layer Calculations"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates climate change resilience values by OHINE region. 

The data layers are used to quantify existence of regulations, implementation + enforcement, and effectiveness + compliance. These three parts are then used to calculate one overall water pollution resilience score for each region. 

1. Regulations: acee_rankings.csv, climate_actions.csv
2. Implementation + Enforcement: renew_consum.csv, carbon_em.csv
3. Effectiveness + Compliance: renew_consum.csv, carbon_em.csv

Data Time Ranges:

acee_rankings: 2005-2017

climate_actions: NA (year column done in the climate actions markdown)

renew_consum:2005-2017

carbon_em: 2005-2016

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(Hmisc)
library(stringi)
library(stringr)
library(readxl)
library(tabulizer)
```

Read in Data
```{r read_data}
acee_rankings_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/climate_change/data/acee_rankings.csv") %>% 
  select(-X) 

climate_actions_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/climate_change/data/climate_actions.csv") %>% 
  select(-X) 

renew_consum_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/climate_change/data/renew_consum.csv") %>% 
  select(-X)

carbon_em_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/climate_change/data/carbon_em.csv") %>% 
  select(-X)
```

## Tidy data for merging

Creating a df to add on to the water pollution scores to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  3, "Gulf of Maine",
  2, "Georges Bank",
  4, "Mid-Atlantic Bight",
  1, "Offshore") 

miss_rgns <- miss_rgns %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(score = NA) %>% 
  unique()
```

ACEE 
```{r acee_tidy}
acee_rankings <- acee_rankings_short %>% 
  select(year, state, score_eee) %>% 
  rename(score = score_eee) %>% 
  rename(rgn_name = state) %>% 
  mutate(rgn_id = case_when(
    rgn_name == "Maine" ~ "6", 
    rgn_name == "New Hampshire" ~ "9", 
    rgn_name == "Massachusetts-North" ~ "7",
    rgn_name == "Massachusetts-South" ~ "8",
    rgn_name == "Rhode Island" ~ "11",
    rgn_name == "Connecticut" ~ "5",
    rgn_name == "New York" ~ "10")) %>% 
  rbind(miss_rgns) %>% 
  rename(acee_rankings_score = score)
```

Climate Actions
```{r clim_act_tidy}
climate_actions <- climate_actions_short %>% 
  select(year, State, score) %>% 
  mutate(rgn_id1 = case_when(
    State == "Maine" ~ "6", 
    State == "New Hampshire" ~ "9", 
    State == "Massachusetts" ~ "7",
    State == "Rhode Island" ~ "11",
    State == "Connecticut" ~ "5",
    State == "New York" ~ "10"),
    rgn_id2 = case_when(
    State == "Massachusetts" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(-filler, - State) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  rbind(miss_rgns) %>% 
  rename(climate_actions_score = score)
```

Renewable Consumption
```{r consum_tidy}
renew_consum <- renew_consum_short %>% 
  select(year, state, score) %>% 
  mutate(rgn_id1 = case_when(
    state == "Maine" ~ "6", 
    state == "New Hampshire" ~ "9", 
    state == "Massachusetts" ~ "7",
    state == "Rhode Island" ~ "11",
    state == "Connecticut" ~ "5",
    state == "New York" ~ "10"),
    rgn_id2 = case_when(
    state == "Massachusetts" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  rbind(miss_rgns) %>% 
  rename(renew_consum_score = score)
```

Carbon Emissions
Once the data for 2017 is released, this part can be removed
```{r carb_em_missing}
##emmissions scores
#create a seperate df to bind to the main one for year 2017 beacuse no data for that
em_2017<- tribble(
  ~year, ~rgn_name, ~rgn_id, ~score,
  2017, "Maine", 6, 1,
  2017, "New Hampshire", 9, 1, 
  2017, "Massachusetts-North", 7, 1,
  2017, "Rhode Island", 11, 0.874,  
  2017, "Connecticut", 5, 1, 
  2017, "New York", 10, 0.958, 
  2017, "Massachusetts-South", 8, 1
  ) %>% 
  as.data.frame() 
```

```{r carb_em_tidy}
carbon_em <- carbon_em_short %>% 
  select(year, state, score) %>% 
  mutate(rgn_id1 = case_when(
    state == "Maine" ~ "6", 
    state == "New Hampshire" ~ "9", 
    state == "Massachusetts" ~ "7",
    state == "Rhode Island" ~ "11",
    state == "Connecticut" ~ "5",
    state == "New York" ~ "10"),
    rgn_id2 = case_when(
    state == "Massachusetts" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  rbind(miss_rgns, em_2017) %>% 
  rename(carbon_em_score = score)
```


## Calculate scores
Combine all the data together and calculate a regulation, enforcement, compliance, and overall resilience score
```{r combine_climate_data}
res_cc <- left_join(acee_rankings, climate_actions, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(carbon_em, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(renew_consum, by = c("year", "rgn_id", "rgn_name")) %>% 
  rowwise() %>% 
  mutate(regulation = mean(c(acee_rankings_score, climate_actions_score)),
         enforcement = mean(c(carbon_em_score, renew_consum_score)),
         compliance = mean(c(carbon_em_score, renew_consum_score)),
         res_score = mean(c(regulation, enforcement, compliance))) 
```

Save necessary information for the water pollution layer
```{r save_layer}
res_climate_change <- res_cc %>% 
  select(year, rgn_id, rgn_name, res_score) %>% 
  rename(score = res_score)

write.csv(res_climate_change, file.path(dir_git,"prep/resilience/ecological/climate_change/data/res_climate_change.csv"))
```

Visualization
```{r graph_cc_res}
res_climate_change_graph <- ggplot(res_climate_change) +
  geom_line(aes(x=year, y= score, color = rgn_name))+
  ggtitle("Northeast Climate Change Resilience Scores") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 

res_climate_change_graph
ggsave("figs/res_climate_change_graph.jpg", width=7, height=5, dpi=300)
```


Write as a resilience layer to ne-scores
```{r write_to_scores}
write.csv(res_climate_change, file.path(dir_calc, "layers/res_climate_change.csv"))
```




