---
title: 'OHI-Northeast: Graphs for Michelle Bachman'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/', message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

library(tidyverse)
library(sf)
library(mapview)
library(cowplot)
```


***

# Data Source 

We use the United States Geological Survey Protected Areas Database for the US. Version 2.0.

**Reference**: 
U.S. Geological Survey (USGS) Gap Analysis Project (GAP), 2018, Protected Areas Database of the United States (PAD-US): U.S. Geological Survey data release, https://doi.org/10.5066/P955KPLE.

**Downloaded**: June 5th, 2019

**Description**:  

**Time range**: Protected areas list established date back to early 1900's where available

**Format**:  Spatial Geodatabase


***

Look at the areas we included and didn't include as protected areas for this layer (map for Michelle Bachman)
```{r}
marine <- read_sf(dsn = file.path(dir_anx, "_raw_data/USGS/PADUS2_0_Shapefiles/PADUS2_0Marine.shp")) %>%
  st_transform(us_alb) %>%
  st_crop(rgns)

states <- read_sf(dsn = file.path(dir_prep, "spatial/shapefiles/state_waters.shp")) %>%
  st_transform(us_alb)

## Fishery Management Area
a <- marine  %>% 
  filter(Loc_Ds %in% c("Fishery Management Areas"))
#mapview(a)
ag <- ggplot(a) +
 geom_sf(states, mapping = aes(), color= "gray88", fill= "transparent") +
  geom_sf(mapping = aes(fill=Loc_Ds), data = NULL, stat = "sf",
  position = "identity", show.legend = FALSE) +
  theme_bw()+
  theme(panel.grid.major = element_line(colour = "transparent"))

a2 <- marine  %>% 
  filter(Loc_Ds %in% c("Fishery Management Area"))
#mapview(a)
a2g <- ggplot(a2) +
 geom_sf(states, mapping = aes(), color= "gray88", fill= "transparent") +
  geom_sf(mapping = aes(fill=Loc_Ds), data = NULL, stat = "sf",
  position = "identity", show.legend = FALSE) +
  theme_bw()+
  theme(panel.grid.major = element_line(colour = "transparent"))

## Closure Areas
b <- marine %>% 
  filter(Loc_Ds %in% c("Closure Area"))
bg <- ggplot(b) +
  geom_sf(states, mapping = aes(), color= "gray88", fill= "transparent")+
  geom_sf(mapping = aes(fill=Loc_Ds), data = NULL, stat = "sf",
  position = "identity",  show.legend = FALSE) +
  theme_bw()+
  theme(panel.grid.major = element_line(colour = "transparent"))

## Shellfihs management areas
c <- marine %>% 
  filter(Loc_Ds %in% c("Shellfish Management Area"))
cg <- ggplot(c) +
  geom_sf(states, mapping = aes(), color= "gray88", fill= "transparent")+
  geom_sf(mapping = aes(fill=Loc_Ds), data = NULL, stat = "sf",
  position = "identity",  show.legend = FALSE) + 
  theme_bw()+
  theme(panel.grid.major = element_line(colour = "transparent"))

## Essential Fish Habitat Conservation Area"
d <- marine %>% 
  filter(Loc_Ds %in% c("Essential Fish Habitat Conservation Area"))
dg <- ggplot(d) +
  geom_sf(states, mapping = aes(), color= "gray88", fill= "transparent")+
  geom_sf(mapping = aes(fill=Loc_Ds), data = NULL, stat = "sf",
  position = "identity",  show.legend = FALSE) + 
  theme_bw()+
  theme(panel.grid.major = element_line(colour = "transparent"))


## Conservation area
e <- marine %>% 
  filter(Loc_Ds %in% c("Conservation Area"))
eg <- ggplot(e) +
  geom_sf(states, mapping = aes(), color= "gray88", fill= "transparent")+
  geom_sf(mapping = aes(fill=Loc_Ds), data = NULL, stat = "sf",
  position = "identity",  show.legend = FALSE) + 
    theme_bw()+
  theme(panel.grid.major = element_line(colour = "transparent"))

## Gear restricted area
f <- marine %>% 
  filter(Loc_Ds %in% c("Gear Restricted Area"))
fg <- ggplot(f) +
  geom_sf(states, mapping = aes(), color= "gray88", fill= "transparent")+
  geom_sf(mapping = aes(fill=Loc_Ds), data = NULL, stat = "sf",
  position = "identity",  show.legend = FALSE) + 
  theme_bw()+
  theme(panel.grid.major = element_line(colour = "transparent"))

## Special Area Management Plan
g <- marine %>% 
  filter(Loc_Ds %in% c("Special Area Management Plan"))
gg <- ggplot(g) +
  geom_sf(states, mapping = aes(), color= "gray88", fill= "transparent")+
  geom_sf(mapping = aes(fill=Loc_Ds), data = NULL, stat = "sf",
  position = "identity",  show.legend = FALSE) + 
  theme_bw()+
  theme(panel.grid.major = element_line(colour = "transparent"))

inc <- marine %>% 
  filter(!Loc_Ds %in% c("Fishery Management Area", "Closure Area", "Fishery Management Areas", "Shellfish Management Area", "Essential Fish Habitat Conservation Area", "Conservation Area", "Gear Restricted Area", "Special Area Management Plan"), #the Conservation Areas are for Mussel Seed
         !is.na(Loc_Ds))
inc_g <- ggplot(inc) +
  geom_sf(states, mapping = aes(),color= "gray88", fill= "transparent")+
  geom_sf(mapping = aes(fill=Loc_Ds), data = NULL, stat = "sf",
  position = "identity",  show.legend = FALSE) + 
  theme_bw()+
  theme(panel.grid.major = element_line(colour = "transparent"))

                            
all <- cowplot::plot_grid(inc_g, ag,a2g, bg,cg,dg,eg,fg,gg,
                          labels= c("Areas included in LSP", "Fishery Management Areas(excl)","Fishery Management Area(excl)", "Closure Area(excl)","Shellfish Management Area(excl)", "Essential Fish Habitat Conservation Area(excl)", "Conservation Area(excl)", "Gear Restricted Area(excl)", "Special Area Management Plan(excl)"),
                          label_size = 8, 
                          ncol = 3)
#all

ggsave("figs/michelle_graphs.jpg", width=10, height=8, dpi=300)
```

Create csv file

```{r}
lsp_areas <- marine %>% 
  mutate(Included = ifelse(Loc_Ds == "Fishery Management Areas" | Loc_Ds == "Fishery Management Area" | Loc_Ds =="Closure Area" | Loc_Ds =="Shellfish Management Area" | Loc_Ds =="Essential Fish Habitat Conservation Area" | Loc_Ds =="Conservation Area" | Loc_Ds =="Gear Restricted Area" | Loc_Ds =="Special Area Management Plan", "No", "Yes")) %>% 
  filter(!is.na(Loc_Ds) & d_State_Nm != "New Jersey") %>% 
  as_data_frame() %>% 
  select(Included, d_State_Nm, Unit_Nm, d_Mang_Nam, d_Mang_Typ, Loc_Ds) %>% 
  rename('Management Type' = d_Mang_Typ) %>% 
  rename('Management Name' = d_Mang_Nam) %>% 
  rename(State = d_State_Nm) %>% 
  rename("Location Designation (our filter)" = Loc_Ds) %>% 
  rename('Area Name' = Unit_Nm)

write.csv(lsp_areas, file.path(dir_prep, "prep/sop/lsp/data/lsp_areas.csv"))


```


















