---
title: 'OHI-Northeast: Lasting Special Places - Terrestrial'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary
This script creates the designated terrestrial (land) areas layer for the Lasting Special Places subgoal. This layer measures the amount of designated and/or protected area within 1km of the coastline for each state in the Northeast region.

The output layer is saved to `ne-scores/region/layers` for use in the Toolbox.

``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/', message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

library(tidyverse)
library(sf)
library(mapview)
```


***

# Data Source 

We use the United States Geological Survey Protected Areas Database for the US. Version 2.0.

**Reference**: 
U.S. Geological Survey (USGS) Gap Analysis Project (GAP), 2018, Protected Areas Database of the United States (PAD-US): U.S. Geological Survey data release, https://doi.org/10.5066/P955KPLE.

**Downloaded**: June 5th, 2019

**Description**:  

**Time range**: Protected areas list established date back to early 1900's where available

**Format**:  Spatial Geodatabase


***

# Load Data

Load our inland buffer so we can extract just the areas within 1km of the coastline.
```{r buffers}
inland_buffer <- read_sf("~/github/ne-prep/spatial/shapefiles/ohine_inland_1km.shp")
```

Let's look at the available layers in the USGS PADUS database

```{r look_at_gdb_layers}
st_layers(file.path(dir_anx, "_raw_data/USGS/PADUS2_0_GDB_Arc10x/PADUS2_0.gdb"))
```

After exploring the data descriptions we know we want all ecept the Proclamation layer. When I load these, I'm also cropping to our region to reduce object size.

```{r load_three_land_layers}
desig <- read_sf(dsn = file.path(dir_anx, "_raw_data/USGS/PADUS2_0_Shapefiles/PADUS2_0Designation.shp")) %>%
  st_transform(us_alb) %>%
  st_crop(rgns) %>%
  st_intersection(inland_buffer)

ease <- read_sf(dsn = file.path(dir_anx, "_raw_data/USGS/PADUS2_0_Shapefiles/PADUS2_0Easement.shp")) %>%
  st_transform(us_alb) %>%
  st_crop(rgns)  %>%
  st_intersection(inland_buffer)

#fee is causing geometry errors. so i filter just for our states first and the error goes away. The error must exist outside or region of interest
fee <- read_sf(dsn = file.path(dir_anx, "_raw_data/USGS/PADUS2_0_Shapefiles/PADUS2_0Fee.shp")) %>%
  filter(State_Nm %in% c("NY", "RI", "CT", "MA", "NH", "ME")) %>%
  st_transform(us_alb) %>%
  st_crop(rgns)  %>%
  st_intersection(inland_buffer)
```

# Calculate total area protected

We want to know the proportion of area protected within each region. 
```{r}
#select the same columns from each of the three land layers
ease_clean <- ease %>%
  select(Loc_Ds, Loc_Nm, Date_Est, rgn_name, rgn_id)

fee_clean <- fee %>%
  select(Loc_Ds, Loc_Nm, Date_Est, rgn_name, rgn_id)

desig_clean <- desig %>%
  select(Loc_Ds, Loc_Nm, Date_Est, rgn_name, rgn_id)

#combine them into `land`

land <- ease_clean %>%
  rbind(fee_clean) %>%
  rbind(desig_clean) %>%
  unique() %>% #remove duplicates
  lwgeom::st_make_valid()
```

We want to measure amount protected each year. The data has this information stored in the `Date_Est` column. To account for this, we can iterate over each year (2005-2017). Some designated areas have an `NA` for Date Established. Since we have no other information to use for these areas, we assume they have been established since before 2005.

```{r calculate_land_protection}
#create empty dataframe to fill with each iteration in the for loop
land_protected_by_rgn <- data.frame()

#for each year from 2005 - 2017
for(i in 2005:2017){
  
  #filter dataset to only include areas with Date_Est <= year i, and NAs
  s_yr <- filter(land, Date_Est <= i | is.na(Date_Est)) %>% #all NA included as well
    group_by(rgn_id) %>%
    summarise() %>% #this takes care of overlapping geometries so we don't double count areas
    ungroup() %>%
    mutate(prot_area_m = st_area(geometry)) %>% #calculate total protected area
    left_join(inland_buffer %>% st_set_geometry(NULL)) %>% #join to inland buffer 
    mutate(prop_area = prot_area_m/area, #calculate proportion of land in 1km region protected
           type = "land",
           year = i) %>%
    st_set_geometry(NULL)
  
  #fill out dataframe
  land_protected_by_rgn <- bind_rows(s_yr, land_protected_by_rgn)
}
```


# Results

```{r land_protected_prop_by_rgn}
ggplot(land_protected_by_rgn, aes(x = year, y = prop_area, color = rgn_name)) +
  geom_line() +
  theme_bw() +
  labs(y = "Proportion of inland 1km protected")
```


# Save Layer

```{r save_to_toolbox}
#save land
write.csv(land_protected_by_rgn %>% select(rgn_id, year, prop_area), file = "~/github/ne-scores/region/layers/lsp_protected_land.csv")
```






