---
title: 'OHI-Northeast: Lasting Special Places - Marine'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary
This script creates the designated marine areas layer for the Lasting Special Places subgoal. This layer measures the amount of marine area within state waters that is designated and/or protected.

The output layer is saved to `ne-scores/region/layers` for use in the Toolbox.

``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/', message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

library(tidyverse)
library(sf)
library(mapview)
```


***

# Data Source 

We use the United States Geological Survey Protected Areas Database for the US. Version 2.0.

**Reference**: 
U.S. Geological Survey (USGS) Gap Analysis Project (GAP), 2018, Protected Areas Database of the United States (PAD-US): U.S. Geological Survey data release, https://doi.org/10.5066/P955KPLE.

**Downloaded**: June 5th, 2019

**Description**:  

**Time range**: Protected areas list established date back to early 1900's where available

**Format**:  Spatial Geodatabase


***

# Load Data

Load the state waters shapefile.
```{r buffer}
#load buffer
state_waters <- rgns %>% filter(rgn_id > 4)
```

Let's look at the available layers in the USGS PADUS database

```{r look_at_gdb_layers}
st_layers(file.path(dir_anx, "_raw_data/USGS/PADUS2_0_GDB_Arc10x/PADUS2_0.gdb"))
```

After exploring the data descriptions we know we just want the Marine layer. When I load this layer, I'm also cropping to our region to reduce object size.

We are also going to remove all areas that are designated as fishery/shellfish management areas as well as the Special Area Management Plans. This designations are not associated with protection or designation as a special place, and therefore should not be included.

```{r load_and_clean_marine_layer}
marine <- read_sf(dsn = file.path(dir_anx, "_raw_data/USGS/PADUS2_0_Shapefiles/PADUS2_0Marine.shp")) %>%
  st_transform(us_alb) %>%
  st_crop(rgns)  %>%
  st_intersection(state_waters) %>%
  filter(!Loc_Ds %in% c("Fishery Management Area", "Closure Area", "Fishery Management Areas", "Shellfish Management Area", "Essential Fish Habitat Conservation Area", "Conservation Area", "Gear Restricted Area", "Special Area Management Plan"), #the Conservation Areas are for Mussel Seed
         !is.na(Loc_Ds)) 
```

# Calculate total area protected

We want to know the proportion of area protected within each region. 

First let's combine polygons so we don't double count areas. For example, multiple shipwrecks are included within the larger MA North Shore Sanctuary. 

All areas included in the Marine layer were established before our time period (or have NA for Date_Est) so we don't need to do this by year. 
```{r}
marine_protected_by_rgn <- marine %>%
  select(rgn_id, Date_Est, Loc_Ds, Loc_Nm) %>%
  mutate(rgn_id = ifelse(c(rgn_id == 11 & Loc_Nm == "Cape and Islands Ocean Sanctuary"), 8, rgn_id)) %>% #one sliver of Cape and Islands Ocean Sanctuary was being assigned to Rhode Island but is definitely MA
  group_by(rgn_id) %>%
  summarise() %>%
  ungroup() %>%
  mutate(area = st_area(geometry)/1000000) %>% #calculate total protected area
  left_join(rgn_data) %>%
  mutate(prop_area = as.numeric(area)/area_km2, #calculate proportion of state waters protected
         type = "marine",
         year = 2005) %>%
    st_set_geometry(NULL) %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n)
```

# Results

```{r marine_protected_prop_by_rgn}
ggplot(marine_protected_by_rgn, aes(x = year, y = prop_area, color = rgn_name)) +
  geom_line() +
  theme_bw() +
  labs(y = "Proportion of State waters protected")
```

# Save Layer

```{r save_to_toolbox}
#save marine
write.csv(marine_protected_by_rgn %>% select(rgn_id, year, prop_area), file = "~/github/ne-scores/region/layers/lsp_protected_marine.csv")
```






