---
title: 'OHI-Northeast: Get eelgrass spatial information'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary
This script creates eelgrass shapefiles for us in developing the eelgrass habitat layer.

# Data

All spatial data, both current and historical eelgrass locations, were downloaded from the [Northeast Ocean Data Portal](https://www.northeastoceandata.org/data-explorer/?habitat|biological).

# Setup

```{r setup, message = F, warning = F}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')

library(magrittr)
library(sf)
```

# Load eelgrass spatial data

The spatial data was downloaded from the Northeast Ocean Data Portal. All available data was downloaded, historical and current extent of eelgrass. This is then combined to create one single shapefile that indicates presence of eelgrass at any point in recent history (back to the 1980's). I read each one in individually to manually check them against their metadata. Some of them require additional cleaning. Some have "widgeongrass" as well - which we exclude.

```{r}
#list of each individual folder containing a shapefile of historical eelgrass data
raw_list <- list.files(file.path(dir_anx, "_raw_data/NEOceanDataPortal/historical_eelgrass"))

#load as shapefiles
#unfortunatley we have to go file by file since these are not standardized
  
poly1 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/SAV06/SAV06.shp")) %>%
  filter(str_detect(habitat, "eelgrass")) %>%
  select(geometry) %>%
  st_transform(crs = us_alb)

#this one only has eelgrass/widgeon grass
poly2 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/SAV09/SAV09.shp")) %>%
  filter(habitat %in% c("eelgrass", "eelgrass/widgeon grass")) %>%
  select(geometry) %>%
  st_transform(crs = us_alb) 


poly3 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/EELGRASS_BEDS_2006_POLY/EELGRASS_BEDS_2006_POLY.shp")) %>% select(geometry) %>%
  st_transform(crs = us_alb) 

poly4 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/EELGRASS_BEDS_2002_POLY/EELGRASS_BEDS_2002_POLY.shp")) %>% select(geometry) %>%
  st_transform(crs = us_alb) 

poly5 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/EELGRASS_BEDS_2009_POLY/EELGRASS_BEDS_2009_POLY.shp")) %>% select(geometry) %>%
  st_transform(crs = us_alb) 

poly6 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/EELGRASS_BEDS_HISTORIC_POLY/EELGRASS_BEDS_HISTORIC_POLY.shp")) %>% select(geometry) %>%
  st_transform(crs = us_alb) 

poly7 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/EELGRASS2006_7_POLY/EELGRASS2006_7_POLY.shp")) %>% 
  filter(HABITAT == "eelgrass") %>%
  select(geometry) %>%
  st_transform(crs = us_alb) 

#9 Mb
poly8 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/Eelgrass2010_ME/Eelgrass2010_ME.shp")) %>% select(geometry) %>%
  st_transform(crs = us_alb) 

## very large (14.5 Mb)
poly9 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/eelgrass97/eelgrass97.shp")) %>% 
  filter(COVER > 0) %>% #remove locations of 0 cover
  select(geometry) %>%
  st_transform(crs = us_alb)

poly10 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/fordpoly/fordpoly.shp")) %>% select(geometry) %>%
  st_transform(crs = us_alb)

poly11 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/narrbayproj/narrbayproj.shp")) %>% 
  filter(VEGETATION == "EELGRASS") %>%
  select(geometry) %>%
  st_transform(crs = us_alb)

poly12 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/nbeppolys/nbeppolys.shp")) %>% select(geometry)  %>%
  st_transform(crs = us_alb)

poly13 <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/urinrs/urinrs.shp")) %>% select(geometry) %>%
  st_transform(crs = us_alb)


#combine all polygons

#25.3 Mb poly3 was giving us issues
all_polys <- rbind(poly1, poly2, poly3, poly4, poly5, poly6, poly7, poly8, poly9, poly10, poly11, poly12, poly13) 

st_write(all_polys, dsn = file.path(dir_anx,"hab/past_eelgrass_locations"), layer = "past_eelgrass_locations", driver = "ESRI Shapefile", delete_layer = T)
```

One of the files is a lines object. I want to keep this but it needs to be saved on it's own

```{r}
#this is a lines object
line_eelgrass <- read_sf(file.path(dir_anx,"/_raw_data/NEOceanDataPortal/historical_eelgrass/crmclines/crmclines.shp")) %>% select(geometry) %>%
  st_transform(crs = us_alb) %>%
  st_write(dsn = file.path(dir_anx,"hab/line_eelgrass_locations"), layer = "line_eelgrass_locations", driver = "ESRI Shapefile", delete_layer = T)
```

The Great Bay Estuary shapefiles all have the same format so we can use a function to grab all of these shapefiles
```{r}
gbe_years <- raw_list[str_detect(raw_list, "Eelgrass[0-9]+$")]

gbe <- c()
for(i in 1:length(gbe_years)){
  print(i)
  
  file <- gbe_years[i]
  
  poly <- read_sf(paste0(file.path(dir_anx),"/_raw_data/NEOceanDataPortal/historical_eelgrass/", file, "/", file, ".shp"))
  
  names(poly) %<>% tolower
  
  poly2 <- poly %>% filter(density != "not mapped") %>% select(geometry) %>% st_transform(crs = us_alb)
  
  gbe <- rbind(poly2, gbe)
  
}

#combine the GBE polygons
gbe_single <- st_union(gbe)

#st_write(gbe_single, dsn = file.path(dir_anx,"hab/gbe_eelgrass_locations"), layer = "gbe_eelgrass_locations", driver = "ESRI Shapefile")


# 7.6 Mb
current_eelgrass <- st_read(dsn = file.path(dir_anx,"_raw_data/NEOceanDataPortal/Habitat/Habitat.gdb"), layer = "EelgrassBeds") %>%
  select(geometry = Shape) %>%
  st_transform(crs = us_alb) %>%
  st_union()

#st_write(current_eelgrass, dsn = file.path(dir_anx,"hab/current_eelgrass_locations"), layer = "current_eelgrass_locations", driver = "ESRI Shapefile")

```
