---
title: 'OHI-Northeast: Salt Marsh Habitat Data'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary
This script creates the Salt Marsh habitat data layer for the Habitats goal (a sub-goal of Biodiversity). Using spatial land cover data from NOAA's [Coastal Change Analysis Program (C-CAP)](https://coast.noaa.gov/digitalcoast/data/ccapregional.html). 

# Data

Data was downloaded [from this site](https://coast.noaa.gov/ccapftp/#/) for each state for the years 1996, 2001, 2006, and 2010.

# Setup

```{r setup, message = F, warning = F}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)
source('~/github/ne-prep/src/R/common.R')

# Multiprocessing
library(parallel)
#for gapfilling
library(zoo)
```

# Load in C-CAP data

The C-CAP Regional Land Cover Classifications that we are using include:   
- Estuarine Forested Wetland (16)  
- Estuarine Scrub/Shrub Wetland (17)  
- Estuarine Emergent Wetland (18)  
- Estuarine Aquatic Bed (23)  

We may potentially need to include:    
- Palustrine Forested Wetlands (13)  
- Palustrine Scrub/Shrub Wetlands (14)  
- Palustrine Emergent Wetlands (15)  
- Palustrine Aquatic Bed (22)  

The rasters are provided as `.img` files. Fortunately the `raster` package can read these as rasters.

```{r load_files}
#load all files in this folder held on Mazu
ccap_files <- list.files(file.path(dir_anx, "_raw_data/NOAA_C-CAP"), full.names = T, recursive = T)

#get the rasters (.img)
files <- ccap_files[str_detect(ccap_files, ".img")]
```

# Remove non-wetland cells

The `get_wetland_cells` reads in each raster, and sets all cells that are not equal to 16, 17, 18 or 23 as `NA`, then saves the file to our server.
```{r get_wetland_cells, eval = F}
#define all values that represent estuarine wetland habitats
wetlands <- c(16, 17, 18, 23)

get_wetland_cells <- function(file){
  
  name <- substr(file, 80, 86)
  r    <- raster(file)
  
  r[!r %in% wetlands] <- NA
  writeRaster(r, filename = paste0(dir_anx, "/hab/salt_marsh/", name,"_wetland_cells.tif"), overwrite = T)
  
}

mclapply(files[9:10], get_wetland_cells, mc.cores = 6)
```

# Calculate habitat loss over time

Now we want to look at change over time and calculate the proportion lost. Since all of these cells are the same size, we can simply get the number of cells in each time period and then calculate % loss or gain. The `get_cells_over_time` function does just this.

```{r calculate_habitat_loss}
states <- c("/nh", "/ny", "/me", "/ri", "/ct", "/ma") #added the slash because "me" was bringing all rasters

files2 <- list.files(file.path(dir_anx, "hab/salt_marsh"), full.names = T)

get_cells_over_time <- function(state){
  
  state_files <- files2[str_detect(files2, state)] %>%
    stack()
  
  #create dataframe
  vals <- getValues(state_files) %>% 
    colSums(na.rm = T) %>%
    data.frame() %>%
    rownames_to_column() %>%
    mutate(state = str_replace(state, "/", ""),
           year  = substr(rowname, 4, 7)) %>%
    rename(num_wetland_cells = ".") %>%
    select(-rowname)
  
  return(vals)
}
```

####

This section is on hold for now. Eventually we may want to figure out a way to assign inland cells to either MA region. But until we can figure that out, both MA regions get the same data.
```{r massachusetts, eval = F}
# need to do massachusetts separately
ma_rast <- files2[str_detect(files2, "/ma")] %>%
    stack()

#get MA

ma_shp <- read_sf("~/github/ne-prep/spatial/shapefiles/coastal_rgns_3km_inland_buffer.shp") %>%
  filter(rgn_id %in% c(7,8)) %>%
  st_transform(crs = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") #transform to the same CRS as the C-CAP data.

a <- raster::extract(ma_rast, ma_shp)
```

####

By using `map_df` we can apply the `get_cells_over_time()` function for each state and produce a dataframe that shows the number of cells per time period per state.

```{r salt_marsh_percent_change_since_1996}
state_cells <- map_df(states, get_cells_over_time) 

#we want to compare each time period to the 1996 time period, not the previous one
perc_change_df <- state_cells %>%
  group_by(state) %>%
  mutate(first  = dplyr::first(num_wetland_cells),
         change = (1-(num_wetland_cells/first))*100) %>% #calculate % change from 1996
  ungroup() %>%
  mutate(state = toupper(state))

ggplot(perc_change_df, aes(x = as.numeric(year), y = change, color = state)) +
  geom_line() +
  theme_bw() +
  labs(x = "Year",
       y = "% Habitat loss (negative values mean gains in habitat)")
```

We want to add in historical loss from the 1777-1854 time period. We have values from the 80's/90's. We can add what we just calculated to these for a couple more data points.

```{r historic_loss}
hist_loss <- data.frame(state_name = c("Rhode Island", "Massachusetts", "New Hampshire", "Maine", "Connecticut", "New York"),
                        state = c("RI", "MA", "NH", "ME", "CT", "NY"),
                        hist_perc_loss = c(53, 41, 18, 0, 27, 48))

full_df <- perc_change_df %>%
  left_join(hist_loss) %>%
  mutate(total_loss = change + hist_perc_loss,
         year = as.numeric(year)) %>%
  select(state, state_name, year, total_loss)

#save for dashboard
write_csv(full_df, "data/salt_marsh_percent_change.csv")
```

# Save to toolbox

We need to save this data for the toolbox. We only have data points for a couple years in our time series. We will do a linear interpolation between points to gapfill `NA`. We also convert % loss to scores simply by inverting percent loss. So 75% loss gets a score of 25.

```{r salt_marsh_total_perc_loss_by_rgn}
toolbox_layer <- full_df %>%
  group_by(state) %>%
  complete(year = 1996:2010) %>% #only completing to 2010 for linear interpolation
  mutate(perc_loss = na.approx(total_loss)) %>%
  complete(year = 1996:2017) %>%
  mutate(perc_loss = na.locf(perc_loss)) %>% #now we carry forward the 2010 value for all years to 2017
  ungroup() %>%
  left_join(rgns, by = c("state" = "state_abv")) %>%
  select(rgn_id, rgn_name, year, perc_loss) 
```

# Results

```{r}
ggplot(toolbox_layer, aes(x = year, y = perc_loss, color = rgn_name)) +
  geom_line() +
  theme_bw() +
  labs(x = "Year",
       y = "Percent (%) lost",
       title = "Salt Marsh")
```


Add in offshore regions with NA then save
```{r save}
out <- toolbox_layer %>%
  complete(rgn_id = 1:11, year) %>%
  select(rgn_id, year, perc_loss) %>%
  mutate(habitat = "saltmarsh")
write_csv(out, file.path(dir_calc, "layers/hab_salt_marsh.csv"))
```



















