---
title: 'OHI-Northeast: Carbon Storage sub-goal'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 2
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---


``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R
library(ggrepel)
```

# Summary

The Carbon Storage sub-goal measures the health of carbon storing marine habitats. In the Northeast, this includes Salt Marsh and Eelgrass habitats. Both of these habitats are also included in the Biodiversity - Habitats subgoal. We can use those same layers here. Both of these habitats are able to capture carbon at a near equal rate, so we don't include carbon storage weights. Instead we simply average the health of the habitats across each region to get a sense of how well this habitat service is doing.

***

# Data Source

**Eelgrass**
All spatial data, both current and historical eelgrass locations, were downloaded from the [Northeast Ocean Data Portal](https://www.northeastoceandata.org/data-explorer/?habitat|biological).

**Salt Marsh**

Spatial land cover data was downloaded [from this site](https://coast.noaa.gov/ccapftp/#/) for each state in the Northeast for the years 1996, 2001, 2006, and 2010.

***
  
# Methods

## Load Data

```{r load_data}
eelgrass   <- read_csv("~/github/ne-scores/region/layers/hab_eelgrass.csv") %>%
  mutate(habitat = "eelgrass") %>%
  select(rgn_name, year, score, habitat, rgn_id)

#read in the salt marsh data and add a column that converts the perc_loss to a score between 0 and 100
salt_marsh <- read_csv("~/github/ne-scores/region/layers/hab_salt_marsh.csv") %>%
  rowwise() %>%
  mutate(habitat = "saltmarsh",
          score  = min(100, (100 - perc_loss))) %>% #prevents score from going above 100
  left_join(rgns) %>%
  select(rgn_name, year, score, habitat, rgn_id)
```

```{r eelgrass_score_plot}
ggplot(eelgrass, aes(x = year, y = score, color = rgn_name)) +
  geom_line() +
  theme_bw() +
  geom_text_repel(data = subset(eelgrass, year == max(year)),
                  aes(label = rgn_name),
                  size = 3,
                  hjust = 0,
                  direction = "y",
                  nudge_x = 2,
                  segment.color = NA) +
  labs(x = "Year",
       y = "Score",
       title = "Eelgrass") +
  theme(legend.position = "none")

```
```{r saltmarsh_score_plot}
ggplot(salt_marsh, aes(x = year, y = score, color = rgn_name)) +
  geom_line() +
  theme_bw() +
  geom_text_repel(data = subset(salt_marsh, year == max(year)),
                  aes(label = rgn_name),
                  size = 3,
                  nudge_x = 45,
                  segment.color = NA) +
  labs(x = "Year",
       y = "Score",
       title = "Salt Marsh") +
  theme(legend.position = "none")
```

## Average across habitats

```{r calculate_cs_scores}
combo <- eelgrass %>%
  bind_rows(salt_marsh) %>%
  group_by(rgn_id, rgn_name, year) %>%
  summarize(cs_score = mean(score)) %>%
  filter(year > 2005) #eelgrass data starts at 2006 so can only measure CS for overlapping years
```


***

# Results

```{r carbon_storage_rgn_scores}
ggplot(combo %>% filter(!is.na(cs_score)), aes(x = year, y = cs_score, color = rgn_name)) +
  geom_line(show.legend = F) +
  theme_bw() +
  labs(x = "Year",
       y = "Score",
       title = "Carbon Storage") +
  geom_text_repel(
    data = subset(combo, year == max(year)),
    aes(label = rgn_name),
    size = 3,
    hjust = 0,
    direction = "y",
    nudge_x = 2,
    segment.color = NA,
    show.legend = FALSE
  )
```


***

# Save to toolbox

```{r save_to_toolbox}
write.csv(combo, file = "~/github/ne-scores/region/layers/hs_carbon_storage.csv")
```


***



