---
title: 'OHI-Northeast: Calculate proportion of catch for food or bait'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Setup

```{r setup, message = F, warning = F}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)
source('~/github/ne-prep/src/R/common.R')
library(readxl)
```

# Data

We received **Dealer Data** from NOAA that reports the amount of harvested species that results in food, bait or other markets.

```{r}
data <- read_excel(file.path(dir_anx, "_raw_data/NOAA_NMFS/Afflerbach_UCSB_Dealer Data All Species_MAR 2019.xlsx"))
```

## Clean data

```{r}
clean <- data %>%
  rename(year = YEAR,
         state = STATE,
         species = SPECIES,
         market = DISPOSITION,
         pounds_landed = `LBS (LANDED WEIGHT)`,
         pounds_live = `LBS (LIVE WEIGHT)`,
         value = VALUE) %>%
  mutate(species = str_replace(species, ",", ", ")) %>%
  filter(year > 2002)
```

# Visualize by state

Let's just look at species that have more than one market

```{r}
df1 <- clean %>%
  group_by(state, year, species) %>%
  mutate(markets = n(),
         total_pounds_live = sum(pounds_live)) %>%
  ungroup() %>%
  mutate(prop = pounds_live/total_pounds_live) 

#we want to keep all records where there are either 2 markets, or 1 market but market != "FOOD"

nonfood_df <- df1 %>%
  mutate(keep = case_when(
    markets > 1 ~ 1,
    markets == 1 & market != "FOOD" ~ 1,
    TRUE ~ 0
  )) %>%
  filter(keep == 1,
         species != "CONFIDENTIAL SPECIES") #this isn't going to help us
```
What species have more than one market (that aren't "CONFIDENTIAL") or their only market is not FOOD

```{r}
sp <- unique(nonfood_df$species)
sp
```

```{r, eval = F}

for(i in 1:length(sp)){
  
  spp <- sp[i]
  sp_df <- filter(df1, species == spp)
  
ggplot(sp_df, aes(x = year, y = prop, fill = market)) +
  geom_col() +
  facet_wrap(~state) +
  theme_bw() +
  labs(y = "Proportion",
       x = "Year",
       title = spp)

ggsave(paste0("figs/", spp, "_market_prop.pdf"), width = 6, height = 4)
}
```

```{r}
library(trelliscopejs)


df2 <- df1 %>% filter(species %in% sp)

ggplot(df2, aes(x = year, y = prop, fill = market)) +
  geom_col() +
  facet_wrap(~state) +
  theme_bw() +
  labs(y = "Proportion",
       x = "Year") +
  facet_trelliscope(~ species, scales = "free_y", self_contained = TRUE, width = 600, height = 400)
```


# Just get region proportions

Splitting these out by state may actually make things more complicated since our catch data is by statistical area. So instead I'll calculate regional proportions.

```{r}
rgn_prop <- clean %>%
  filter(species != "CONFIDENTIAL SPECIES",
         !market %in% c("REASON NOT SPECIFIED", "KEPT, DISPOSITION UNKNOWN", "REASON NOT SPECIFIED", "UNKNOWN")) %>%
  group_by(year, species) %>%
  mutate(total_pounds_live = sum(pounds_live)) %>%
  ungroup() %>%
  select(year, species, market, pounds_live, total_pounds_live) %>%
  group_by(market, species, year) %>%
  mutate(pounds_live_by_market = sum(pounds_live)) %>%
  select(year, species, market, pounds_live_by_market, total_pounds_live) %>%
  distinct() %>%
  mutate(prop = pounds_live_by_market/total_pounds_live) %>%
  filter(prop < 1) #we really only want to keep all records where proportion is less than 1
```


# Save

Save the proportions for these species in a .csv to be used in the Fisheries data prep.

```{r}
out <- df2 %>%
  select(year, state, species, market, prop) 

write_csv(out, "data/fish_catch_food_prop_by_state.csv")

write_csv(rgn_prop, "data/fish_catch_food_prop_rgn.csv")
```









