---
title: "Spatializing fish catch in the Northeast"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary

This script calculates catch by OHI region using landings data provided by NOAA.

# Data

**Downloaded**: July, 2019

**Description**:  Commercial fish landings by statistical area

**Time range**: 1996-2017. Data provided annually

**Format**:  Excel spreadsheet

***

# Setup

```{r, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.path = "figs/")

library(tidyverse)
library(readxl)
library(sf)
#install.packages("viridis")
library(viridis)

source("~/github/ne-prep/src/R/common.R")
```


# Data Cleaning

Cleaning the raw data a bit by fixing column names and turning `stat_area` numeric. 
```{r}
raw <- read_excel(file.path(dir_anx, "_raw_data/NOAA_NMFS/catch_by_stat_area/Afflerbach_UCSB_Landings by Stat Area_JUL 2019 Updated.xlsx"))

clean <- raw %>%
  rename(year = YEAR,
         stat_area = `STAT\r\nAREA`,
         species = SPECIES,
         pounds = `LBS LANDED \r\n(HAIL WT)`,
         stock_id = `STOCK ID`,
         stock = `STOCK NAME`) %>%
  mutate(stat_area = as.numeric(stat_area))

head(clean)
```

There are a total of `r length(unique(clean$species))` reported species, which include groupings such as "Confidential Species" or "Clam, species not specified". Some of these species, such as Atlantic Cod, have multiple stocks.

# Spatial data

Since the data is provided by statistical landing area, we can use this information to infer what OHI region's encompass or overlap with these areas. We have downloaded the shapefile for Statistical Areas from [this public FTP NOAA site](ftp://ftp.nefsc.noaa.gov/pub/gis/).

## Statistical areas

Load in the statistical areas and add area of each polygon as a column.

```{r statistical_areas}
stat_shp <- sf::read_sf(file.path(dir_anx, "spatial/Statistical_Areas_2010_withNames.shp")) %>%
  st_set_crs(p4s_nad83) %>% #set the CRS
  st_transform(crs = crs(rgns))  #transform to our NE specific CRS

stat_shp$stat_area <- st_area(stat_shp) #add area as column


ggplot(stat_shp) +
  geom_sf() +
  theme_bw() +
  labs(title = "Statistical areas") +
  theme(legend.position = "none")
```

We overlay statistical areas with our regions to select just the statistical areas that overlap with our regions.

```{r ohi_stat_areas}
ohi_stat_areas          <- st_intersection(rgns, stat_shp) #intersects statistical areas with OHI regions
ohi_stat_areas$ohi_area <- st_area(ohi_stat_areas) #calculate area of each overlapped polygon

ggplot(ohi_stat_areas) +
  geom_sf(aes(fill = rgn_name)) +
  geom_sf(data = ne_states, fill = "beige") +
  theme_bw() +
  labs(title = "Statistical areas overlapped with OHI regions") 
```

Calculate proportion of each statistical area in our OHI regions. For statistical areas that overlap with OHI regions, we can use proportional area overlap to adjust catch. We assume that catch is evenly distributed across each statistical area.

```{r prop_stat_areas_in_ohi_rgns}
calc_prop_area <- ohi_stat_areas %>%
  group_by(Id) %>%
  mutate(ohi_rgn_prop_area = ohi_area/stat_area) #this column tells us how much of each OHI sub-region falls within the statistical area in our region

ggplot(calc_prop_area) +
  geom_sf(aes(fill = ohi_rgn_prop_area)) +
  geom_sf(data = ne_states, fill = "beige") +
  theme_bw() +
  labs(title = "Proportion of each \nstatistical area in OHI region") 
```

# Catch per OHI region

Now we calculate the total catch per species and year for each of the OHI regions.

First let's filter the catch data to just the statistical areas in our region. We don't care about the catch outside of these statistical areas.

```{r}
region_catch <- clean %>%
  filter(stat_area %in% ohi_stat_areas$Id) %>%
  left_join(calc_prop_area, by = c("stat_area" = "Id")) %>%
  mutate(catch = pounds*ohi_rgn_prop_area) %>% #adjusting catch by the proportional area with overlap
  select(-area_km2, -FULL_NAME, -SHORT_NAME, -stat_area.y, -ohi_area, -NAFODIV) %>%
  group_by(species, stock_id, stock, rgn_id, year, rgn_name) %>%
  summarize(catch = sum(catch)) %>%
  ungroup() %>%
  mutate(display_name = ifelse(is.na(stock_id), species, stock_id))

head(region_catch, n = 20)
```

```{r}
write_csv(region_catch, "data/nmfs_spatial_catch_by_ohi_rgn.csv") #this file is used in fis_species_lookuptable.Rmd to create the species lookup table
```


## Total catch over time

```{r total_northeast_catch}
total_catch <- region_catch %>%
  group_by(year) %>%
  summarize(catch = sum(catch))

ggplot(total_catch, aes(x = year, y = catch)) +
  geom_area() +
  labs(x = "",
       y = "Pounds",
       title= "Total catch in the Northeast") +
  theme_bw()
```

## Visualize catch by region

```{r}
p <- ggplot(region_catch, aes(x = year, y = catch, fill = display_name)) +
  facet_wrap(~rgn_name, scales = "free_y") +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(legend.position = 'none',
        axis.text = element_text(size = 6),
        axis.text.x = element_text(angle = 45))

plotly::ggplotly(p)
```

# Gapfill

The data shared with us includes records of 0 catch. But there is still missing data. As an example, let's look at ALEWIFE.

```{r}
ale <- clean %>%
  filter(species == "ALEWIFE")

unique(ale$year)
```

Ok clearly we are missing data for 2001, 2002, 04-05, 2008-11. We don't know if these are 0's or missing data. We need to **gapfill** this missing data. When a species/state combination has missing data for a year, we can not assume it has a catch of 0. Since we calculate a rolling average of catch, NAs will remain as NA's and the average will rely on just one or two years of catch. This is done to account for any wild fluctuations in catch year to year.

```{r}
gf_data <- region_catch %>%
  group_by(rgn_id, rgn_name, species, stock, stock_id, display_name) %>%
  complete(year = 1998:2017) %>%
  arrange(year) %>%
  mutate(mean_catch = zoo::rollapply(catch, 3, mean, fill = NA, align = 'right')) %>% ## create a new column `mean_catch` with rolling mean of 3 yrs
  filter(year > 2004) %>%
  select(year, rgn_id, rgn_name, species, stock, stock_id, mean_catch, display_name) %>%
  ungroup()

p <- ggplot(gf_data, aes(x = year, y = mean_catch, fill = display_name)) +
  geom_bar(stat = "identity") +
  facet_wrap(~rgn_name, scales = "free_y") +
  theme_bw() +
  theme(legend.text = element_text(size = 6),
        legend.position = "below",
        axis.text = element_text(size = 6),
        axis.text.x = element_text(angle = 45))

plotly::ggplotly(p)
```


Let's look at total regional catch for each species (not stock)

```{r}
#calculate total regional catch per species
species_catch <- gf_data %>%
  group_by(species, year) %>%
  summarize(sp_catch = sum(mean_catch, na.rm=T)) %>%
  ungroup() %>%
  group_by(year) %>%
  mutate(yr_catch = sum(sp_catch),
         catch_prop = sp_catch/yr_catch) %>%
  ungroup() %>%
  filter(year > 2004) 
```

```{r}
p <- ggplot(species_catch, aes(x = year, y = catch_prop, fill = species)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(legend.text = element_text(size = 6))

plotly::ggplotly(p)
```

Clearly atlantic herring is making up the majority of catch! Atlantic herring is primarily a bait fishery, so we need to account for that since this goal is only measuring catch meant for human consumption. We adjust for this below.

# Remove non-food catch

Some of these species are harvested for food as well as other markets like pet food or bait. We want to make sure this goal captures catch meant for human consumption. We have data from NOAA that identifies the amount of catch per species, state and year meant for food, bait, and other markets. This data was cleaned in `prop_catch_food_bait.Rmd`.

```{r tot_catch_as_bait}
prop_data <- read_csv("data/fish_catch_food_prop_rgn.csv")

ggplot(prop_data %>% filter(market == "BAIT"), aes(x = year, y = pounds_live_by_market, color = species)) +
  geom_line() +
  theme_bw() +
  theme(legend.title = element_blank()) +
  labs(x = "",
       y = "Catch (pounds)",
       title = "Total amount of species catch sold as bait \n(not for direct human consumption)")


```

```{r remove_bait_proportion}
toolbox_data <- gf_data %>%
  left_join(prop_data) %>%
  mutate(prop = ifelse(is.na(prop),1,prop),
         mean_catch_times_prop = mean_catch*prop) %>%
  filter(!market %in% c("BAIT", "NO MARKET", "CANNED PET FOOD"), #remove bait, pet food and no market records.
         !is.na(mean_catch)) %>%  #remove records with no catch (don't need them)
  select(-market, -pounds_live_by_market, -total_pounds_live, -prop)
```

```{r total_catch_by_ohi_region}

tot <- toolbox_data %>%
  group_by(year, rgn_name) %>%
  summarize(catch = sum(mean_catch_times_prop))

ggplot(tot, aes(x = year, y = catch, fill = rgn_name)) +
  geom_area() +
  scale_fill_viridis_d() +
  theme_bw() +
  theme(legend.title =  element_blank()) +
  labs(x = "",
       y = "Pounds",
       title = "Total catch in pounds meant for human consumption, by OHI region")
```

# Clean up display names

The raw data from NMFS comes with species names displayed like "SCALLOP, SEA" instead of Sea Scallop. This little fix changes the species names to a more user-friendly version.

```{r clean_up_names}
names <- read_csv("data/display_names.csv")

toolbox_data <- toolbox_data %>%
  left_join(names)

```


# Save to toolbox

```{r save_to_toolbox}
# save to toolbox
write.csv(toolbox_data, file = file.path(dir_calc, "layers/fis_meancatch.csv"))
```









