---
title: "Compare RAM and NOAA stock assessment data"
author: "Jamie Afflerbach"
date: "2/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Load data

```{r}
nmfs <- read.csv("data/nmfs_stock_assessment_data.csv") %>% select(-X)
ram  <- read.csv("data/fis_ram_stock_assessment_data.csv") %>% select(-X)
```

Prep for matching

```{r}
nmfs_clean <- nmfs %>% 
  separate(stock, into = c("species", "location"), sep = " - ") %>%
  spread(key = indicator, value = value) %>%
  rename(nmfs_ffmsy = f_fmsy, nmfs_bbmsy = b_bmsy) %>%
  mutate(species = tolower(species),
         location = str_replace(location, " \\| Asmt & Status", ""),
         location = str_replace(location, " \\| Asmt", "")) %>%
  mutate(species = case_when(
    species == "goosefish" ~ "monkfish",
    TRUE ~ as.character(species)
  ))
```

```{r}
ram_clean <- ram %>%
  rename(location = areaname) %>%
  mutate(species = tolower(commonname),
         location = str_replace(as.character(location), "/Mid Atlantic", "/ Mid-Atlantic"),
    indicator = case_when(
    metric == "B/Bmsy" ~ "b_bmsy",
    metric == "F/Fmsy" ~ "f_fmsy"
  )) %>%
  select(year, species, location, indicator, value) %>%
  spread(indicator, value) %>%
  rename(ram_bbmsy = b_bmsy, 
         ram_ffmsy = f_fmsy) %>%
  filter(year > 2000)

#join

both <- nmfs_clean %>%
  full_join(ram_clean, by = c("species", "year", "location"))
```

Plot NMFS then RAM then compare

```{r}
ggplot(nmfs_clean %>% gather(indicator, value, -species, -location, -year), 
       aes(x = year, y = value, color = indicator)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line() +
  facet_wrap(~species+location, scales = "free_y") +
  theme_bw()
```

 
```{r}
ggplot(ram_clean %>% gather(indicator, value, -species, -location, -year), 
       aes(x = year, y = value, color = indicator)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line() +
  facet_wrap(~species+location, scales = "free_y") +
  theme_bw()
```
 
 
```{r, fig.width = 14, fig.height = 10}
plot_df <- both %>%
  gather(indicator, value, -species, -location, -year) %>%
  filter(!is.na(value))


ggplot(plot_df, aes(x = year, y = value, color = indicator)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line() +
  facet_wrap(~species+location, scales = "free_y") +
  theme_bw()
ggsave(file = "figs/ram_nmfs_stock_assessment_comparison.pdf")

```
 

Why is Atlantic menhaden in RAM but not NMFS
Why is there no Scup in NMFS
Why is there such a big difference in Winter flounder for George's bank b/w the indicators (years 2011, 2014)
Why is there such a big difference between the two data sources for Summer flounder?
Why isn't monkfish, spiny dogfish or striped bass in the NMFS data?








