---
title: "New Hampshire fisheries scores"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Setup

```{r setup, message = F, warning = F}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)
source('~/github/ne-prep/src/R/common.R')

```

# Load data

```{r}

#catch
catch <- read_csv("data/nmfs_spatial_catch_by_ohi_rgn.csv") %>% filter(rgn_id == 9)

#proportion
prop <- read_csv("data/fish_catch_food_prop_rgn.csv")

#mean catch
mean_catch <- read_csv("~/github/ne-scores/region/layers/fis_meancatch.csv") %>%
  filter(rgn_id == 9)
 
#stock scores
stock_scores <- read_csv("~/github/ne-scores/region/layers/fis_stockscores.csv") %>%
  filter(rgn_id == 9) %>%
  rename(species = nmfs_original_species)

```

# Visualize

New Hampshire catch

This data is created by multiplying the amount of catch reported for statistical area 513 by 1.75%. Our NH region is just 1.75% of th entire stat area 513. So we attribute 1.75% of all catch reported in 513 for NH.

```{r}

ggplot(catch, aes(x = year, y = catch, fill = display_name)) +
  geom_col() +
  theme_bw() +
  theme(legend.position = "none")

plotly::ggplotly()
```


We implement a rolling mean for catch to try and adjust for variability. Right now, we just use a 3 year rolling mean. So the catch for 2015 is actually the average catch from 2013, 2014 and 2015.

```{r}
ggplot(mean_catch, aes(x = year, y = mean_catch, fill = display_name)) +
  geom_col() +
  theme_bw() +
  theme(legend.position = "none")

plotly::ggplotly()
```

Regional proportion for species that are harvested for non-food uses as well. Not all species are included here. Just those that have some proportion of their total dedicated towards non-human consumption. 

Notice some species are not in every year. This **has not been gapfilled**. 

```{r}
#filter prop for food
food <- prop %>% filter(market == "FOOD")

ggplot(food, aes(x = year, y = prop)) +
  geom_col() +
  theme_bw() +
  theme(legend.position = "none") +
  facet_wrap(~species)

plotly::ggplotly()
```

To see if it matters, here I'm plotting these 13 species for NH

```{r}
nh_spp <- catch %>% filter(species %in% food$species)

ggplot(nh_spp, aes(x = year, y = catch)) +
  geom_col() +
  facet_wrap(~display_name) +
  theme_bw()
```

As suspected, I don't think these food vs bait proportions matter except for Herring, and we have all years for Herring.

Here is the mean catch data after it is multiplied by proportion for food. We really see the drop in 2010. 

```{r}
ggplot(mean_catch, aes(x = year, y = mean_catch_times_prop, fill = display_name)) +
  geom_col() +
  theme_bw() +
  theme(legend.position = "none")

plotly::ggplotly()
```

Now lets look at catch over time along with stock scores

```{r}

catch_status <- catch %>%
    left_join(stock_scores) %>%
    filter(!is.na(score)) %>% #remove stocks with no stock scores **THIS MIGHT NEED TO CHANGE IF WE WANT TO KEEP THESE STOCKS AND GAPFILL INSTEAD**
    group_by(year, rgn_id) %>%
    mutate(SumCatch = sum(catch, na.rm=T)) %>% #calculate total catch per region per year
    ungroup() %>%
    rowwise() %>%
    mutate(wprop = catch / SumCatch,  #calculate proportional catch
           weighted_score = sum(score * wprop)) 

```

```{r}
ggplot(catch_status, aes(x = year, y = wprop, fill = score)) +
  geom_col() +
  scale_fill_gradient2(low = "red", high = "forestgreen", mid = "yellow", midpoint = 0.5) +
  theme_bw()
```


```{r, fig.width=10, fig.height=6}
ggplot(catch_status, aes(x = year, y = score, color = species, size = wprop)) +
  geom_line() +
  theme_bw() +
  theme(legend.position = "bottom")
```

