---
title: 'OHI-Northeast: Fishing Engagement'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/', message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

library(readxl)
```

# Summary

This script creates the Fishing Engagement data layer for use in the Sense of Place goal. We use two different data sets from NOAA's Soc

***

# Data Source

**NOAA Social Vulnerability**
**Downloaded**: August 2019

**Description**: Calculates Social Vulnerability based on social vulnerability, fishing vulnerability, inundation risk, and gentrification vulnerability  

**Time range**: 2009-2016

**Format**: Tabular

***

# Load Data

NOAA Social Vulnerability Index
```{r}
raw_svi_2009 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=1) %>%
  select(year, GEO_NAME, ComEng_ct,  RecRel_ct)
raw_svi_2010 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=2) %>%
  select(year, GEO_NAME, ComEng_ct,  RecRel_ct)
raw_svi_2011 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=3) %>%
  select(year, GEO_NAME, ComEng_ct,  RecRel_ct)
raw_svi_2012 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=4) %>%
  select(year, GEO_NAME, ComEng_ct,  RecRel_ct)
raw_svi_2013 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=5) %>%
  select(year, GEO_NAME, ComEng_ct,  RecRel_ct)
raw_svi_2014 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=6) %>%
  select(year, GEO_NAME, ComEng_ct,  RecRel_ct)
raw_svi_2015 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=7) %>%
  select(year, GEO_NAME, ComEng_ct = ComEng_15ct,  RecRel_ct)
raw_svi_2016 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=8) %>%
  select(year, GEO_NAME, ComEng_ct,  RecRel_ct)
raw_svi_categories <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=11)
```

Load Massachusetts counties to split out the North/South regions
```{r}
MA_counties <- read.csv("~/github/ne-prep/src/tables/MA_counties.csv") %>% 
  select(-X)
MA_cities <- read_csv(file.path(dir_anx, "_raw_data/MA_gov/list-cities-massachusetts.csv")) %>% select(Name, County) %>% 
  rename(city = Name)
```

# Data Clean

## Assign MA communities

```{r}
ma_communities <- raw_svi_2009 %>% 
  select(GEO_NAME) %>% 
  separate(GEO_NAME, c("city", "state"), sep= ", ") %>% 
  filter(state == "MA") %>% 
  left_join(MA_cities, by = c("city")) %>% 
  rename(city_community = city)
```


Communities in this dataset were manually assigned to either Massachuasetts-North or Massachusetts-South for the social resilience layer.
```{r}
MA_counties_missing <- read.csv(file.path(dir_anx, "_raw_data/MA_gov/county_missing_filled.csv"))
```


Combine MA counties with their region Ids
```{r}
ma_rgns_by_counties <- ma_communities %>% 
  filter(!is.na(County)) %>% 
  left_join(MA_counties, by = c("County")) %>% 
  rbind(MA_counties_missing) 
```

Do the same data tidying for every year (2009-2016) and combine into a single dataframe
```{r}
out <- data.frame()
for(i in 2009:2016){
  df <- eval(as.name(paste0("raw_svi_", i)))
  
  clean <- df %>%
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  filter(state %in% c("MA", "ME", "NY", "RI", "CT", "NH")) %>% 
  left_join(rgn_data, by = c("state" = "state_abv")) %>%
  left_join(ma_rgns_by_counties) %>%
  mutate(rgn_id = case_when(
    state == "MA" & !is.na(County) ~ rgn_id,
    state != "MA" ~ rgn_id
  )) %>%
  filter(!is.na(rgn_id)) %>%
  select(year, city_community, state, ComEng_ct, RecRel_ct, rgn_name, rgn_id)
  
  out <- bind_rows(out, clean)
}
```

Now we want to calculate scores between 0 and 1 and then average by region id and year.

```{r}
rescaled_df <- out %>%
  gather(key = "layer", value = "index_score", -year, -city_community, -state, -rgn_name, -rgn_id) %>%
  mutate(layer = ifelse(layer == "ComEng_ct", "commercial", "recreational"), 
         score = index_score/4) %>%
  group_by(year, rgn_name, rgn_id, layer) %>%
  summarize(mean_score = mean(score))
  
```

# Visualize

```{r}
ggplot(rescaled_df, aes(x = year, y = mean_score, color = rgn_name)) +
  geom_line() +
  facet_wrap(~layer) +
  theme_bw()
```


# Save layer to toolbox

```{r save}
```



