---
title: 'OHI-Northeast: Fishing pressure layers'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')

#libraries
library(raster)
library(readxl)
```

#Summary

***

#Data Source

**Reference**: Watson, R. A. and Tidd, A. 2018. Mapping nearly a century and a half of global marine fishing: 1869–2015. Marine Policy, 93, pp. 171-177. [(Paper URL)](https://doi.org/10.1016/j.marpol.2018.04.023)

**Downloaded**: July 3, 2019 from [IMAS portal](http://data.imas.utas.edu.au/portal/search?uuid=ff1274e1-c0ab-411b-a8a2-5a12eb27f2c0) - click on download tab, step 3

**Description**:  Global fisheries landings data per cell separated by Industrial versus Non-Industrial catch, IUU, and discards.

**Native data resolution**:   

**Time range**: 1950 - 2015

**Format**:  CSV format

**Additional Information**: [Metadata](http://metadata.imas.utas.edu.au/geonetwork/srv/eng/metadata.show), [Supplementary Material](https://ars.els-cdn.com/content/image/1-s2.0-S0308597X18300605-mmc1.docx)


***
  
# Load data

The Watson fishing data is held on Mazu for the global OHI. 

```{r}
#list all files in the watson raw data folder
files <- list.files(file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2019"), full.names = T)
```

We want the industrial and non-industrial catch from 2000 onwards.

```{r}
all_files <- files[str_detect(files, paste(c("2000_2004", "2005_2009", "2010_2014", "2015_2015"), collapse = '|'))]
```

Load `.rds` files

```{r}
index_ind  <- read_csv(file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2019/IndexInd.csv"))
index_nind <- read_csv(file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2019/IndexNInd.csv"))
cells      <- read_excel(path = file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2019/Codes.xlsx"), sheet = "Cell") 
```


### Get Northeast cells

We don't need all data, just the data for our region. Using the cells information, crop it to our region of interest and keep a vector of `ne_cells` to use for querying data.

```{r}
cells_raster <- cells %>%
  rename(x = LonCentre, y = LatCentre, z = Cell) %>% #I renamed these xyz to use in the rasterFromXYZ() function below
  select(x,y,z) %>%
  rasterFromXYZ() %>%
  crop(wgs_ext)

#need to set CRS
crs(cells_raster) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" 

plot(cells_raster)

ne_cells <- getValues(cells_raster)

#vector of all cell ids in our region of interest
ne_cell_ids <- ne_cells[!is.na(ne_cells)]
```


Tidy Fisheries Files:
1. Combine the Master Index and Spatial Cells with the CatchInd and CatchNInd files.
2. Save each year of data as a separate file into: "globalprep/prs_fish/v2019/int/annual_catch"

**Function for Combining & Saving Files**: Create function to read in each file name, combine with Index and Cells data frame, and save each year of catch data into a separate file in mazu.

```{r tidy, eval=FALSE}
## Set function to read in each file, combine with Index.csv and Cells.csv, and save each year of data back into mazu
combine_fis <- function(x) {
  #x <- all_files[1]
 
  #get the right index file depending on industrial or non industrial
 if(str_detect(basename(x), "CatchInd")){
    index <- index_ind
  }else{
    index <- index_nind
  }
  
  ## Read in the catch data
  ## Create a total Landings column
  ## Join with master and spatial cells file
  df <- readRDS(x) %>%
    dplyr::filter(Cell %in% ne_cell_ids) %>%
    dplyr::mutate(Landings = Reported+IUU) %>% 
    dplyr::left_join(index, by = "ID") %>% 
    select(ID, Cell, Landings, Year = IYear, Cell, Discards = Discards.x)
  
  ## Save each individual year as a single file
  five_years <- sort(unique(df$Year)) 
  
  for(yr in five_years){
    print(yr) # will show you your progress
    #yr = 2014
    single_yr_df <- df %>%
      filter(Year == yr)
    
    ## Save files with prefix CatchInd or CatchNInd
    ## Remove the suffix starting with '_' to get CatchInd or CatchNInd
    ind_Nind <- basename(x) %>% 
      tools::file_path_sans_ext() %>% 
      str_remove("d.*") # remove everything after the first underscore
    
    write_csv(single_yr_df, paste0(dir_anx, "/prs_fish/data/", ind_Nind, "d_", yr, ".csv"))
    
  }
}

# apply to all files

map(all_files, combine_fis)
```

Create annual bycatch and landings raster layers. Combine non-industrial and industrial catch per cell, we do not differentiate.

```{r}
make_catch_rasters <- function(year){
  
  ind  <- read_csv(paste0(dir_anx, "/prs_fish/data/CatchInd_", year, ".csv"))
  nind <- read_csv(paste0(dir_anx, "/prs_fish/data/CatchNInd_", year, ".csv"))
  
  both <- bind_rows(ind, nind) %>%
    group_by(Cell) %>%
    summarize(landings = sum(Landings),
              bycatch  = sum(Discards))
  
  #subs cells_raster with landings and bycatch
  #reproject & resample to 1km (the catch is already at 1 km - in Version 4.0 it comes as catch/km2)
  land_rast <- subs(cells_raster, both, by = "Cell", which = "landings") %>%
    projectRaster(ocean_ne) %>%
    resample(ocean_ne) %>%
    mask(zones)
  
  by_rast   <- subs(cells_raster, both, by = "Cell", which = "bycatch") %>%
    projectRaster(ocean_ne) %>%
    resample(ocean_ne) %>%
    mask(zones)
  
  writeRaster(land_rast, filename = paste0(dir_anx, "/prs_fish/data/landings_", year, ".tif"))
  writeRaster(by_rast, filename = paste0(dir_anx, "/prs_fish/data/bycatch_", year, ".tif"))
}

#apply to all years

map(c(2000:2015), make_catch_rasters)
```

# Make gif for landings and bycatch over time

```{r}

```

# Get reference point

We will use a reference point of the highest + 10% across the time period

```{r reference_point}
#landings
land_stack    <- list.files(file.path(dir_anx, "prs_fish/data/"), pattern = "landings_", full.names = T) %>% stack()
landings_ref  <- max(getValues(land_stack), na.rm = T)*1.1

#bycatch
bycatch_stack <- list.files(file.path(dir_anx, "prs_fish/data/"), pattern = "bycatch_", full.names = T) %>% stack()
bycatch_ref   <- max(getValues(bycatch_stack), na.rm = T)*1.1

```

# Rescale between 0 and 1

```{r rescale}

rescale_rasts <- function(year, type){
  
  r <- raster(paste0(dir_anx, "/prs_fish/data/", type, "_", year, ".tif"))
  
  if(type == "landings"){
    ref <- landings_ref
  }else{
    ref <- bycatch_ref
  }
  rescaled_r <- r/ref
  
  writeRaster(rescaled_r, filename = paste0(dir_anx, "/prs_fish/data/", type, "_rescaled_", year, ".tif"), overwrite = T)
}

d <- expand.grid(year = c(2000:2015), type = c("bycatch", "landings"))
map2(d$year, d$type, rescale_rasts)
```


# Run zonal statistics for each ohi region

Stack each year then run `zonal` statistics to get the mean.

```{r}
#landings
land_stack <- list.files(file.path(dir_anx, "prs_fish/data/"), pattern = "landings_rescaled", full.names = T) %>% stack()
land_z     <- zonal(land_stack, zones) %>% #zonal's default statistic is to take the mean
  as.data.frame() %>%
  gather(key = "year", value = "pressure_score", -zone) %>%
  separate(year, c("type", "year"), sep = "_rescaled_") %>%
  rename(region_id = zone) %>%
  select(-type)

#bycatch
by_stack <- list.files(file.path(dir_anx, "prs_fish/data/"), pattern = "bycatch_rescaled", full.names = T) %>% stack()
by_z     <- zonal(by_stack, zones) %>% #zonal's default statistic is to take the mean
  as.data.frame() %>%
  gather(key = "year", value = "pressure_score", -zone) %>%
  separate(year, c("type", "year"), sep = "_rescaled_") %>%
  rename(region_id = zone) %>%
  select(-type)
```

# Save to toolbox

```{r}
write_csv(land_z, "~/github/ne-scores/region/layers/prs_landings.csv")
write_csv(by_z, "~/github/ne-scores/region/layers/prs_bycatch.csv")
```









***
