---
title: 'OHI-Northeast: Ocean Acidification Pressure Layer'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    fig_caption: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohi-northeast/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary

This script calculates the ocean acidification pressure score.

# Data:

1. East Coast Ocean Acidification Product Suite https://www.coral.noaa.gov/accrete/east-coast-oaps.html

We have downloaded aragonite data from the site. 

Time frame: 2014-2017

The data is downloaded as seperate NetCDF files for each year. Each file has one variable (OmegaAr) with three dimensions (lat, long, and time). The unit for time is months since Janurary 2014 (0-11), but when we convert this data into an array later the month dimesions "turns in" the normal 1-12 because it counting the number of data sets for that dimension.

Since we do not have historic data, we used a base of aragaonite staturation of 1. This is widely used in as the threshold level.

Equation for calculating aragonite satuation change:
$$\Delta \Omega_{year} = \frac{(\Omega_{base} - \Omega_{year})}{(\Omega_{base} - 1)}$$

Note that current is subtracted from the baseline; this way, a reduction in $\Omega$ becomes a positive pressure value. It is then normalized by the current mean state; so a decrease in $\Omega$ while the current state is high indicates less pressure than the same decrease when the current state is near 1. 

$\Delta \Omega_{year}$ is then modified to account for increases in aragonite saturation state (pressure = 0) and arag sat state less than 1 (pressure = 1).

# Methods

## Setup (libraries and loading data)

The main R libraries needed for this analysis are the `raster`, and `ncdf4` packages.
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 8, fig.height = 6, fig.path = 'figs/', message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R') ### an OHINE specific version of common.R

#libraries

library(raster)
library(ncdf4)
library(maps)
library(RColorBrewer)
library(googleVis)
library(maps)
library(parallel)
library(foreach)
library(doParallel)
library(fasterize)
library(rasterVis)
library(here)
library(raster)
library(gstat)
library(scales)

#define paths for the raw data and OA folder held on git-annex on our NCEAS server, Mazu
raw_dir     = file.path(dir_M,'git-annex/neprep/_raw_data')
glob_dir    = file.path(dir_M,'git-annex/globalprep/prs_oa')
ne_dir  = file.path(dir_M,'git-annex/neprep/prs_oa')

cols      = colorRampPalette(brewer.pal(9, 'Spectral'))(255) # rainbow color scheme
```
  

*Helpful website for using NetCDF files http://geog.uoregon.edu/bartlein/courses/geog607/Rmd/netCDF_01.htm*

We need to extract the arrays individually from each year and then convert into a raster file. If we end up using this data I want to figure out how to make this a loop fo future ease of additional years of data.
```{r creating_arays}
## read in the files for each year
nc_2014 = nc_open(file.path(raw_dir,'NOAA_OA/omega_Ar_East_Coast_2014.nc'))
nc_2015 = nc_open(file.path(raw_dir,'NOAA_OA/omega_Ar_East_Coast_2015.nc'))
nc_2016 = nc_open(file.path(raw_dir,'NOAA_OA/omega_Ar_East_Coast_2016.nc'))
nc_2017 = nc_open(file.path(raw_dir,'NOAA_OA/omega_Ar_East_Coast_2017.nc'))
  
## grab the lat and longs; only need to do this once because the lats and longs are the same for every year
lat <- ncvar_get(nc_2014,varid='lat')
long <- ncvar_get(nc_2014,varid='lon')

## Need to grab the fill value ie: the value used when no data is avaliable so. it is -999 for 2014 and -32767 for 2015-2017
fillvalue_2014 <- ncatt_get(nc_2014, "OmegaAr", "_FillValue")
fillvalue_2015 <- ncatt_get(nc_2015, "OmegaAr", "_FillValue")
fillvalue_2016 <- ncatt_get(nc_2016, "OmegaAr", "_FillValue")
fillvalue_2017 <- ncatt_get(nc_2017, "OmegaAr", "_FillValue")

## create an array from the netcdf file for each year
oa.array_2014 <- ncvar_get(nc_2014, "OmegaAr")
oa.array_2015 <- ncvar_get(nc_2015, "OmegaAr")
oa.array_2016 <- ncvar_get(nc_2016, "OmegaAr")
oa.array_2017 <- ncvar_get(nc_2017, "OmegaAr")

# close each net cdf file
nc_close(nc_2014)
nc_close(nc_2015)
nc_close(nc_2016)
nc_close(nc_2017)

# Replace the fill values in each year with NAs
oa.array_2014[oa.array_2014 == fillvalue_2014$value] <- NA
oa.array_2015[oa.array_2015 == fillvalue_2015$value] <- NA
oa.array_2016[oa.array_2016 == fillvalue_2016$value] <- NA
oa.array_2017[oa.array_2017 == fillvalue_2017$value] <- NA
```

The following code:
1. Rasterizes the monthly data
2. Reprojects, crops, resamples, and interpolates using nearest neighbor
3. Saves as monthly tiff files

Need to do this for each year

2014
```{r rastering_monthly_2014}
registerDoParallel(6)

  foreach(month = c(1:12))  %dopar% {
  
  #month = 12
  
  slice <- oa.array_2014[, , month] 
  
  r <- raster(t(slice), xmn=min(long), xmx=max(long), ymn=min(lat), ymx=max(lat), crs=CRS(p4s_wgs84)) %>% 
  flip(direction = 'y') %>%
  projectRaster(crs = us_alb, over = FALSE, progress="text") %>% 
  raster::crop(ne_ext) %>% 
  raster::resample(zones, method = "ngb") %>% 
    writeRaster(filename = sprintf("%s/int/noaa_monthly/v2014/noaa_2014month_%s.tif", file.path(dir_anx, "prs_oa"), month), 
                overwrite = TRUE)
  }
```


## Results nearest neighbor for 2014 just to look (had also done idw interpolation, but chose this one.)
```{r gif_results}
library(animation)

monthly_2014 <- list.files(file.path(dir_anx, 'prs_oa/int/noaa_monthly/v2014'), full.names = T) %>%stack()
                         
names(monthly_2014) <- paste0("2014_month_", substr(names(monthly_2014),12, 13))

animation::saveGIF({
  for(i in 1:nlayers(monthly_2014)){
     plot(ocean_ne,col='cornsilk2',  main=names(monthly_2014[[i]]),axes=F,legend=F)
      # don't forget to fix the zlimits
    plot(monthly_2014[[i]], zlim=c(0,2.3),axes=F, col=cols, add=T) 
      
  }
}, movie.name = 'oa_monthly_2014_nn.gif')
```

2015
```{r rastering_monthly_2015}
registerDoParallel(6)

  foreach(month = c(1:12))  %dopar% {
  
  #month = 12
  
  slice <- oa.array_2015[, , month] 
  
  r <- raster(t(slice), xmn=min(long), xmx=max(long), ymn=min(lat), ymx=max(lat), crs=CRS(p4s_wgs84)) %>% 
  flip(direction = 'y') %>%
  projectRaster(crs = us_alb, over = FALSE, progress="text") %>% 
  raster::crop(ne_ext) %>% 
  raster::resample(zones, method = "ngb") %>% 
    writeRaster(filename = sprintf("%s/int/noaa_monthly/v2015/noaa_2015month_%s.tif", file.path(dir_anx, "prs_oa"), month), 
                overwrite = TRUE)
  }
```

2016
```{r rastering_monthly_2016}
registerDoParallel(6)

  foreach(month = c(01:12))  %dopar% {
  
  #month = 12
  
  slice <- oa.array_2016[, , month] 
  
  r <- raster(t(slice), xmn=min(long), xmx=max(long), ymn=min(lat), ymx=max(lat), crs=CRS(p4s_wgs84)) %>% 
  flip(direction = 'y') %>%
  projectRaster(crs = us_alb, over = FALSE, progress="text") %>% 
  raster::crop(ne_ext) %>% 
  raster::resample(zones, method = "ngb") %>% 
    writeRaster(filename = sprintf("%s/int/noaa_monthly/v2016/noaa_2016month_%s.tif", file.path(dir_anx, "prs_oa"), month), 
                overwrite = TRUE)
  }
```

2017
```{r rastering_monthly_2017}
registerDoParallel(6)

  foreach(month = c(1:12))  %dopar% {
  
  #month = 12
  
  slice <- oa.array_2017[, , month] 
  
  r <- raster(t(slice), xmn=min(long), xmx=max(long), ymn=min(lat), ymx=max(lat), crs=CRS(p4s_wgs84)) %>% 
  flip(direction = 'y') %>%
  projectRaster(crs = us_alb, over = FALSE, progress="text") %>% 
  raster::crop(ne_ext) %>% 
  raster::resample(zones, method = "ngb") %>% 
    writeRaster(filename = sprintf("%s/int/noaa_monthly/v2017/noaa_2017month_%s.tif", file.path(dir_anx, "prs_oa"), month), 
                overwrite = TRUE)
  }
```

Now that we have all the monthly files as tiff files and in the right projection and extent we can calulate an annual arag saturation.

First we need to load and save all the files
```{r read_in_monthly_files}
monthly_files <- c(list.files(file.path(dir_anx, "prs_oa/int/noaa_monthly/v2014"),
                       full.names = TRUE),
              list.files(file.path(dir_anx, "prs_oa/int/noaa_monthly/v2015"),
                       full.names = TRUE),
              list.files(file.path(dir_anx, "prs_oa/int/noaa_monthly/v2016"),
                       full.names = TRUE),
              list.files(file.path(dir_anx, "prs_oa/int/noaa_monthly/v2017"),
                       full.names = TRUE))
```


## Calculate Annual Means

```{r calculate_annual_means}
maxyr <- substr(monthly_files, 60, 63) %>% 
  as.numeric() %>% 
  max()

## stack all rasters for this year, and calc annual mean, then write as raster
registerDoParallel(6)
foreach(yr = c(2014:maxyr)) %dopar% {
  
  files <- monthly_files[str_detect(monthly_files, as.character(yr))]
  
  rast_annual_mean <- stack(files) %>%
    calc(mean, na.rm = TRUE) %>%
    writeRaster(filename = sprintf("%s/int/noaa_annual_mean/oa_annual_%s.tif", file.path(dir_anx, "prs_oa"), yr), 
                overwrite = TRUE)
}
```

## Rescale from 0 to 1

Thresholds:

- 1 aragonite saturation = a pressure of 1. This is a widely accepted saturation lethal threshold for pteropods

- 1.5 aragonite saturation = a pressure of 0.5. This saturation level is when mild dissolution begins to occurr 
(source: https://www.frontiersin.org/articles/10.3389/fmars.2019.00227/full)

- 2.5 aragonite saturation = a pressure of 0. This is the lowest saturation level where few organims are affected source(https://pubs.geoscienceworld.org/gsa/geology/article/37/12/1131/103987/Marine-calcifiers-exhibit-mixed-responses-to-CO2)

This means that we are going to use two different trends for rescaling. One between the first two threshold, and a second between the last two. These linear relationships were calcaculated externally and should not change unless the thresholds change. If this happens they would need to be recalculated
The first linear regression is:
y= -1x + 2

The second linear regression is:
y= -0.5x + 1.25

The maximum aragonite saturation values by year in our regions:
2014: 1.85
2015: 1.85
2016: 2.75
2017: 2.63

```{r reference_point}
ref_1 = 1.5
ref_2 = 2.5 

files = c(list.files(file.path(dir_anx, "prs_oa/int/noaa_annual_mean"), full.names=TRUE))

registerDoParallel(10) 

foreach(file = files) %dopar% { 
  
  file = files[1]
  yr <- str_sub(file, -8,-5)

   rescale <- raster::raster(file) %>%
    calc(fun = function(x){ifelse(x <=1, 1, x)}) %>% 
    calc(fun = function(x){ifelse(x >=1 & x <= ref_1, -1*x + 2, x)}) %>% 
    calc(fun = function(x){ifelse(x >= ref_1 & x <= ref_2, -0.5*x + 1.25, x)}) %>% 
    calc(fun = function(x){ifelse(x>= ref_2, 0, x)})                          
                                  
 plot(rescale)                                     
                        
  writeRaster(rescale, filename=paste0(dir_anx, '/prs_oa/test/oa_rescaled_', yr, ".tif"),overwrite=TRUE)
}
```


Now we want to interpolate using the nearest neighbor technique. We are using the nearest neighbor technique to interpolate
```{r interpolate_noaa} 
#register parallel cores

registerDoParallel(24)

files <- list.files(file.path(dir_anx,'prs_oa/int/noaa_rescaled'), full.names=TRUE)

foreach(file = files) %dopar%{

  #file = files[1]
  r  <- raster(file) #oa raster
  yr <- substr(file, nchar(file)-7, nchar(file)-4)
  xy <- data.frame(xyFromCell(r, 1:ncell(r))) #get xy coords into dataframe
  v  <- getValues(r)                          # get cell values
  tmpdf <- cbind(xy, v)%>%
    filter(!is.na(v))  #create dataframe of x,y, and values. remove NAs (throws error since these are cells we are interpolating over)
  
  mg <- gstat(id = "v", formula = v~1, locations = ~x+y, data=tmpdf,
            nmax=7, set=list(idp = 2)) #define model.
  z <- interpolate(r, mg, progress='text')                            #interpolate across NA cells

  writeRaster(z, filename=paste0(dir_anx, '/prs_oa/int/noaa_rescaled_int/oa_rescaled_int_', yr, ".tif"),overwrite=TRUE)
}
```

Next we are going to resample and mask the raster
```{r}
files <- list.files(file.path(dir_anx,'prs_oa/int/noaa_rescaled_int'), full.names=TRUE)

registerDoParallel(24)

foreach(file = files)%dopar% {
  
#  file <- files[4]
  yr <- substr(file,nchar(file)-7,nchar(file)-4)
  cat(yr)
  x <- raster(file)%>%
  raster::resample(zones, method = 'ngb')%>%
  raster::mask(zones)
       
  writeRaster(x, filename = paste0(dir_anx, '/prs_oa/output/new/oa_prs_layer_', yr, ".tif"), overwrite=TRUE)
}
```

## Results
Create a gif using the new data 2014-2017
```{r gif_results_newdata}
library(animation)

oa_layers <- list.files(file.path(dir_anx, 'prs_oa/output/new'), full.names = T) %>%stack()
                         
names(oa_layers) <- paste0("OA_Pressure_Score_", substr(names(oa_layers),14, 17))

animation::saveGIF({
  for(i in 1:nlayers(oa_layers)){
     plot(ocean_ne,col='cornsilk2',  main=names(oa_layers[[i]]),axes=F,legend=F)
      # don't forget to fix the zlimits
    plot(oa_layers[[i]], zlim=c(0,1),axes=F, col=cols, add=T) 
      
  }
}, movie.name = 'OA_Pressures_scores.gif')
```

### Region scores
Extract data for each region
```{r extract_regional_data}
# read in raster files
pressure_stack <- lapply(list.files(file.path(dir_anx, 'prs_oa/output/new'), full.names=T), raster) %>%
  brick()

# extract data for each region:
regions_stats <- zonal(pressure_stack,  zones, fun="mean", na.rm=TRUE, progress="text")
regions_stats2 <- data.frame(regions_stats)

data <- base::merge(rgn_data, regions_stats, all.y=TRUE, by.x="rgn_id", by.y="zone") %>%
          dplyr::select(-area_km2)%>%
  gather("year", "pressure_score",starts_with("oa_")) 

oa_data <- data %>%
  mutate(year = substr(year, 14, 17)) %>%
  mutate(year = as.numeric(year))

write.csv(oa_data, file.path(dir_anx, "prs_oa/output/oa.csv"), row.names=FALSE)
```


### 2017 scores
```{r map_scores}
last_year <- oa_data%>%
  filter(year==2017)

#map_scores is a function to plot a tmap map of the scores
map_scores(last_year, score_var = last_year$pressure_score, scale_label = "Pressure Score", map_title = "Ocean Acidification Pressure")
```

Changes over time
```{r}
oa_overtime <-ggplot(oa_data)+
  geom_line(aes(x=year, y = pressure_score, color=state))+
  ggtitle("OA Pressure Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 
oa_overtime
```

### Save layer
```{r tidy_for_layer}
oa_prs <- oa_data %>% 
  select(year, rgn_id, rgn_name, pressure_score) %>% 
  rename(score = pressure_score)
```


```{r write_to_scores}
write.csv(oa_prs, file.path(dir_calc, "layers/cc_acid.csv"))
```


Visualzing Actual concentrations of aragonite 
```{r}
# interpolate
registerDoParallel(24)

files <- list.files(file.path(dir_anx,'prs_oa/int/noaa_annual_mean'), full.names=TRUE)

foreach(file = files) %dopar%{

  #file = files[1]
  r  <- raster(file) #oa raster
  yr <- substr(file, nchar(file)-7, nchar(file)-4)
  xy <- data.frame(xyFromCell(r, 1:ncell(r))) #get xy coords into dataframe
  v  <- getValues(r)                          # get cell values
  tmpdf <- cbind(xy, v)%>%
    filter(!is.na(v))  #create dataframe of x,y, and values. remove NAs (throws error since these are cells we are interpolating over)
  
  mg <- gstat(id = "v", formula = v~1, locations = ~x+y, data=tmpdf,
            nmax=7, set=list(idp = 2)) #define model.
  z <- interpolate(r, mg, progress='text')                            #interpolate across NA cells

  writeRaster(z, filename=paste0(dir_anx, '/prs_oa/int/noaa_annual_int/Aragonite_Concentration_', yr, ".tif"),overwrite=TRUE)
}

files <- list.files(file.path(dir_anx,'prs_oa/int/noaa_annual_int'), full.names=TRUE)

#mask
registerDoParallel(24)

foreach(file = files)%dopar% {

  yr <- substr(file,nchar(file)-7,nchar(file)-4)
  cat(yr)
  x <- raster(file)%>%
  raster::resample(zones, method = 'ngb')%>%
  raster::mask(zones)
       
  writeRaster(x, filename = paste0(dir_anx, '/prs_oa/int/noaa_annual_int/Aragonite_Concentration_', yr, ".tif"), overwrite=TRUE)
}

library(animation)

oa_layers <- list.files(file.path(dir_anx, 'prs_oa/int/noaa_annual_int'), full.names = T) %>%stack()
                         
#names(oa_layers) <- paste0("Aragonite_Concentration_", substr(names(oa_layers),25, 28))

animation::saveGIF({
  for(i in 1:nlayers(oa_layers)){
     plot(ocean_ne,col='cornsilk2',  main=names(oa_layers[[i]]),axes=F,legend=F)
      # don't forget to fix the zlimits, zlim=c(-0.5,3)
    plot(oa_layers[[i]],zlim=c(0,3), axes=F, col=cols, add=T) 
      
  }
}, movie.name = 'Aragonite_Concentration.gif')
```

