---
title: 'OHI-Northeast: Mariculture Production Layer'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(tidyverse)
library(kableExtra)
```

## Summary

This script creates the mariculture production layer. This layer measures the total amount of biomass produced through mariculture by state and species. Sustainability scores for each species come from the Seafood Watch program (e.g. [Oysters](https://www.seafoodwatch.org/seafood-recommendations/groups/oysters?q=Oysters&t=Oysters&location=domestic&method=farmed&type=eastern&o=1190987422)) and are used to weight production (i.e. the more production of highly sustainable species will have a more positive impact on final scores).

We also provide an alternate way to score this layer that weights by area of state waters.

## Data

Production data was gathered from available state reports (e.g. CRMC's "[Aquaculture in Rhode Island](http://www.crmc.ri.gov/aquaculture/aquareport15.pdf)", [Massachusetts DF&G reports](chrome-extension://oemmndcbldboiebfnladdacbdfmadadm/https://www.mass.gov/files/documents/2017/08/30/2012-dmf-annual-report.pdf)). Since this data often exists within PDF's and various formats, all production data was entered by hand into a google sheet and subsequently brought into this analysis.

***

## Data cleaning

Since this data often exists within PDF's and various formats, all production data was entered by hand into a google sheet and subsequently brought into this analysis.

```{r production}
#load data
data <- read_csv("data/production.csv", 
                 col_names = FALSE)

#set row 3 as column names
colnames(data) <- data[3,]

#change column names. We have two Pounds columns. I'm differentiating by adding a 2 to the second one, and then fix this a couple rows down
colnames(data) <- c("Year", "Region", "Species", "Tons", "Individuals", "Bags", "Bushels", "Pieces", "Pounds", "Value", "Source ID", NA, "Pounds", "Tons conversion")

#remove first three rows which don't contain information we need
data <- data[4:nrow(data),1:10]

#some weird characters in the Tons column
data$Tons <- gsub("~","", data$Tons)
data$Tons <- gsub(" \\+", "", data$Tons)

value_only <- data %>%
  filter(!is.na(Value) & is.na(Tons) & is.na(Individuals) & is.na(Bags) & is.na(Bushels) & is.na(Pieces) & is.na(Pounds))

#remove these rows from data_clean
data<- setdiff(data, value_only)

data %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover","condensed")) %>%
  scroll_box(width = "100%", height = "400px")
```

<br>

Some of the species are entered with different common names. For example we have "American Oyster", "Eastern Oyster" and "Oysters". This chunk just fixes some of that for oysters, scallops and clams.

```{r}
#convert character to numeric. Clean up species names and convert tons to pounds
data_clean <- data %>%
  mutate(Tons = as.numeric(Tons),
         Individuals = as.numeric(gsub(",","", Individuals)),
         Bags = as.numeric(gsub(",","", Bags)),
         Bushels = as.numeric(gsub(",", "", Bushels)),
         Pieces  = as.numeric(gsub(",", "", Pieces)),
         Pounds  = as.numeric(gsub(",", "", Pounds))) %>%
  mutate(species = 
           case_when(
             Species == "American Oyster" ~ "Oysters",
             Species == "Eastern Oyster" ~ "Oysters",
             Species == "bay scallops" ~ "Scallops",
             Species == "soft shell clams" ~ "Soft Shell Clams"
           )) %>%
  mutate(Species_name_to_use = ifelse(is.na(species), Species, species),
         pounds = Tons*2000,
         Pounds = ifelse(is.na(Pounds), pounds, Pounds)) %>%
  filter(!is.na(Year)) #there are two records for CT where there is no year..
```

### What species do not have tons or pounds information?

A lot of the data is entered as **Individuals**, **Bags** or **Bushels** for these species 

```{r no_tons_or_lbs}
miss_lbs <- data_clean %>%
  filter(is.na(Pounds))

unique(miss_lbs$Species_name_to_use)
```

Unfortunately, we need some sort of weight metric. We can convert these units to pounds using a conversion table that we put together.


#### Weight Conversion Table
```{r}
#load the weight conversion table
conv <- read.csv("data/weight_conversion_table.csv")

conv %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover","condensed"), full_width = F)
```

### Convert to pounds

This chunk uses the conversion table to convert all units to pounds.

```{r}
spp <- unique(conv$Species)

convert_to_pounds <- function(sp) {
sp_conv <- filter(conv, Species == sp)
  
  out <- data_clean %>%
  filter(Species_name_to_use == sp) %>%
  mutate(pounds = 
           case_when(
             !is.na(Individuals) ~ Individuals/sp_conv$Individual,
             !is.na(Bags) ~ Bags/sp_conv$Bag,
             !is.na(Bushels) ~ Bushels/sp_conv$Bushel,
             !is.na(Pounds) ~ Pounds,
             !is.na(Pieces) ~ Pieces/sp_conv$Individual
           )) %>%
    select(Year, Region, Species, Species_name_to_use, pounds)
  
  return(out)
}

hclam <- convert_to_pounds("Hard Clam")
oyster <- convert_to_pounds("Oysters")
quahog <- convert_to_pounds("Quahog")
scallop <- convert_to_pounds("Scallops")
ss_clams <- convert_to_pounds("Soft Shell Clams")
clams <- convert_to_pounds("Clams")

combo <- bind_rows(hclam, oyster, quahog, scallop, ss_clams, clams)
```

### Fix Maine finfish data

Due to weird data reporting for Salmon and Trout, we need to fix Maine's finfish. The `maine_finfish.csv` dataset is used to replace all finfish from Maine.

```{r}
#read in the Maine finfish data
maine_ff <- read_csv("data/maine_finfish.csv") %>%
  rename(pounds = `Whole Pounds`)

ggplot(maine_ff, aes(x = Year, y = pounds, color = Species)) +
  geom_line() +
  theme_bw() +
  labs(title = "Maine finfish production",
       y = "Pounds")
```

Now we have the data cleaned up and can look at production by species and state over time.

```{r prod_by_species, fig.width = 8, fig.height = 6}
remaining_sp <- data_clean %>%
    filter(!is.na(pounds)) %>%
    select(Year, Region, Species, Species_name_to_use, pounds) 

production_df <- full_join(combo, remaining_sp) %>%
    mutate(Year = as.numeric(Year)) %>%
    filter(!Species_name_to_use %in% c("Atlantic Salmon and Trout", "Atlantic Salmon", "Shellfish", "Mussels, clams, quahogs, and other shellfish (not including oysters)")) %>%
    select(Year, Region, Species = Species_name_to_use, pounds) %>%
  bind_rows(maine_ff)

ggplot(production_df, aes(x = Year, y = pounds, color = Species)) +
  geom_line() +
  theme_bw() +
  ylab("Production (pounds)") +
  ggtitle("Mariculture production in the Northeast") +
  facet_wrap(~Region, scales = "free") +
  theme(legend.position = "bottom",
        legend.title=element_blank())
```

## Weight by sustainability scores

Seafood watch provides sustainability scores between 0 and 10. We use these scores to weight production by species.

#### Sustainability Scores
```{r}
#sustainability scores from Seafood Watch
sust_scores <- read_csv("data/species_sust_scores.csv")
sust_scores %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover","condensed", full_width = F))
```

```{r prod_weighted_by_sust_scores}
weighted_prod <- production_df %>%
  left_join(sust_scores, by = c("Species" = "species")) %>%
  mutate(sustainabilityscore = 
           ifelse(Species %in% c("Clams", "Quahog"), 7.01, sustainabilityscore), #assign the same sust score for Hard Clam to Clams and Quahog
         weight_prod = pounds * sustainabilityscore) %>%
  group_by(Year, Region) %>%
  summarize(total_prod = sum(weight_prod))

ggplot(weighted_prod, aes(x = Year, y = total_prod, color = Region)) +
  geom_line() +
  theme_bw() +
  labs(y = "Production * Sustainability Score",
       title = "Total production per state, weighted by sustainability scores")
```

## Calculate annual growth rates

Now that we have production weighted by the sustainability scores we calculate annual growth rates in total production over time. All values above 1 indicate an increase in production. The goal is to be at or above 1 in order to be maintaining growth in production.

```{r prod_growth_rate, fig.width = 6, fig.height = 4}
gr <- weighted_prod %>%
  group_by(Region) %>%
  mutate(prod_avg_3yr = zoo::rollapply(total_prod, 3, FUN = mean, align = "right", na.rm = F, partial = T), #calculate the mean for three years
         last_years_prod = lag(prod_avg_3yr, n = 1),
         growth = total_prod/last_years_prod) %>%
  filter(Year > 2004)

ggplot(gr, aes(x = Year, y = growth, color = Region)) +
  geom_line() +
  geom_hline(yintercept = 1, color = "black") +
  theme_bw() +
  facet_wrap(~Region) +
  labs(y = "Growth rate (above 1 indicates increase)",
       title = "Mariculture production growth") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

## Score

To score each state, we compare the current growth rate to a target of 4% annual growth rate. 

```{r mar_status_by_rgn}

target <- 1.04 #4% growth rate target

score_mar <- gr %>%
  mutate(score = growth/target,
         final_score = 
           case_when(
             score >= 1 ~ 100,
             score < 1 ~ score*100
             ))

ggplot(score_mar, aes(x = Year, y = final_score, color = Region)) +
  geom_line() +
  theme_bw() +
  labs(y = "Score",
       title = "Mariculture scores by state")

```


## Alternate - weight by area

An alternate option for scoring this goal is to weight by area. Since These regions have vastly different operational area, we might want to account for "farmable area".

As of now, we do not have farmed area or farmable area estimates by state. Just for proof of concept, we can use the full state waters as weight. So here we still apply the sustainability scores to the production values, and then divide by state waters (km2).

### Apply area weighting

```{r}
area_weight <- weighted_prod %>%
  left_join(rgn_data, by = c("Region" = "state")) %>%
  mutate(area = as.numeric(as.character(area_km2))) %>%
  group_by(Region) %>%
  mutate(area = sum(area)) %>%
  select(-rgn_name,-rgn_id,-area_km2) %>%
  ungroup() %>%
  distinct() %>%
  mutate(area_wgt_prod = total_prod/area)

ggplot(area_weight, aes(x = Year, y = area_wgt_prod, color = Region)) +
  geom_line() +
  theme_bw() +
  labs(y = "Area weighted production")
```

### Score

```{r}
gr <- area_weight %>%
  group_by(Region) %>%
  mutate(prod_avg_3yr = zoo::rollapply(area_wgt_prod, 3, FUN = mean, align = "right", na.rm = F, partial = T), #calculate the mean for three years
         last_years_prod = lag(prod_avg_3yr, n = 1),
         growth = area_wgt_prod/last_years_prod) %>%
  filter(Year > 2004) %>%
  ungroup() %>%
  mutate(prod_x_4perc = last_years_prod * 1.04) %>%
  group_by(Year) %>%
  mutate(target = max(prod_x_4perc, na.rm = T),
         score1 = area_wgt_prod/target,
         score2 = ifelse(score1 >= 1, 100, score1*100))

ggplot(gr, aes(x = Year, y = score2, color = Region)) +
  geom_line() +
  theme_bw() +
  labs(y = "Score",
       title = "Mariculture scores by state (including area weighting)")
```


