---
title: 'OHI-Northeast: Aquaculture Sustainability Layer'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(tidyverse)
library(kableExtra)
library(readxl)
```

# Summary

This script creates the mariculture sustainability layer. Sustainability scores for each species come from the Seafood Watch program (e.g. [Oysters](https://www.seafoodwatch.org/seafood-recommendations/groups/oysters?q=Oysters&t=Oysters&location=domestic&method=farmed&type=eastern&o=1190987422)) and are used to weight production (i.e. the more production of highly sustainable species will have a more positive impact on final scores).

***

# Scoring sustainability

Seafood watch provides sustainability scores between 0 and 10. We use these scores to weight production by species. For OHI we want these scores to be on a scale from 0 - 1.


```{r}
#sustainability scores from Seafood Watch
sfw_sustainability <- read_excel(file.path(dir_anx, "_raw_data/SeafoodWatch/Seafood_Watch_USAquaculture82319.xlsx"))
```

This is aquaculture data for the entire United States. What species does this include? We will need to filter just for the species in our data (`data/production.csv`).

```{r}
spp <- unique(sfw_sustainability$`Common name`)
spp
```

### Oysters

We know that Eastern Oysters and American Oysters are farmed in the region. 


```{r oyster_score}
oysters <- spp[str_detect(spp, "yster")] #left off the o to get capital and lowercase
oysters
```

Looks like they have Eastern oyster but not American oyster (not sure if this is different than Eastern).

Scores for all oyster species

```{r}
oyst_scores <- sfw_sustainability %>%
  filter(`Common name` %in% oysters) %>%
  select(`Common name`,`Overall Score`)

oyst_scores %>%
  DT::datatable(.)
```
Ok since they are all given a score of **7.98** then we use that value.

### Atlantic salmon

```{r}
salm <- sfw_sustainability %>%
  filter(`Common name` == "Atlantic salmon")  %>%
  select(`Common name`, Method, `Overall Score`)

salm %>%
  DT::datatable(.)
```


The score for Atlantic salmon grown by Indoor recirculating tank was updated at one point. But for our purposes, we are only interested in the Marine net pen data. Since this score has not been updated, we will use the value **4.82** for Atlantic salmon sustainability.

### Scallops

We have bay scallops and scallops in our production data.

```{r}
scallops <- spp[str_detect(spp, "callop")] 
scallops
```

```{r}
sfw_sustainability %>%
  filter(`Common name` %in% scallops)  %>%
  select(`Common name`, Method, `Overall Score`) %>%
  DT::datatable(.)
```

Again, we see that all scallop species have the same overall score of **8.86**. We will use that value.

### Mussels

We have mussels and blue mussels in our production data.

```{r}
mussels <- spp[str_detect(spp, "ussel")] 
mussels
```

```{r}
sfw_sustainability %>%
  filter(`Common name` %in% mussels) %>%
  select(`Common name`, Method, `Overall Score`) %>%
  DT::datatable(.)
```
Again, we see that all mussel species have the same overall score of **8.11**. We will use that value.

### Clams

We have production values for Hard Clams/Quahogs and soft shell Clams.

```{r}
clams <- spp[str_detect(spp, "lam")] 
clams
```

```{r}
sfw_sustainability %>%
  filter(`Common name` %in% clams) %>%
  select(`Common name`, Method, `Overall Score`) %>%
  DT::datatable(.)
```


Seafood Watch doesn't have Quahog or "Clams" but those are both found in the production data. We will assign those two species the same value as Hard Clam.

```{r}
clam_scores <-  sfw_sustainability %>%
  filter(`Common name` %in% c("Hard clam (unspecified)", "Softshell clam")) %>%
  select(`Common name`, Method, `Overall Score`, `Report Status`, `Publish Date`, `Inactive Date`, `Start year`)

DT::datatable(clam_scores)
  
```

It looks like the sustainability scores for Hard clams were updated in 2018. Since this is outside of our time period, we will use the values **8.39** for Hard clams/quahogs, and **7** for Softshell clams.

# Combine sustainability scores

```{r}
sust_scores <- data.frame(species = c("Hard Clam", "Soft Shell Clams", "Blue Mussel", "Oysters", "Scallops", "Atlantic Salmon (marine net pen)"),
                          sustainabilityscore = c(8.39, 7, 8.11, 7.98, 8.86, 4.82)) %>%
  mutate(rescaled = sustainabilityscore/10)

DT::datatable(sust_scores)

#save to data folder
write_csv(sust_scores, "data/species_sust_scores.csv")
```

# Save for toolbox

```{r}
write_csv(sust_scores, "~/github/ne-scores/region/layers/mar_sust_scores.csv")
```


