---
title: 'OHI-Northeast: Eelgrass habitat layer'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary
This script creates the Eelgrass habitat data layer for the Habitats goal (a sub-goal of Biodiversity). 

# Data

There is not great eelgrass data. We identify all locations where eelgrass beds have been recorded since the 1980's and then overlay two water quality indicators.

# Setup

```{r setup, message = F, warning = F}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')

library(magrittr)
library(sf)
library(leaflet)
library(htmlwidgets)
```

# Load eelgrass spatial data

The shapefiles were cleaned and saved in the `eelgrass.Rmd` script.

```{r load_eelgrass_shapefiles}
current <- read_sf(file.path(dir_anx, "hab/current_eelgrass_locations/current_eelgrass_locations.shp")) 
past    <- read_sf(file.path(dir_anx, "hab/past_eelgrass_locations/past_eelgrass_locations.shp")) 
gbe     <- read_sf(file.path(dir_anx, "hab/gbe_eelgrass_locations/gbe_eelgrass_locations.shp")) %>%   lwgeom::st_make_valid() #make the geometry valid
lines   <- read_sf(file.path(dir_anx, "hab/line_eelgrass_locations/line_eelgrass_locations.shp")) 
```

# Grab NCCA site sample data layers

```{r get_ncca_site_info}

#load site info for 05/06 time period
site_info_1 <- readxl::read_excel(file.path(dir_anx, "_raw_data/EPA/nca_siteinformationdata_narschallenge.xlsx")) %>% 
  filter(PSTL_CODE %in% c("MA", "ME", "RI", "NY", "NH", "CT"),
         SAMPYEAR > 2001) %>%
  select(SITE_ID, PSTL_CODE, LON_DD, LAT_DD, WGT_NCA_56)

#load site info for 2010 time period
site_info_2 <- read_csv(file.path(dir_anx, "_raw_data/EPA/assessed_ncca2010_siteinfo.revised.06212016.csv")) %>%
  filter(STATE %in% c("MA", "ME", "RI", "NY", "NH", "CT"),
         NCA_REGION == "East Coast")  %>%
  select(SITE_ID, STATE, LON_DD = ALON_DD, LAT_DD = ALAT_DD, WGT_NCCA10)
```

# Overlap

We are going to overlap the sites with the eelgrass polygons and identify those sites that intersect. We might want to use a buffer

```{r}
#spatialize sites from 05/06 dataset
site_1 <- site_info_1 %>%
  select(SITE_ID, LON_DD, LAT_DD, STATE = PSTL_CODE) %>%
  distinct() %>%
  st_as_sf(coords = c("LON_DD", "LAT_DD"), crs = 4326) %>%
  st_transform(crs = st_crs(current))

#spatialize sites from 2010 dataset
site_2 <- site_info_2 %>%
  select(SITE_ID, STATE, LAT_DD, LON_DD) %>%
  distinct() %>%
  st_as_sf(coords = c("LON_DD", "LAT_DD"), crs = 4326) %>%
  st_transform(crs = st_crs(current))

#combine
sites <- rbind(site_1, site_2) 
```

There are a total of `r length(unique(sites$SITE_ID))` water quality monitoring sites. 

Create an interactive .html map to look at alignment of sites and beds. 

```{r interactive_map, eval = F}
#map the data layers  
interactive_map <-  mapview(past, col.regions = "darkred", layer.name = "Past eelgrass locations") + 
  mapview(gbe, col.regions = "darkorange", layer.name = "Great Bay Estuary") + 
  mapview(current, col.regions = "darkgreen", layer.name = "Current eelgrass locations") +
  mapview(sites, layer.name = "Monitoring sites") 

## save interactive map
#mapshot(interactive_map, url = "eelgrass_map.html")
```

Create a 5km buffer around sites

```{r site_buffer}
sites_w_buffer <- st_buffer(sites, 5000)
```


Intersect eelgrass beds and monitoring sites

```{r intersect}
#intersect
current_sites <- st_intersection(sites_w_buffer, current)
past_sites    <- st_intersection(sites_w_buffer, past)
gbe_sites     <- st_intersection(sites_w_buffer, gbe)
#lines_sites   <- st_intersection(sites, lines) #gives us nothing

#combine
eelgrass_sites_int <- rbind(current_sites, past_sites, gbe_sites) 

eelgrass_sites <- filter(sites, SITE_ID %in% eelgrass_sites_int$SITE_ID)

mapview(current) + mapview(past) + mapview(eelgrass_sites)
```

Now that we have our `r length(unique(eelgrass_sites$SITE_ID))` sites we can grab the data.

```{r}
#load WQI data for 05/06
water_old <- readxl::read_excel(file.path(dir_anx, "_raw_data/EPA/nca_waterchemdata_narschallenge.xlsx")) %>% 
  filter(PSTL_CODE %in% c("MA", "ME", "RI", "NY", "NH", "CT"),
         SITE_ID %in% eelgrass_sites$SITE_ID) %>%
  select(SITE_ID, nitrogen = `DIN Cond`, clarity = `Trans Cond`, LON_DD, LAT_DD, PSTL_CODE, Year = SAMPYEAR) %>%
  left_join(site_info_1) #this adds in the site weights

#load WQI data for 2010
water_new <- read_csv(file.path(dir_anx, "_raw_data/EPA/ncca2010_water_quality_indicator_status.csv")) %>%
  filter(PSTL_CODE %in% c("MA", "ME", "RI", "NY", "NH", "CT"),
         NCCA_REG == "Northeast",
         SITE_ID %in% eelgrass_sites$SITE_ID) %>%
  select(SITE_ID, nitrogen = DIN_COND, clarity = TRANS_COND, LON_DD, LAT_DD, PSTL_CODE) %>%
  mutate(Year = 2010) %>%
  left_join(site_info_2, by = "SITE_ID") %>% #have to join just by SITE_ID. When joining with LAT and LON the WGTs dont show... it's weird
  select(SITE_ID, nitrogen, clarity, LON_DD = LON_DD.x, LAT_DD = LAT_DD.x, PSTL_CODE, Year, STATE, WGT_NCCA10)
```

Combining both datasets and assigning the time periods for each. If the year is 2000 or 2001, that's time period 1. 2005-2006 is time period 2 and 2010 is time period 3.

```{r combine_data}
wqi <- water_old %>%
  bind_rows(water_new) %>%
  mutate(time_period = 
           case_when(
             Year %in% c(2005, 2006) ~ 1,
             Year == 2010 ~ 2
           ))
```


Now assign OHI region names or ids to each of the sites. For the most part, the State in the WQI data is the region, but for Massachusetts we need to be specific about what region the sites belong to. The best way to do this is manually so I use `mapview::mapview()` to identify sites in MA and assign to either Massachusetts-Gulf of Maine or Massachusetts-Virginian.

Look first just at sites identified by the `MA` postal code.

```{r ma_sites}
#filter wqi for sites only in MA
ma <- wqi %>%
  filter(PSTL_CODE == "MA")

#create a simple features (sf) object from the coordinates contained in the dataset
ma_points = st_as_sf(ma, coords = c("LON_DD", "LAT_DD"), 
                 crs = 4326)
#plot
ggplot() +
  geom_sf(data = rgns %>% filter(rgn_id %in% c(7,8)), aes(fill = rgn_id)) +
  geom_sf(data = ma_points, color = 'darkgreen') +
  theme_bw()
```

```{r}
int <- ma_points %>%
  st_transform('+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs') %>%
  st_intersection(rgns)


#find sites that are not in the intersection
s <- setdiff(ma_points$SITE_ID, int$SITE_ID)

#limit to just the missing sites
miss <- ma_points %>%
  filter(SITE_ID %in% s)

miss_gf <- miss %>%
  data.frame() %>%
  mutate(rgn_id = 
           case_when(
             SITE_ID == "NCCA10-1023" ~ 7,
             SITE_ID == "MA05-0006" ~ 7,
             SITE_ID == "MA06-0048" ~ 7,
             SITE_ID == "MA06-0026" ~ 8,
             SITE_ID == "NCCA10-1009" ~ 8,
             SITE_ID == "MA06-0042" ~ 8,
             SITE_ID == "NCCA10-1011" ~ 8,
             SITE_ID == "MA05-0009" ~ 8,
             SITE_ID == "NCCA10-1001" ~ 8,
             SITE_ID == "NCCA10-1058" ~ 8,
             SITE_ID == "NCCA10-2001" ~ 8
           )) %>%
  select(-geometry)
```

Take the info from `int` that we want and combine with these missing sites

```{r}
ma_df <- int %>%
  data.frame() %>%
  select(-area_km2, -geometry, -rgn_name) %>%
  bind_rows(miss_gf)
```

Additional data cleaning that adds the Massachusetts sites back in 

```{r clean_wqi_data}
wqi_clean <- wqi %>%
  select(SITE_ID, nitrogen, clarity, PSTL_CODE, Year, time_period, LON_DD, LAT_DD, WGT_NCA_56, WGT_NCCA10) %>%
  mutate(rgn_name = 
           case_when(
             PSTL_CODE == "ME" ~ "Maine",
             PSTL_CODE == "RI" ~ "Rhode Island",
             PSTL_CODE == "NY" ~ "New York",
             PSTL_CODE == "CT" ~ "Connecticut",
             PSTL_CODE == "NH" ~ "New Hampshire",
             PSTL_CODE == "MA" ~ "NA"
           )) %>%
  left_join(rgn_data) %>%
  left_join(ma_df, by = c("SITE_ID", "time_period", "Year", "nitrogen", "clarity", "PSTL_CODE")) %>%
  mutate(rgn_id = ifelse(is.na(rgn_id.x), rgn_id.y, rgn_id.x)) %>%
  select(-rgn_id.x, -rgn_id.y, -rgn_name, -state.x, -state.y, -area_km2, -WGT_NCA_56.y, -WGT_NCCA10.y) %>%
  left_join(rgn_data) %>%
  select(-area_km2) %>%
  rename(WGT_NCA_56 = WGT_NCA_56.x,
         WGT_NCCA10 = WGT_NCCA10.x) %>%
  filter(Year > 2004) %>% #remove the early years
  as.data.frame()
```

Let's see how the status of each of our indicators is for these sites

```{r}
nit_wc_sites <- eelgrass_sites %>% left_join(wqi_clean, by = "SITE_ID") %>%
  mutate(year = case_when(
    time_period == 1 ~ "2005-2006",
    time_period == 2 ~ "2010"
  ))

#nitrogen
ggplot() +
  geom_sf(data = ne_states %>% filter(!STUSPS %in% c("PA", "VT", "NJ", "NY")), fill = NA) +
  geom_sf(data = nit_wc_sites, aes(color = nitrogen, fill = nitrogen)) +
  facet_wrap(~year) +
  theme_bw()

#water clarity
ggplot() +
  geom_sf(data = ne_states %>% filter(!STUSPS %in% c("PA", "VT", "NJ", "NY")), fill = NA) +
  geom_sf(data = nit_wc_sites, aes(color = clarity, fill = clarity)) +
  facet_wrap(~year) +
  theme_bw()

```


# Get scores using `spsurvey`


Use the [`spsurvey`](https://www.rdocumentation.org/packages/spsurvey/versions/3.4) package to run `cat.analysis` and produce a sub-population estimate that incorporates site weights. 

We also are going to split out the nitrogen and water clarity data.

We remove all sites that have `MISSING` as a status. In this data, New York has 5%, New Hampshire has 3.5% and Maine has 2% missing.

In 2010 only New York has missing sites, a total of 3.35% of area.

**For now I will ignore the MISSING sites**

```{r cat_analysis_0506}
library(spsurvey)
```

```{r}
#select just 0506 data
wqi_0506_din <- wqi_clean %>% 
  select(-clarity) %>%
  rename(siteID = SITE_ID) %>%
  filter(Year < 2007, nitrogen != "MISSING")

# Create the required data frames
sites <- wqi_0506_din %>%
          select(siteID) %>%
          mutate(use = rep(TRUE, nrow(.)))
subpop <- wqi_0506_din %>%
          select(siteID, rgn_name)
design <- wqi_0506_din %>%
          select(siteID,
                 wgt = WGT_NCA_56,
                 xcoord = LON_DD,
                 ycoord = LAT_DD)
data <- wqi_0506_din %>%
        select(siteID, nitrogen)

## from this output we only want to grab. Estimate.P is the percent of area (e.g. 98%) for the coastal waters of a subpopulation (e.g. Maine) in a particular category of condition (e.g. Good)
Water_Column_Status_Estimates_0506_din <- cat.analysis(sites, subpop, design, data) %>%
  select(rgn_name = Subpopulation, Category, perc = Estimate.P) %>%
  mutate(cat_score = 
      case_when(
        Category == "POOR" ~ 0,
        Category == "FAIR" ~ 50,
        Category == "GOOD" ~ 100),
      year = 2006)
```

Do the same for water clarity

```{r}
#select just 0506 data
wqi_0506_wc <- wqi_clean %>% 
  select(-nitrogen) %>%
  rename(siteID = SITE_ID) %>%
  filter(Year < 2007, clarity != "MISSING")

# Create the required data frames
sites <- wqi_0506_wc %>%
          select(siteID) %>%
          mutate(use = rep(TRUE, nrow(.)))
subpop <- wqi_0506_wc %>%
          select(siteID, rgn_name)
design <- wqi_0506_wc %>%
          select(siteID,
                 wgt = WGT_NCA_56,
                 xcoord = LON_DD,
                 ycoord = LAT_DD)
data <- wqi_0506_wc %>%
        select(siteID, clarity)

## from this output we only want to grab. Estimate.P is the percent of area (e.g. 98%) for the coastal waters of a subpopulation (e.g. Maine) in a particular category of condition (e.g. Good)
Water_Column_Status_Estimates_0506_wc <- cat.analysis(sites, subpop, design, data) %>%
  select(rgn_name = Subpopulation, Category, perc = Estimate.P) %>%
  mutate(cat_score = 
      case_when(
        Category == "POOR" ~ 0,
        Category == "FAIR" ~ 50,
        Category == "GOOD" ~ 100),
      year = 2006)
```


Do the same `cat.analysis` but for 2010 data. First for nitrogen.

```{r cat_analysis_2010}
# Create the required data frames
wqi_2010_din <- wqi_clean %>% 
  filter(Year == 2010, nitrogen != "MISSING") %>%
  rename(siteID = SITE_ID)

sites <- wqi_2010_din %>%
          select(siteID) %>%
          mutate(use = rep(TRUE, nrow(.)))
subpop <- wqi_2010_din %>%
          select(siteID, rgn_name)
design <- wqi_2010_din %>%
          select(siteID,
                 wgt = WGT_NCCA10,
                 xcoord = LON_DD,
                 ycoord = LAT_DD)
data <- wqi_2010_din %>%
        select(siteID, nitrogen)

## from this output we only want to grab. Estimate.P is the percent of area (e.g. 98%) for the coastal waters of a subpopulation (e.g. Maine) in a particular category of condition (e.g. Good)
Water_Column_Status_Estimates_2010_din <- cat.analysis(sites, subpop, design, data) %>%
  select(rgn_name = Subpopulation, Category, perc = Estimate.P) %>%
  mutate(cat_score = 
      case_when(
        Category == "POOR" ~ 0,
        Category == "FAIR" ~ 50,
        Category == "GOOD" ~ 100),
      year = 2010) 
```

Then for water clarity.

```{r cat_analysis_2010}
# Create the required data frames
wqi_2010_wc <- wqi_clean %>% 
  filter(Year == 2010, clarity != "MISSING") %>%
  rename(siteID = SITE_ID)

sites <- wqi_2010_wc %>%
          select(siteID) %>%
          mutate(use = rep(TRUE, nrow(.)))
subpop <- wqi_2010_wc %>%
          select(siteID, rgn_name)
design <- wqi_2010_wc %>%
          select(siteID,
                 wgt = WGT_NCCA10,
                 xcoord = LON_DD,
                 ycoord = LAT_DD)
data <- wqi_2010_wc %>%
        select(siteID, clarity)

## from this output we only want to grab. Estimate.P is the percent of area (e.g. 98%) for the coastal waters of a subpopulation (e.g. Maine) in a particular category of condition (e.g. Good)
Water_Column_Status_Estimates_2010_wc <- cat.analysis(sites, subpop, design, data) %>%
  select(rgn_name = Subpopulation, Category, perc = Estimate.P) %>%
  mutate(cat_score = 
      case_when(
        Category == "POOR" ~ 0,
        Category == "FAIR" ~ 50,
        Category == "GOOD" ~ 100),
      year = 2010) 
```

## Create layer

Aggregate by region and get WQI scores

```{r wqi_layer_rgn_scores}
#nitrogen
wqi_rgns_din <- Water_Column_Status_Estimates_0506_din %>%
  rbind(Water_Column_Status_Estimates_2010_din) %>%
  mutate(score = cat_score * (perc/100)) %>%
  filter(Category != "Total") %>%
  group_by(rgn_name, year) %>%
  summarize(wqi_score = sum(score))

#water clarity
wqi_rgns_wc <- Water_Column_Status_Estimates_0506_wc %>%
  rbind(Water_Column_Status_Estimates_2010_wc) %>%
  mutate(score = cat_score * (perc/100)) %>%
  filter(Category != "Total") %>%
  group_by(rgn_name, year) %>%
  summarize(wqi_score = sum(score))

write_csv(wqi_rgns_din, "data/wqi_din_score_rgns.csv")
write_csv(wqi_rgns_wc, "data/wqi_clarity_score_rgns.csv")

ggplot(wqi_rgns_din, aes(x = year, y = wqi_score, color = rgn_name)) +
  geom_line() +
  theme_bw() +
  labs(x = "Year",
       y = "Score",
       color = "Region",
       title = "Dissolved Inorganic Nitrogen")

ggplot(wqi_rgns_wc, aes(x = year, y = wqi_score, color = rgn_name)) +
  geom_line() +
  theme_bw() +
  labs(x = "Year",
       y = "Score",
       color = "Region",
       title = "Water Clarity")
```

## Gap-filling

The plot has guessed the values between 2006 and 2010 but we need to actually assign these before saving this layer. We will assume a simple linear interpolation between years.

```{r gapfill, eval = F}
df_gf_din <- wqi_rgns_din %>%
  complete(year = 2006:2010) %>%
  group_by(rgn_name) %>%
  mutate(score = zoo::na.approx(wqi_score),
         gapfilled = ifelse(is.na(wqi_score), 1, 0),
         metric = "nitrogen")

write_csv(df_gf_din, "data/wqi_din_score_rgn_gf.csv")

df_gf_wc <- wqi_rgns_wc %>%
  complete(year = 2006:2010) %>%
  group_by(rgn_name) %>%
  mutate(score = zoo::na.approx(wqi_score),
         gapfilled = ifelse(is.na(wqi_score), 1, 0),
         metric = "clarity")

write_csv(df_gf_wc, "data/wqi_wc_score_rgn_gf.csv")
```

## Calculate score




## Save layer for toolbox
Average the two layers per region

I need to attach region ID's to each row. I also add the offshore regions 1:4 with NA values for the toolbox to run. 

```{r save_layer, eval = F}
other_rgns <- data.frame(year     = rep(2006:2010, each = 4),
                         rgn_id   = c(1,2,3,4),
                         rgn_name = c("Offshore", "Georges Bank", "Gulf of Maine", "Mid-Atlantic Bight"),
                         score = NA,
                         metric = NA)

out <- df_gf_din %>%
  bind_rows(df_gf_wc) %>%
  left_join(rgn_data) %>%
  select(year, rgn_id, rgn_name, score, metric) %>%
  bind_rows(other_rgns) %>%
  group_by(year, rgn_id, rgn_name) %>%
  summarize(score = mean(score))

ggplot(out %>% filter(!is.na(score)), aes(x = year, y = score, color = rgn_name)) +
  geom_line() +
  theme_bw() +
  labs(x = "Year",
       y = "Score",
       color = "Region",
       title = "Eelgrass habitat")

#save
  write_csv(out, file.path(dir_calc, "layers/hab_eelgrass.csv"))
```



