---
title: 'OHI-Northeast: Maine Coastal Access'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

#libraries
library(tidyverse)
```

---

# Data Source 

**Reference**: NOAA [Environmental Sensitivity Index](https://response.restoration.noaa.gov/maps-and-spatial-data/download-esi-maps-and-gis-data.html)

**Downloaded**: August, 2019

**Description**:  Spatial point data identifying biological and human use resources. 

**Native data resolution**: Point data

**Time range**: NA. There are datasets from different time periods between 2001 and 2016 but we combine all datasets as a single representative layer, therefore the time does not play a role in the final coastal access layer.

**Format**:  Geodatabase/shapefile/point file
---


# Load Data

Inland buffer

We are going to use the 1 mile inland buffer as well as the rgns to identify coastal points. 
```{r}
inland_buffer <- read_sf("~/github/ne-prep/spatial/shapefiles/ohine_inland_1km.shp", quiet = T) %>%
  filter(rgn_name == "Maine") %>%
  select(rgn_name, rgn_id)
```

Combine inland buffer with connecticut state waters

```{r}
me_shape <- rgns %>%
  filter(rgn_name == "Maine") %>%
  select(rgn_name, rgn_id) %>%
  rbind(inland_buffer)
```


### Maine 2016

**Socioeconomic Point Layer**  
A - AIRPORT  
AV - ABANDONED VESSEL
B - BEACH*  
C - CAMPGROUND*  
RE - RENEWABLE ENERGY  
HWR  HISTORICAL WRECK  
HP - HELIPORT  
NOAA - NOAA FACILITY  
EPAF - EPA FACILITY  

```{r}
me_2016_soc <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/ME_NH_2016_GDB/Maine_NewHampshire_2016_ESI.gdb"), layer = "SOCECON_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(me_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE %in% c("B", "C")) %>%
  select(TYPE, rgn_name, rgn_id) %>%
  mutate(year = 2016)
```

**Navigational Marine Point layer**  
A2 - ACCESS POINT*  
BR - BOAT RAMP*  
LD - LOCK AND DAM  
M - MARINA*  
PT - PORT*  
RM - RM  

```{r}
me_2016_nav <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/ME_NH_2016_GDB/Maine_NewHampshire_2016_ESI.gdb"), layer = "NAV_MARINE_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(me_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE %in% c("A2", "BR", "M", "PT")) %>%
  select(TYPE, rgn_name, rgn_id) %>%
  mutate(year = 2016)
```

**Managed Point layer**  
NL = National Landmark*  
NG = National Guard  

```{r}
me_2016_mgd <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/ME_NH_2016_GDB/Maine_NewHampshire_2016_ESI.gdb"), layer = "MANAGED_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(me_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE == "NL") %>%
  select(TYPE, rgn_name, rgn_id) %>%
  mutate(year = 2016)
```

Combine the layers

```{r}
me_2016 <- me_2016_nav %>%
  rbind(me_2016_mgd) %>%
  rbind(me_2016_soc)
```

----

Lets look at the 2016 points

```{r}
library(mapview)
mapview(me_2016, zcol = "TYPE")
```

# Save

```{r save}
write_sf(me_2016, "data/coastal_access_shapefiles/me_coast_access.shp", quiet = TRUE, delete_layer = TRUE)
```



