---
title: 'OHI-Northeast: New Hampshire Coastal Access'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

#libraries
library(tidyverse)
```

---

#Data Source 

**Reference**: NOAA [Environmental Sensitivity Index](https://response.restoration.noaa.gov/maps-and-spatial-data/download-esi-maps-and-gis-data.html)

**Downloaded**: August, 2019

**Description**:  Spatial point data identifying biological and human use resources. 

**Native data resolution**: Point data

**Time range**: NA. There are datasets from different time periods between 2001 and 2016 but we combine all datasets as a single representative layer, therefore the time does not play a role in the final coastal access layer.

**Format**:  Geodatabase/shapefile/point file

***

## Load Data

Inland buffer

We are going to use the 1 mile inland buffer as well as the rgns to identify coastal points. 
```{r}
inland_buffer <- read_sf("~/github/ne-prep/spatial/shapefiles/ohine_inland_1km.shp", quiet = T) %>%
  filter(rgn_name == "New Hampshire") %>%
  select(rgn_name, rgn_id)
```

Combine inland buffer with connecticut state waters

```{r}
nh_shape <- rgns %>%
  filter(rgn_name == "New Hampshire") %>%
  select(rgn_name, rgn_id) %>%
  rbind(inland_buffer)
```


### New Hampshire 2016

**Socioeconomic Point Layer**

CODES
A - AIRPORT
AV - ABANDONED VESSEL
B - BEACH
C - CAMPGROUND
RE - RENEWABLE ENERGY
HWR  HISTORICAL WRECK
HP - HELIPORT
NOAA - NOAA FACILITY
EPAF - EPA FACILITY

```{r}
nh_2016_soc <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/ME_NH_2016_GDB/Maine_NewHampshire_2016_ESI.gdb"), layer = "SOCECON_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(nh_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE %in% c("B", "C")) %>%
  select(TYPE, rgn_name, rgn_id) %>%
  mutate(year = 2016)
```

**Navigational Marine Point layer**

CODES
A2 - ACCESS POINT
BR - BOAT RAMP
LD - LOCK AND DAM
M - MARINA
PT - PORT
RM - RM

```{r}
nh_2016_nav <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/ME_NH_2016_GDB/Maine_NewHampshire_2016_ESI.gdb"), layer = "NAV_MARINE_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(nh_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE %in% c("A2", "BR", "M", "PT")) %>%
  select(TYPE, rgn_name, rgn_id) %>%
  mutate(year = 2016)
```

**Managed Point layer**
CODES
NL = National Landmark
NG = National Guard


```{r}
nh_2016_mgd <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/ME_NH_2016_GDB/Maine_NewHampshire_2016_ESI.gdb"), layer = "MANAGED_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(nh_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE == "NL") %>%
  select(TYPE, rgn_name, rgn_id) %>%
  mutate(year = 2016)
```

## New Hampshire 2004 

> This data set contains human-use resource data for airports, aquaculture sites, archaeological and historic sites, the U.S. Coast Guard station location, groundwater well sites, lock and dam locations, marinas, recreational beaches, recreational fishing, staging (access) areas, popular surfing locations, water intake sites, bridges, and state borders in New Hampshire. Vector points in this data set represent human-use site locations. Vector lines in this data set represent bridges and state borders. Location specific type and source information are stored in relational data tables (described below) designed to be used in conjunction with this spatial data layer. This data set comprises a portion of the Environmental Sensitivity Index (ESI). 

TYPE column  
A - AIRPORT  
HP - HELIPORT  
LD - LOCK OR DAM  
AQ - AQUACULTURE  
B - BEACH*  
RF - RECREATIONAL FISHING  
CG - COAST GUARD  
HS - HISTORICAL SITE*  
M - MARINA*  
WI - WATER INTAKE  
W - WELL  
ST - STAGING SITE  
S2 - SURFING*  

```{r}
nh_2004 <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/NewHamp_2004_Shapefiles/AVPROJ/shape/soceconp.shp"), quiet = T) %>% 
  filter(TYPE %in% c("B", "HS", "M", "S2")) %>% 
  st_set_crs("+init=epsg:4269") %>%
  st_transform(us_alb)
```


Combine the layers

```{r}
nh_2016 <- nh_2016_nav %>%
  rbind(nh_2016_mgd) %>%
  rbind(nh_2016_soc) %>%
  select(-rgn_name)
```

----

Lets look at the 2016 and 2004 points

```{r}
library(mapview)
mapview(nh_2016, zcol = "TYPE") + mapview(nh_2004, zcol = "TYPE")
```


# Combine and Save

We are going to combine all points from each layer and then save as a single Connecticu coastal access shapefile

```{r}
all <- nh_2004 %>%
  mutate(rgn_id = 9,
         year = 2004) %>%
  select(TYPE, rgn_id, year) %>%
  rename(SHAPE = geometry) %>%
  rbind(nh_2016)

write_sf(all, "data/coastal_access_shapefiles/nh_coast_access.shp", quiet = TRUE, delete_layer = TRUE)
```










