---
title: 'OHI-Northeast: NY Beach Access Layer'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

#libraries
library(tidyverse)
```

#Summary

---

#Data Source 

**Reference**: NOAA [Environmental Sensitivity Index](https://response.restoration.noaa.gov/maps-and-spatial-data/download-esi-maps-and-gis-data.html)

**Downloaded**: August, 2019

**Description**:  Spatial point data identifying biological and human use resources. 

**Native data resolution**: Point data

**Time range**: NA. There are datasets from different time periods between 2001 and 2016 but we combine all datasets as a single representative layer, therefore the time does not play a role in the final coastal access layer.

**Format**:  Geodatabase/shapefile/point file

***
  

# Load Data

Inland buffer

We are going to use the 1 mile inland buffer as well as the rgns to identify coastal points. 
```{r}
inland_buffer <- read_sf("~/github/ne-prep/spatial/shapefiles/ohine_inland_1km.shp") %>%
  filter(rgn_name == "New York") %>%
  select(rgn_name, rgn_id)
```

Combine inland buffer with connecticut state waters

```{r}
ny_shape <- rgns %>%
  filter(rgn_name == "New York") %>%
  select(rgn_name, rgn_id) %>%
  rbind(inland_buffer)
```


### Rhode Island, Connecticut, and New York-New Jersey Metropolitan Area*
*Note this doesn't include Long Island...

2001
Socioeconomic Resource Points and Lines
"This data set contains human-use resource data for bridges, state borders, airports, aquaculture sites, beaches, boat ramps, commercial fishing sites, Coast Guard, ferries, historical sites, marinas, water intakes, and wash over sites in Rhode Island, Connecticut, New York, and New Jersey. Vector points in this data set represent human-use site locations. This data set comprises a portion of the Environmental Sensitivity Index (ESI) data for Rhode Island, Connecticut, New York, and New Jersey. ESI data characterize the marine and coastal environments and wildlife by their sensitivity to spilled oil. The ESI data include information for three main components: shoreline habitats, sensitive biological resources, and human-use resources. See also the MGT (Management Area Polygons) data layer, part of the larger Rhode Island, Connecticut, New York, and New Jersey ESI database, for additional human-use information."

TYPE column
A - AIRPORT
AQ - AQUACULTURE
B - BEACH
BR - BOATRAMP
CF - COMMERCIAL FISHING
CG - COAST GUARD
F - FERRY
HS - HISTORICAL CSITE
M - MARINA
WI - WATER INTAKE
WO - WASHOVER

```{r load_esi_data}
socecon_2001 <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/RI_CT_NY_NJ_2001_Shapefiles/shape/soceconp.shp"), quiet = T)
```

Filter

```{r}
ny_se_2001 <- socecon_2001 %>% 
  filter(TYPE %in% c("B", "BR", "M", "F", "HS")) %>% 
  st_set_crs("+init=epsg:4269")
```

Select NY after intersecting with our rgns

```{r}
ny_2001_points <- ny_se_2001 %>%
  st_transform(us_alb) %>%
  st_intersection(ny_shape) %>%
  mutate(year = 2001,
         TYPE = as.character(TYPE)) %>%
  select(TYPE, rgn_name, rgn_id, year)
```

---

### Long Island

Long Island 2009 data

From the metadata:

> This data set contains human-use resource data for aquaculture sites, beaches, commercial fishing, historic sites, marinas, recreational fishing areas, and water intakes for Long Island, New York. Vector points and lines in this data set represent
the human-use site locations. Location-specific type and source information are stored in relational data tables (described below) designed to be used in conjunction with this spatial data layer. This data set comprises a portion of the Environmental Sensitivity Index (ESI) data for Long Island. ESI data characterize the marine and coastal environments and
wildlife by their sensitivity to spilled oil. The ESI data include information for three main components: shoreline habitats, sensitive biological resources, and human-use resources. See also the MGT (Management Area Polygons) data layer, part of the larger Long Island ESI database, for additional human-use information.

TYPE column
AQ - AQUACULTURE
B - BEACH
CF - COMMERCIAL FISHING
RF - RECREATIONAL FISHING
HS - HISTORICAL SITE
M - MARINA
WI - WATER INTAKE

```{r}
li_2009_socecon <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/LongIs_NY_2009_Shapefiles/avproj/shape/soceconp.shp"), quiet = T) %>%
  st_set_crs("+init=epsg:4269") %>%
  st_transform(us_alb) %>%
  st_intersection(ny_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE %in% c("B", "HS", "M")) %>%
  select(TYPE, rgn_name, rgn_id)
```


Long Island Sound 2016 data. Filter for socioeconomic and managed area point layers.

CODES

Socioeconomic Point (those with * are what we use)  
AV - Abandoned Vessel  
LF - Landfill  
HWR - Historic Wreck  
HS - Historical Site*  
B - Beach*  

```{r}
li_2016_socecon <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/LongIslandSound_2016_GDB/LongIslandSound_ESI_2016.gdb"), layer = "SOCECON_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(ny_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE %in% c("B", "HS")) %>%
  select(TYPE, rgn_name, rgn_id)
```

Managed points  
NL = National Landmark  

No need to filter here since this layer includes national landmarks

```{r}
li_2016_mgd <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/LongIslandSound_2016_GDB/LongIslandSound_ESI_2016.gdb"), layer = "MANAGED_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(ny_shape) %>%
  select(TYPE, rgn_name, rgn_id)
```

**Navigational Marine Points** 
M = Marina  
F = Ferry  
BR = Boat Ramp  
A2 = Access point  
PT = Port  

We consider all of these to be coastal access points

```{r}
li_2016_nav <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/LongIslandSound_2016_GDB/LongIslandSound_ESI_2016.gdb"), layer = "NAV_MARINE_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(ny_shape) %>%
  select(TYPE, rgn_name, rgn_id)
```

Combine the 2016 layers

```{r}
li_2016 <- li_2016_nav %>%
  rbind(li_2016_mgd) %>%
  rbind(li_2016_socecon) %>%
  mutate(year = 2016)
```


---

## New York 2016 Data

**Socioeconomic Point Codes** we only use those marked with *
A = Airport  
AS = Archaeological site  
AV - Abandoned Vessel  
LF - Landfill  
HS - Historical Site*  
B = Beach*  
RE = Renewable Energy  
S2 = Surfing*  
OF = Oil Facility  
C = Campground*  

```{r}
ny_2016_socecon <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/NY_NJ_2016_GDB/NY_NJ_ESI_2016.gdb"), layer = "SOCECON_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(ny_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE %in% c("B", "HS", "C", "S2")) %>%
  select(TYPE, rgn_name, rgn_id)
```


**Navigational Marine Points**  
M = Marina*  
F = Ferry*  
BR = Boat Ramp*  
A2 = Access point*  
PT = Port*  
DV = Diving Site  
AN = Anchorage*  
LD = Lock and Dam  

```{r}
ny_2016_nav <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/NY_NJ_2016_GDB/NY_NJ_ESI_2016.gdb"), layer = "NAV_MARINE_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(ny_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(!TYPE %in% c("DV", "LD")) %>% #keep all but lock and dam
  select(TYPE, rgn_name, rgn_id)
```


Combine the 2016 layers

```{r}
ny_2016 <- ny_2016_nav %>%
  rbind(ny_2016_socecon) %>%
  mutate(year = 2016)
```

----

Lets look at the 2016 and 2001 points


```{r, fig.width = 8}
library(mapview)
mapview(ny_2016, col.regions = "red") + mapview(ny_2001_points, col.regions = "blue") + mapview(li_2016, col.regions = "green") + mapview(li_2009_socecon, col.regions = "orange")
```

# Combine and Save

We are going to combine all points from each layer and then save as a single New York coastal access shapefile

```{r}
all <- ny_2016 %>% 
  rename(Shape = SHAPE) %>%
  rbind(li_2016)

write_sf(all, "data/coastal_access_shapefiles/ny_coast_access.shp", quiet = TRUE, delete_layer = TRUE)
```







