---
title: 'OHI-Northeast: Rhode Island Coastal Access'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

#libraries
library(tidyverse)
```

# Data Source
**Reference**: NOAA [Environmental Sensitivity Index](https://response.restoration.noaa.gov/maps-and-spatial-data/download-esi-maps-and-gis-data.html)

**Downloaded**: August, 2019

**Description**:  Spatial point data identifying biological and human use resources. 

**Native data resolution**: Point data

**Time range**: NA. There are datasets from different time periods between 2001 and 2016 but we combine all datasets as a single representative layer, therefore the time does not play a role in the final coastal access layer.

**Format**:  Geodatabase/shapefile/point file

## Load Data

Inland buffer

We are going to use the 1 mile inland buffer as well as the rgns to identify coastal points. 
```{r}
inland_buffer <- st_read("~/github/ne-prep/spatial/shapefiles/ohine_inland_1km.shp", quiet = T) %>%
  filter(rgn_name == "Rhode Island") %>%
  select(rgn_name, rgn_id)
```

Combine inland buffer with connecticut state waters

```{r}
ri_shape <- rgns %>%
  filter(state_name == "Rhode Island") %>%
  select(rgn_name, rgn_id) %>%
  rbind(inland_buffer)
```


### Rhode Island 2016

**Socioeconomic Point Layer**    
HWR - HISTORICAL WRECK  
HS - HISTORICAL SITE*
B - BEACH*  
A - AIRPORT  
AS - ARCHAEOLOGICAL SITE  
F3 - FACILITY  


```{r}
ri_2016_soc <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/Mass_RhodeIsland_2016_GDB/MA_RI_ESI_2016.gdb"), layer = "SOCECON_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(ri_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE %in% c("HS", "B")) %>%
  select(TYPE, rgn_name, rgn_id) %>%
  mutate(year = 2016)
```

**Navigational Marine Point layer**  
A2 - ACCESS POINT*  
BR - BOAT RAMP*  
LD - LOCK AND DAM  
M - MARINA*  
F - FERRY*  
PT - PORT*  

```{r}
ri_2016_nav <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/Mass_RhodeIsland_2016_GDB/MA_RI_ESI_2016.gdb"), layer = "NAV_MARINE_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(ri_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE != "LD") %>%
  select(TYPE, rgn_name, rgn_id) %>%
  mutate(year = 2016)
```

**Managed Point layer**

```{r}
ri_2016_mgd <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/Mass_RhodeIsland_2016_GDB/MA_RI_ESI_2016.gdb"), layer = "MANAGED_POINT", quiet = T) %>%
  st_transform(us_alb) %>%
  st_intersection(ri_shape) 
```

There is just a single MI point in the MANAGED_POINT layer which I believe to be MILITARY...

Combine the layers

```{r}
ri_2016 <- ri_2016_nav %>%
  rbind(ri_2016_soc)
```

---

### Rhode Island 2001 points


**Socioeconomic Points**

Metadata: https://response.restoration.noaa.gov/sites/default/files/esimaps/gisdata/Mass_2001_Meta.pdf

M - MARINA/ANCHORAGE*  
BR - BOAT RAMP*  
B - BEACH*  
WI - WATER INTAKE  
CF - COMMERCIAL FISHING  
AQ - AQUACULTURE  
F - FERRY*   
WO - WASHOVER  
HS - HISTORICAL SITE*  
A - AIRPORT  

```{r}
ri_2001_soc <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/RI_CT_NY_NJ_2001_Shapefiles/shape/soceconp.shp"), quiet = T) %>% 
  st_set_crs("+init=epsg:4269") %>%
  st_transform(us_alb) %>%
  st_intersection(ri_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE %in% c("M", "HS", "BR", "B", "F"))
```

----

Lets look at the 2016 and 2001 points

```{r, fig.width = 8, fig.height = 8}
library(mapview)
mapview(ri_2016, col.regions = "blue") + mapview(ri_2001_soc, col.regions = "red")
```


# Combine and Save

We are going to combine all points from each layer and then save as a single Massachusetts coastal access shapefile

```{r}
all <-ri_2001_soc %>%
  select(TYPE, rgn_name, rgn_id) %>%
  mutate(year = 2001) %>%
  rename(SHAPE = geometry) %>%
  rbind(ri_2016)

write_sf(all, "data/coastal_access_shapefiles/ri_coast_access.shp", quiet = TRUE, delete_layer = TRUE)
```