---
title: "OHIEC: Resilience -Social"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates social resilience values by OHINE region. 

**3. League of Conservation Voters Scorecards**
**Downloaded**: August

**Description**: Scores the actions of senate and house members based on how they vote on environmentally related topics.

**Time range**: 1971-2018

**Format**:  Tabular


``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R
dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')
```

Load Libraries
``` {r}
library(csv)
library(dplyr)
library(tidyverse)
library(readxl)
library(gridExtra)
library(scales)
```


League of Conservation Voters (LCV)

Load in and tidy house of representative scores
```{r}
house_05 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2005-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2005 Score") %>% 
  mutate(congress = "house",
         year = 2005,
         district = District) %>% 
  rename(score = "2005 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_06 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2006-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2006 Score") %>% 
  mutate(congress = "house",
         year = 2006,
         district = District) %>% 
  rename(score = "2006 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_07 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2007-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2007 Score") %>% 
  mutate(congress = "house",
         year = 2007,
         district = District) %>% 
  rename(score = "2007 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_08 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2008-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2008 Score") %>% 
  mutate(congress = "house",
         year = 2008,
         district = District) %>% 
  rename(score = "2008 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_09 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2009-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2009 Score") %>% 
  mutate(congress = "house",
         year = 2009,
         district = District) %>% 
  rename(score = "2009 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_10 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2010-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2010 Score") %>% 
  mutate(congress = "house",
         year = 2010,
         district = District) %>% 
  rename(score = "2010 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_11 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2011-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2011 Score") %>% 
  mutate(congress = "house",
         year = 2011,
         district = District) %>% 
  rename(score = "2011 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_12 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2012-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2012 Score") %>% 
  mutate(congress = "house",
         year = 2012,
         district = District) %>% 
  rename(score = "2012 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_13 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2013-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2013 Score") %>% 
  mutate(congress = "house",
         year = 2013,
         district = District) %>% 
  rename(score = "2013 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_14 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2014-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2014 Score") %>% 
  mutate(congress = "house",
         year = 2014,
         district = District) %>% 
  rename(score = "2014 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_15 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2015-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2015 Score") %>% 
  mutate(congress = "house",
         year = 2015,
         district = District) %>% 
  rename(score = "2015 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_16 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2016-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2016 Score") %>% 
  mutate(congress = "house",
         year = 2016,
         district = District) %>% 
  rename(score = "2016 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")

house_17 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/house/2017-house-scorecard-grid-export.csv"), skip = 6) %>% 
  select(District, Party, "Member of Congress", "2017 Score") %>% 
  mutate(congress = "house",
         year = 2017,
         district = District) %>% 
  rename(score = "2017 Score",
         party = Party,
         name = "Member of Congress") %>% 
  separate(District, c("state", "dis_num"), sep = "-")
```


Combine
```{r}
house <- rbind(house_05, house_06, house_07, house_08, house_09, house_10, house_11, house_12, house_13, house_14, house_15, house_16, house_17)%>% 
  select(-dis_num, -district) %>% 
  group_by(year, state) %>% 
  mutate(score= as.numeric(score)) %>% 
  dplyr::summarise(house_score = mean(score, na.rm=TRUE)) %>% 
  filter(state == "ME" | state == "NH" | state == "MA" | state == "RI" | state == "CT" | state == "NY") 
```

Load in and tidy senate scores
```{r}
senate_05 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2005-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2005 Score") %>% 
  mutate(congress = "senate",
         year = 2005) %>% 
  rename(score = "2005 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_06 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2006-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2006 Score") %>% 
  mutate(congress = "senate",
         year = 2006) %>% 
  rename(score = "2006 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_06 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2006-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2006 Score") %>% 
  mutate(congress = "senate",
         year = 2006) %>% 
  rename(score = "2006 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_07 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2007-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2007 Score") %>% 
  mutate(congress = "senate",
         year = 2007) %>% 
  rename(score = "2007 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_08 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2008-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2008 Score") %>% 
  mutate(congress = "senate",
         year = 2008) %>% 
  rename(score = "2008 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_09 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2009-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2009 Score") %>% 
  mutate(congress = "senate",
         year = 2009) %>% 
  rename(score = "2009 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_10 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2010-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2010 Score") %>% 
  mutate(congress = "senate",
         year = 2010) %>% 
  rename(score = "2010 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_11 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2011-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2011 Score") %>% 
  mutate(congress = "senate",
         year = 2011) %>% 
  rename(score = "2011 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_12 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2012-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2012 Score") %>% 
  mutate(congress = "senate",
         year = 2012) %>% 
  rename(score = "2012 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_13 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2013-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2013 Score") %>% 
  mutate(congress = "senate",
         year = 2013) %>% 
  rename(score = "2013 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_14 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2014-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2014 Score") %>% 
  mutate(congress = "senate",
         year = 2014) %>% 
  rename(score = "2014 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_15 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2015-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2015 Score") %>% 
  mutate(congress = "senate",
         year = 2015) %>% 
  rename(score = "2015 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_16 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2016-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2016 Score") %>% 
  mutate(congress = "senate",
         year = 2016) %>% 
  rename(score = "2016 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 

senate_17 <- read_csv(file.path(dir_anx, "_raw_data/LCV/LCV_scorecards/senate/2017-senate-scorecard-grid-export.csv"), skip = 6) %>% 
  select(State, Party, "Member of Congress", "2017 Score") %>% 
  mutate(congress = "senate",
         year = 2017) %>% 
  rename(score = "2017 Score",
         party = Party,
         name = "Member of Congress",
         state = State) 
```

Combine senate
```{r}
senate <- rbind(senate_05, senate_06, senate_07, senate_08, senate_09, senate_10, senate_11, senate_12, senate_13, senate_14, senate_15, senate_16, senate_17) %>% 
  group_by(year, state) %>% 
  mutate(score= as.numeric(score)) %>% 
  dplyr::summarise(senate_score = mean(score, na.rm=TRUE)) %>% 
  filter(state == "ME" | state == "NH" | state == "MA" | state == "RI" | state == "CT" | state == "NY") 
```

Combine all congress and tidy
```{r}
congress <- left_join(senate, house, by = c("year", "state")) %>% 
  rowwise() %>% 
  mutate(score = mean(senate_score,house_score))
```

Visualize
```{r}
congress_graph <- ggplot(congress) +
  geom_line(aes(x=year, y= score, color = state)) +
  ggtitle("LCV Raw Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

congress_graph
```

Convert to Scores
```{r}
lcv <- congress %>% 
   mutate(score_lcv = scales::rescale(score, to = c(0,1), from=c(0,100)),
         rgn_id1 = case_when(
           state == "ME" ~ "6", 
           state == "NH" ~ "9", 
           state == "MA" ~ "7",
           state == "RI" ~ "11",
           state == "CT" ~ "5",
           state == "NY" ~ "10"),
         rgn_id2 = case_when(
           state == "MA" ~ "8")) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  select(year, rgn_id, rgn_name, score_lcv) 

write.csv(lcv, file = "data/lcv.csv")
```

Visualize
```{r}
lcv_graph <- ggplot(lcv) +
  geom_line(aes(x=year, y= score_lcv, color = rgn_name)) +
  ggtitle("LCV Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

lcv_graph

ggsave("figs/lcv_graph.jpg", width=7, height=5, dpi=300)
```


