---
title: "OHIEC: Social Resilience -Social"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates social resilience values by OHINE region. 


**3. NOAA Social Vulnerability**
**Downloaded**: August 2019

**Description**: Calculates Social Vulnerability based on social vulnerability, fishing vulnerability, inundation risk, and gentrification vulnerability. https://www.st.nmfs.noaa.gov/humandimensions/social-indicators/map  

The raw data was emailed to us by Lisa Colburn (lisa.l.colburn@noaa.gov)

**Time range**: 2009-2016

**Format**: Tabular

# About
Each community is scored in 4 indicators: Social Vulnerability, Gentrification Pressure, Sea Level Rise Vulnerability, Fishing Engagement and Reliance. The factors used to calculate individual scores are:

LabFrc_ct	Labor Force categorical ranking
HsChr_ct	Housing Characteristics categorical ranking
Pvrty_ct	Poverty categorical ranking
PopCom_ct	Population Composition categorical ranking
PerDis_ct	Personal Disruption categorical ranking
HsDis_ct	Housing Disruption categorical ranking
RetMig_ct	Retiree Migration categorical ranking
UrbSpl_ct	Urban Sprawl categorical ranking
ComEng_ct	Commercial Fishing Engagement categorical ranking
ComRel_ct	Commercial Reliance categorical ranking
RecEng_ct	Recreational Engagement categorical ranking
RecRel_ct	Recreational Reliance categorical ranking
SLR_ct	Inundation Risk categorical ranking
SS_ct	Storm Surge Risk categorical Ranking (2015 and 2016 only)

Total overall vulnerability index score was then calculated by taking the average of all 4 indicator scores.

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R
dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(dplyr)
library(tidyverse)
library(readxl)
library(gridExtra)
library(scales)
library(raster)
library(rgdal)
library(gdalUtils)
library(fasterize)
library(stringr)
```


NOAA Social Vulnerability Index
```{r load_data}
raw_svi_2009 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=1)
raw_svi_2010 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=2)
raw_svi_2011 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=3)
raw_svi_2012 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=4)
raw_svi_2013 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=5)
raw_svi_2014 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=6)
raw_svi_2015 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=7)
raw_svi_2016 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=8)
raw_svi_categories <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=11)

MA_counties <- read.csv("~/github/ne-prep/src/tables/MA_counties.csv") %>% 
  select(-X)
MA_cities <- read_csv(file.path(dir_anx, "_raw_data/MA_gov/list-cities-massachusetts.csv")) %>% select(Name, County) %>% 
  rename(city = Name)
```

## Tidying

First we need to divide the regions of Massachusetts North and South. The vulnerability index looked at communities not counties, but most of them do match up to counties. For the rest I manually went and and looked where they were and assigned the community to a region 

Look at Massachusetts "counties" in the SVI and see if they match up with the actual counties
```{r ma_com}
ma_communities <- raw_svi_2009 %>% 
  select(GEO_NAME) %>% 
  separate(GEO_NAME, c("city", "state"), sep= ", ") %>% 
  filter(state == "MA") %>% 
  left_join(MA_cities, by = c("city")) %>% 
  rename(city_community = city)

#pull out the communities with NA for county
county_missing <- ma_communities %>% 
  filter(is.na(County))

write.csv(county_missing, file = "data/county_missing.csv")

#between this line and above I manually went in and assigned North/South to the NA communitites.
#this file is the above file with the counties filled in
MA_counties_missing <- read.csv(file.path(dir_anx, "_raw_data/MA_gov/county_missing_filled.csv"))
```

Look to see where the division of north and south massachusetts is on the cape
```{r regions}
regions_map <- ggplot(rgns) +
  geom_sf(aes(fill= rgn_name))
regions_map
```

Combine 
```{r combine_missing_counties}
ma_rgns_by_counties <- ma_communities %>% 
  filter(!is.na(County)) %>% 
  left_join(MA_counties, by =c("County")) %>% 
  rbind(MA_counties_missing) 
 
write.csv(ma_rgns_by_counties, file = "data/ma_rgns_by_counties.csv")
```

Now that the communities have the proper regions associated with them, we can calculate the score of the 4 indices (Social Vulnerability, Gentrification Pressure, Sea Level Rise Vulnerability, Fishing Engagement and Reliance)

Do the same data tidying for every year (2009-2016)
```{r tidy_2009}
# tidy and find score averages
calc_09 <- raw_svi_2009 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_09 <- calc_09%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_09 <- calc_09 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2009 <- rbind(MA_09,noMA_09)
```

```{r tidy_2010}
calc_10 <- raw_svi_2010 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk))) 

noMA_10 <- calc_10%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_10 <- calc_10 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2010 <- rbind(MA_10,noMA_10)
```

```{r tidy_2011}
calc_11 <- raw_svi_2011 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_11 <- calc_11%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_11 <- calc_11 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2011 <- rbind(MA_11,noMA_11)
```

```{r tidy_2012}
calc_12 <- raw_svi_2012 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_12 <- calc_12%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_12 <- calc_12 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2012 <- rbind(MA_12,noMA_12)
```

```{r tidy_2013}
calc_13 <- raw_svi_2013 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk))) 

noMA_13 <- calc_13%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_13 <- calc_13 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2013 <- rbind(MA_13,noMA_13)
```

```{r tidy_2014}
calc_14 <- raw_svi_2014 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_14 <- calc_14%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_14 <- calc_14 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2014 <- rbind(MA_14, noMA_14)
```

```{r tidy_2015}
calc_15 <- raw_svi_2015 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_15ct, ComRel_15ct, RecEng_ct, RecRel_ct, SLR_ct) %>% # SS_ct even though says there is
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_15ct, ComRel_15ct, RecEng_ct, RecRel_ct )),
         inundation_risk = mean(SLR_ct)) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_15 <- calc_15%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_15 <- calc_15 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2015 <- rbind(MA_15, noMA_15)
```


```{r tidy_2016}
calc_16 <- raw_svi_2016 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% # SS_ct even though says there is
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = mean(SLR_ct)) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_16 <- calc_16%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_16 <- calc_16 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2016 <- rbind(MA_16,noMA_16)
```

Create a df with years 2005-2008 to be added to and gap filled 
```{r create_df_for_gapfill}
gapfil <- c("6", "9", "7", "8", "11", "5", "10") %>% 
  as.data.frame() %>% 
  rename(rgn_id = ".") %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  uncount(4, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(state = NA,
         city_community = NA,
         social_vul = NA,
         gentrification_vul = NA,
         fishing_vul = NA,
         inundation_risk =NA,
         vulnerability_score= NA)
```


## Calculate scores and graphs

Read in the NOAA community population data by region
```{r}
noaa_com_populations <- read_csv(file.path(dir_git, "prep/resilience/social/data/noaa_com_populations.csv")) %>% 
  select(-X1) %>% 
  filter(rgn_id != 12)
```


Combine all years together and calculate the resilience scores for this data for each subregion
```{r combine_all_years}
# join all years together
svi_bind <- rbind(svi_2009, svi_2010, svi_2011, svi_2012, svi_2013, svi_2014, svi_2015, svi_2016, gapfil)

#calculate social score  
svi_social <- svi_bind %>% 
  group_by(state, year, rgn_id, rgn_name) %>% 
  summarise(score_social = mean(social_vul))%>% 
  ungroup() %>% 
  mutate(social_score_scale = scales::rescale(score_social, to = c(0,1)),
         res_social_score = (1-social_score_scale)) %>% 
  select(-state) %>% 
  filter(rgn_id != is.na(rgn_id)) %>% 
  group_by(rgn_name) %>% 
  arrange(year) %>% 
  fill(res_social_score, .direction = "up") 

#calculate gentrification score  
svi_gen <- svi_bind %>% 
  group_by(state, year, rgn_id, rgn_name) %>% 
  summarise(score_gen = mean(gentrification_vul)) %>% 
  ungroup() %>% 
  mutate(gen_score_scale = scales::rescale(score_gen, to = c(0,1)),
         res_gen_score = (1-gen_score_scale)) %>% 
  select(-state) %>% 
  filter(rgn_id != is.na(rgn_id))%>% 
  group_by(rgn_name) %>% 
  arrange(year) %>% 
  fill(res_gen_score, .direction = "up") 

#calculate fishing engagement score  
svi_fish <- svi_bind %>% 
  group_by(state, year, rgn_id, rgn_name) %>% 
  summarise(score_fish = mean(fishing_vul)) %>% 
  ungroup() %>% 
  mutate(fish_score_scale = scales::rescale(score_fish, to = c(0,1)),
         res_fish_score = (1-fish_score_scale)) %>% 
  select(-state) %>% 
  filter(rgn_id != is.na(rgn_id))%>% 
  group_by(rgn_name) %>% 
  arrange(year) %>% 
  fill(res_fish_score, .direction = "up") 

#calculate inundation risk score  
svi_inund <- svi_bind %>% 
  group_by(state, year, rgn_id, rgn_name) %>% 
  summarise(score_inund = mean(inundation_risk))%>% 
  ungroup() %>% 
  mutate(score_inund_scale = scales::rescale(score_inund, to = c(0,1)),
         res_inund_score = (1-score_inund_scale))%>% 
  select(-state) %>% 
  filter(rgn_id != is.na(rgn_id))%>% 
  group_by(rgn_name) %>% 
  arrange(year) %>% 
  fill(res_inund_score, .direction = "up") 

#calculate overall score, weighting all 4 peices equally, and then weighting the score by population size by communities in the region
noaa_vul_int <- left_join(svi_social, svi_gen, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(svi_inund, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(svi_fish, by = c("year", "rgn_id", "rgn_name")) %>% 
  select(year, rgn_id, rgn_name, res_inund_score,res_fish_score, res_social_score, res_gen_score) %>% 
  rowwise() %>% 
  mutate(noaa_vul_score = mean(c(res_inund_score, res_fish_score, res_social_score, res_gen_score)),
         rgn_id = as.numeric(rgn_id)) %>%
  select(year, rgn_id, rgn_name, noaa_vul_score) 
```


Do the same thing as above but for the Northeast region as a whole 
```{r combine_all_years}
#calculate social score  
svi_social_rgn12 <- svi_bind %>% 
  group_by(year) %>% 
  summarise(score_social = mean(social_vul))%>% 
  ungroup() %>% 
  mutate(social_score_scale = scales::rescale(score_social, to = c(0,1)),
         res_social_score = (1-social_score_scale)) %>% 
  mutate(rgn_id = 12,
         rgn_name = "Northeast") %>% 
  arrange(year) %>% 
  fill(res_social_score, .direction = "up") 

#calculate gentrification score  
svi_gen_rg12 <- svi_bind %>% 
  group_by(year) %>% 
  summarise(score_gen = mean(gentrification_vul)) %>% 
  ungroup() %>% 
  mutate(gen_score_scale = scales::rescale(score_gen, to = c(0,1)),
         res_gen_score = (1-gen_score_scale)) %>% 
  mutate(rgn_id = 12,
         rgn_name = "Northeast") %>% 
  arrange(year) %>% 
  fill(res_gen_score, .direction = "up") 

#calculate fishing engagement score  
svi_fish_rgn12 <- svi_bind %>% 
  group_by(year) %>% 
  summarise(score_fish = mean(fishing_vul)) %>% 
  ungroup() %>% 
  mutate(fish_score_scale = scales::rescale(score_fish, to = c(0,1)),
         res_fish_score = (1-fish_score_scale)) %>% 
  mutate(rgn_id = 12,
         rgn_name = "Northeast") %>% 
  arrange(year) %>% 
  fill(res_fish_score, .direction = "up") 

#calculate inundation risk score  
svi_inund_rgn12 <- svi_bind %>% 
  group_by(year) %>% 
  summarise(score_inund = mean(inundation_risk))%>% 
  ungroup() %>% 
  mutate(score_inund_scale = scales::rescale(score_inund, to = c(0,1)),
         res_inund_score = (1-score_inund_scale))%>% 
   mutate(rgn_id = 12,
         rgn_name = "Northeast") %>% 
  arrange(year) %>% 
  fill(res_inund_score, .direction = "up") 

#calculate overall score, weighting all 4 peices equally, and then weighting the score by population size by communities in the region
noaa_vul_rgn12 <- left_join(svi_social_rgn12, svi_gen_rg12, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(svi_inund_rgn12, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(svi_fish_rgn12, by = c("year", "rgn_id", "rgn_name")) %>% 
  select(year, rgn_id, rgn_name, res_inund_score,res_fish_score, res_social_score, res_gen_score) %>% 
  rowwise() %>% 
  mutate(noaa_vul_score = mean(c(res_inund_score, res_fish_score, res_social_score, res_gen_score)),
         rgn_id = as.numeric(rgn_id)) %>%
  select(year, rgn_id, rgn_name, noaa_vul_score) 
```

Combine the state and entire northeast dfs

```{r}
noaa_vul <- dplyr::bind_rows(noaa_vul_int,noaa_vul_rgn12)
write.csv(noaa_vul, file = "data/noaa_vul.csv")
```

Visualize (states only)
```{r graph_indicies}
svi_social_vul_graph <- ggplot(svi_social) +
  geom_line(aes(x=year, y= res_social_score, color = rgn_name)) +
  ggtitle("Social Vulnerability Indices") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

svi_gen_vul_graph <- ggplot(svi_gen) +
  geom_line(aes(x=year, y= res_gen_score, color = rgn_name)) +
  ggtitle("Gentrification Vulnerability Indices") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

svi_fish_vul_graph <- ggplot(svi_fish) +
  geom_line(aes(x=year, y= res_fish_score, color = rgn_name)) +
  ggtitle("Fishing Vulnerability Indices") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

svi_inun_vul_graph <- ggplot(svi_inund) +
  geom_line(aes(x=year, y= res_inund_score, color = rgn_name)) +
  ggtitle("Inundation Risk") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

svi_social_vul_graph
ggsave("figs/svi_social_vul_graph.jpg", width=7, height=5, dpi=300)

svi_gen_vul_graph
ggsave("figs/svi_gen_vul_graph.jpg", width=7, height=5, dpi=300)

svi_fish_vul_graph
ggsave("figs/svi_fish_vul_graph.jpg", width=7, height=5, dpi=300)

svi_inun_vul_graph
ggsave("figs/svi_inun_vul_graph.jpg", width=7, height=5, dpi=300)
```

```{r graph_scores}
noaa_vul_graph <- ggplot(noaa_vul) +
  geom_line(aes(x=year, y= noaa_vul_score, color = rgn_name)) +
  ggtitle("NOAA Vul Social Resilience Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

noaa_vul_graph
ggsave("figs/noaa_vul_graph.jpg", width=7, height=5, dpi=300)
```