---
title: "OHIEC: Resilience -Social"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates social resilience values by OHINE region. 


**3. NOAA Social Vulnerability**
**Downloaded**: August 2019

**Description**: Calculates Social Vulnerability based on social vulnerability, fishing vulnerability, inundation risk, and gentrification vulnerability  

**Time range**: 2009-2016

**Format**: Tabular


``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R
dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

```

Load Libraries
``` {r}
library(csv)
library(dplyr)
library(tidyverse)
library(readxl)
library(gridExtra)
library(scales)
```


NOAA Social Vulnerability Index
```{r}
raw_svi_2009 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=1)
raw_svi_2010 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=2)
raw_svi_2011 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=3)
raw_svi_2012 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=4)
raw_svi_2013 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=5)
raw_svi_2014 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=6)
raw_svi_2015 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=7)
raw_svi_2016 <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=8)
raw_svi_categories <- read_excel(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_updated_082119_Juliette Verstaen.xlsx"), sheet=11)

MA_counties <- read.csv("~/github/ne-prep/src/tables/MA_counties.csv") %>% 
  select(-X)
MA_cities <- read_csv(file.path(dir_anx, "_raw_data/MA_gov/list-cities-massachusetts.csv")) %>% select(Name, County) %>% 
  rename(city = Name)
```

Time for data tidying!

Here I group and average by each region by the four categories (Social Vulnerability, Gentrification Pressure, Sea Level Rise Vulnerability, Fishing Engagement and Reliance). The variable definitions are below.
Variable Definitions:
LabFrc_ct	Labor Force categorical ranking
HsChr_ct	Housing Characteristics categorical ranking
Pvrty_ct	Poverty categorical ranking
PopCom_ct	Population Composition categorical ranking
PerDis_ct	Personal Disruption categorical ranking
HsDis_ct	Housing Disruption categorical ranking
RetMig_ct	Retiree Migration categorical ranking
UrbSpl_ct	Urban Sprawl categorical ranking
ComEng_ct	Commercial Fishing Engagement categorical ranking
ComRel_ct	Commercial Reliance categorical ranking
RecEng_ct	Recreational Engagement categorical ranking
RecRel_ct	Recreational Reliance categorical ranking
SLR_ct	Inundation Risk categorical ranking
SS_ct	Storm Surge Risk categorical Ranking (2015 and 2016 only)


First however, I need to divide out the regions of Massachusetts North and South. The vulnerability index looked at communities not counties, but most of them do match up to counties. For the rest I manually went and and looked where they were and assigned the community to a region 

Look at Masachusetts "counties" in the SVI and see if they match up with the actual counties
```{r}
ma_communities <- raw_svi_2009 %>% 
  select(GEO_NAME) %>% 
  separate(GEO_NAME, c("city", "state"), sep= ", ") %>% 
  filter(state == "MA") %>% 
  left_join(MA_cities, by = c("city")) %>% 
  rename(city_community = city)

#pull out the communities with NA for county
county_missing <- ma_communities %>% 
  filter(is.na(County))

write.csv(county_missing, file = "data/county_missing.csv")

#between this line and above I manually went in and assigned North/South to the NA communitites.
#this file is the above file with the counties filled in
MA_counties_missing <- read.csv(file.path(dir_anx, "_raw_data/MA_gov/county_missing_filled.csv"))
```

Look to see where the division of north and south massachusetts is on the cape
```{r}
regions_map <- ggplot(rgns) +
  geom_sf(aes(fill= rgn_name))
regions_map
```

Combine 
```{r}
ma_rgns_by_counties <- ma_communities %>% 
  filter(!is.na(County)) %>% 
  left_join(MA_counties, by =c("County")) %>% 
  rbind(MA_counties_missing) 
```

Now that the communities have the proper regions associated with them, we can calculate the score of the 4 indices (Social Vulnerability, Gentrification Pressure, Sea Level Rise Vulnerability, Fishing Engagement and Reliance)


Do the same data tidying for every year (2009-2016)
```{r}
# tidy and find score averages
calc_09 <- raw_svi_2009 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_09 <- calc_09%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_09 <- calc_09 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2009 <- rbind(MA_09,noMA_09)
```

```{r}
calc_10 <- raw_svi_2010 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk))) 

noMA_10 <- calc_10%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_10 <- calc_10 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2010 <- rbind(MA_10,noMA_10)
```

```{r}
calc_11 <- raw_svi_2011 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_11 <- calc_11%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_11 <- calc_11 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2011 <- rbind(MA_11,noMA_11)
```

```{r}
calc_12 <- raw_svi_2012 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_12 <- calc_12%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_12 <- calc_12 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2012 <- rbind(MA_12,noMA_12)
```

```{r}
calc_13 <- raw_svi_2013 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk))) 

noMA_13 <- calc_13%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_13 <- calc_13 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2013 <- rbind(MA_13,noMA_13)
```

```{r}
calc_14 <- raw_svi_2014 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% #no SS_ct this year 
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = SLR_ct) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_14 <- calc_14%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_14 <- calc_14 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2014 <- rbind(MA_14, noMA_14)
```

```{r}
calc_15 <- raw_svi_2015 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_15ct, ComRel_15ct, RecEng_ct, RecRel_ct, SLR_ct) %>% # SS_ct even though says there is
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_15ct, ComRel_15ct, RecEng_ct, RecRel_ct )),
         inundation_risk = mean(SLR_ct)) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_15 <- calc_15%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_15 <- calc_15 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2015 <- rbind(MA_15, noMA_15)
```


```{r}
calc_16 <- raw_svi_2016 %>% 
  select(year, GEO_NAME, LabFrc_ct, HsChr_ct, Pvrty_ct, PopCom_ct, PerDis_ct, HsDis_ct, RetMig_ct, UrbSpl_ct, ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct, SLR_ct) %>% # SS_ct even though says there is
  separate(GEO_NAME, c("city_community", "state"), sep= ", ") %>% 
  rowwise() %>% 
  mutate(social_vul = mean(c(PerDis_ct, PopCom_ct, Pvrty_ct, LabFrc_ct, HsChr_ct, HsDis_ct)),
         gentrification_vul = mean(c(RetMig_ct, UrbSpl_ct)), #missing Natural Amenities, will ask Lisa about this
         fishing_vul = mean(c(ComEng_ct, ComRel_ct, RecEng_ct, RecRel_ct )),
         inundation_risk = mean(SLR_ct)) %>% 
  select(year, state, city_community, social_vul, gentrification_vul, fishing_vul, inundation_risk) %>% 
  mutate(vulnerability_score = mean(c(social_vul, gentrification_vul, fishing_vul, inundation_risk)))

noMA_16 <- calc_16%>% 
  filter( state != "MA") %>%  #add rgn_id and rgn_name to rbind with MA later
   mutate(rgn_id = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10")) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))

MA_16 <- calc_16 %>% 
  filter(state == "MA") %>% 
  right_join(ma_rgns_by_counties) %>% 
  select(-County)

svi_2016 <- rbind(MA_16,noMA_16)
```

Create a df with years 2005-2008 to be added to and gap filled 
```{r}
gapfil <- c("6", "9", "7", "8", "11", "5", "10") %>% 
  as.data.frame() %>% 
  rename(rgn_id = ".") %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  uncount(4, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(state = NA,
         city_community = NA,
         social_vul = NA,
         gentrification_vul = NA,
         fishing_vul = NA,
         inundation_risk =NA,
         vulnerability_score= NA)
```


Combine all years together and calculate the resilinece scores for this data
```{r}
svi_bind <- rbind(svi_2009, svi_2010, svi_2011, svi_2012, svi_2012, svi_2013, svi_2014, svi_2015, svi_2015, svi_2016, gapfil)

svi_social <- svi_bind %>% 
  group_by(state, year, rgn_id, rgn_name) %>% 
  summarise(score_social = mean(social_vul)) %>% 
  ungroup() %>% 
  mutate(social_score_scale = scales::rescale(score_social, to = c(0,1)),
         res_social_score = (1-social_score_scale)) %>% 
  select(-state) %>% 
  filter(rgn_id != is.na(rgn_id)) %>% 
  group_by(rgn_name) %>% 
  arrange(year) %>% 
  fill(res_social_score, .direction = "up") 

svi_gen <- svi_bind %>% 
  group_by(state, year, rgn_id, rgn_name) %>% 
  summarise(score_gen = mean(gentrification_vul)) %>% 
  ungroup() %>% 
  mutate(gen_score_scale = scales::rescale(score_gen, to = c(0,1)),
         res_gen_score = (1-gen_score_scale)) %>% 
  select(-state) %>% 
  filter(rgn_id != is.na(rgn_id))%>% 
  group_by(rgn_name) %>% 
  arrange(year) %>% 
  fill(res_gen_score, .direction = "up") 

svi_fish <- svi_bind %>% 
  group_by(state, year, rgn_id, rgn_name) %>% 
  summarise(score_fish = mean(fishing_vul)) %>% 
  ungroup() %>% 
  mutate(fish_score_scale = scales::rescale(score_fish, to = c(0,1)),
         res_fish_score = (1-fish_score_scale)) %>% 
  select(-state) %>% 
  filter(rgn_id != is.na(rgn_id))%>% 
  group_by(rgn_name) %>% 
  arrange(year) %>% 
  fill(res_fish_score, .direction = "up") 

svi_inund <- svi_bind %>% 
  group_by(state, year, rgn_id, rgn_name) %>% 
  summarise(score_inund = mean(inundation_risk))%>% 
  ungroup() %>% 
  mutate(score_inund_scale = scales::rescale(score_inund, to = c(0,1)),
         res_inund_score = (1-score_inund_scale))%>% 
  select(-state) %>% 
  filter(rgn_id != is.na(rgn_id))%>% 
  group_by(rgn_name) %>% 
  arrange(year) %>% 
  fill(res_inund_score, .direction = "up") 

noaa_vul <- left_join(svi_social, svi_gen, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(svi_inund, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(svi_fish, by = c("year", "rgn_id", "rgn_name")) %>% 
  select(year, rgn_id, rgn_name, res_inund_score,res_fish_score, res_social_score, res_gen_score) %>% 
  rowwise() %>% 
  mutate(noaa_vul_score = mean(c(res_inund_score, res_fish_score, res_social_score, res_gen_score))) %>%
  select(year, rgn_id, rgn_name, noaa_vul_score) 

write.csv(noaa_vul, file = "data/noaa_vul.csv")
```

Visulalize
```{r}
svi_social_vul_graph <- ggplot(svi_social) +
  geom_line(aes(x=year, y= res_social_score, color = rgn_name)) +
  ggtitle("Social Vulnerability Indices") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

svi_gen_vul_graph <- ggplot(svi_gen) +
  geom_line(aes(x=year, y= res_gen_score, color = rgn_name)) +
  ggtitle("Gentrification Vulnerability Indices") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

svi_fish_vul_graph <- ggplot(svi_fish) +
  geom_line(aes(x=year, y= res_fish_score, color = rgn_name)) +
  ggtitle("Fishing Vulnerability Indices") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

svi_inun_vul_graph <- ggplot(svi_inund) +
  geom_line(aes(x=year, y= res_inund_score, color = rgn_name)) +
  ggtitle("Inundation Risk") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

svi_social_vul_graph
ggsave("figs/svi_social_vul_graph.jpg", width=7, height=5, dpi=300)

svi_gen_vul_graph
ggsave("figs/svi_gen_vul_graph.jpg", width=7, height=5, dpi=300)

svi_fish_vul_graph
ggsave("figs/svi_fish_vul_graph.jpg", width=7, height=5, dpi=300)

svi_inun_vul_graph
ggsave("figs/svi_inun_vul_graph.jpg", width=7, height=5, dpi=300)
```

```{r}
noaa_vul_graph <- ggplot(noaa_vul) +
  geom_line(aes(x=year, y= noaa_vul_score, color = rgn_name)) +
  ggtitle("NOAA Vul Social Resilience Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

noaa_vul_graph
ggsave("figs/noaa_vul_graph.jpg", width=7, height=5, dpi=300)
```