---
title: "OHIEC: NOAA Community shapefiles with regions"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary
This script takes the NOAA community shapefiles and overlays it on top of the 2010 census population rasters, and then intersects the regions.


Results = population data for just these communities

I think I will need to remove MA and do this separatly sicne I only went through and fixed those communities.
```{r}
library(dplyr)
library(tidyverse)
library(readxl)
library(gridExtra)
library(scales)
library(raster)
library(rgdal)
library(gdalUtils)
library(fasterize)
library(stringr)
library(readr)
```


Read in NOAA communities shapefile 
```{r}
shp <- read_sf(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_FINAL_090519/commondata/2016_mapping/2016_National_Indicators_FINAL_082919.shp")) %>%
  filter(STATEABBR %in% c("CT", "NH", "NY", "ME", "MA", "RI")) %>%
  st_transform(crs = us_alb)
mapview(shp)
```


Read in the population tif files for each state, and then create one NE raster
```{r tif_files}
## read in all states rasters
northeast_rasters <- c(file.path(dir_anx, "_raw_data/NASA/usgrid_data_2010/geotiff/nypop10.tif"),
                    file.path(dir_anx, "_raw_data/NASA/usgrid_data_2010/geotiff/mepop10.tif"),
                    file.path(dir_anx, "_raw_data/NASA/usgrid_data_2010/geotiff/nhpop10.tif"),
                    file.path(dir_anx, "_raw_data/NASA/usgrid_data_2010/geotiff/ctpop10.tif"),
                    file.path(dir_anx, "_raw_data/NASA/usgrid_data_2010/geotiff/ripop10.tif"),
                    file.path(dir_anx, "_raw_data/NASA/usgrid_data_2010/geotiff/mapop10.tif"))

## combine all tif files into one 
mosaic_rasters(gdalfile = northeast_rasters, dst_dataset = file.path(dir_anx, "prs_hab/northeast.tif"), of = "GTiff")

## create one ne raster
northeast_rast <- raster::raster(file.path(dir_anx, "prs_hab/northeast.tif")) %>%
  projectRaster(crs = us_alb, progress="text") 

## check it out!
plot(northeast_rast)
```

```{r}
## create blank raster for rasterizing
r <- raster(crs = us_alb, ne_ext, res = 1000)
r[r] <- 1

## convert to a raster
noaa_comm_rast <- fasterize(shp, r) %>% 
  raster::resample(northeast_rast, method = "ngb") ## match the resolution and extent
# the extent cuts off the top part of the buffer because that's out of Maine

plot(noaa_comm_rast)
```

Now we want to keep only the population density data that falls into our noaa communities
```{r crop_area}
## keep only the areas 1km along the shorelines
ne_pop_coast <- northeast_rast %>% 
 raster::mask(noaa_comm_rast) %>% 
 calc(fun = function(x){ifelse(x < 0, 0, x)})  ## there are 3 negative numbers in this data for some reason (-77.9399, -0.2386,	-0.0178) 

plot(ne_pop_coast)
```
```{r zones_raster}
ne_communities <- read_sf(file.path(dir_anx, "_raw_data/NOAA_Social/2009-2016_National_Indicators_FINAL_090519/commondata/2016_mapping/2016_National_Indicators_FINAL_082919.shp")) %>%
  filter(STATEABBR %in% c("CT", "NH", "NY", "ME", "MA", "RI")) %>%
  st_transform(crs = us_alb) %>% 
  mutate(zone = seq(1:nrow(ne_communities)))

ne_communities_zones <- ne_communities %>% 
  st_set_geometry(NULL) %>% 
  select(STATEABBR, MAPNAME, zone)
  
ne_states_rast <- fasterize(ne_communities, r, field = "zone") %>% 
  raster::resample(northeast_rast, method = "ngb") ## match the resolution
  #mask(buffer_rast)
```

Extract population totals for each region
```{r extract_data}
# extract data for each region:
pop_stats <- zonal(ne_pop_coast, ne_states_rast, fun="sum", na.rm=TRUE, progress="text")

community_pop <- data.frame(pop_stats) %>% 
  left_join(ne_communities_zones, by = c("zone")) %>% 
  select(-zone) %>% 
  rename(population_total = sum,
         state = STATEABBR,
         city_community = MAPNAME) 
```

Do the sum calculations for all regions except MA
```{r}
com_pop_noMA <- community_pop %>% 
  filter(state %in% c("NY", "CT", "RI", "NH", "ME")) %>% 
  group_by(state) %>% 
  summarize(sum_com_pop = sum(population_total)) %>% 
  mutate(rgn_name = case_when(state == "NY" ~ "New York",
                              state == "CT" ~ "Connecticut",
                              state == "RI" ~ "Rhode Island", 
                              state == "NH" ~ "New Hampshire",
                              state == "ME" ~ "Maine")) %>% 
  mutate(rgn_id = case_when(rgn_name == "New York" ~ 10,
                            rgn_name == "Connecticut" ~ 5,
                            rgn_name == "Rhode Island" ~ 11,
                            rgn_name == "New Hampshire" ~ 9,
                            rgn_name == "Maine" ~ 6)) %>% 
  select(-state)
```

Do the sum calculations for all regions except MA
Read in csv file with the MA regions divided up North and South
```{r}
ma_rgns_by_counties <- read_csv(file.path(dir_git, "prep/resilience/social/data/ma_rgns_by_counties.csv")) %>% 
  select(-X1)
```

```{r}
com_pop_MA <- community_pop %>% 
  filter(state == "MA") %>%
  left_join(ma_rgns_by_counties, by= c("state", "city_community")) %>% ## there are a few extra communities in the shapefile that we haven't assigned rngs too. we'll do this manually here
   mutate(rgn_id  = ifelse(city_community == "Marshfield", 7,
                          ifelse(city_community == "Wareham", 8,
                                 ifelse(city_community == "Tisbury", 8,
                                        ifelse(city_community == "Scituate", 7,
                                               ifelse(city_community == "Pembroke", 7,
                                                      ifelse(city_community == "Sandwich", 7,
                                                             ifelse(city_community == "Eastham", 7,
                                                                    ifelse(city_community == "Yarmouth", 8,
                                                                           ifelse(city_community =="Madaket", 8,
                                                                                  ifelse(city_community == "Siasconset", 8,
                                                                                  rgn_id))))))))))) %>% 
  mutate(rgn_name = case_when(rgn_id == 8 ~ "Massachusetts-South",
                              rgn_id == 7 ~ "Massachusetts-North")) %>% 
  group_by(rgn_name, rgn_id) %>% 
  summarize(sum_com_pop = sum(population_total))
```


Combine Massachusetts with all the regions
```{r}
noaa_com_populations_int <- dplyr::bind_rows(com_pop_MA, com_pop_noMA)

#Create a row for entire NE pop size
noaa_com_populations <- noaa_com_populations_int %>% 
  ungroup() %>% 
  mutate(entire_pop = sum(sum_com_pop)) %>% 
  select(entire_pop) %>% 
  mutate(rgn_id = 12,
         rgn_name = "Northeast") %>% 
  rename(sum_com_pop = entire_pop) %>% 
  unique() %>% 
  dplyr::bind_rows(noaa_com_populations_int)
  
write.csv(noaa_com_populations, file = file.path(dir_git, "prep/resilience/social/data/noaa_com_populations.csv"))
```












