---
title: "OHIEC: Resilience -Social"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates social resilience values by OHINE region. 

**2. Opportunity Index**
**Downloaded**: July 2019

**Description**:  Calculates opportunity in each state based on economy, education, health, community

**Time range**: 2012, 2014-2017

**Format**:  PDF


``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R
dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

```

Load Libraries
``` {r}
library(csv)
library(dplyr)
library(tidyverse)
library(readxl)
library(gridExtra)
library(scales)
```

Read in Data
```{r}
raw_oi <- read_xlsx(file.path(dir_anx, "_raw_data/OpportunityIndex/opportunity_index_scores.xlsx")) %>% 
  rename(oi_score = score)
```


Opportunity Index Scores

Create a df with years 2005-2008 to be added to and gap filled 
back filling from last year
```{r}
gapfil <- c("6", "9", "7", "8", "11", "5", "10") %>% 
  as.data.frame() %>% 
  rename(rgn_id = ".") %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  uncount(6, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(oi_score = NA)
```

```{r}
opportunity <- raw_oi %>% 
  group_by(state) %>% 
  mutate(year_n = seq(1:7)) %>% 
  ungroup() %>% 
  mutate(oi_score = if_else(is.na(oi_score) & state == "Connecticut", 57.4, oi_score), #gapfilling for 2011 and 2013. using 2012 values
         oi_score = if_else(is.na(oi_score) & state == "Maine", 56.7, oi_score),
         oi_score = if_else(is.na(oi_score) & state == "Massachusetts", 59.1, oi_score),
         oi_score = if_else(is.na(oi_score) & state == "New Hampshire", 60.4, oi_score),
         oi_score = if_else(is.na(oi_score) & state == "New York", 53.9	, oi_score),
         oi_score = if_else(is.na(oi_score) & state == "Rhode Island", 50.9	, oi_score),
         score = oi_score/target) %>% 
  select(year, state, score) %>% 
  mutate(rgn_id1 = case_when(
    state == "Maine" ~ "6", 
    state == "New Hampshire" ~ "9", 
    state == "Massachusetts" ~ "7",
    state == "Rhode Island" ~ "11",
    state == "Connecticut" ~ "5",
    state == "New York" ~ "10"),
    rgn_id2 = case_when(
    state == "Massachusetts" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  rename(opportunity_score= score)

write.csv(opportunity, "data/opportunity.csv")
```

Visualize
```{r}
opportunity_graph <- ggplot(opportunity) +
  geom_line(aes(x=year, y= opportunity_score, color = rgn_name)) +
  ggtitle("Opportunity Index Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

opportunity_graph

ggsave("figs/opportunity_graph.jpg", width=7, height=5, dpi=300)
```