---
title: "OHINE: Social Resilience - Social Resilience Layer Calculations"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This markdown calculates the social resilience score.

Data Time Ranges:

opportunity.csv 2011-2017

state_compet.csv 2005-2017

noaa_vul.csv 2009-2017 (backfilled in other markdown)

lcv.csv 2005-2017


``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R
dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(dplyr)
library(tidyverse)
library(mapview)
```

# Methods

## Set up, libraries, load data
Read in Data
```{r load_data}
opportunity_short <- read.csv("~/github/ne-prep/prep/resilience/social/data/opportunity.csv") %>% 
  select(-X) 

state_compet <- read.csv("~/github/ne-prep/prep/resilience/social/data/state_compet.csv") %>% 
  select(-X)

noaa_vul_short<- read.csv("~/github/ne-prep/prep/resilience/social/data/noaa_vul.csv") %>% 
  select(-X)

lcv <- read.csv("~/github/ne-prep/prep/resilience/social/data/lcv.csv") %>% 
  select(-X)

state_pop <- read_csv("~/github/ne-prep/src/tables/state_pop.csv")%>% 
  select(-X1)
```

## Tidy Data
Add earlier years and backfill for the opportunity index data

Create a df with years 2005-2008 to be added to and gap filled 
back filling from last year
```{r gap_op}
gapfill_op <- c("6", "9", "7", "8", "11", "5", "10") %>% 
  as.data.frame() %>% 
  rename(rgn_id = ".") %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  uncount(6, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(opportunity_score = NA)
```

Add it on to the opportunity dataset
```{r tidy_op}
opportunity_int <- opportunity_short %>% 
  rbind(gapfill_op) %>% 
  group_by(rgn_name) %>% 
  arrange(year) %>% 
  fill(opportunity_score, .direction = "up") %>% 
  ungroup() %>% 
  mutate(rgn_id = as.integer(rgn_id))

# create a region 12 for entire NE calculation
opportunity_rgn12 <- opportunity_int %>% 
#  mutate(rgn_id = as.numeric(rgn_id)) %>% 
  left_join(state_pop, by = c("rgn_name", "rgn_id")) %>% 
  select(year, opportunity_score, state_name, pop_total) %>% 
  unique() %>%  #get rid of the two MA since they would be double counting in this case
  mutate(score_weighted = opportunity_score*pop_total) %>% 
  group_by(year) %>% 
  dplyr:: summarize(sum_pop = sum(pop_total, na.rm =TRUE),
         sum_pop_weight = sum(score_weighted,na.rm =TRUE)) %>% 
  mutate(opportunity_score = sum_pop_weight /sum_pop,
         rgn_id = 12,
         rgn_name = "Northeast") %>% 
  select(year, opportunity_score, rgn_id, rgn_name)

#merge together into 1!
opportunity <- opportunity_int %>% 
  rbind(opportunity_rgn12)
```

Caculate and add Northeast/rgn 12 to State Competiveness
```{r}
state_compet_rgn12 <- state_compet %>% 
  left_join(state_pop, by = c("rgn_name", "rgn_id")) %>% 
  select(year, state_compet_score, state_name, pop_total) %>% 
  unique() %>%  #get rid of the two MA since they would be double counting in this case
  mutate(score_weighted = state_compet_score*pop_total) %>% 
  group_by(year) %>% 
  dplyr:: summarize(sum_pop = sum(pop_total, na.rm =TRUE),
         sum_pop_weight = sum(score_weighted,na.rm =TRUE)) %>% 
  mutate(state_compet_score = sum_pop_weight /sum_pop,
         rgn_id = 12,
         rgn_name = "Northeast") %>% 
  select(year, state_compet_score, rgn_id, rgn_name)

#join region 12 scores to df
state_compet<- state_compet %>% 
  rbind(state_compet_rgn12)
```

Caculate and add Northeast/rgn 12 to LCV scores
```{r}
lcv_rgn12 <- lcv %>% 
  left_join(state_pop, by = c("rgn_name", "rgn_id")) %>% 
  select(year, score_lcv, state_name, pop_total) %>% 
  unique() %>%  #get rid of the two MA since they would be double counting in this case
  mutate(score_weighted = score_lcv*pop_total) %>% 
  group_by(year) %>% 
  dplyr:: summarize(sum_pop = sum(pop_total, na.rm =TRUE),
         sum_pop_weight = sum(score_weighted,na.rm =TRUE)) %>% 
  mutate(score_lcv = sum_pop_weight /sum_pop,
         rgn_id = 12,
         rgn_name = "Northeast") %>% 
  select(year, score_lcv, rgn_id, rgn_name)

#join region 12 scores to df
lcv<- lcv %>% 
  rbind(lcv_rgn12)
```


Add 2017 for NOAA social vulnerability index
```{r gap_noaa}
gapfill_noaa <- c("6", "9", "7", "8", "11", "5", "10", "12") %>% 
  as.data.frame() %>% 
  rename(rgn_id = ".") %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York",
    rgn_id == "12" ~ "Northeast")) %>% 
  mutate(year = 2017, 
         noaa_vul_score = NA)
```

Add to the noaa vulnerability index
```{r tidy_noaa}
noaa_vul <- noaa_vul_short %>% 
  rbind(gapfill_noaa) %>% 
  group_by(rgn_name) %>% 
  arrange(year) %>% 
  fill(noaa_vul_score, .direction = "down")  %>% 
  ungroup() %>% 
  mutate(rgn_id = as.integer(rgn_id))
```

Creating a df to add on to the social resilience scores to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  3, "Gulf of Maine",
  2, "Georges Bank",
  4, "Mid-Atlantic Bight",
  1, "Offshore") 

miss_rgns <- miss_rgns %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(score = NA) %>% 
  unique()
```

## Calculate and graph scores
Merge all four layers together
```{r combine_data}
res_social <- left_join(opportunity, state_compet, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(noaa_vul, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(lcv, by = c("year", "rgn_id", "rgn_name")) %>% 
  rowwise() %>% 
  mutate(score = mean(c(opportunity_score, state_compet_score, noaa_vul_score, score_lcv))) %>% 
  select(year, rgn_id, rgn_name, score) %>% 
  rbind(miss_rgns)
```

Save data layer
```{r save_data}
write.csv(res_social, file = "data/res_social.csv")
```


Visualize
```{r graph}
res_social_graph <- ggplot(res_social) +
  geom_line(aes(x=year, y= score, color = rgn_name)) +
  ggtitle("Social Resilience Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

res_social_graph

ggsave("figs/res_social_graph.jpg", width=7, height=5, dpi=300)
```


Write as a resilience layer to ne-scores
```{r write_to_scores}
write.csv(res_social, file.path(dir_calc, "layers/res_social.csv"))
```

Calculate inverse for pressure layer and save

```{r write_as_pressure}
prs_social <- read_csv(file.path(dir_calc, "layers/res_social.csv")) %>% 
  mutate(score= (1-score)) %>% 
  rename(pressure_score = score)
  
write.csv(prs_social, file.path(dir_calc, "layers/prs_social.csv"))
```












