---
title: "OHIEC: Resilience -Social"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---


# Summary

This script calculates social resilience values by OHINE region. 


# Data

**Reference**:


**Downloaded**: July, 2019

**Description**:  

**Time range**: BHI=2005-2017

**Format**:  BHI=CSV file

1) Beacon Hill Competitiveness Report

``` {r setup,  message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(dplyr)
library(tidyverse)


```


Beacon Hill Institute State Competitiveness Report

"Overall competitiveness scores are based on a simple average of the eight normalized sub-indices (mean=5, SD=1), which is then again normalized to have a mean of 5 and a standard deviation of 1. We then assigned a score to each state, from 0 to 1, where 1 is the maximum possible index score of 7.39 and 0 is the minimum of 2.88

```{r}

raw_bhi <- read_csv(file.path(dir_anx, "_raw_data/BHI/BHITable_long.csv"))

bhi_scores <- raw_bhi %>% 
  filter(state == "Maine" | state == "New Hampshire" | state == "Massachusetts" | state == "Rhode Island" | state == "Connecticut" | state == "New York") %>% 
  mutate(target = 10,
         score = index/target)

#write.csv(bhi_scores, "data/bhi_scores.csv")

bhi_graph <- ggplot(bhi_scores) +
  geom_line(aes(x=year, y= score, color = state)) +
  ggtitle("Rescaled scores State Competitiveness Report") +
  ylab("score") +
  xlab("Year") +
  theme_classic()

#ggsave("figs/bhi_graph.jpg", width=7, height=5, dpi=300)

```


