---
title: "OHIEC: Resilience -Social"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates social resilience values by OHINE region. 


**1. Beacon Hill Competitiveness Report**
**Downloaded**: July 2019

**Description**:  Report assesses each states competivieness by assessing "if it has in place the policies and conditions that ensure and sustain a high level of per capita income and its continued growth" http://beaconhill.org/economic-competitiveness/)

**Time range**: 2005-2017

**Format**: PDF
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R
dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

```

Load Libraries
``` {r}
library(csv)
library(dplyr)
library(tidyverse)
library(readxl)
library(gridExtra)
library(scales)
```

Read in Data
```{r}
raw_bhi <- read_csv(file.path(dir_anx, "_raw_data/BHI/BHITable_long.csv"))
```

Beacon Hill Institute State Competitiveness Report
```{r}
state_compet <- raw_bhi %>% 
  filter(state == "Maine" | state == "New Hampshire" | state == "Massachusetts" | state == "Rhode Island" | state == "Connecticut" | state == "New York") %>% 
  mutate(target = 10,
         score = index/target) %>% 
  mutate(rgn_id1 = case_when(
    state == "Maine" ~ "6", 
    state == "New Hampshire" ~ "9", 
    state == "Massachusetts" ~ "7",
    state == "Rhode Island" ~ "11",
    state == "Connecticut" ~ "5",
    state == "New York" ~ "10"),
    rgn_id2 = case_when(
    state == "Massachusetts" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  select(year, rgn_name, rgn_id, score) %>% 
  rename(state_compet_score = score)

write.csv(state_compet, "data/state_compet.csv")
```

Visualize
```{r}
state_compet_graph <- ggplot(state_compet) +
  geom_line(aes(x=year, y= state_compet_score, color = rgn_name)) +
  ggtitle("Scores State Competitiveness Report") +
  ylab("score") +
  xlab("Year") +
  theme_classic()

ggsave("figs/state_compet_graph.jpg", width=7, height=5, dpi=300)
```



