---
title: 'OHI-Northeast: Aquaculture - Disease Pressure Data Prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the aquaculture disease pressure scores by OHINE region. 

# Data

**Monterey Bay Aquarium Seafood Watch** 

**Downloaded**: August, 2019

**Format**: xlsx

# About

In this script we use the seafood watch data on the disease criteria per species and multiply by production.

Disease scores for each farmed species is criteria 7 in the Seafood Watch data. Each species get a score 0-10 with 10 being the least concern and 0 being high/critical.

## Setup

``` {r setup, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 10, fig.height = 8, fig.path = 'figs/', message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHI-NE specific version of common.R

library(tidyverse)
library(ggplot2)
library(readxl)
library(stringi)
```


```{r load_data}
sf_watch_raw <- read_xlsx(file.path(dir_anx, "_raw_data/SeafoodWatch/Seafood_Watch_USAquaculture82319.xlsx"))
production <- read_csv(file.path(dir_calc, "layers/mar_production.csv")) %>% 
  mutate(Species = tolower(Species)) %>% 
  select(year, rgn_id, Species, production)
```

## Methods

Cleaning up the seafood watch data so to be combined with production data
```{r tidy_sfw}
sf_watch <- sf_watch_raw %>% 
  select('FAO Common name', 'AqCriteria7', 'Report Title') %>% 
  rename(report_title ='Report Title',
         Species = 'FAO Common name') %>% 
  filter(report_title != "Atlantic salmon, Global RAS" & 
         report_title != "Global Recirculating Aquaculture Systems" &
         report_title != "Clams, Global",
         report_title != 'Rainbow trout, US OLD STANDARD') %>% # there are 3 different atlantic salmon scores, we are keeping the when Report Title' == Atlantic Salmon, Atlantic North America; we are also keeping "Clams, Global NEW STANDARD" and removing the old one; same with trout as with clams
  select(-report_title) %>% 
  mutate(Species = stri_replace_all_fixed(Species, "Hard clam (unspecified)", "Hard clam"),
         Species =  tolower(Species)) %>% 
  mutate(Species =  ifelse(Species == "manila clam", "soft shell clams", # all clams have the save score, but choosing to keep these se
                           ifelse(Species == "littleneck clam", "soft shell clams",
                              ifelse(Species == "razor clam", "soft shell clams",
                                ifelse(Species == "venus clam", "soft shell clams",
                                  ifelse(Species == "lyrate hard clam", "soft shell clams",
                                        Species)))))) %>% 
  mutate(Species = ifelse(str_detect(Species, "oyster"), "oysters", Species),
         Species = ifelse(str_detect(Species, "scallop"), "scallops", Species),
         Species = ifelse(str_detect(Species, "mussel"), "blue mussel", Species),
         Species = ifelse(str_detect(Species, "trout"), "trout", Species)) %>% #all oysters, scallops, and mussels have the same score; rainbow trout is farmed in the northeast, change the name for merging
  unique() 
```

## Calculate scores

Using Aquaculture subgoal method of calculating scores:

```{r calculate_score}
disease_prs <- left_join(production, sf_watch, by = ("Species")) %>% 
  mutate(sfw_score = AqCriteria7/10) %>% 
  rowwise() %>% 
  mutate(scored_production = production*sfw_score) %>% 
  ungroup() %>% 
  group_by(year, rgn_id) %>% 
  mutate(sum_scored_pro = sum(scored_production),
         sum_pro = sum(production),
         prs_score = 1 - (sum_scored_pro/sum_pro)) %>% 
  ungroup() %>% 
  mutate(rgn_id = as.factor(rgn_id),
         prs_score = ifelse(prs_score == "NaN", 0, prs_score)) %>% 
   filter(year >= 2005 & year <=2017) %>% 
  select(year, rgn_id, prs_score) %>% 
  unique()

write.csv(disease_prs, file = "data/disease_prs.csv")
```


Graph
```{r graph}
disease_prs_graph <-ggplot(disease_prs)+
  geom_line(aes(x=year, y = prs_score, color=rgn_id))+
  ggtitle("Aquaculture disease pressure score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 
disease_prs_graph

ggsave(file.path(dir_git, "prep/pressures/species/figs/disease_prs_graph.png"), width=7, height=5, dpi=300)
```

Clean up the data so it can be read into the toolbox
Creating a df to add on to the water pollution scores to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  3, "Gulf of Maine",
  2, "Georges Bank",
  4, "Mid-Atlantic Bight",
  1, "Offshore") 

miss_rgns <- miss_rgns %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(score = NA) %>% 
  unique()
```


```{r}
prs_aqua_disease <- disease_prs %>% 
    mutate(rgn_name =ifelse(rgn_id==5, "Connecticut",
                            ifelse(rgn_id==6, "Maine",
                                   ifelse(rgn_id==7, "Massachusetts-North",
                                          ifelse(rgn_id==8, "Massachusetts-South",
                                                 ifelse(rgn_id==9, "New Hampshire",
                                                        ifelse(rgn_id==10, "New York",
                                                               "Rhode Island"))))))) %>% 
  rename(score = prs_score) %>% 
  rbind(miss_rgns)
```


Save layer to scores
```{r}
write.csv(prs_aqua_disease, file.path(dir_calc, "layers/prs_aqua_disease.csv"))
```















