---
title: 'OHI-Northeast: Aquaculture - Disease Pressure Data Prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the aquaculture disease pressure scores by OHINE region. 

# Data

**Monterrey Bay Aquarium Seafood Watch** 

**Downloaded**: August, 2019

**Format**: xlsx

# About

In this script we use the seafood watch data for the disease criteria per species and scale by production.

Disease scores for each farmed species is criteria 7 in the Seafood Watch data. In the Seafood Watch data, each species get a score 0-10 with 10 being the least concern and 0 being high/critical.

## Setup

``` {r setup, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 10, fig.height = 8, fig.path = 'figs/', message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHI-NE specific version of common.R

library(tidyverse)
library(ggplot2)
library(readxl)
library(stringi)

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')
```

Here we load the raw Seafood watch data, and the cleaned/ gapfilled aquaculture production data created for the aquaculture goal.
```{r load_data}
sf_watch_raw <- read_xlsx(file.path(dir_anx, "_raw_data/SeafoodWatch/Seafood_Watch_USAquaculture82319.xlsx"))

production_raw <- read_csv("https://raw.githubusercontent.com/OHI-Northeast/ne-scores/master/region/layers/mar_production.csv")

production <- production_raw %>% 
  mutate(Species = tolower(Species)) %>% 
  select(year, rgn_id, Species, production)
```

## Methods

### Cleaning up the seafood watch data so to be combined with production data.
The data includes scores from previous methods of calculating the criteria, so we are removing those here. We are also fixing the names of the species to match up with the species in our production data. For example, with this current Seafood Watch data it is easy to do because all clams have the same sustainability score. If that changes in the future, differences in scores across clams, oyster, scallop, and mussel species need to be accounted for. For trout, the same score as salmon is assigned, because we know that they are farmed in the same way.
```{r tidy_sfw}
sf_watch <- sf_watch_raw %>% 
  select('FAO Common name', 'AqCriteria7', 'Report Title') %>% 
  rename(report_title ='Report Title',
         Species = 'FAO Common name') %>% 
  filter(report_title != "Atlantic salmon, Global RAS" & 
         report_title != "Global Recirculating Aquaculture Systems" &
         report_title != "Clams, Global",
         report_title != 'Rainbow trout, US OLD STANDARD') %>% # there are 3 different atlantic salmon scores, we are keeping the when Report Title' == Atlantic Salmon, Atlantic North America; we are also keeping "Clams, Global NEW STANDARD" and removing the old one; same with trout as with clams
  select(-report_title) %>% 
  mutate(Species =  tolower(Species),
         Species = ifelse(Species == "lyrate hard clam", "hard clam",
                          ifelse(Species == "hard clam (unspecified)", "hard clam",
                                        Species)),
         Species =  ifelse(Species == "manila clam", "soft shell clams", # all clams have the save score, but choosing to keep these se
                           ifelse(Species == "littleneck clam", "soft shell clams",
                              ifelse(Species == "razor clam", "soft shell clams",
                                ifelse(Species == "venus clam", "soft shell clams",
                                  ifelse(Species == "lyrate hard clam", "soft shell clams",
                                        Species)))))) %>% 
  mutate(Species = ifelse(str_detect(Species, "oyster"), "oysters", Species),
         Species = ifelse(str_detect(Species, "scallop"), "scallops", Species),
         Species = ifelse(str_detect(Species, "mussel"), "blue mussel", Species),
         Species = ifelse(str_detect(Species, "trout"), "trout", Species), #all oysters, scallops, and mussels have the same score; rainbow trout is farmed in the northeast, change the name for merging
         AqCriteria7 = ifelse(Species == "trout", 4, AqCriteria7)) %>% #this data set uses rainbow trout which is a freshwater trout. our trout is not freshwater, so we are using our knowlege of similarities to slamon aquaculture
  unique() 
```

## Calculate scores

To calculate the aquaculture escapes pressure on each region we will scale it by production and the area of the state waters. 
The first thing we need to do is read in the state waters shapefiles and grab the total area.
```{r state_waters_area}
area_state_waters <- read_sf(file.path(dir_git, "spatial/shapefiles/state_waters.shp")) %>% 
  as_data_frame() %>% 
  select(NAME10, Area_km) %>% 
  rename(state= NAME10) %>% 
  mutate(rgn_id1 = case_when(
    state == "Maine" ~ "6" , 
    state == "New Hampshire" ~ "9", 
    state == "Massachusetts" ~ "7",
    state == "Rhode Island" ~ "11",
    state == "Connecticut" ~ "5",
    state == "New York" ~ "10"),
    rgn_id2 = case_when(
    state == "Massachusetts" ~ "8"
    )) %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != is.na(rgn_id)) %>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"),
    rgn_id = as.numeric(rgn_id))
```

Now that we have the state waters areas, we can calculate the scores. The way we will do this is:
1. rescale the Seafood Watch scores from 0-1
2. Multiply the SFW score by the production to scale it by amount produced
3. Calculate the total scaled production in each region, each year
4. Scale the production by state water area
5. Calculate a reference point of two times the maximum pressure
6. Calculate the pressure using the reference point

```{r calculate_prs_score}
disease_prs <- left_join(production, sf_watch, by = ("Species")) %>% 
  left_join(area_state_waters, by = c("rgn_id"))  %>%
  mutate(production = ifelse(production == "NaN", NA , production)) %>% 
  rowwise() %>% 
  mutate(sfw_score = AqCriteria7/10,
         scaled_production = (production*sfw_score)) %>% 
  group_by(year, rgn_name, Area_km, rgn_id) %>% 
  summarise(sum_scale_prod = sum(scaled_production, na.rm = TRUE)) %>% 
  mutate(scaled_area_prod = sum_scale_prod/Area_km) %>% 
  ungroup() %>% 
  mutate(ref = 2*max(scaled_area_prod, na.rm = TRUE)) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(prs_score = scaled_area_prod/ref)
```

Graph
```{r graph}
disease_prs_graph <-ggplot(disease_prs)+
  geom_line(aes(x=year, y = prs_score, color=rgn_name))+
  ggtitle("Aquaculture escape pressure score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 
disease_prs_graph

ggsave(file.path(dir_git, "prep/pressures/species/figs/disease_prs_graph.png"), width=7, height=5, dpi=300)
```


Clean up the data so it can be read into the toolbox
Creating a df to add on to the water pollution scores to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  3, "Gulf of Maine",
  2, "Georges Bank",
  4, "Mid-Atlantic Bight",
  1, "Offshore") 

miss_rgns <- miss_rgns %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(score = NA) %>% 
  unique()
```

Create a layer for calculating scores
```{r create_layer}
prs_aqua_disease <- disease_prs %>% 
    mutate(rgn_name =ifelse(rgn_id==5, "Connecticut",
                            ifelse(rgn_id==6, "Maine",
                                   ifelse(rgn_id==7, "Massachusetts-North",
                                          ifelse(rgn_id==8, "Massachusetts-South",
                                                 ifelse(rgn_id==9, "New Hampshire",
                                                        ifelse(rgn_id==10, "New York",
                                                               "Rhode Island"))))))) %>% 
  rename(score = prs_score) %>% 
  select(year, rgn_name, rgn_id, score) %>% 
  filter(year >= 2005 & year<= 2017) %>% 
  rbind(miss_rgns)
```

Save layer to scores
```{r write_scores}
write.csv(prs_aqua_disease, file.path(dir_calc, "layers/prs_disease.csv"))
```





