---
title: "OHINE: Species Pollution Resilience - Nonindigenous Aquatic Species USGS Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---
# Summary

This script calculates the invasive species pressure by OHINE region.

# Data

**Nonindigenous Aquatic Species (NAS) USGS** 

**Downloaded**: September, 2019

**Description**: This dataset has occurrence and precise location data of new nonindigenous species https://nas.er.usgs.gov/

**Time range**: 1700-2019

**Format**: csv file


# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R
dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(tidyverse)
library(dplyr)
```


```{r load_data}
nas_raw <- read_csv(file.path(dir_anx, "_raw_data/NAS/NAS-Data-Download.csv")) 
```

## Explore data for pressures

Exploring total number of occurrences that have become established for the species pollution pressure layer
Now we want to see how many species are observed in a new area in each to quantify invasive species pressures
```{r occurances}
nas_est_subrgns <- nas_raw %>% 
  select('Scientific Name', 'Country', 'State', 'Year', 'Status') %>% 
  rename(sci_name = "Scientific Name") %>% 
  filter(Status %in% c('established', 'locally established', 'eradicated','extirpated'),
         Country %in% 'United States of America',
         State %in% c("ME", "MA", "NY", "RI", "CT", "NH"),
         Year >= 1970 & Year <= 2018) %>% 
  select(-Country) %>% 
  unique() %>% 
  mutate(counts = 1) %>% 
  group_by(Year, State) %>% 
  summarize(total_est = sum(counts)) %>% 
  ungroup() %>% 
  mutate(max_est = max(total_est),
         score = total_est/max_est) 
```

Now do the same thing but calculating for the region as a whole
```{r}
nas_est_rgn12 <- nas_raw %>% 
  select('Scientific Name', 'Country', 'State', 'Year', 'Status') %>% 
  rename(sci_name = "Scientific Name") %>% 
  filter(Status %in% c('established', 'locally established', 'eradicated','extirpated'),
         Country %in% 'United States of America',
         State %in% c("ME", "MA", "NY", "RI", "CT", "NH"),
         Year >= 1970 & Year <= 2018) %>% 
  select(-Country) %>% 
  unique() %>% 
  mutate(counts = 1) %>% 
  group_by(Year) %>% 
  summarize(total_est = sum(counts)) %>% 
  ungroup() %>% 
  mutate(max_est = max(total_est),
         score = total_est/max_est,
         State = "Northeast") 
```

combine the two
```{r}
nas_est <- nas_est_subrgns %>% 
  rbind(nas_est_rgn12)
```



Fish color palette for fun!
```{r fish_palette}
library(fishualize)
cols = fish(6, option = "Thalassoma_bifasciatum")
```


Look at occurrences
```{r occurances_viz}
est_gr <- ggplot(nas_est) +
  geom_line(aes(x= Year, y=total_est, color=State))+
  scale_color_fish_d(option = "Thalassoma_bifasciatum", direction = -1)+
  theme_classic() +
  labs(y = "Number Species", x = "State") +
  ggtitle("Invasive Species")

est_gr 

ggsave("figs/established_nas.jpg", width=7, height=5, dpi=300)
```

Look at scores
```{r scores_viz}
scores <- ggplot(nas_est) +
  geom_line(aes(x= Year, y=score, color=State))+
 # scale_color_fish_d(option = "Scarus_globiceps", direction = -1)+
  theme_classic() +
  labs(y = "Score", x = "State") +
  ggtitle("Invasive Species Scores")

scores 

ggsave("figs/scores.jpg", width=7, height=5, dpi=300)
```

Save a more simplified df for future score calculations
```{r}
invasive_prs_int <- nas_est %>% 
  select(State, Year, score) %>% 
  rename(year = Year) %>% 
  filter(year >= 2005 & year <= 2017)
```

## Cleaning up the layer for scores

4 entries missing in the nas data, which means that 4 times no species occurrences were logged:
RI 2008, 2011, 2012,2013. Even though Rhode Island has a score of 0 for every entry, this is because of very low invasive species occurrences  (1= lowest (and the mode) and 7=highest). So here I will give RI a score of 1 because the way that the data is represented this means that no new invasive species were introduced.

```{r ri_missing}
nas_ri_miss <- tribble(
  ~year, ~rgn_name, ~rgn_id, ~score,
  2008, "Rhode Island", 11, 0,  
  2012, "Rhode Island", 11, 0, 
  2013, "Rhode Island", 11, 0,
  ) %>% 
  as.data.frame() 
```

We also need to add on the offshore missing regions
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  3, "Gulf of Maine",
  2, "Georges Bank",
  4, "Mid-Atlantic Bight",
  1, "Offshore") 

miss_rgns <- miss_rgns %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(score = NA) %>% 
  unique()
```

```{r cleaning_up_layer}
sp_alien <- invasive_prs_int %>% 
  mutate(rgn_id1 = case_when(
    State == "ME" ~ "6", 
    State == "NH" ~ "9", 
    State == "MA" ~ "7",
    State == "RI" ~ "11",
    State == "CT" ~ "5",
    State == "NY" ~ "10",
    State == "Northeast" ~ "12"),
    rgn_id2 = case_when(
    State == "MA" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(-filler, - State) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York",
    rgn_id == "12" ~ "Northeast")) %>% 
  rbind(miss_rgns) %>% 
  rbind(nas_ri_miss)
```

```{r scores_graph}
scores <- ggplot(sp_alien) +
  geom_line(aes(x= year, y=score, color=rgn_name))+
 # scale_color_fish_d(option = "Scarus_globiceps", direction = -1)+
  theme_classic() +
  labs(y = "Score", x = "State") +
  ggtitle("Invasive Species Scores")

scores 

ggsave("figs/sp_alien_scores.jpg", width=7, height=5, dpi=300)
```

Write as a pressure layer to ne-scores
```{r write_to_scores}
write.csv(sp_alien, file.path(dir_calc, "layers/sp_alien.csv"))
```








