---
title: "Gather all species rasters"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Dataportal rasters

We need to get the cellIDs for each dataportal raster

```{r, eval = F}
cellid_rast <- raster("~/github/ne-prep/spatial/ocean_rasters/ne_cellids.tif")

#dataframe that links cellIDs with the OHI region they are in
rgn_cells <- read.csv("~/github/ne-prep/spatial/rgn_cellids.csv") %>%
  select(-X)

files <- list.files(file.path(dir_anx, "bio/portal_spp_rasters"), full.names = T)

## this function maps each species presence/absence raster and overlays the cellIDs to return a dataframe of cellIDs where the species is found, and a column with SID (species ID)
cellid_extract <- function(file){
  # update the progress bar (tick()) and print progress (print())
  pb$tick()$print()
  #get SID (species ID)
  sp <- substr(file, 58, nchar(file)-4)
  df <- data.frame(pres = 1, sp = sp)
  
  #read in raster
  r <- raster(file)
  
  #use the raster::zonal() function to get a dataframe that contains cellids and SID as two columns
  z <- zonal(r, cellid_rast, na.rm=T) %>%
    as.data.frame() %>%
    filter(!is.na(mean)) %>%
    rename(cellID = zone) %>%
    mutate(species = sp) %>%
    inner_join(rgn_cells) %>%
    select(-mean)
}

pb <- progress_estimated(length(files)) #progress bar

sid_cells <- purrr::map_df(files, cellid_extract)
```


We need to link these species up with the associated SID from IUCN. In the `ne_spp` data we list what species have NE portal maps. 

```{r}
#list of IUCN species with rangemaps that are found in the OHI Northeast region
ne_spp <- read_csv("data/1_iucn_spp_in_ne.csv") %>%
  filter(neportal == 1) 

a <- species %>%
  mutate(common = str_replace(full_name, "_", " "),
         sciname = taxize::comm2sci(common, db = "itis")) 

b <- ne_spp %>%
  left_join(a, by = "sciname") %>%
  left_join(a, by = c("common.x" = "common"))

#next step is to get specific SID for each of the Northeast species!
c <-b %>% 
  select(iucn_sid, sciname.x, common.y) %>% 
  distinct() %>%
  filter(!is.na(common.y)) #remove species that are only in IUCN maps

View(c)
```

There are some species with maps from the portal that are listed multiple times. For example we have a map from Duke that is for "Cusk" or "Spot" but these are associated with multiple Cusk and Spot species maps from IUCN.

```{r}
dups <- c$common.y[duplicated(c$common.y)]
unique(dups)

dup_df <- filter(c, common.y %in% dups) %>%
  rename(portal_map = common.y,
         IUCN_map = sciname.x)
View(dup_df)
```

Are there species with maps in the portal that are not in IUCN. And do those species have statuses? If so we need to add them in.
