---
title: "OHINE: Resilience - Discharge Facilities With Violations Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

Summary

This script calculates the percentage of discharge facilities with violations by OHINE region. 

**EPA ECHO database** 

**Downloaded**: July, 2019

**Description**:  
Data is located in the ICIS-NPDES National Data Set https://echo.epa.gov/tools/data-downloads 


**Time range**: 
  a. ICIS_FACILITIES.csv: no date range
  b. NPDES_QNCR_HISTORY.csv:1973-2019

**Format**:  CSV file

# About

Each region was scored by calculating the percentage of discharge facilities violations, taking the inverse, and then dividing by target (100% facilities with no violations) 

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(dplyr)
library(tidyverse)
```

*Note: not differentiating between types of violations (of which there are 4)*

Load data
```{r load_data}
raw_report_history <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/NPDES_QNCR_HISTORY.csv"))
raw_facilities <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/ICIS_FACILITIES.csv"))
```

## Tidy data
```{r viol_hist}
violations_hist <- raw_report_history %>% 
  mutate(state = substr(NPDES_ID, 0, 2),
         year = substr(YEARQTR, 0, 4))%>% 
  filter(year >=2005 & year <=2017) %>% 
  filter(state == "ME" | state == "NH" | state == "MA" | state =="RI" | state == "CT" | state == "NY") %>% 
  mutate(violation = 1) %>% #can assign a 1 to all columns becuase all rows included here mean there was a violation of some sort
  select(year, state, NPDES_ID, violation) %>% 
  distinct() %>%  # want to know if a violation that year, doesn't matter what quarter or if multiple violations in one year
  group_by(year, state) %>% 
  count() %>% 
  rename(fac_violations= n)
```


```{r fac_load}
fac <- raw_facilities %>% 
  select(NPDES_ID) %>% 
  mutate("2005" = "2005") %>% # creating a list of all the facilities for each year during our time range
  mutate("2006" = "2006") %>% 
  mutate("2007" = "2007") %>% 
  mutate("2008" = "2008") %>% 
  mutate("2009" = "2009") %>% 
  mutate("2010" = "2010") %>% 
  mutate("2011" = "2011") %>% 
  mutate("2012" = "2012") %>% 
  mutate("2013" = "2013") %>% 
  mutate("2014" = "2014") %>% 
  mutate("2015" = "2015") %>% 
  mutate("2016" = "2016") %>% 
  mutate("2017" = "2017") %>% 
  gather(year, year2, 2:14) %>% 
  select(NPDES_ID, year)%>% 
  mutate(state = substr(NPDES_ID, 0, 2)) 
```

```{r num_facilities_state}
#total number of facilities in each state: should be the same every year since used the meta data of all facilities. Could improve this by finding a list of facilities each year to account for some going out of business and new ones arriving

num_fac <- fac %>% 
  select(state, year, NPDES_ID) %>% 
  filter(state == "ME" | state == "NH" | state == "MA" | state =="RI" | state == "CT" | state == "NY") %>% 
  distinct() %>% 
  group_by(year, state) %>% 
  tally() %>% 
  ungroup() %>% 
  rename(fac_total = n) 
```

## Calculate scores, graph, save 
```{r percent_facilities_with_violations_by_state}
facilities_violations <- left_join(num_fac, violations_hist, by = c("year", "state"))%>% 
  mutate(percent_w_violations = fac_violations/fac_total,
         percent_no_violation = 1- percent_w_violations) %>% #to avoid dividing by a target of 0 violationfor later score calculations 
  mutate(year = as.numeric(year))

#save data
write.csv(facilities_violations, "data/facilities_violations.csv")
```

Visualization
```{r graph}
facilities_violations_graph <- ggplot(facilities_violations) +
  geom_line(aes(x= year, y = percent_w_violations, color= state)) +
  ggtitle("Percentage of Facilities with Violations in NE") +
  ylab("Percent") +
  xlab("Year") +
  theme_classic()

facilities_violations_graph
##save fig
ggsave("figs/facilities_violations_graph.jpg", width=7, height=5, dpi=300)
```


