---
title: "OHINE: Resilience - Discharge Facilities Inspected Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the percentage of discharge facilities that are inspected at least once a year by OHINE region. 

# Data

**EPA ECHO database** 

**Downloaded**: July, 2019

**Description**:  
Data is located in the ICIS-NPDES National Data Set https://echo.epa.gov/tools/data-downloads 

**Time range**: 
  a. ICIS_FACILITIES.csv: no date range
  b. NPDES_INSPECTIONS.csv: 1967-2019

**Format**:  CSV file

# About

Each region was scored by calculating the percentage of discharge facilities inspected and then dividing that by the 50% target. 
Target = EPA established ‘National Goals’: 50% for inspector coverage

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(dplyr)
library(tidyverse)
```

Load Data
```{r load_data}
raw_inspec <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/NPDES_INSPECTIONS.csv"))
raw_facilities <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/ICIS_FACILITIES.csv"))

#all the state annual DMRs
#2009
raw_maine_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/maine/ME_FY2009_NPDES_DMRS.csv"))
raw_connecticut_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/connecticut/CT_FY2009_NPDES_DMRS.csv"))
raw_massachusetts_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/massachusets/MA_FY2009_NPDES_DMRS.csv"))
raw_newhampshire_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/new_hampshire/NH_FY2009_NPDES_DMRS.csv"))
raw_newyork_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/new_york/NY_FY2009_NPDES_DMRS.csv"))
raw_rhodeisland_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/rhode_island/RI_FY2009_NPDES_DMRS.csv"))

#2010
raw_maine_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/maine/ME_FY2010_NPDES_DMRS.csv"))
raw_connecticut_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/connecticut/CT_FY2010_NPDES_DMRS.csv"))
raw_massachusetts_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/massachusetts/MA_FY2010_NPDES_DMRS.csv"))
raw_newhampshire_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/new_hampshire/NH_FY2010_NPDES_DMRS.csv"))
raw_newyork_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/new_york/NY_FY2010_NPDES_DMRS.csv"))
raw_rhodeisland_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/rhode_island/RI_FY2010_NPDES_DMRS.csv"))

#2011
raw_maine_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/maine/ME_FY2011_NPDES_DMRS.csv"))
raw_connecticut_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/connecticut/CT_FY2011_NPDES_DMRS.csv"))
raw_massachusetts_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/massachusetts/MA_FY2011_NPDES_DMRS.csv"))
raw_newhampshire_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/new_hampshire/NH_FY2011_NPDES_DMRS.csv"))
raw_newyork_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/new_york/NY_FY2011_NPDES_DMRS.csv"))
raw_rhodeisland_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/rhode_island/RI_FY2011_NPDES_DMRS.csv"))

#2012
raw_maine_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/maine/ME_FY2012_NPDES_DMRS.csv"))
raw_connecticut_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/connecticut/CT_FY2012_NPDES_DMRS.csv"))
raw_massachusetts_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/massachusetts/MA_FY2012_NPDES_DMRS.csv"))
raw_newhampshire_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/new_hampshire/NH_FY2012_NPDES_DMRS.csv"))
raw_newyork_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/new_york/NY_FY2012_NPDES_DMRS.csv"))
raw_rhodeisland_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/rhode_island/RI_FY2012_NPDES_DMRS.csv"))

#2013
raw_maine_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/maine/ME_FY2013_NPDES_DMRS.csv"))
raw_connecticut_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/connecticut/CT_FY2013_NPDES_DMRS.csv"))
raw_massachusetts_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/massachusetts/MA_FY2013_NPDES_DMRS.csv"))
raw_newhampshire_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/new_hampshire/NH_FY2013_NPDES_DMRS.csv"))
raw_newyork_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/new_york/NY_FY2013_NPDES_DMRS.csv"))
raw_rhodeisland_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/rhode_island/RI_FY2013_NPDES_DMRS.csv"))

#2014
raw_maine_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/maine/ME_FY2014_NPDES_DMRS.csv"))
raw_connecticut_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/connecticut/CT_FY2014_NPDES_DMRS.csv"))
raw_massachusetts_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/massachusetts/MA_FY2014_NPDES_DMRS.csv"))
raw_newhampshire_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/new_hampshire/NH_FY2014_NPDES_DMRS.csv"))
raw_newyork_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/new_york/NY_FY2014_NPDES_DMRS.csv"))
raw_rhodeisland_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/rhode_island/RI_FY2014_NPDES_DMRS.csv"))


#2015
raw_maine_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/maine/ME_FY2015_NPDES_DMRS.csv"))
raw_connecticut_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/connecticut/CT_FY2015_NPDES_DMRS.csv"))
raw_massachusetts_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/massachusetts/MA_FY2015_NPDES_DMRS.csv"))
raw_newhampshire_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/new_hampshire/NH_FY2015_NPDES_DMRS.csv"))
raw_newyork_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/new_york/NY_FY2015_NPDES_DMRS.csv"))
raw_rhodeisland_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/rhode_island/RI_FY2015_NPDES_DMRS.csv"))

#2016
raw_maine_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/maine/ME_FY2016_NPDES_DMRS.csv"))
raw_connecticut_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/connecticut/CT_FY2016_NPDES_DMRS.csv"))
raw_massachusetts_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/massachusetts/MA_FY2016_NPDES_DMRS.csv"))
raw_newhampshire_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/new_hampshire/NH_FY2016_NPDES_DMRS.csv"))
raw_newyork_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/new_york/NY_FY2016_NPDES_DMRS.csv"))
raw_rhodeisland_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/rhode_island/RI_FY2016_NPDES_DMRS.csv"))


#2017
raw_maine_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/maine/ME_FY2017_NPDES_DMRS.csv"))
raw_connecticut_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/connecticut/CT_FY2017_NPDES_DMRS.csv"))
raw_massachusetts_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/massachusetts/MA_FY2017_NPDES_DMRS.csv"))
raw_newhampshire_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/new_hampshire/NH_FY2017_NPDES_DMRS.csv"))
raw_newyork_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/new_york/NY_FY2017_NPDES_DMRS.csv"))
raw_rhodeisland_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/rhode_island/RI_FY2017_NPDES_DMRS.csv"))
```

#Methods

```{r inspec_tidy}
inspec <- raw_inspec %>% 
  select(NPDES_ID, ACTUAL_END_DATE) %>% #choose end instead of start because many NAs in "ACTUAL_START_DATE" and not in end
  separate(ACTUAL_END_DATE, c("month", "day", "year"), sep = "/") %>% 
  select(NPDES_ID, year)%>%
  filter(year >=2005 & year <=2017) %>% 
  distinct() %>% # just want to know if the facility had been inspected that year, doesn't matter if it was more than once
  mutate(inspec = "1") %>%  # yes(1) or no(0) there was an inspection
  select(NPDES_ID, year, inspec) 

```

```{r fac_tidy}
fac <- raw_facilities %>% 
  select(NPDES_ID) %>% 
  mutate("2005" = "2005") %>% # creating a list of all the facilities for each year during our time range
  mutate("2006" = "2006") %>% 
  mutate("2007" = "2007") %>% 
  mutate("2008" = "2008") %>% 
  mutate("2009" = "2009") %>% 
  mutate("2010" = "2010") %>% 
  mutate("2011" = "2011") %>% 
  mutate("2012" = "2012") %>% 
  mutate("2013" = "2013") %>% 
  mutate("2014" = "2014") %>% 
  mutate("2015" = "2015") %>% 
  mutate("2016" = "2016") %>% 
  mutate("2017" = "2017") %>% 
  gather(year, year2, 2:14) %>% 
  select(NPDES_ID, year)%>% 
  mutate(state = substr(NPDES_ID, 0, 2)) 
```

```{r inspec_clean}
inspections_clean <- left_join(fac, inspec, by = c("NPDES_ID", "year")) %>% 
  replace(., is.na(.), "0") %>% 
  mutate(state = substr(NPDES_ID, 0, 2)) %>% 
  filter(state == "ME" | state == "NH" | state == "MA" | state =="RI" | state == "CT" | state == "NY")
```

Inspections stats
The total number of facilities in each state should be the same every year since used the meta data of all facilities. Could improve this by finding a list of facilities each year to account for some going out of business and new ones arriving
```{r fac_clean}
num_fac <- fac %>% 
  select(state, year, NPDES_ID) %>% 
  filter(state == "ME" | state == "NH" | state == "MA" | state =="RI" | state == "CT" | state == "NY") %>% 
  distinct() %>% 
  group_by(year, state) %>% 
  tally() %>% 
  ungroup() %>% 
  rename(fac_total = n) 
```

Calculate the total number of facilities inspected at least once that year by state
```{r calc_inspec}
num_fac_inspec <- inspections_clean %>% 
  filter(inspec == 1) %>% 
  select(state, year, NPDES_ID) %>% 
  group_by(year, state) %>% 
  count() %>% 
  ungroup() %>% 
  rename(fac_inspection = n)
```

Merging and writing tidy csv
```{r merge_tidy}
facilities_inspected <- left_join(num_fac, num_fac_inspec, by = c("year", "state")) %>% 
  mutate(percent_w_inspections = fac_inspection/fac_total,
         year = as.numeric(year))

##save file
write.csv(facilities_inspected, "data/facilities_inspected.csv")
```

Visualization
```{r graph_inspections}
inspections_graph <- ggplot(facilities_inspected) +
  geom_line(aes(x= year, y = percent_w_inspections, color= state))+
 ggtitle("Percentage of Facilities Inspected in NE") +
  ylab("Percent") +
  xlab("Year") +
  theme_classic()+
  geom_hline(yintercept=0.5)
inspections_graph

##save fig
ggsave("figs/facilities_inspected_graph.jpg", width=7, height=5, dpi=300)
```