---
title: "OHINE: Water Pollution Resilience - Reports Submitted by Discharge Facilities Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the percentage of expected discharge monitoring reports received by OHINE region. This data is used to measure water pollution resilience implementation & enforcement.

# Data
**EPA ECHO database**
Permit Limit and DMR Data by Jurisdiction

**Downloaded**: July, 2019

**Description**:  
Data is located in the ICIS-NPDES National Data Set https://echo.epa.gov/tools/data-downloads 

**Time range**: 
Permit Limit and DMR Data by Jurisdiction: 2009-2019

**Format**:  CSV file

# About
Each region was scored by calculating the percentage of received discharge monitoring reports that was expected by the target of 95%.
Target = EPA established ‘National Goals’: 95% for DMR submissions

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(readr)
library(dplyr)
library(tidyverse)
```

Load Data
```{r load_data}
raw_facilities <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/ICIS_FACILITIES.csv"))

#all the state annual DMRs
#2009
raw_maine_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/maine/ME_FY2009_NPDES_DMRS.csv"))
raw_connecticut_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/connecticut/CT_FY2009_NPDES_DMRS.csv"))
raw_massachusetts_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/massachusets/MA_FY2009_NPDES_DMRS.csv"))
raw_newhampshire_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/new_hampshire/NH_FY2009_NPDES_DMRS.csv"))
raw_newyork_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/new_york/NY_FY2009_NPDES_DMRS.csv"))
raw_rhodeisland_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/rhode_island/RI_FY2009_NPDES_DMRS.csv"))

#2010
raw_maine_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/maine/ME_FY2010_NPDES_DMRS.csv"))
raw_connecticut_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/connecticut/CT_FY2010_NPDES_DMRS.csv"))
raw_massachusetts_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/massachusetts/MA_FY2010_NPDES_DMRS.csv"))
raw_newhampshire_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/new_hampshire/NH_FY2010_NPDES_DMRS.csv"))
raw_newyork_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/new_york/NY_FY2010_NPDES_DMRS.csv"))
raw_rhodeisland_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/rhode_island/RI_FY2010_NPDES_DMRS.csv"))

#2011
raw_maine_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/maine/ME_FY2011_NPDES_DMRS.csv"))
raw_connecticut_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/connecticut/CT_FY2011_NPDES_DMRS.csv"))
raw_massachusetts_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/massachusetts/MA_FY2011_NPDES_DMRS.csv"))
raw_newhampshire_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/new_hampshire/NH_FY2011_NPDES_DMRS.csv"))
raw_newyork_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/new_york/NY_FY2011_NPDES_DMRS.csv"))
raw_rhodeisland_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/rhode_island/RI_FY2011_NPDES_DMRS.csv"))

#2012
raw_maine_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/maine/ME_FY2012_NPDES_DMRS.csv"))
raw_connecticut_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/connecticut/CT_FY2012_NPDES_DMRS.csv"))
raw_massachusetts_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/massachusetts/MA_FY2012_NPDES_DMRS.csv"))
raw_newhampshire_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/new_hampshire/NH_FY2012_NPDES_DMRS.csv"))
raw_newyork_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/new_york/NY_FY2012_NPDES_DMRS.csv"))
raw_rhodeisland_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/rhode_island/RI_FY2012_NPDES_DMRS.csv"))

#2013
raw_maine_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/maine/ME_FY2013_NPDES_DMRS.csv"))
raw_connecticut_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/connecticut/CT_FY2013_NPDES_DMRS.csv"))
raw_massachusetts_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/massachusetts/MA_FY2013_NPDES_DMRS.csv"))
raw_newhampshire_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/new_hampshire/NH_FY2013_NPDES_DMRS.csv"))
raw_newyork_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/new_york/NY_FY2013_NPDES_DMRS.csv"))
raw_rhodeisland_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/rhode_island/RI_FY2013_NPDES_DMRS.csv"))

#2014
raw_maine_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/maine/ME_FY2014_NPDES_DMRS.csv"))
raw_connecticut_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/connecticut/CT_FY2014_NPDES_DMRS.csv"))
raw_massachusetts_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/massachusetts/MA_FY2014_NPDES_DMRS.csv"))
raw_newhampshire_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/new_hampshire/NH_FY2014_NPDES_DMRS.csv"))
raw_newyork_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/new_york/NY_FY2014_NPDES_DMRS.csv"))
raw_rhodeisland_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/rhode_island/RI_FY2014_NPDES_DMRS.csv"))


#2015
raw_maine_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/maine/ME_FY2015_NPDES_DMRS.csv"))
raw_connecticut_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/connecticut/CT_FY2015_NPDES_DMRS.csv"))
raw_massachusetts_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/massachusetts/MA_FY2015_NPDES_DMRS.csv"))
raw_newhampshire_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/new_hampshire/NH_FY2015_NPDES_DMRS.csv"))
raw_newyork_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/new_york/NY_FY2015_NPDES_DMRS.csv"))
raw_rhodeisland_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/rhode_island/RI_FY2015_NPDES_DMRS.csv"))

#2016
raw_maine_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/maine/ME_FY2016_NPDES_DMRS.csv"))
raw_connecticut_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/connecticut/CT_FY2016_NPDES_DMRS.csv"))
raw_massachusetts_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/massachusetts/MA_FY2016_NPDES_DMRS.csv"))
raw_newhampshire_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/new_hampshire/NH_FY2016_NPDES_DMRS.csv"))
raw_newyork_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/new_york/NY_FY2016_NPDES_DMRS.csv"))
raw_rhodeisland_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/rhode_island/RI_FY2016_NPDES_DMRS.csv"))


#2017
raw_maine_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/maine/ME_FY2017_NPDES_DMRS.csv"))
raw_connecticut_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/connecticut/CT_FY2017_NPDES_DMRS.csv"))
raw_massachusetts_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/massachusetts/MA_FY2017_NPDES_DMRS.csv"))
raw_newhampshire_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/new_hampshire/NH_FY2017_NPDES_DMRS.csv"))
raw_newyork_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/new_york/NY_FY2017_NPDES_DMRS.csv"))
raw_rhodeisland_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/rhode_island/RI_FY2017_NPDES_DMRS.csv"))
```

## Report Submissions
Here we need to figure out how many reports were submitted and how many were expected but not recieved. We can do this by using the RNC_DETECTION_CODE column. Here N = Non-Receipt of DMR/Schedule Report. A lack of code with the submitted report means that the report was submitted with no issues.

All regions 2009
```{r}
all_rgns_09 <- rbind(raw_maine_2009, raw_connecticut_2009, raw_massachusetts_2009, raw_newhampshire_2009, raw_newyork_2009,raw_rhodeisland_2009 ) %>% 
  mutate(state = substr(EXTERNAL_PERMIT_NMBR, start = 1, stop =2)) %>% 
  mutate(year = 2009,
         submitted = case_when(RNC_DETECTION_CODE == "N" ~ 0,
                               RNC_DETECTION_CODE != "N" ~ 1,
                               is.na(RNC_DETECTION_CODE) ~ 1),
         expected = 1) %>% 
  select(state, year, submitted, expected) 
```

All regions 2010
```{r}
all_rgns_10 <- rbind(raw_maine_2010, raw_connecticut_2010, raw_massachusetts_2010, raw_newhampshire_2010, raw_newyork_2010,raw_rhodeisland_2010) %>% 
  mutate(state = substr(EXTERNAL_PERMIT_NMBR, start = 1, stop =2)) %>% 
  mutate(year = 2010,
         submitted = case_when(RNC_DETECTION_CODE == "N" ~ 0,
                               RNC_DETECTION_CODE != "N" ~ 1,
                               is.na(RNC_DETECTION_CODE) ~ 1),
         expected = 1) %>% 
  select(state, year, submitted, expected) 
```

All regions 2011
```{r}
all_rgns_11 <- rbind(raw_maine_2011, raw_connecticut_2011, raw_massachusetts_2011, raw_newhampshire_2011, raw_newyork_2011,raw_rhodeisland_2011) %>% 
  mutate(state = substr(EXTERNAL_PERMIT_NMBR, start = 1, stop =2)) %>% 
  mutate(year = 2011,
         submitted = case_when(RNC_DETECTION_CODE == "N" ~ 0,
                               RNC_DETECTION_CODE != "N" ~ 1,
                               is.na(RNC_DETECTION_CODE) ~ 1),
         expected = 1) %>% 
  select(state, year, submitted, expected) 
```

All regions 2012
```{r}
all_rgns_12 <- rbind(raw_maine_2012, raw_connecticut_2012, raw_massachusetts_2012, raw_newhampshire_2012, raw_newyork_2012,raw_rhodeisland_2012) %>% 
  mutate(state = substr(EXTERNAL_PERMIT_NMBR, start = 1, stop =2)) %>% 
  mutate(year = 2012,
         submitted = case_when(RNC_DETECTION_CODE == "N" ~ 0,
                               RNC_DETECTION_CODE != "N" ~ 1,
                               is.na(RNC_DETECTION_CODE) ~ 1),
         expected = 1) %>% 
  select(state, year, submitted, expected) 
```

All regions 2013
```{r}
all_rgns_13 <- rbind(raw_maine_2013, raw_connecticut_2013, raw_massachusetts_2013, raw_newhampshire_2013, raw_newyork_2013,raw_rhodeisland_2013) %>% 
  mutate(state = substr(EXTERNAL_PERMIT_NMBR, start = 1, stop =2)) %>% 
  mutate(year = 2013,
         submitted = case_when(RNC_DETECTION_CODE == "N" ~ 0,
                               RNC_DETECTION_CODE != "N" ~ 1,
                               is.na(RNC_DETECTION_CODE) ~ 1),
         expected = 1) %>% 
  select(state, year, submitted, expected) 
```

All regions 2014
```{r}
all_rgns_14 <- rbind(raw_maine_2014, raw_connecticut_2014, raw_massachusetts_2014, raw_newhampshire_2014, raw_newyork_2014,raw_rhodeisland_2014) %>% 
  mutate(state = substr(EXTERNAL_PERMIT_NMBR, start = 1, stop =2)) %>% 
  mutate(year = 2014,
         submitted = case_when(RNC_DETECTION_CODE == "N" ~ 0,
                               RNC_DETECTION_CODE != "N" ~ 1,
                               is.na(RNC_DETECTION_CODE) ~ 1),
         expected = 1) %>% 
  select(state, year, submitted, expected) 
```

All regions 2015
```{r}
all_rgns_15 <- rbind(raw_maine_2015, raw_connecticut_2015, raw_massachusetts_2015, raw_newhampshire_2015, raw_newyork_2015,raw_rhodeisland_2015) %>% 
  mutate(state = substr(EXTERNAL_PERMIT_NMBR, start = 1, stop =2)) %>% 
  mutate(year = 2015,
         submitted = case_when(RNC_DETECTION_CODE == "N" ~ 0,
                               RNC_DETECTION_CODE != "N" ~ 1,
                               is.na(RNC_DETECTION_CODE) ~ 1),
         expected = 1) %>% 
  select(state, year, submitted, expected) 
```

All regions 2016
```{r}
all_rgns_16 <- rbind(raw_maine_2016, raw_connecticut_2016, raw_massachusetts_2016, raw_newhampshire_2016, raw_newyork_2016,raw_rhodeisland_2016) %>% 
  mutate(state = substr(EXTERNAL_PERMIT_NMBR, start = 1, stop =2)) %>% 
  mutate(year = 2016,
         submitted = case_when(RNC_DETECTION_CODE == "N" ~ 0,
                               RNC_DETECTION_CODE != "N" ~ 1,
                               is.na(RNC_DETECTION_CODE) ~ 1),
         expected = 1) %>% 
  select(state, year, submitted, expected) 
```

All regions 2017
```{r}
all_rgns_17 <- rbind(raw_maine_2017, raw_connecticut_2017, raw_massachusetts_2017, raw_newhampshire_2017, raw_newyork_2017,raw_rhodeisland_2017) %>% 
  mutate(state = substr(EXTERNAL_PERMIT_NMBR, start = 1, stop =2)) %>% 
  mutate(year = 2017,
         submitted = case_when(RNC_DETECTION_CODE == "N" ~ 0,
                               RNC_DETECTION_CODE != "N" ~ 1,
                               is.na(RNC_DETECTION_CODE) ~ 1),
         expected = 1) %>% 
  select(state, year, submitted, expected) 
```

all regions
```{r}
all_rgns <- rbind(all_rgns_17, all_rgns_16, all_rgns_15, all_rgns_14, all_rgns_13, all_rgns_12, all_rgns_11, all_rgns_10, all_rgns_09) %>% 
  group_by(year, state) %>% 
  summarize(recieved = sum(submitted),
            expect = sum(expected)) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(percent_recieved =  recieved/expect)
```

Do the same calculations for the entire Northeast region as a whole
```{r}
rgn12 <- rbind(all_rgns_17, all_rgns_16, all_rgns_15, all_rgns_14, all_rgns_13, all_rgns_12, all_rgns_11, all_rgns_10, all_rgns_09) %>% 
  group_by(year) %>% 
  summarize(recieved = sum(submitted),
            expect = sum(expected)) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(percent_recieved =  recieved/expect,
         state = "Northeast")
```

Combine the two data frames togther
```{r}
reports_submitted <-rbind(all_rgns, rgn12)
  
write.csv(reports_submitted,"data/reports_submitted.csv")  
```

Visualization
```{r graph}
reports_submitted_graph <- ggplot(reports_submitted) +
  geom_line(aes(x= year, y=percent_recieved , color=state))+
  ggtitle("Percentage of Reports Submitted in NE") +
  ylab("Percent") +
  xlab("Year") +
  theme_classic() +
  geom_hline(yintercept=0.95)

reports_submitted_graph

##save fig
ggsave("figs/reports_submitted_graph.jpg", width=7, height=5, dpi=300)
```

