---
title: "OHINE: Resilience - Reports Submitted by Discharge Facilities Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the percentage of expected discharge monitoring reports received by OHINE region. 

# Data
**EPA ECHO database**
Permit Limit and DMR Data by Jurisdiction

**Downloaded**: July, 2019

**Description**:  
Data is located in the ICIS-NPDES National Data Set https://echo.epa.gov/tools/data-downloads 

**Time range**: 
Permit Limit and DMR Data by Jurisdiction: 2009-2019

**Format**:  CSV file

# About
Each region was scored by calculating the percentage of received discharge monitoring reports that was expected by the target of 95%.
Target = EPA established ‘National Goals’: 95% for DMR submissions

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(dplyr)
library(tidyverse)
```

Load Data
```{r load_data}
raw_facilities <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/ICIS_FACILITIES.csv"))

#all the state annual DMRs
#2009
raw_maine_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/maine/ME_FY2009_NPDES_DMRS.csv"))
raw_connecticut_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/connecticut/CT_FY2009_NPDES_DMRS.csv"))
raw_massachusetts_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/massachusets/MA_FY2009_NPDES_DMRS.csv"))
raw_newhampshire_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/new_hampshire/NH_FY2009_NPDES_DMRS.csv"))
raw_newyork_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/new_york/NY_FY2009_NPDES_DMRS.csv"))
raw_rhodeisland_2009 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2009/rhode_island/RI_FY2009_NPDES_DMRS.csv"))

#2010
raw_maine_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/maine/ME_FY2010_NPDES_DMRS.csv"))
raw_connecticut_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/connecticut/CT_FY2010_NPDES_DMRS.csv"))
raw_massachusetts_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/massachusetts/MA_FY2010_NPDES_DMRS.csv"))
raw_newhampshire_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/new_hampshire/NH_FY2010_NPDES_DMRS.csv"))
raw_newyork_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/new_york/NY_FY2010_NPDES_DMRS.csv"))
raw_rhodeisland_2010 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2010/rhode_island/RI_FY2010_NPDES_DMRS.csv"))

#2011
raw_maine_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/maine/ME_FY2011_NPDES_DMRS.csv"))
raw_connecticut_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/connecticut/CT_FY2011_NPDES_DMRS.csv"))
raw_massachusetts_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/massachusetts/MA_FY2011_NPDES_DMRS.csv"))
raw_newhampshire_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/new_hampshire/NH_FY2011_NPDES_DMRS.csv"))
raw_newyork_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/new_york/NY_FY2011_NPDES_DMRS.csv"))
raw_rhodeisland_2011 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2011/rhode_island/RI_FY2011_NPDES_DMRS.csv"))

#2012
raw_maine_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/maine/ME_FY2012_NPDES_DMRS.csv"))
raw_connecticut_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/connecticut/CT_FY2012_NPDES_DMRS.csv"))
raw_massachusetts_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/massachusetts/MA_FY2012_NPDES_DMRS.csv"))
raw_newhampshire_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/new_hampshire/NH_FY2012_NPDES_DMRS.csv"))
raw_newyork_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/new_york/NY_FY2012_NPDES_DMRS.csv"))
raw_rhodeisland_2012 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2012/rhode_island/RI_FY2012_NPDES_DMRS.csv"))

#2013
raw_maine_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/maine/ME_FY2013_NPDES_DMRS.csv"))
raw_connecticut_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/connecticut/CT_FY2013_NPDES_DMRS.csv"))
raw_massachusetts_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/massachusetts/MA_FY2013_NPDES_DMRS.csv"))
raw_newhampshire_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/new_hampshire/NH_FY2013_NPDES_DMRS.csv"))
raw_newyork_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/new_york/NY_FY2013_NPDES_DMRS.csv"))
raw_rhodeisland_2013 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2013/rhode_island/RI_FY2013_NPDES_DMRS.csv"))

#2014
raw_maine_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/maine/ME_FY2014_NPDES_DMRS.csv"))
raw_connecticut_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/connecticut/CT_FY2014_NPDES_DMRS.csv"))
raw_massachusetts_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/massachusetts/MA_FY2014_NPDES_DMRS.csv"))
raw_newhampshire_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/new_hampshire/NH_FY2014_NPDES_DMRS.csv"))
raw_newyork_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/new_york/NY_FY2014_NPDES_DMRS.csv"))
raw_rhodeisland_2014 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2014/rhode_island/RI_FY2014_NPDES_DMRS.csv"))


#2015
raw_maine_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/maine/ME_FY2015_NPDES_DMRS.csv"))
raw_connecticut_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/connecticut/CT_FY2015_NPDES_DMRS.csv"))
raw_massachusetts_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/massachusetts/MA_FY2015_NPDES_DMRS.csv"))
raw_newhampshire_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/new_hampshire/NH_FY2015_NPDES_DMRS.csv"))
raw_newyork_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/new_york/NY_FY2015_NPDES_DMRS.csv"))
raw_rhodeisland_2015 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2015/rhode_island/RI_FY2015_NPDES_DMRS.csv"))

#2016
raw_maine_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/maine/ME_FY2016_NPDES_DMRS.csv"))
raw_connecticut_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/connecticut/CT_FY2016_NPDES_DMRS.csv"))
raw_massachusetts_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/massachusetts/MA_FY2016_NPDES_DMRS.csv"))
raw_newhampshire_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/new_hampshire/NH_FY2016_NPDES_DMRS.csv"))
raw_newyork_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/new_york/NY_FY2016_NPDES_DMRS.csv"))
raw_rhodeisland_2016 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2016/rhode_island/RI_FY2016_NPDES_DMRS.csv"))


#2017
raw_maine_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/maine/ME_FY2017_NPDES_DMRS.csv"))
raw_connecticut_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/connecticut/CT_FY2017_NPDES_DMRS.csv"))
raw_massachusetts_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/massachusetts/MA_FY2017_NPDES_DMRS.csv"))
raw_newhampshire_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/new_hampshire/NH_FY2017_NPDES_DMRS.csv"))
raw_newyork_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/new_york/NY_FY2017_NPDES_DMRS.csv"))
raw_rhodeisland_2017 <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/DMR_Annual/2017/rhode_island/RI_FY2017_NPDES_DMRS.csv"))
```

## Report Submissions
Here we need to figure out how many reports were submitted and how many were expected but not recieved. We can do this by using the RNC_DETECTION_CODE column. Here N = Non-Receipt of DMR/Schedule Report. Do this for every year for each state

2009
```{r dmr_2009}
## Maine
maine_2009_sub <- raw_maine_2009 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

ME_2009 <- 1-(count(maine_2009_sub)/count(raw_maine_2009))
#expected 85510 reports
#did not receive 2832
#0.9668811 received

## New Hampshire
newhampshire_2009_sub <- raw_newhampshire_2009 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

NH_2009 <- 1-(count(newhampshire_2009_sub)/count(raw_newhampshire_2009))
    #expected 40841 reports
    #did not receive 115
    #0.9971842 received

## Massachusetts
massachusetts_2009_sub <- raw_massachusetts_2009 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

MA_2009 <- 1-(count(massachusetts_2009_sub)/count(raw_massachusetts_2009))
#expected 102756 reports
# did not receive 1626
#0.9841761 received


## Rhode Island
rhodeisland_2009_sub <- raw_rhodeisland_2009 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

RI_2009 <- 1-(count(rhodeisland_2009_sub)/count(raw_rhodeisland_2009))
#expected 45477 reports
# did not receive 0
#1 received

## Connecticut
connecticut_2009_sub <- raw_connecticut_2009 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

CT_2009 <- 1-(count(connecticut_2009_sub)/count(raw_connecticut_2009))
#expected 130445 reports
# did not receive 0
#1 received


## New York
newyork_2009_sub <- raw_newyork_2009 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

NY_2009 <- 1-(count(newyork_2009_sub)/count(raw_newyork_2009))
#expected 517016 reports
# did not receive 1380
#0.9973308 received
```

2010
```{r dmr_2010}
## Maine
maine_2010_sub <- raw_maine_2010 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

ME_2010 <- 1-(count(maine_2010_sub)/count(raw_maine_2010))
#expected 82388 reports
# did not receive 2311
#0.9719498 received


## New Hampshire
newhampshire_2010_sub <- raw_newhampshire_2010 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

NH_2010 <- 1-(count(newhampshire_2010_sub)/count(raw_newhampshire_2010))
#expected 40820 reports
# did not receive 99
#0.9975747 received

## Massachusetts
massachusetts_2010_sub <- raw_massachusetts_2010 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

MA_2010 <- 1-(count(massachusetts_2010_sub)/count(raw_massachusetts_2010))
#expected 103549 reports
# did not receive 0
#1 received


## Rhode Island
rhodeisland_2010_sub <- raw_rhodeisland_2010 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

RI_2010 <- 1-(count(rhodeisland_2010_sub)/count(raw_rhodeisland_2010))
#expected 45642 reports
# did not receive 0
#1 received

## Connecticut
connecticut_2010_sub <- raw_connecticut_2010 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

CT_2010 <- 1-(count(connecticut_2010_sub)/count(raw_connecticut_2010))
#expected 128342 reports
# did not receive 0
#1 received


## New York
newyork_2010_sub <- raw_newyork_2010 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

NY_2010 <- 1-(count(newyork_2010_sub)/count(raw_newyork_2010))
#expected 523307 reports
# did not receive 1610
#0.9969234 received
```

2011
```{r dmr_2011}
## Maine
maine_2011_sub <- raw_maine_2011 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

ME_2011 <- 1-(count(maine_2011_sub)/count(raw_maine_2011))
#expected 82716 reports
# did not receive 1367
#0.9834736 received

  
## New Hampshire
newhampshire_2011_sub <- raw_newhampshire_2011 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

NH_2011 <- 1-(count(newhampshire_2011_sub)/count(raw_newhampshire_2011))
#expected 40210 reports
# did not receive 0
#0.9975747 received

## Massachusetts
massachusetts_2011_sub <- raw_massachusetts_2011 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

MA_2011 <- 1-(count(massachusetts_2011_sub)/count(raw_massachusetts_2011))
#expected 103150 reports
# did not receive 0
#1 received


## Rhode Island
rhodeisland_2011_sub <- raw_rhodeisland_2011 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

RI_2011 <- 1-(count(rhodeisland_2011_sub)/count(raw_rhodeisland_2011))
#expected 46621 reports
# did not receive 0
#1 received

## Connecticut
connecticut_2011_sub <- raw_connecticut_2011 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

CT_2011 <- 1-(count(connecticut_2011_sub)/count(raw_connecticut_2011))
#expected 128656 reports
# did not receive 0
#1 received


## New York
newyork_2011_sub <- raw_newyork_2011 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

NY_2011 <- 1-(count(newyork_2011_sub)/count(raw_newyork_2011))
#expected 534640 reports
# did not receive 1492
#0.9972093 received
```

2012
```{r dmr_2012}
## Maine
maine_2012_sub <- raw_maine_2012 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 80818 reports
# did not receive 695
ME_2012 <- 1-(count(maine_2012_sub)/count(raw_maine_2012))
#0.9914004 received

  
## New Hampshire
newhampshire_2012_sub <- raw_newhampshire_2012 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

NH_2012 <- 1-(count(newhampshire_2012_sub)/count(raw_newhampshire_2012))
#expected 42033 reports
# did not receive 0
#0.9975747 received

## Massachusetts
massachusetts_2012_sub <- raw_massachusetts_2012 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

MA_2012 <- 1-(count(massachusetts_2012_sub)/count(raw_massachusetts_2012))
#expected 103575 reports
# did not receive 1477
#0.9857398 received


## Rhode Island
rhodeisland_2012_sub <- raw_rhodeisland_2012 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

RI_2012 <- 1-(count(rhodeisland_2012_sub)/count(raw_rhodeisland_2012))
#expected 42751 reports
# did not receive 0
#1 received

## Connecticut
connecticut_2012_sub <- raw_connecticut_2012 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

CT_2012 <- 1-(count(connecticut_2012_sub)/count(raw_connecticut_2012))
#expected 127430 reports
# did not receive 0
#1 received


## New York
newyork_2012_sub <- raw_newyork_2012 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

NY_2012 <- 1-(count(newyork_2012_sub)/count(raw_newyork_2012))
#expected 544661 reports
# did not receive 1864
#0.9965777 received
```

2013
```{r dmr_2013}
## Maine
maine_2013_sub <- raw_maine_2013 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 82573 reports
# did not receive 424
ME_2013 <- 1-(count(maine_2013_sub)/count(raw_maine_2013))
#0.9948651 received

  
## New Hampshire
newhampshire_2013_sub <- raw_newhampshire_2013 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 43505 reports
# did not receive 0
NH_2013 <- 1-(count(newhampshire_2013_sub)/count(raw_newhampshire_2013))
#0.9975747 received

## Massachusetts
massachusetts_2013_sub <- raw_massachusetts_2013 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 105635 reports
# did not receive 0
MA_2013 <- 1-(count(massachusetts_2013_sub)/count(raw_massachusetts_2013))
#0.9857398 received


## Rhode Island
rhodeisland_2013_sub <- raw_rhodeisland_2013 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 37956 reports
# did not receive 0
RI_2013 <- 1-(count(rhodeisland_2013_sub)/count(raw_rhodeisland_2013))
#1 received

## Connecticut
connecticut_2013_sub <- raw_connecticut_2013 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 129114 reports
# did not receive 920
CT_2013 <- 1-(count(connecticut_2013_sub)/count(raw_connecticut_2013))
#0.9928745 received


## New York
newyork_2013_sub <- raw_newyork_2013 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 545149 reports
# did not receive 2230
NY_2013 <- 1-(count(newyork_2013_sub)/count(raw_newyork_2013))
#0.9959094 received
```

2014
```{r dmr_2014}
## Maine
maine_2014_sub <- raw_maine_2014 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 82143 reports
# did not receive 238
ME_2014 <- 1-(count(maine_2014_sub)/count(raw_maine_2014))
#0.9971026 received

  
## New Hampshire
newhampshire_2014_sub <- raw_newhampshire_2014 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 43494 reports
# did not receive 0
NH_2014 <- 1-(count(newhampshire_2014_sub)/count(raw_newhampshire_2014))
#0.9975747 received

## Massachusetts
massachusetts_2014_sub <- raw_massachusetts_2014 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 108101 reports
# did not receive 1245
MA_2014 <- 1-(count(massachusetts_2014_sub)/count(raw_massachusetts_2014))
#0.988483 received


## Rhode Island
rhodeisland_2014_sub <- raw_rhodeisland_2014 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 34257 reports
# did not receive 1032
RI_2014 <- 1-(count(rhodeisland_2014_sub)/count(raw_rhodeisland_2014))
#0.9698748 received

## Connecticut
connecticut_2014_sub <- raw_connecticut_2014 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 132920 reports
# did not receive 1280
CT_2014 <- 1-(count(connecticut_2014_sub)/count(raw_connecticut_2014))
#0.9903701 received


## New York
newyork_2014_sub <- raw_newyork_2014 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 545446 reports
# did not receive 2140
NY_2014 <- 1-(count(newyork_2014_sub)/count(raw_newyork_2014))
#0.9959094 received
```

2015
```{r dmr_2015}
## Maine
maine_2015_sub <- raw_maine_2015 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 81137 reports
# did not receive 344
ME_2015 <- 1-(count(maine_2015_sub)/count(raw_maine_2015))
#0.9957603 received

  
## New Hampshire
newhampshire_2015_sub <- raw_newhampshire_2015 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 41484 reports
# did not receive 0
NH_2015 <- 1-(count(newhampshire_2015_sub)/count(raw_newhampshire_2015))
#0.9975747 received

## Massachusetts
massachusetts_2015_sub <- raw_massachusetts_2015 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 110400 reports
# did not receive 1494
MA_2015 <- 1-(count(massachusetts_2015_sub)/count(raw_massachusetts_2015))
#0.988483 received


## Rhode Island
rhodeisland_2015_sub <- raw_rhodeisland_2015 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 34740 reports
# did not receive 1096
RI_2015 <- 1-(count(rhodeisland_2015_sub)/count(raw_rhodeisland_2015))
#0.9684514 received

## Connecticut
connecticut_2015_sub <- raw_connecticut_2015 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 144541 reports
# did not receive 1236
CT_2015 <- 1-(count(connecticut_2015_sub)/count(raw_connecticut_2015))
#0.9914488 received


## New York
newyork_2015_sub <- raw_newyork_2015 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 549224 reports
# did not receive 2117
NY_2015 <- 1-(count(newyork_2015_sub)/count(raw_newyork_2015))
#0.9959094 received
```

2016
```{r dmr_2016}
## Maine
maine_2016_sub <- raw_maine_2016 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 82597 reports
# did not receive 431
ME_2016 <- 1-(count(maine_2016_sub)/count(raw_maine_2016))
#0.9957603 received

## New Hampshire
newhampshire_2016_sub <- raw_newhampshire_2016 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 45762 reports
# did not receive 0
NH_2016 <- 1-(count(newhampshire_2016_sub)/count(raw_newhampshire_2016))
#0.9975747 received


## Massachusetts
massachusetts_2016_sub <- raw_massachusetts_2016 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 121687 reports
# did not receive 1480
MA_2016 <- 1-(count(massachusetts_2016_sub)/count(raw_massachusetts_2016))
#0.988483 received


## Rhode Island
rhodeisland_2016_sub <- raw_rhodeisland_2016 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 33917 reports
# did not receive 827
RI_2016 <- 1-(count(rhodeisland_2016_sub)/count(raw_rhodeisland_2016))
#0.9756169 received


## Connecticut
connecticut_2016_sub <- raw_connecticut_2016 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 155873 reports
# did not receive 1512
CT_2016 <- 1-(count(connecticut_2016_sub)/count(raw_connecticut_2016))
#0.9914488 received


## New York
newyork_2016_sub <- raw_newyork_2016 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 560331 reports
# did not receive 3730
NY_2016 <- 1-(count(newyork_2016_sub)/count(raw_newyork_2016))
#0.993 received
```

2017
```{r dmr_2017}
## Maine
maine_2017_sub <- raw_maine_2017 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 83635 reports
# did not receive 388
ME_2017 <- 1-(count(maine_2017_sub)/count(raw_maine_2017))
#0.9953608 received

  
## New Hampshire
newhampshire_2017_sub <- raw_newhampshire_2017 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 47424 reports
# did not receive 0
NH_2017 <- 1-(count(newhampshire_2017_sub)/count(raw_newhampshire_2017))
#1 received

## Massachusetts
massachusetts_2017_sub <- raw_massachusetts_2017 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 124349 reports
# did not receive 728
MA_2017 <- 1-(count(massachusetts_2017_sub)/count(raw_massachusetts_2017))
#0.9941455 received


## Rhode Island
rhodeisland_2017_sub <- raw_rhodeisland_2017 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 33742 reports
# did not receive 705
RI_2017 <- 1-(count(rhodeisland_2017_sub)/count(raw_rhodeisland_2017))
#0.9756169 received

## Connecticut
connecticut_2017_sub <- raw_connecticut_2017 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 166786 reports
# did not receive 1686
CT_2017 <- 1-(count(connecticut_2017_sub)/count(raw_connecticut_2017))
#0.9914488 received


## New York
newyork_2017_sub <- raw_newyork_2017 %>% 
  select(EXTERNAL_PERMIT_NMBR, RNC_DETECTION_CODE) %>% 
  filter(RNC_DETECTION_CODE == "N")

#expected 565392 reports
# did not receive 4728
NY_2017 <- 1-(count(newyork_2017_sub)/count(raw_newyork_2017))
#0.9959094 received
```

## Calculating Scores and graphing
Creating a comprehensive dataset
```{r combine_dmrs}
states <- c("ME", "NH", "MA", "RI", "CT", "NY")

dmr_2009 <- rbind(ME_2009, NH_2009, MA_2009, RI_2009, CT_2009, NY_2009) %>% 
  cbind(states) %>% 
  as.data.frame() %>% 
  mutate(year= 2009) %>% 
  rename(percent_submitted=n, state = states, year=year)

dmr_2010 <- rbind(ME_2010, NH_2010, MA_2010, RI_2010, CT_2010, NY_2010) %>% 
  cbind(states) %>% 
  as.data.frame() %>% 
  mutate(year= 2010) %>% 
  rename(percent_submitted=n, state = states, year=year)
    
dmr_2011 <- rbind(ME_2011, NH_2011, MA_2011, RI_2011, CT_2011, NY_2011) %>% 
  cbind(states) %>% 
  as.data.frame() %>% 
  mutate(year= 2011) %>% 
  rename(percent_submitted=n, state = states, year=year)

dmr_2012 <- rbind(ME_2012, NH_2012, MA_2012, RI_2012, CT_2012, NY_2012) %>% 
  cbind(states) %>% 
  as.data.frame() %>% 
  mutate(year= 2012) %>% 
  rename(percent_submitted=n, state = states, year=year)
    
dmr_2013 <- rbind(ME_2013, NH_2013, MA_2013, RI_2013, CT_2013, NY_2013) %>% 
  cbind(states) %>% 
  as.data.frame() %>% 
  mutate(year= 2013) %>% 
  rename(percent_submitted=n, state = states, year=year)
    
dmr_2014 <- rbind(ME_2014, NH_2014, MA_2014, RI_2014, CT_2014, NY_2014) %>% 
  cbind(states) %>% 
  as.data.frame() %>% 
  mutate(year= 2014) %>% 
  rename(percent_submitted=n, state = states, year=year)
    
dmr_2015 <- rbind(ME_2015, NH_2015, MA_2015, RI_2015, CT_2015, NY_2015) %>% 
  cbind(states) %>% 
  as.data.frame() %>% 
  mutate(year= 2015) %>% 
  rename(percent_submitted=n, state = states, year=year)
    
dmr_2016 <- rbind(ME_2016, NH_2016, MA_2016, RI_2016, CT_2016, NY_2016) %>% 
  cbind(states) %>% 
  as.data.frame() %>% 
  mutate(year= 2016) %>% 
  rename(percent_submitted=n, state = states, year=year)
    
dmr_2017 <- rbind(ME_2017, NH_2017, MA_2017, RI_2017, CT_2017, NY_2017) %>% 
  cbind(states) %>% 
  as.data.frame() %>% 
  mutate(year= 2017) %>% 
  rename(percent_submitted=n, state = states, year=year)
```
    
```{r caculate_scores}    
reports_submitted <- rbind(dmr_2009, dmr_2010, dmr_2011, dmr_2012, dmr_2013, dmr_2014, dmr_2015, dmr_2016, dmr_2017) %>% 
  mutate(percent_submitted = as.character(percent_submitted)) %>% 
  mutate(percent_submitted = as.numeric(percent_submitted))

#save data
write.csv(reports_submitted,"data/reports_submitted.csv")
```

Visualization
```{r graph}
reports_submitted_graph <- ggplot(reports_submitted) +
  geom_line(aes(x= year, y= percent_submitted, color=state))+
  ggtitle("Percentage of Reports Submitted in NE") +
  ylab("Percent") +
  xlab("Year") +
  theme_classic() +
  geom_hline(yintercept=0.95)

reports_submitted_graph

##save fig
ggsave("figs/reports_submitted_graph.jpg", width=7, height=5, dpi=300)

```

