---
title: "OHIEC: Resilience - Water Pollution"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates water pollution resilience values by OHINE region. 

The data layers are used to quantify existence of regulations, implementation + enforcement, and effectiveness + compliance. These three parts are then used to calculate one overall water pollution resilience score for each region. 

1. Regulations: discharge_reg
2. Implementation + Enforcement: facilities_inspected, reports submitted
3. Effectiveness + Compliance: facilities_violations

Data Time Ranges:
discharge_reg: 2005-2017
facilities_inspected: 2005-2017
reports_submitted: 2009-2017
facilities_violations: 2005-2017

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(dplyr)
library(tidyverse)
library(readxl)
```


Read in Data
```{r}
discharge_reg <- read_xlsx(file.path(dir_anx, "_raw_data/manually_created/discharge_reg.xlsx"))

facilities_inspected_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/water_pollution/data/facilities_inspected.csv") %>% 
  select(-X) 

reports_submitted_short<- read.csv("~/github/ne-prep/prep/resilience/ecological/water_pollution/data/reports_submitted.csv") %>% 
  select(-X)

facilities_violations_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/water_pollution/data/facilities_violations.csv") %>% 
  select(-X)
```


Creating a df to add on to the water pollution scores to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  3, "Gulf of Maine",
  2, "Georges Bank",
  4, "Mid-Atlantic Bight",
  1, "Offshore") 

miss_rgns <- miss_rgns %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(score = NA) %>% 
  unique()
```

Cleaning up the data to be combined and used as a layer

```{r}
facilities_inspected <-facilities_inspected_short %>% 
  select(year, state , percent_w_inspections) %>% 
  mutate(rgn_id1 = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "MA" ~ "7",
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10"),
    rgn_id2 = case_when(
    state == "MA" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  rename(score = percent_w_inspections) %>% 
  rbind(miss_rgns) %>% 
  rename(facilities_inspected_score = score)
```


Creating a df to add on to the reorts submitted data to have the years 2005-2008
```{r missing_years_df}
miss_years <- c("Maine","New Hampshire", "Massachusetts-North", "Massachusetts-South","Rhode Island", "Connecticut", "New York") %>% 
  as.data.frame() %>% 
  uncount(4, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  rename(rgn_name = ".") %>% 
  mutate(score = NA) %>% 
  unique()
```


```{r}
reports_submitted <- reports_submitted_short %>% 
  rename(score = percent_submitted) %>% 
   mutate(rgn_id1 = case_when(
    state == "ME" ~ "Maine", 
    state == "NH" ~ "New Hampshire", 
    state == "MA" ~ "Mass",
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10"),
    rgn_id2 = case_when(
    state == "MA" ~ "8"
    )) 
  
  
  
```

















#Load Data
```{r}

raw_inspec <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/NPDES_INSPECTIONS.csv"))
raw_report_history <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/NPDES_QNCR_HISTORY.csv"))
raw_facilities <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/ICIS_FACILITIES.csv"))
```

Gapfilling for years 2005-2008
to make the dmr submission csv match/bind with the other two

```{r}

y5 <- rep(2005, times= 6) 
y6 <- rep(2006, times= 6) 
y7 <- rep(2007, times= 6) 
y8 <- rep(2008, times= 6) 
year <- c(y5, y6, y7, y8)
  
state <- rep(c("ME", "NH", "MA", "RI", "CT", "NY"), times = 4)

gapfill <- cbind(year, state) %>% 
  as.data.frame() %>% 
  mutate(percent_submitted = case_when(
    state == "ME" ~ "0.9668811",
    state == "NH" ~ "0.9971842",
    state == "MA" ~ "0.9841761",
    state == "RI" ~ "1",
    state == "CT" ~ "1",
    state == "NY" ~ "0.9973308")) %>% 
  select(percent_submitted, state, year) %>% 
  mutate(percent_submitted = as.numeric(percent_submitted),
         year = as.character(year))

dmr_gapfill_submissions <- rbind(gapfill, dmr_submissions)  %>% 
  mutate(year = as.numeric(year))

write.csv(dmr_gapfill_submissions,"data/dmr_gapfill_submissions.csv")

```


#**Compliance**
##Violations
Percentage of facilities with compliance regualtions by year by state
Note: not differentiating between types of violations (of which there are 4)

```{r}

violations_hist <- raw_report_history %>% 
  mutate(state = substr(NPDES_ID, 0, 2),
         year = substr(YEARQTR, 0, 4))%>% 
  filter(year >=2005 & year <=2017) %>% 
  filter(state == "ME" | state == "NH" | state == "MA" | state =="RI" | state == "CT" | state == "NY") %>% 
  mutate(violation = 1) %>% #can assign a 1 to all columns becuase all rows included here mean there was a violation of some sort
  select(year, state, NPDES_ID, violation) %>% 
  distinct() %>%  # want to know if a violation that year, doesn't matter what quarter or if multiple violations in one year
  group_by(year, state) %>% 
  count() %>% 
  rename(fac_violations= n)

```

```{r}

## percent of facilities with violations by state

violations <- left_join(num_fac, violations_hist, by = c("year", "state"))%>% 
  mutate(percent_w_violations = fac_violations/fac_total,
         percent_no_violation = 1- percent_w_violations) %>% #to avoid dividing by a target of 0 violationfor later score calculations 
  mutate(year = as.numeric(year))

#save data
write.csv(violations, "data/violations.csv")
```

Visualization
```{r}

violations_graph <- ggplot(violations) +
  geom_line(aes(x= year, y = percent_w_violations, color= state)) +
  ggtitle("Percentage of Facilities with Violations in NE") +
  ylab("Percent") +
  xlab("Year") +
  theme_classic()

violations_graph
##save fig
ggsave("figs/violations_graph.jpg", width=7, height=5, dpi=300)

```



#Creating a single data frame for water pollution resilience metric


```{r}

wp_res_metrics <- left_join(dmr_gapfill_submissions, violations, by = c("year", "state")) %>% 
  left_join(.,facilities_inspected, by = c("year", "state")) %>% 
  select(year, state, percent_submitted, percent_no_violation, percent_w_inspections)%>% 
  mutate(target_inspections = 0.5,
         target_no_violations = 1,
         target_submitted = 0.95,
         target_regulation = 1,
         percent_regulation = 1) %>% 
  select(year, state, percent_regulation, target_regulation, percent_submitted, target_submitted, percent_no_violation,target_no_violations, percent_w_inspections, target_inspections) %>% 
  mutate(rgn_id1 = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "MA" ~ "7",
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10"),
    rgn_id2 = case_when(
    state == "MA" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"))
  

write.csv(wp_res_metrics, "data/wp_res_metrics.csv")

```


Calculating Scores:

For the water pollution resilience scores (wp_res_scores) a score of 1 would be the best

4 equally weighted metrics used in the calculation:
1. Score of 1 for having regulations in place (ie: the EPA); no regulation would be 0
2. Score between 0-1 for inspections of discharge facilities. Calculated by taking the percent of facilities inspected at least once/the EPA's target go 50%. If reached target a score of 1 would be acheived
3. Score between 0-1 for discharge facilitiy violations. Calculated by taking the percent of facilities that had no violations/ the EPA's target of 100% no violations. If reached target a score of 1 would be acheived
4. Score between 0-1 for submissions of discharge monitoring reports (DMR). Calculated by taking the percent of DMRs submitted/ % the EPA's target of 95% submission of expected DMRs. If reached target a score of 1 would be acheived

Creating a df to add on to the water pollution scores to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  3, "Gulf of Maine",
  2, "Georges Bank",
  4, "Mid-Atlantic Bight",
  1, "Offshore") 

miss_rgns <- miss_rgns %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(score = NA) %>% 
  unique()
```

```{r}
wp_res_score <- wp_res_metrics %>% 
  mutate(regulation_score = percent_regulation/target_regulation,
         inspection_score = percent_w_inspections/target_inspections,
         noviolation_score = percent_no_violation/target_no_violations,
         submitted_score = percent_submitted/target_submitted,
         submitted_score = if_else(submitted_score >1, 1, submitted_score)) %>%  # capping the score at 1
  select(year, rgn_name, rgn_id, regulation_score, inspection_score, noviolation_score, submitted_score) %>% 
  group_by(year, rgn_name, rgn_id) %>% 
  summarise(wp_res_score = sum(inspection_score, submitted_score, noviolation_score, regulation_score)/4) %>% 
  ungroup() %>% 
  rename(score = wp_res_score) %>% 
  rbind(miss_rgns)

write.csv(wp_res_score, "data/wp_res_score.csv")
```

Visualization
```{r}

wp_res_scores_graph <- ggplot(wp_res_score) +
  geom_line(aes(x=year, y= score, color = rgn_name))+
  ggtitle("North East Water Pollution Resilience Scores") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 

wp_res_scores_graph
ggsave("figs/wp_res_scores_graph.jpg", width=7, height=5, dpi=300)

```

Write as a pressure layer to ne-scores
```{r}
write.csv(wp_res_score, file.path(dir_calc, "layers/res_water_pollution.csv"))
```



Playing around with Water Quality monitoring data/ frequency 

put data in the EPA ECHO, wither remove when done or put in its own folder bc its not epa echo

```{r}
test <- read_csv(file.path(dir_anx, "_raw_data/EPA_ECHO/activityall_ny.csv"))

colnames(test)

t1 <- test %>% 
  select(ActivityStartDate,ActivityConductingOrganizationText, ActivityTypeCode, HydrologicEvent, HydrologicCondition) %>% 
  unique() %>% 
  separate(ActivityStartDate, c("year", "month", "day"), sep = "-") %>% 
  filter(HydrologicEvent %in% c("Routine sample")) %>% 
  group_by(year) %>% 
  count() %>% 
  ungroup() %>% 
  rename(times_sampled = n) %>% 
  mutate(target = max(times_sampled),
         score = times_sampled/target) 

```












