---
title: "OHINE: Water Pollution Resilience - Calculating the Water Pollution Resilience Layer"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates water pollution resilience values by OHINE region. 

The data layers are used to quantify existence of regulations, implementation + enforcement, and effectiveness + compliance. These three parts are then used to calculate one overall water pollution resilience score for each region. 

1. Regulations: discharge_reg
2. Implementation + Enforcement: facilities_inspected, reports submitted
3. Effectiveness + Compliance: facilities_violations

Data Time Ranges:

discharge_reg: 2005-2017

facilities_inspected: 2005-2017

reports_submitted: 2009-2017

facilities_violations: 2005-2017

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(dplyr)
library(tidyverse)
library(readxl)
```


Read in Data
```{r load_data}
discharge_reg <- read_xlsx(file.path(dir_anx, "_raw_data/manually_created/discharge_reg.xlsx"))

facilities_inspected_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/water_pollution/data/facilities_inspected.csv") %>% 
  select(-X) 

reports_submitted_short<- read.csv("~/github/ne-prep/prep/resilience/ecological/water_pollution/data/reports_submitted.csv") %>% 
  select(-X)

facilities_violations_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/water_pollution/data/facilities_violations.csv") %>% 
  select(-X)
```


Creating a df to add on to the water pollution scores to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  3, "Gulf of Maine",
  2, "Georges Bank",
  4, "Mid-Atlantic Bight",
  1, "Offshore") 

miss_rgns <- miss_rgns %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(score = NA) %>% 
  unique()
```

## Tidying of data inputs
Cleaning up the data to be combined and used as a layer

Facilities inspected
```{r fac_in_tidy}
facilities_inspected <-facilities_inspected_short %>% 
  select(year, state , percent_w_inspections) %>% 
  mutate(rgn_id1 = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "MA" ~ "7",
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10"),
    rgn_id2 = case_when(
    state == "MA" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  mutate(target = 0.5,
         score = percent_w_inspections/target,
         rgn_id = as.numeric(rgn_id)) %>% 
  select(year, rgn_id, rgn_name, score) %>% 
  rbind(miss_rgns) %>% 
  rename(facilities_inspected_score = score)
```


Creating a df to add on to the reorts submitted data to have the years 2005-2008
```{r missing_years_df}
miss_years <- c("ME","NH", "MA","RI", "CT", "NY") %>% 
  as.data.frame() %>% 
  uncount(4, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  rename(state = ".") %>% 
  mutate(score = NA) %>% 
  unique()
```

Reports Submitted
```{r rep_sub_tidy}
reports_submitted <- reports_submitted_short %>% 
  mutate(target = 0.95,
         score = percent_submitted/target) %>%
  select(year, state, score) %>% 
  rbind(miss_years) %>% 
  mutate(rgn_id1 = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "MA" ~ "7",
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10"),
    rgn_id2 = case_when(
    state == "MA" ~ "8"
    )) %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != is.na(rgn_id))%>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York"),
    rgn_id = as.numeric(rgn_id)) %>% 
  arrange(rgn_name, year) %>% 
  fill(score, .direction = c("up")) %>%  #using the first year score in 2009 as the score for 2005-2008
  rbind(miss_rgns) %>% 
  rename(reports_submitted_score = score)
```

Facilities Violations
```{r fac_vio_tidy}
facilities_violations <- facilities_violations_short %>% 
  mutate(rgn_id1 = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "MA" ~ "7",
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10"),
    rgn_id2 = case_when(
    state == "MA" ~ "8"
    )) %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != is.na(rgn_id))%>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  mutate(target= 1,
         score = percent_no_violation/target,
        rgn_id = as.numeric(rgn_id)) %>% 
  select(year, rgn_id, rgn_name, score) %>% 
  rbind(miss_rgns) %>% 
  rename(facilities_violations_score = score)
```

## Calculate Scores
Combine all the data together and calcualte a regulation, enforcement, compliance, and overall resilience score
```{r calc_scores}
res_wp <- left_join(discharge_reg, facilities_inspected, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(facilities_violations, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(reports_submitted, by = c("year", "rgn_id", "rgn_name")) %>% 
  rowwise() %>% 
  mutate(enforcement = mean(c(facilities_inspected_score, reports_submitted_score))) %>% 
  rename(regulation = discharge_reg_score) %>% 
  rename(compliance = facilities_violations_score) %>% 
  mutate(res_score = mean(c(regulation, enforcement, compliance)))
```

Save necessary information for the water pollution layer
```{r save_data}
res_water_pollution <- res_wp %>% 
  select(year, rgn_id, rgn_name, res_score) %>% 
  rename(score = res_score)

write.csv(res_water_pollution, file.path(dir_git,"prep/resilience/ecological/water_pollution/data/res_water_pollution.csv"))
```

Visualization
```{r graph}
res_water_pollution_graph <- ggplot(res_water_pollution) +
  geom_line(aes(x=year, y= score, color = rgn_name))+
  ggtitle("North East Water Pollution Resilience Scores") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 

res_water_pollution_graph
ggsave("figs/res_water_pollution_graph.jpg", width=7, height=5, dpi=300)
```


Write as a resilience layer to ne-scores
```{r write_to_scores}
write.csv(res_water_pollution, file.path(dir_calc, "layers/res_water_pollution.csv"))
```



