---
title: "OHIEC: Resilience - Climate Change Pressure prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates fishing pressure resilience values by OHINE region. 

# Data

**Reference**: 

**Downloaded**: 

**Description**:  

**Time range**: 

**Format**: 

Regulations:
1) extract info from America's Pledge Report
2) American Council for Energy Efficiency:
Rankings: https://database.aceee.org/state/

Effectiveness:
US. energy information administration:
1) Total CO2 emissions https://www.epa.gov/statelocalenergy/state-co2-emissions-fossil-fuel-combustion
2) Total power generation:https://www.eia.gov/electricity/data/state/

Targets for effectiveness from Beyond Carbon https://www.beyondcarbon.org/look-up-your-state/


``` {r setup,  message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(Hmisc)
library(stringi)
library(readxl)

```


Data
```{r}

eee <- read_excel(file.path(dir_anx, "_raw_data/EEE/eee_scores.xlsx"))
raw_gen <- read_excel(file.path(dir_anx, "_raw_data/EIA/eia_annual_generation_state.xls"))
raw_clim_act <- read_excel(file.path(dir_anx, "_raw_data/Americas_Pledge/states_climate_regs_summary.xlsx"), skip=2, sheet = 2)

```

**Regualtions**
Energy Efficiency Scores
```{r}

#calculate
eee_score <- eee %>% 
  mutate(score = eee_score/eee_target) %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n)

#save
write.csv(eee_score, file = "data/eee_score")

```

Visualize
```{r}
eee_score_graph <-ggplot(eee_score)+
  geom_line(aes(x=year, y = score, color=state))+
  ggtitle("Energy Efficiency Score") +
  ylab("Score") +
  xlab("State") +
  theme_classic() 
eee_score_graph

ggsave("figs/eee_score_graph.jpg", width=7, height=5, dpi=300)
```

Climate Friendly Actions
total of 30 possible actions to be taken. 
Set target as 30. A state would need to have taken some sort of action in each of the following categories to get 100%

1) Integrating transportation & land-use in comprehensive plans	
2) Dedicated funding streams for public transit
3) Adopting legislation in line with Complete Streets objectives
4) Financial incentives for high efficiency vehicles
5) Conservation easement tax credits
6) California’s vehicle emission standards
7) Rules and incentives to reduce food waste
8) Most recent building energy codes
9) Appliance and equipment energy efficiency standards
10) Carbon pricing	
11) Zero Emission Vehicle mandate
12) Cost-sharing programs to improve forest systems
13) Wildfire protection incentives
14) Coal mine methane standards
15) Low carbon fuel standard
16) Methane standards for existing oil and natural gas facilities
17) Setting methane emission reduction targets
18) Freight-specific energy efficiency performance metrics
19) HFC management program requiring all leaks to be repaired
20) Financial incentives for CCS
21) Zero-emission credits for nuclear
22) GHG Emission Targets
23) Renewable energy portfolio standards or goals
24) Property Assessed Clean Energy
25) Combined heat and power financing and incentives
26) Landfill gas energy project bond, grant, loan, or rebate programs
27) Energy efficiency resource standard or goals
28) Property tax programs to support sustainable forests
29) Efficient vehicle requirement for public fleet procurement
31) Freight plan  with multimodal freight strategies


```{r}

clim_act <- raw_clim_act %>% 
  filter(State == "Maine" | State == "Massachusetts" | State == "New Hampshire" | State == "Rhode Island" | State == "Connecticut" | State == "New York" | State == "Total # of actions considered") %>% 
  select(-"Energy efficiency", -"Total") %>% 
  mutate(target = 23,
         total = rowSums(.[2:9]),
         score = total/target) %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  filter(State != "Total # of actions considered")

write.csv(clim_act, file = "data/clim_act")

```

```{r}

clim_act_graph <-ggplot(clim_act)+
  geom_line(aes(x=year, y = score, color=State))+
  ggtitle("Climate Friendly Actions Score") +
  ylab("Score") +
  xlab("State") +
  theme_classic() 
clim_act_graph

ggsave("figs/clim_act_graph.jpg", width=7, height=5, dpi=300)
```


**Effectivness**

##Clean Energy Cosumption

**Renewable/Clean Energy Targets:**
Drawn from http://www.ncsl.org/research/energy/renewable-portfolio-standards.aspx#me
note: all targets are part os "renewable energy portfolio standards" so I won't be including nuclear energy into calcualting if states are on track for targets

1. Maine = 40% by 2017
2. New Hampshire = 25.2% by 2025
3. Connecticut = 44% by 2030
4. Massachusetts = 35% 2030
5. New York = 50% by 2030
6. Rhode Island = 38.5% by 2035

https://www.eia.gov/beta/states/states/ny/data/dashboard/total-energy

Load in the energy consumption data for each state. All numbers are in Billion Btu
```{r}
ct_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ct_consumption.csv"), skip = 1)
me_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/me_consumption.csv"), skip = 1)
ma_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ma_consumption.csv"), skip = 1)
ri_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ri_consumption.csv"), skip = 1)
nh_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/nh_consumption.csv"), skip = 1)
ny_con_raw <- read_csv(file.path(dir_anx, "_raw_data/EIA/energy_consumption/ny_consumption.csv"), skip = 1)
```

Connecticut
```{r}
ct_con <- ct_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Connecticut",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Connecticut",
         nuc = "Nuclear energy consumed for electricity generation, total, Connecticut",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Connecticut",
         renew = "Renewable energy total consumption, Connecticut",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Connecticut",
         net_elec_import = "Net imports of electricity into the United States, Connecticut") %>% 
  mutate(state = "Connecticut") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Maine
```{r}
me_con <- me_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Maine",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Maine",
         nuc = "Nuclear energy consumed for electricity generation, total, Maine",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Maine",
         renew = "Renewable energy total consumption, Maine",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Maine",
         net_elec_import = "Net imports of electricity into the United States, Maine") %>% 
  mutate(state = "Maine") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Massachusetts
```{r}
ma_con <- ma_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Massachusetts",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Massachusetts",
         nuc = "Nuclear energy consumed for electricity generation, total, Massachusetts",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Massachusetts",
         renew = "Renewable energy total consumption, Massachusetts",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Massachusetts",
         net_elec_import = "Net imports of electricity into the United States, Massachusetts") %>% 
  mutate(state = "Massachusetts") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Rhode Island
```{r}
ri_con <- ri_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, Rhode Island",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), Rhode Island",
         nuc = "Nuclear energy consumed for electricity generation, total, Rhode Island",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, Rhode Island",
         renew = "Renewable energy total consumption, Rhode Island",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), Rhode Island",
         net_elec_import = "Net imports of electricity into the United States, Rhode Island") %>% 
  mutate(state = "Rhode Island") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

New Hampshire
```{r}
nh_con <- nh_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, New Hampshire",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), New Hampshire",
         nuc = "Nuclear energy consumed for electricity generation, total, New Hampshire",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, New Hampshire",
         renew = "Renewable energy total consumption, New Hampshire",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), New Hampshire",
         net_elec_import = "Net imports of electricity into the United States, New Hampshire") %>% 
  mutate(state = "New Hampshire") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

New York
```{r}
ny_con <- ny_con_raw %>% 
  select(-"Series Name_1",-"Series Name_2",-"Series Name_3",-"Series Name_4",-"Series Name_5",-"Series Name_6") %>% 
  rename(year = "Series Name",
         coal = "Coal total consumption, New York",
         nat_gas = "Natural gas total consumption (excluding supplemental gaseous fuels), New York",
         nuc = "Nuclear energy consumed for electricity generation, total, New York",
         petrol = "All petroleum products total consumption, excluding fuel ethanol blended into motor gasoline, New York",
         renew = "Renewable energy total consumption, New York",
         net_elec_flow = "Net interstate flow of electricity and associated losses (negative indicates flow out of state), New York",
         net_elec_import = "Net imports of electricity into the United States, New York") %>% 
  mutate(state = "New York") %>% 
  filter(year <= "2017") %>% 
  select(-net_elec_flow, -net_elec_import)
```

Combine all states energy consumption data into one and calculate the percent of renewable. Also calculating percent with clean(ie: nuclear) for fun/its interesting
```{r}
consum <- rbind(me_con, ma_con, ct_con, nh_con, ny_con, ri_con) %>% 
  mutate(coal = as.numeric(coal),
         nat_gas = as.numeric(nat_gas),
         nuc = as.numeric(nuc),
         petrol = as.numeric(petrol),
         renew = as.numeric(renew)) %>% 
  rowwise() %>% 
  mutate(total = sum(coal, nat_gas, nuc, petrol, renew),
         non_renew = sum(coal, nat_gas, nuc, petrol),
         per_renew = renew/total,
         year = as.numeric(year),
         clean = sum(nuc, renew),
         per_clean= clean/total) %>% 
  ungroup()
```

Visualize
```{r}
# renewable only
con_renew_graph <- ggplot(consum)+
  geom_line(aes(x=year, y= per_renew, color= state))+
 ggtitle("Percent of renewable energy used (no nuclear)") +
  ylab("Percent") +
  xlab("Year") +
  theme_classic()

con_renew_graph

#clean(nuclear) and renewable
con_clean_graph <- ggplot(consum)+
  geom_line(aes(x=year, y= per_clean, color= state))+
 ggtitle("Percent of clean/renewable energy used") +
  ylab("Percent") +
  xlab("Year") +
  theme_classic()

con_clean_graph
```

Now that we have the actual percentage of energy consumption that comes from renewable, we can do linear regressions using the state set targets as the end points. With this we can calculate if each state is on track to hitting their renewable energy target goal for a given year.

Maine
```{r}
#MA 40% clean by 2017
me <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Maine") %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
me_target <- tribble(
  ~year, ~per_clean,
  1, 0.277,
  14, 0.4
)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_me <- lm(per_clean ~ year, data = me_target)
#intercept = 0.26754 ; coef = 0.00946 
summary(lm_me)

plot(per_clean ~ year, data = me_target)
abline(lm_me)

###trying to use equation no loop
me_clean_target <- me %>% 
  rowwise() %>% 
  mutate(target = (0.00946*year_n + 0.26754 ),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter( year != "2004") %>% 
  select(year, state, per_renew, target) 
```

Massachusetts
```{r}
#MA 35% clean by 2030
ma <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Massachusetts")  %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ma_target <- tribble(
  ~year, ~per_clean,
  1, 		0.0387	,
  26, .35)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ma <- lm(per_clean ~ year, data = ma_target)
#coef = 0.0125    ; int = 0.0262  

ma_clean_target <- ma %>% 
  rowwise() %>% 
  mutate(target = (0.0125*year_n+ 0.0262   ),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter( year != "2004") %>% 
  select(year, state, per_renew, target)
```

New Hampshire
```{r}
#NH 25.2% clean by 2025
nh <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "New Hampshire") %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
nh_target <- tribble(
  ~year, ~per_clean,
  1, 0.0770,
  21, .252
)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_nh <- lm(per_clean ~ year, data = nh_target)
#coef=  0.00875 int=  0.06825 

nh_clean_target <- nh %>% 
  mutate(target = ( 0.00875    *year_n + 0.06825 ),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
   filter( year != "2004") %>% 
  select(year, state, per_renew, target)
```


Rhode Island
```{r}
#RI 38.5% clean by 2035
ri <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Rhode Island") %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ri_target <- tribble(
  ~year, ~per_clean,
  0, 0.0256	,
  31, .385)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ri <- lm(per_clean ~ year, data = ri_target)
#coef=0.0256  , int = 0.0116  

ri_clean_target <- ri %>% 
   mutate(target = ( 0.0116  *year_n+0.0256),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_renew, target)
```


New York
```{r}
#NY 50% clean by 2030
ny <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "New York") %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ny_target <- tribble(
  ~year, ~per_clean,
  0, 	0.0950	,
  26, .50)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ny <- lm(per_clean ~ year, data = ny_target)
#coef=0.0156   , int =   0.0950

ny_clean_target <- ny %>% 
  mutate(target = (0.0156*year_n +0.0950 ),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_renew, target)
```


Connecticut
```{r}
#CT 44% clean by 2030
ct <- consum %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  arrange(year) %>% 
  filter(state == "Connecticut")  %>% 
  mutate(year_n = seq(1:14))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ct_target <- tribble(
  ~year, ~per_clean,
  0, 0.0490,
  26, .44
)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ct <- lm(per_clean ~ year, data = ct_target)
#coef= 0.015  , int= 0.049 

ct_clean_target <- ct %>% 
  mutate(target = (0.015*year_n + 0.049 ),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_renew, target)
```



Combine all states df with target
```{r}
energy_target <- rbind(me_clean_target, nh_clean_target, ma_clean_target, ri_clean_target, ct_clean_target, ny_clean_target)
```


Calculate scores
```{r}
power_gen_scores <- energy_target %>% 
  mutate(score = per_renew/target,
         score = if_else(score >1, 1, score)) 
#save
write.csv(power_gen_scores, file = "data/power_gen_scores_nuc")
```

Visualize scores
```{r}  

graph_clean_energy_score_nuc <-ggplot(power_gen_scores)+
  geom_line(aes(x=year, y = score, color=state))+
  ggtitle("Clean Energy Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 
graph_clean_energy_score_nuc
ggsave("figs/graph_clean_energy_score.jpg", width=7, height=5, dpi=300)

```

##Carbon Emmissions

**Carbon Emmission Targets:**
Targets pulled from https://www.c2es.org/document/greenhouse-gas-emissions-targets/
Rhode Island is missing from this map so pulled from http://www.planning.ri.gov/documents/LU/energy/energy15.pdf

Maine= 80% reduction below 1990 levels by 2050
New Hampshire= 80% reduction below 1990 levels by 2050
Connecticut= 80% reduction below 2001 levels by 2050
Massachusetts= 80% reduction below 1990 levels by 2050
New York= 100% reduction below 1990 levels by 2050
Rhode Island= 80% reduction below 1990 levels by 2050

Converting table from a pdf from EPA website to a df
```{r}

library(tabulizer)
library(dplyr)

location <- 'https://www.epa.gov/sites/production/files/2017-09/documents/co2ffc_2015.pdf'
out <- extract_tables(location)

#page 1
out1 <- out[[1]] %>%  
  as.data.frame() %>% 
  rename(state = V1, blank = V2, "1990" = V3, "1991"= V4, "1992"= V5, "1993"=V6, "1994"=V7, "1995"=V8, "1996"=V9, "1997"=V10, "1998"=V11, "1999"=V12, "2000"=V13, "2001"=V14, "2002"=V15, "2003"=V16, "2004"=V17, "2005"=V18, "2006"=V19, "2007"=V20, "2008"=V21, "2009"=V22, "2010"=V23, "2011"=V24, "2012"=V25, "2013"=V26, "2014"=V27, "2015"=V28, "2016"=V29) %>% 
  filter(state != "State Sector",
         state != "Commercial",
         state != "Industrial", 
         state != "Residential",
         state != "Transportation",
         state != "Electric Power") %>% 
  select(-blank)
  
#page 2  
out2 <- out[[2]] %>% 
  as.data.frame() %>% 
  rename(state = V1, blank = V2, "1990" = V3, "1991"= V4, "1992"= V5, "1993"=V6, "1994"=V7, "1995"=V8, "1996"=V9, "1997"=V10, "1998"=V11, "1999"=V12, "2000"=V13, "2001"=V14, "2002"=V15, "2003"=V16, "2004"=V17, "2005"=V18, "2006"=V19, "2007"=V20, "2008"=V21, "2009"=V22, "2010"=V23, "2011"=V24, "2012"=V25, "2013"=V26, "2014"=V27, "2015"=V28, "2016"=V29) %>% 
  filter(blank != "Commercial",
         blank != "Industrial", 
         blank != "Residential",
         blank != "Transportation",
         blank != "Electric Power")%>% 
  select(-blank)

#page 3
out3 <- out[[3]] %>% 
   as.data.frame() %>% 
  rename(state = V1, "1990" = V2, "1991"= V3, "1992"= V4, "1993"=V5, "1994"=V6, "1995"=V7, "1996"=V8, "1997"=V9, "1998"=V10, "1999"=V11, "2000"=V12, "2001"=V13, "2002"=V14, "2003"=V15, "2004"=V16, "2005"=V17, "2006"=V18, "2007"=V19, "2008"=V20, "2009"=V21, "2010"=V22, "2011"=V23, "2012"=V24, "2013"=V25, "2014"=V26, "2015"=V27, "2016"=V28) %>% 
  filter(state != "Commercial",
         state != "Industrial", 
         state != "Residential",
         state != "Transportation",
         state != "Electric Power")

#page 4
out4 <- out[[4]] %>%  
   as.data.frame() %>% 
  rename(state = V1, "1990" = V2, "1991"= V3, "1992"= V4, "1993"=V5, "1994"=V6, "1995"=V7, "1996"=V8, "1997"=V9, "1998"=V10, "1999"=V11, "2000"=V12, "2001"=V13, "2002"=V14, "2003"=V15, "2004"=V16, "2005"=V17, "2006"=V18, "2007"=V19, "2008"=V20, "2009"=V21, "2010"=V22, "2011"=V23, "2012"=V24, "2013"=V25, "2014"=V26, "2015"=V27, "2016"=V28) %>% 
  filter(state != "Commercial",
         state != "Industrial", 
         state != "Residential",
         state != "Transportation",
         state != "Electric Power")

#page 5
out5 <- out[[5]] %>%  
     as.data.frame() %>% 
  rename(state = V1, "1990" = V2, "1991"= V3, "1992"= V4, "1993"=V5, "1994"=V6, "1995"=V7, "1996"=V8, "1997"=V9, "1998"=V10, "1999"=V11, "2000"=V12, "2001"=V13, "2002"=V14, "2003"=V15, "2004"=V16, "2005"=V17, "2006"=V18, "2007"=V19, "2008"=V20, "2009"=V21, "2010"=V22, "2011"=V23, "2012"=V24, "2013"=V25, "2014"=V26, "2015"=V27, "2016"=V28) %>% 
  filter(state != "Commercial",
         state != "Industrial", 
         state != "Residential",
         state != "Transportation",
         state != "Electric Power")
```


Combining and Tidying
```{r}

emissions <- rbind(out1, out2, out3, out4, out5) %>% 
  gather(year, emission, 2:28) %>% 
   filter(state == "Maine" | state == "New Hampshire" | state == "Massachusetts" | state == "New Hampshire" | state == "Rhode Island" | state == "Connecticut" | state == "New York") %>% 
  mutate(year = as.numeric(year),
         emission = as.numeric(emission)) %>% 
  mutate(end_target = case_when(state == "Maine" ~ (23.64 - 0.8*(23.64)), #targets for clean energy from Beyond Carbon
                            state == "New Hampshire" ~ (14.59 - 0.8*(14.59)),
                            state == "Connecticut" ~ (41.53 - 0.8*(41.53)),
                            state == "Massachusetts" ~ (83.09 - 0.8*(83.09)),
                            state == "New York" ~ (207.82 - 0.8*(207.82)),
                            state == "Rhode Island" ~ (8.91 - 0.8*(8.91))))

#save
write.csv(emissions, file = "data/emissions")

```

Visualize emmissions
```{r}

graph_em <-ggplot(emissions)+
  geom_line(aes(x=year, y = emission, color=state))+
  ggtitle("Total Carbon Emissions") +
  ylab("Emissions (million metric tons)") +
  xlab("Year") +
  theme_classic() 

graph_em 
ggsave("figs/graph_em .jpg", width=7, height=5, dpi=300)

```

Regressions for trend targets


New Hampshire
```{r}

#NH 80% reduction below 1990 levels by 2050
nh_em <- emissions %>% 
  filter(state == "New Hampshire") %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  mutate(year_n = seq(0:12))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
nh_em_target <- tribble(
  ~year, ~emission,
  0, 	21.8	,
  46, 2.92)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_nh_em <- lm(emission~ year, data = nh_em_target)
#coe =  -0.41   ; int= 21.80

nh_em_target <- nh_em %>% 
  mutate(target = ( -0.41  *year_n +21.80),
         emission = round(emission, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, emission, target)


```

Massachusetts
```{r}

#MA 80% reduction below 1990 levels by 2050
ma_em <- emissions %>% 
  filter(state == "Massachusetts") %>% 
  filter(year >= 2004 & year <= 2017)%>% 
  mutate(year_n = seq(0:12))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ma_em_target <- tribble(
  ~year, ~emission,
  0, 82.6	,
  46, 16.6)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ma_em <- lm(emission~ year, data = ma_em_target)
#coe=-1.43   ; int =82.60

ma_em_target <- ma_em %>% 
  mutate(target = (-1.43  *year_n +82.60 ),
         emission = round(emission, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, emission, target)

```

Rhode Island
```{r}

#RI 80% reduction below 1990 levels by 2050
ri_em <- emissions %>% 
  filter(state == "Rhode Island") %>% 
  filter(year >= 2004 & year <= 2017)%>% 
  mutate(year_n = seq(0:12))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ri_em_target <- tribble(
  ~year, ~emission,
  0, 10.97,
  46, 1.78
)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ri_em <- lm(emission~ year, data = ri_em_target)
#coe= -0.2 ; int = 11.0 

ri_em_target<- ri_em %>% 
    mutate(target = (-0.2*year_n+11.0 ),
         emission = round(emission, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, emission, target)

```


Connecticut
```{r}

#CT 80% reduction below 2001 levels by 2050
ct_em <- emissions %>% 
  filter(state == "Connecticut") %>% 
  filter(year >= 2004 & year <= 2017)%>% 
  mutate(year_n = seq(0:12))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ct_em_target <- tribble(
  ~year, ~emission,
  0, 44.5	,
  46, 8.31)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ct_em <- lm(emission~ year, data = ct_em_target)
#coef=-0.787; int = 44.500 

ct_em_target <- ct_em %>% 
  mutate(target = (-0.787 *year_n+44.500),
         emission = round(emission, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, emission, target)

```


New York
```{r}

#NY 80% reduction below 1990 levels by 2050
ny_em <- emissions %>% 
  filter(state == "New York") %>% 
  filter(year >= 2004 & year <= 2017) %>% 
  mutate(year_n = seq(0:12))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ny_em_target <- tribble(
  ~year, ~emission,
  0, 215,
  46, 41.6)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ny_em <- lm(emission~ year, data = ny_em_target)
#coe= -3.77 ; int = 215.00  

ny_em_target <- ny_em %>% 
  mutate(target = (-3.77 *year_n+215.00   ),
         emission = round(emission, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, emission, target)

```

Maine
  NO GOAL YEAR FOR EMISSION REDUCION 
```{r}

#MA 80% reduction below 2003 levels
me_em_target <- emissions %>% 
  filter(state == "Maine") %>% 
  filter(year >= 2005 & year <= 2017) %>% 
  rename(target = end_target)

```


Combine all states df with target
```{r}
emissions_target <- rbind(nh_em_target, ma_em_target, ri_em_target, ct_em_target, ny_em_target, me_em_target)

```


Calculate scores
```{r}
emissions_scores <- emissions_target %>% 
  mutate(score = target/emission,
         score = if_else(score >1, 1, score)) 
#save
write.csv(emissions_scores, file = "data/emissions_scores")

```

Visualize scores
```{r}  

graph_emissions_scores <-ggplot(emissions_scores)+
  geom_line(aes(x=year, y = score, color=state))+
  ggtitle("Carbon emission scores") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 
graph_emissions_scores
ggsave("figs/graph_emissions_scores.jpg", width=7, height=5, dpi=300)

```

Calculate Climate Change Resilience Score

eee_score
emissions_scores
clim_act
power_gen_scores

Tidy each piece so can combine
```{r}
##eee scores
eee_score_t <- eee_score %>% 
  mutate(state = str_to_title(gsub(",", " ", state)),
         rgn_id1 = case_when(state == "Maine" ~ 6,
                            state == "New Hampshire" ~ 9,
                            state == "Massachusetts" ~ 7,
                            state == "Rhode Island" ~ 11,
                            state == "Connecticut" ~ 5,
                            state == "New York" ~ 10),
         rgn_id8 = case_when(state == "Massachusetts" ~ 8)) %>% 
  rename(rgn_name = state) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id8, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(year, rgn_name, rgn_id, score) %>% 
  rename(score_eee = score) %>% 
  mutate(rgn_name = ifelse(rgn_name == "Massachusetts" & rgn_id == 7, "Massachusetts-North",
                           ifelse(rgn_name == "Massachusetts" & rgn_id == 8, "Massachusetts-South", rgn_name))) 

##emmissions scores
#create a seperate df to bind to the main one for year 2017 beacuse no data for that
em_2017<- tribble(
  ~year, ~rgn_name, ~rgn_id,
  2017, "Maine", 6,  
  2017, "New Hampshire", 9,  
  2017, "Massachusetts-North", 7, 
  2017, "Rhode Island", 11,  
  2017, "Connecticut", 5,  
  2017, "New York", 10,  
  2017, "Massachusetts-South", 8
  ) %>% 
  as.data.frame() %>% 
  add_column(score= NA)

emissions_scores_t <- emissions_scores %>% 
  mutate(rgn_id1 = case_when(state == "Maine" ~ 6,
                            state == "New Hampshire" ~ 9,
                            state == "Massachusetts" ~ 7,
                            state == "Rhode Island" ~ 11,
                            state == "Connecticut" ~ 5,
                            state == "New York" ~ 10),
         rgn_id8 = case_when(state == "Massachusetts" ~ 8)) %>% 
  rename(rgn_name = state) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id8, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(year, rgn_name, rgn_id, score) %>% 
  mutate(rgn_name = ifelse(rgn_id == 1, "Offshore", 
                       ifelse(rgn_id==2, "Georges Bank",
                               ifelse(rgn_id==3, "Gulf of Maine",
                                      ifelse(rgn_id==4, "Mid-Atlantic Bight",
                                             ifelse(rgn_id==5, "Connecticut",
                                                    ifelse(rgn_id==6, "Maine",
                                                           ifelse(rgn_id==7, "Massachusetts-North",
                                                                  ifelse(rgn_id==8, "Massachusetts-South",
                                                                         ifelse(rgn_id==9, "New Hampshire",
                                                                                ifelse(rgn_id==10, "New York",
                                                                                       "Rhode Island")))))))))))%>% 
  rbind(em_2017) %>% 
  arrange(rgn_name) %>% 
  fill(score, .direction = "down") %>% 
  rename(score_em = score) 

##power generation scores
power_gen_scores_t <- power_gen_scores %>% 
  mutate(rgn_id1 = case_when(
    state == "ME" ~ "6", 
    state == "NH" ~ "9", 
    state == "MA" ~ "7",
    state == "RI" ~ "11",
    state == "CT" ~ "5",
    state == "NY" ~ "10"),
    rgn_id2 = case_when(
    state == "MA" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(-filler, - state) %>% 
  mutate(rgn_name = case_when(
    rgn_id == "6" ~ "Maine", 
    rgn_id == "9" ~ "New Hampshire", 
    rgn_id == "7" ~ "Massachusetts-North",
    rgn_id == "8" ~ "Massachusetts-South",
    rgn_id == "11" ~ "Rhode Island",
    rgn_id == "5" ~ "Connecticut",
    rgn_id == "10" ~ "New York")) %>% 
  select(year, rgn_name, rgn_id, score) %>% 
  rename(score_power = score) 

##climate actions scores
clim_act_t <- clim_act %>% 
  select(year, State, score) %>% 
  rename(state=State) %>% 
  mutate(rgn_id1 = case_when(state == "Maine" ~ 6,
                            state == "New Hampshire" ~ 9,
                            state == "Massachusetts" ~ 7,
                            state == "Rhode Island" ~ 11,
                            state == "Connecticut" ~ 5,
                            state == "New York" ~ 10),
         rgn_id8 = case_when(state == "Massachusetts" ~ 8)) %>% 
  rename(rgn_name = state) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id8, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(year, rgn_name, rgn_id, score) %>% 
   mutate(rgn_name = ifelse(rgn_name == "Massachusetts" & rgn_id == 7, "Massachusetts-North",
                           ifelse(rgn_name == "Massachusetts" & rgn_id == 8, "Massachusetts-South", rgn_name))) %>% 
  rename(score_act = score) 
  
```

Combine all scores!
```{r}
cc_res_score <- left_join(emissions_scores_t, clim_act_t, by = c("year", "rgn_name", "rgn_id")) %>% 
  left_join(power_gen_scores_t, by = c("year", "rgn_name", "rgn_id")) %>% 
  left_join(eee_score_t, by = c("year", "rgn_name", "rgn_id"))%>% 
  rowwise() %>% 
  mutate(score = mean(c(score_power, score_act, score_em, score_eee))) %>% 
  select(year, rgn_name, rgn_id, score)

write.csv(cc_res_score, file = "data/cc_res_score")
```

Visualize
```{r}
cc_res_graph <- ggplot(cc_res_score)+
  geom_line(aes(x=year, y= score, color= rgn_name))+
 ggtitle("Climate Change Resilience Scores") +
  ylab("Score") +
  xlab("Year") +
  theme_classic()

cc_res_graph

##save fig
ggsave("figs/cc_res_graph.jpg", width=7, height=5, dpi=300)
```


finish this up for the rest of the states
compare with what was doing before


##Old code when using energy production to calculate if states are hitting their clean energy targets

Calculate total energy production in each state each year
```{r}  
gen_total <- raw_gen %>% 
  filter(year >= 2004 & year <=2017) %>% 
  filter(state == "ME" | state == "NH" | state == "MA" | state == "RI" | state == "CT" | state == "NY") %>% 
  filter(energy_source != "Total") %>%
  select(year, state, gen_mwh) %>% 
  group_by(year, state) %>% 
  dplyr::summarise(total_power = sum(gen_mwh)) %>% 
  ungroup()
```

Calculate total clean/renewable energy production in each state each year
```{r}  
gen_clean <- raw_gen %>% 
  filter(year >= 2004 & year <=2017) %>% 
  filter(state == "ME" | state == "NH" | state == "MA" | state == "RI" | state == "CT" | state == "NY") %>% 
  filter(energy_source == "Hydroelectric Conventional"| energy_source == "Wind" | energy_source == "Solar Thermal and Photovoltaic" | energy_source == "Pumped Storage") %>%
  select(year, state, gen_mwh) %>% 
  group_by(year, state) %>% 
  dplyr::summarise(total_power_clean = sum(gen_mwh))%>% 
  ungroup()
```

Combine data sets 
```{r}  
power_gen <- left_join(gen_total, gen_clean, by= c("year", "state")) %>% 
  mutate(per_clean = total_power_clean/total_power)
```

Visualize percent of clean energy production
```{r}  
graph_clean_pro <-ggplot(power_gen)+
  geom_line(aes(x=year, y = per_clean, color=state))+
  ggtitle("Percent of Clean/Renewable Energy Production") +
  ylab("Percent") +
  xlab("State") +
  theme_classic() 
graph_clean_pro

ggsave("figs/graph_clean_pro_nuclear.jpg", width=7, height=5, dpi=300)
```

Maine
```{r}

#MA 40% clean by 2017
me <- power_gen %>% 
  filter(state == "ME") %>% 
  mutate(year_n = seq(0:13))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
me_target <- tribble(
  ~year, ~per_clean,
  0, 0.180,
  13, 0.4
)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_me <- lm(per_clean ~ year, data = me_target)
#intercept = 0.1800 ; coef =0.0169 
summary(lm_me)

plot(per_clean ~ year, data = me_target)
abline(lm_me)

###trying to use equation no loop
me_clean_target <- me %>% 
  rowwise() %>% 
  mutate(target = (0.0169*year_n + 0.1800 ),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter( year != "2004") %>% 
  select(year, state, per_clean, target) 


```

Massachusets
```{r}

#MA 35% clean by 2030
ma <- power_gen %>% 
  filter(state == "MA")  %>% 
  mutate(year_n = seq(0:13),
         per_clean = if_else(per_clean <0, 0, per_clean)) 

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ma_target <- tribble(
  ~year, ~per_clean,
  0, 0.0105	,
  26, .35)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ma <- lm(per_clean ~ year, data = ma_target)
#coef = 0.0131  ; int = 0.0105

ma_clean_target <- ma %>% 
  rowwise() %>% 
  mutate(target = (0.0131 *year_n+0.0105 ),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter( year != "2004") %>% 
  select(year, state, per_clean, target)

```

New Hampshire
```{r}

#NH 25.2% clean by 2025
nh <- power_gen %>% 
  filter(state == "NH") %>% 
  mutate(year_n = seq(0:13))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
nh_target <- tribble(
  ~year, ~per_clean,
  0, 0.0551,
  21, .252
)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_nh <- lm(per_clean ~ year, data = nh_target)
#coef=  0.00938 int= 0.05510

nh_clean_target <- nh %>% 
  mutate(target = ( 0.00938*year_n +0.05510),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
   filter( year != "2004") %>% 
  select(year, state, per_clean, target)

```


Rhode Island
```{r}

#RI 38.5% clean by 2035
ri <- power_gen %>% 
  filter(state == "RI") %>% 
  mutate(year_n = seq(0:13))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ri_target <- tribble(
  ~year, ~per_clean,
  0, 0.001106	,
  31, .385)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ri <- lm(per_clean ~ year, data = ri_target)
#coef=0.01238  , int = 0.00111

ri_clean_target <- ri %>% 
   mutate(target = ( 0.01238*year_n+0.00111),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_clean, target)

```


New York
```{r}

#NY 50% clean by 2030
ny <- power_gen %>% 
  filter(state == "NY") %>% 
  mutate(year_n = seq(0:13))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ny_target <- tribble(
  ~year, ~per_clean,
  0, 	0.169	,
  26, .50)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ny <- lm(per_clean ~ year, data = ny_target)
#coef=0.0127   , int =  0.1690 

ny_clean_target <- ny %>% 
  mutate(target = (0.0127*year_n +0.1690 ),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_clean, target)

```


Connecticut
```{r}

#CT 44% clean by 2030
ct <- power_gen %>% 
  filter(state == "CT")  %>% 
  mutate(year_n = seq(0:13))

#create tibble with the start (value in 2005) and end (goal in whichever year) to do a linear regression 
ct_target <- tribble(
  ~year, ~per_clean,
  0, 0.01441,
  26, .44
)

#super simple regression for equation to calculate the necessary percent clean emmisions each year to reach target
lm_ct <- lm(per_clean ~ year, data = ct_target)
#coef= 0.0164, int=0.0144

ct_clean_target <- ct %>% 
  mutate(target = (0.0164*year_n +0.0144),
         per_clean = round(per_clean, digits= 2),
         target = round(target, digits= 2)) %>% 
  filter(year != "2004") %>% 
  select(year, state, per_clean, target)

```



Combine all states df with target
```{r}
energy_target <- rbind(me_clean_target, nh_clean_target, ma_clean_target, ri_clean_target, ct_clean_target, ny_clean_target)

```


Calculate scores
```{r}
power_gen_scores <- energy_target %>% 
  mutate(score = per_clean/target,
         score = if_else(score >1, 1, score)) 
#save
write.csv(power_gen_scores, file = "data/power_gen_scores_nuc")

```

Visualize scores
```{r}  

graph_clean_energy_score_nuc <-ggplot(power_gen_scores)+
  geom_line(aes(x=year, y = score, color=state))+
  ggtitle("Clean Energy Score") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 
graph_clean_energy_score_nuc
ggsave("figs/graph_clean_energy_score.jpg", width=7, height=5, dpi=300)

```







