---
title: "OHIEC: Resilience - Climate Change Pressure prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates fishing pressure resilience values by OHINE region. 

# Data

**Reference**: 

**Downloaded**: 

**Description**:  

**Time range**: 

**Format**: 

Regulations:
1) extract info from America's Pledge Report
2) American Council for Energy Efficiency:
Rankings: https://database.aceee.org/state/maine

Effectiveness:
US. energy information administration:
1) Total CO2 emissions https://www.eia.gov/environment/emissions/state/analysis/
2) Total power generation:https://www.eia.gov/electricity/data/state/



``` {r setup,  message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(Hmisc)
library(stringi)
library(readxl)

```


Data
```{r}

eee <- read_excel(file.path(dir_anx, "_raw_data/EEE/eee_scores.xlsx"))
raw_gen <- read_excel(file.path(dir_anx, "_raw_data/EIA/eia_annual_generation_state.xls"))
raw_em <- read_excel(file.path(dir_anx, "_raw_data/EIA/eia_co2_emmisions.xlsx"))

```


Energy Efficiency Scores
```{r}

eee_score <- eee %>% 
  mutate(score = eee_score/eee_target)

eee_score <-ggplot(eee_score)+
  geom_point(aes(x=state, y = score), color="green4", size=2)+
  ggtitle("Energy Efficiency Score") +
  ylab("Score") +
  xlab("State") +
  theme_classic() 
eee_score   


```


**Effectivness**
Power Generation
```{r}

gen_total <- raw_gen %>% 
  filter(year >= 2005 & year <=2017) %>% 
  filter(state == "ME" | state == "NH" | state == "MA" | state == "RI" | state == "CT" | state == "NY") %>% 
  filter(energy_source != "Total") %>%
  select(year, state, gen_mwh) %>% 
  group_by(year, state) %>% 
  dplyr::summarise(total_power = sum(gen_mwh))

gen_clean <- raw_gen %>% 
  filter(year >= 2005 & year <=2017) %>% 
  filter(state == "ME" | state == "NH" | state == "MA" | state == "RI" | state == "CT" | state == "NY") %>% 
  filter(energy_source == "Hydroelectric Conventional"| energy_source == "Wind" | energy_source == "Solar Thermal and Photovoltaic" | energy_source == "Pumped Storage") %>%
  select(year, state, gen_mwh) %>% 
  group_by(year, state) %>% 
  dplyr::summarise(total_power = sum(gen_mwh)) %>% 
  rename(total_power_clean = total_power )

power_gen <- left_join(gen_total, gen_clean, by= c("year", "state")) %>% 
  mutate(per_clean = total_power_clean/total_power)

graph <-ggplot(power_gen)+
  geom_line(aes(x=year, y = per_clean, color=state))
graph


power_gen_scores <- left_join(gen_total, gen_clean, by= c("year", "state")) %>% 
  mutate(per_clean = total_power_clean/total_power,
         target = case_when(state == "ME" ~ 0.4, #targets for clean energy from Beyond Carbon
                            state == "NH" ~ 0.252,
                            state == "CT" ~ 0.44,
                            state == "MA" ~ 0.35,
                            state == "NY" ~ 0.5,
                            state == "RI" ~ 0.385),
         score = per_clean/target,
         score = if_else(score >1, 1, score)) 
  
graph_score <-ggplot(power_gen_scores)+
  geom_line(aes(x=year, y = score, color=state))+
  ggtitle("Clean Energy Score (2005-2017)") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 
graph_score         


```

Lots of Pumped storage production is negative. What to do about that? Is it when the water gets pumped back up? Seems like the negative values outweigh the positive


Types of energy source:
Coal
Natural Gas
Nuclear
Other
Other Biomass
Other Gases
Petroleum
Wood and Wood Derived Fuels
[below I considered clean energy; need to look into the pumped storage because most of it is negative]
Pumped Storage
Hydroelectric Conventional
Wind
Solar Thermal and Photovoltaic

State Clean Energy Targets (source = Beyond Carbon): 
Maine: 40% renewable/clean by 2017
New Hampshire: 25.5% renewable/clean by 2025
Connecticut: 44% renewable/clean by 2030
Massachusetts: 35% renewable/clean by 2030
New York: 50% renewable/clean by 2030
Rhode Island: 38.5% renewable/clean by 2035


Carbon Emmissions
```{r}

em <- raw_em %>% 
  select(-Percent, -Absolute) %>% 
  gather(year, emission, 2:13) %>% 
  filter(State == "Maine" | State == "New Hampshire" | State == "Mssachusetts" | State == "New Hampshire" | State == "Rhode Island" | State == "Connecticut" | State == "New York") %>% 
  mutate(year = as.numeric(year))

```


```{r}

graph_em <-ggplot(em)+
  geom_line(aes(x=year, y = emission, color=State))+
  ggtitle("Total Carbon Emissions (2005-2017)") +
  ylab("Emissions (million metric tons") +
  xlab("Year") +
  theme_classic() 

graph_em  

```

Carbon Emmission Targets
Maine: 80% reduction below 2003 levels
New Hampshire: 80% reduction below 1990 levels by 2050
Connecticut: 80% reduction below 2001 levels by 2050
Massachusetts: 80% reduction below 1990 levels by 2050
New York: 80% reduction below 1990 levels by 2050
Rhode Island: 80% reduction below 1990 levels by 2050

Determine State Targets
look up what each states emisiosn were in those years










