---
title: 'OHI-Northeast: Coastal Hardening Data Prep Layer'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---
# Data  

**Reference**: NOAA [Environmental Response Management Application](https://erma.noaa.gov/atlantic/erma.html#/layers=1+13763+16973+35410+491&x=-78.22196&y=38.23326&z=6&panel=layer)

**Downloaded**: October, 2019

**Description**:  Spatial line data identifying type of habitat and human made armoring. 

**Native data resolution**: Line data

**Time range**: NA. This data set has been updated in 2016, and the historical data is avaliable, but it's incomplete.

**Format**:  Geodatabase/shapefile/line file

This script uses the same methods for calcualting coastal hardening as https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/150065
However we redid the analysis because of new new avaliability and different specific regions of interest.

***

# Methods
## Setup
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(tidyverse)
library(mapview)
library(sf)
```

## Grabbing spatial data

We will need an buffer because some rivers are harded very far from the coastline. We will use the state waters shapefile and then add a 1 km buffer around it to grb the coastline. We are creating a buffer of 1 km inland.
```{r create_1km_buffer}
coastal_buffer <- read_sf("~/github/ne-prep/spatial/shapefiles/state_waters.shp")  %>% 
  st_transform(us_alb) %>% 
  st_buffer(1000) %>%  #data in meters
  st_union() %>% 
  st_as_sf() %>% 
  mutate(coastal = "coastal")

mapview(coastal_buffer)
```

Read in the shapefile we will use to intersect our data to add region names
While this buffer is not perfect (some overlap along state lines), this buffer does capture the majority of the coastline now.
```{r state_water_shape}
# combine state and state water shapefiles
state <-  st_read(dsn=file.path(dir_git, "spatial/shapefiles/states.shp")) %>% 
  st_transform(us_alb) %>% 
  select(NAME) %>% 
  rename(rgn_name = NAME) %>% 
  st_buffer(200)

statewaters <- st_read(dsn=file.path(dir_git, "spatial/shapefiles/state_waters.shp"))%>% 
  st_transform(us_alb) %>% 
  select(NAME10) %>% 
  rename(rgn_name = NAME10)

state_and_statewaters <- rbind(state, statewaters) %>% 
  mutate(rgn_id = case_when(rgn_name == "Massachusetts" ~ 7,
                            rgn_name == "New Hampshire" ~ 9,
                            rgn_name == "Connecticut" ~ 5,
                            rgn_name == "Maine" ~ 6,
                            rgn_name == "Rhode Island" ~ 10,
                            rgn_name == "New York" ~ 11)) %>% 
  group_by(rgn_id) %>% 
  summarise(hacky = sum(rgn_id)) %>% 
  mutate(rgn_name = case_when(rgn_id == 7 ~"Massachusetts" ,
                            rgn_id == 9 ~ "New Hampshire" ,
                            rgn_id == 5 ~ "Connecticut",
                            rgn_id == 6 ~ "Maine",
                            rgn_id == 10 ~ "Rhode Island",
                            rgn_id == 11 ~ "New York")) %>% 
  filter(rgn_name != is.na(rgn_name)) %>% 
  st_intersection(coastal_buffer) %>% 
  select(-hacky, -coastal)

mapview(state_and_statewaters)
```

Create shapefile for NE
```{r northeast_erma}
erma <- st_read(dsn=file.path(dir_anx, "_raw_data/ERMA/layer_35355/35776/National_ESI_Shoreline_Atlantic_20170608.shp")) %>%
  st_transform(us_alb) %>% 
  filter(ATLAS_NAME %in% c("NY/NJ METRO, HUDSON, SOUTH LONG ISLAND", "LONG ISLAND SOUND", "MASSACHUSETTS/RHODE ISLAND", "MAINE/NEW HAMPSHIRE")) %>%
  st_intersection(state_and_statewaters) 

#save it since it takes so long to run
write_sf(erma, file.path(dir_anx, "prs_hab/coastal_hardening/shapefiles/erma.shp"))
#mapview(erma) ## very big! only do if you need to redo it
```

Double check Maine. The paper that did a similar analysis stated that no hardened coastline exists only as polygons not lines, however it looks like it's been updated since then.
```{r check_maine}
maine <- st_read(dsn=file.path(dir_anx, "_raw_data/ERMA/layer_35355/35776/National_ESI_Shoreline_Atlantic_20170608.shp")) %>%
  st_transform(us_alb) %>% 
  filter(ATLAS_NAME %in% c("MAINE/NEW HAMPSHIRE")) 
mapview(maine)
```

## Calculate percent armored 
Here we are going to do the calculations without MA because more work is required to separate the state regions.
```{r calculate_percent_hard}
coast_arm_subrgns <-  read_sf(file.path(dir_anx, "prs_hab/coastal_hardening/shapefiles/erma.shp")) %>% 
  select(GENERALIZE, rgn_name, rgn_id, geometry) %>% 
  filter(rgn_name != "Massachusetts") %>% 
  mutate(armored = ifelse(str_detect(GENERALIZE, "Armored", negate= TRUE), "not armored", "armored"),
         armored_km = ifelse(armored == "armored", st_length(geometry)/1000, NA),
         coastline_km = st_length(geometry)/1000) %>% # the distances are in meters
  group_by(rgn_name) %>% 
  summarise(hardened_coastline = sum(armored_km, na.rm = TRUE), ## here it looks like its in meter because [m], but we converted it to km. we can read in the data orginially and tell it to read it as km, but it takes a very long time to redo.
            total_coastline = sum(coastline_km, na.rm = TRUE)) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(percent_hard = hardened_coastline/total_coastline)

```

Calculate percent armored for MA regions
We need to create a buffer for N/S MA beach land regions
```{r ma_armored}
## creat buffer
ma_buffer <- read_sf("~/github/ne-prep/spatial/shapefiles/ne_ohi_rgns.shp")  %>% 
  st_transform(us_alb) %>% 
  filter(rgn_id %in% c(7,8)) %>% 
  st_buffer(804.672)

ma <-  read_sf(file.path(dir_anx, "prs_hab/coastal_hardening/shapefiles/erma.shp")) %>% 
  select(GENERALIZE, rgn_name, rgn_id, geometry) %>%
  rename(state = rgn_name) %>% 
  filter(state == "Massachusetts") %>% 
  st_intersection(ma_buffer) %>% 
   mutate(armored = ifelse(str_detect(GENERALIZE, "Armored", negate= TRUE), "not armored", "armored"),
         armored_km = ifelse(armored == "armored", st_length(geometry)/1000, NA),
         coastline_km = st_length(geometry)/1000) %>% # the distances are in meters
  group_by(rgn_name) %>% 
  summarise(hardened_coastline = sum(armored_km, na.rm = TRUE),
            total_coastline = sum(coastline_km, na.rm = TRUE)) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(percent_hard = hardened_coastline/total_coastline)
```

Calculate for the whole region. We need to do this separatly because we are looking at the Northeast region as a whole, not calculating each subregion separately.
```{r calculate_percent_hard}
coast_arm_rgn12 <-  read_sf(file.path(dir_anx, "prs_hab/coastal_hardening/shapefiles/erma.shp")) %>% 
  select(GENERALIZE, rgn_name, rgn_id, geometry) %>% 
  mutate(armored = ifelse(str_detect(GENERALIZE, "Armored", negate= TRUE), "not armored", "armored"),
         armored_km = ifelse(armored == "armored", st_length(geometry)/1000, NA),
         coastline_km = st_length(geometry)/1000,
         rgn_name = "Northeast") %>% # the distances are in meters
  group_by(rgn_name) %>% 
  summarise(hardened_coastline = sum(armored_km, na.rm = TRUE), ## here it looks like its in meter because [m], but we converted it to km. we can read in the data orginially and tell it to read it as km, but it takes a very long time to redo.
            total_coastline = sum(coastline_km, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(percent_hard = hardened_coastline/total_coastline)
```

Combine subregions (no MA), MA, and Northeast calculations
```{r combine_all_regions}
coast_arm <- coast_arm_subrgns %>% 
  dplyr::bind_rows(ma) %>% 
  dplyr::bind_rows(coast_arm_rgn12) %>% 
  mutate(rgn_id = case_when(rgn_name == "Connecticut" ~ 5,
                            rgn_name == "Maine" ~ 6,
                            rgn_name == "New Hampshire" ~ 9,
                            rgn_name == "New York" ~ 10,
                            rgn_name == "Rhode Island" ~ 11,
                            rgn_name == "Massachusetts-North" ~ 7,
                            rgn_name == "Massachusetts-South" ~ 8,
                            rgn_name == "Northeast" ~ 12)) %>% 
  select(rgn_name, rgn_id, percent_hard)
```

Map with binary armored or not to double check that it looks good and nothing crazy is happening
```{r map_armed_or_not}
arm_map <-  read_sf(file.path(dir_anx, "prs_hab/coastal_hardening/shapefiles/erma.shp")) %>% 
  select(GENERALIZE, rgn_name, rgn_id,  geometry) %>% 
  mutate(armored = ifelse(str_detect(GENERALIZE, "Armored", negate= TRUE), "not armored", "armored")) %>% 
  select(rgn_name, armored)

#map <- mapview(arm_map, zcol = "armored")
#mapview::mapshot(map, url = "armored_map.html")
```


Visualize the data
```{r viz_data}
arm_graph <- ggplot(coast_arm) +
  geom_bar(aes(x= rgn_name, y = percent_hard), stat = "identity")+
  coord_flip() + 
  theme_bw()+ 
  ggtitle("Percent of armored coastline (within 1 km of shore)")

arm_graph
ggsave("figs/arm_graph.jpg", width=7, height=5, dpi=300)
```

# Calculate scores
Creating a df to add on to the climate change scores to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name, ~score,
  3, "Gulf of Maine", NA,
  2, "Georges Bank", NA,
  4, "Mid-Atlantic Bight", NA,
  1, "Offshore", NA) 
```

Tidy df for scores
```{r create_scores_layer}
prs_hardening <- coast_arm %>% 
  mutate(year = 2017)
```


Save to scores
```{r write_to_scores}
write.csv(prs_hardening, file.path(dir_calc, "layers/prs_hardening.csv"))
```








