---
title: 'OHI-Northeast: Coastal Hardening Data Prep Layer'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(tidyverse)
library(mapview)
library(sf)
```

#Data Source 

**Reference**: NOAA [Environmental Response Management Application](https://erma.noaa.gov/atlantic/erma.html#/layers=1+13763+16973+35410+491&x=-78.22196&y=38.23326&z=6&panel=layer)

**Downloaded**: October, 2019

**Description**:  Spatial line data identifying type of habitat and human made armoring. 

**Native data resolution**: Line data

**Time range**: NA. This data set has been updated in 2016, and the historical data is avaliable, but it's incomplete.

**Format**:  Geodatabase/shapefile/line file

***


# Load Data

This script uses the same methods for calcualting coastal hardening as https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/150065

Downloaded ERMA data instead of the ESI data. Will look into this now to see if it's better than the ESI data

We will need an buffer because some rivers are harded very far from the coastline. We will use the state waters shapefile and then add a 1 km buffer around it to grb the coastline. We are creating a buffer of 1 km inland
```{r create_1km_buffer}
coastal_buffer <- read_sf("~/github/ne-prep/spatial/shapefiles/state_waters.shp")  %>% 
  st_transform(us_alb) %>% 
  st_buffer(1000) %>%  #data in meters
  st_union() %>% 
  st_as_sf() %>% 
  mutate(coastal = "coastal")

mapview(coastal_buffer)
```

Read in the shapefile we will use to intersect our data to add region names
This buffer is not perfect. There is some overlap along rivers state lines and is still missing a few areas along the coast. Howvever, this buffer does capture the majority of the coastline now
```{r state_water_shape}
# combine state and state water shapefiles
state <-  st_read(dsn=file.path(dir_git, "spatial/shapefiles/states.shp")) %>% 
  st_transform(us_alb) %>% 
  select(NAME) %>% 
  rename(rgn_name = NAME) %>% 
  st_buffer(200)

statewaters <- st_read(dsn=file.path(dir_git, "spatial/shapefiles/state_waters.shp"))%>% 
  st_transform(us_alb) %>% 
  select(NAME10) %>% 
  rename(rgn_name = NAME10)

state_and_statewaters <- rbind(state, statewaters) %>% 
  mutate(rgn_id = case_when(rgn_name == "Massachusetts" ~ 7,
                            rgn_name == "New Hampshire" ~ 9,
                            rgn_name == "Connecticut" ~ 5,
                            rgn_name == "Maine" ~ 6,
                            rgn_name == "Rhode Island" ~ 10,
                            rgn_name == "New York" ~ 11)) %>% 
  group_by(rgn_id) %>% 
  summarise(hacky = sum(rgn_id)) %>% 
  mutate(rgn_name = case_when(rgn_id == 7 ~"Massachusetts" ,
                            rgn_id == 9 ~ "New Hampshire" ,
                            rgn_id == 5 ~ "Connecticut",
                            rgn_id == 6 ~ "Maine",
                            rgn_id == 10 ~ "Rhode Island",
                            rgn_id == 11 ~ "New York")) %>% 
  filter(rgn_name != is.na(rgn_name)) %>% 
  st_intersection(coastal_buffer) %>% 
  select(-hacky, -coastal)

mapview(state_and_statewaters)
```


```{r northeast_erma}
erma <- st_read(dsn=file.path(dir_anx, "_raw_data/ERMA/layer_35355/35776/National_ESI_Shoreline_Atlantic_20170608.shp")) %>%
  st_transform(us_alb) %>% 
  filter(ATLAS_NAME %in% c("NY/NJ METRO, HUDSON, SOUTH LONG ISLAND", "LONG ISLAND SOUND", "MASSACHUSETTS/RHODE ISLAND", "MAINE/NEW HAMPSHIRE")) %>%
  st_intersection(state_and_statewaters) 

#save it since it takes so long to run
write_sf(erma, file.path(dir_anx, "prs_hab/coastal_hardening/shapefiles/erma.shp"))
#mapview(erma) ## very big! only do if you need to redo it
```

Double check Maine. The paper that did a similar analysis stated that no hardened coastline exists only as polygons not lines, however it looks like it's been updated since then.
```{r check_maine}
maine <- st_read(dsn=file.path(dir_anx, "_raw_data/ERMA/layer_35355/35776/National_ESI_Shoreline_Atlantic_20170608.shp")) %>%
  st_transform(us_alb) %>% 
  filter(ATLAS_NAME %in% c("MAINE/NEW HAMPSHIRE")) 
mapview(maine)
```


Calculate percent armored 
```{r calculate_percent_hard}
coast_arm <-  read_sf("~/github/ne-prep/prep/pressures/habitat/shapefiles/erma.shp") %>% 
  select(GENERALIZE, rgn_name, rgn_id, geometry) %>% 
  mutate(armored = ifelse(str_detect(GENERALIZE, "Armored", negate= TRUE), "not armored", "armored"),
         armored_km = ifelse(armored == "armored", st_length(geometry)/1000, NA),
         coastline_km = st_length(geometry)/1000) %>% # the distances are in meters
  group_by(rgn_name) %>% 
  summarize(hardened_coastline = sum(armored_km, na.rm = TRUE),
            total_coastline = sum(coastline_km, na.rm = TRUE)) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(percent_hard = hardened_coastline/total_coastline)
 # filter(rgn_name != "Mid-Atlantic Bight" & rgn_name !="Gulf of Maine") # the buffer got into the ends of hte state waters into "offshore"

coast_arm
```

Visualize the data

```{r viz_data}
arm_graph <- ggplot(coast_arm) +
  geom_bar(aes(x= rgn_name, y = percent_hard), stat = "identity")+
  coord_flip() + 
  theme_bw()+ 
  ggtitle("Percent of armored coastline (within 1 km of shore)")

arm_graph
```


```{r map_armed_or_not}
arm_map <-  read_sf("~/github/ne-prep/prep/pressures/habitat/shapefiles/erma.shp") %>% 
  select(GENERALIZE, rgn_name, rgn_id,  geometry) %>% 
  mutate(armored = ifelse(str_detect(GENERALIZE, "Armored", negate= TRUE), "not armored", "armored")) %>% 
  select(rgn_name, armored)

map <- mapview(arm_map, zcol = "armored")
#mapview::mapshot(map, url = "armored_map.html")
```

Tidy df for scores
```{r}

```


Save to scores
```{r write_to_scores}
write.csv(res_species_pollution, file.path(dir_calc, "layers/res_alien.csv"))
```








