---
title: 'OHI-Northeast: Trampling Pressure Data Prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary
 
This code calculates the effects of beach trampling using the average population density along the coast as a proxy.

# Data

GeoTiff files were downloaded from https://sedac.ciesin.columbia.edu/data/set/usgrid-summary-file1-2010/metadata
Octover 15th 2019

The data is at 1km2 resolution. Since we are only looking at total population that means #people/km2

citation:
Center for International Earth Science Information Network - CIESIN - Columbia University. 2017. U.S. Census Grids (Summary File 1), 2010. Palisades, NY: NASA Socioeconomic Data and Applications Center (SEDAC). https://doi.org/10.7927/H40Z716C. Accessed DAY MONTH YEAR. 

# Methods

## Setup

``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 10, fig.height = 8, fig.path = 'figs/', message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHI-NE specific version of common.R

library(tidyverse)
library(ggplot2)
library(raster)
library(rgdal)
library(gdalUtils)
library(fasterize)

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')
```


Read in the population tif files for each state, and then create one NE raster

MOVED HTE NORTHEAST TIFF. REWRITE TO SAVE IN THE RIGHT PLACE
```{r read_in_data}
## read in all states rasters
northeast_rasters <- c(file.path(dir_anx, "_raw_data/USGS/usgrid_data_2010/geotiff/nypop10.tif"),
                    file.path(dir_anx, "_raw_data/USGS/usgrid_data_2010/geotiff/mepop10.tif"),
                    file.path(dir_anx, "_raw_data/USGS/usgrid_data_2010/geotiff/nhpop10.tif"),
                    file.path(dir_anx, "_raw_data/USGS/usgrid_data_2010/geotiff/ctpop10.tif"),
                    file.path(dir_anx, "_raw_data/USGS/usgrid_data_2010/geotiff/ripop10.tif"),
                    file.path(dir_anx, "_raw_data/USGS/usgrid_data_2010/geotiff/mapop10.tif"))

## combine all tif files into one 
mosaic_rasters(gdalfile = northeast_rasters, dst_dataset = "northeast.tif", of = "GTiff")

## create one ne raster
northeast_rast <- raster::raster("northeast.tif") %>%
  projectRaster(crs = us_alb, progress="text") 
#%>% crop(ne_ext)

## check it out!
plot(northeast_rast)

ne_vals <-northeast_rast%>% 
  getValues() %>%
  na.omit() 

hist(ne_vals)
```


We are creating a coastal buffer that includes our coastline and 25 miles inland. I choose 25 miles because that is what Casey did in BC, but we can discuss if this makes sense.
```{r create_buffer}
## create a buffer 1 km along all state waters to get the beaches
buffer_1km_coastal <-  read_sf("~/github/ne-prep/spatial/shapefiles/state_waters.shp")  %>% 
  st_transform(us_alb) %>% 
  st_buffer(40233.6) %>%  ## data in meters; using a 25 miles buffer
  st_union() %>% 
  st_as_sf() 

plot(buffer_1km_coastal)

## create blank raster for rasterizing
r <- raster(crs = us_alb, ne_ext, res = 1000)
r[r] <- 1

## convert to a raster
buffer_rast <- fasterize(buffer_1km_coastal, r) %>% 
  raster::resample(northeast_rast, method = "ngb") ## match the resolution and extent

plot(buffer_rast)
```

Now we want to keep only the population density data that falls into our 25 miles within the coastline raster
```{r crop_area}
## keep only the areas 1km along the shorelines
ne_pop_coast <- northeast_rast %>% 
 raster::mask(buffer_rast) %>% 
 calc(fun = function(x){ifelse(x < 0, 0, x)})  ## there are 3 negative numbers in this data for some reason (-77.9399, -0.2386,	-0.0178)

plot(ne_pop_coast)
summary(ne_pop_coast) # The max value is very large, this is because of NYC
```


## Region score
Create a raster with our states rgn_ids
figure out how to separate out the MA county thats two regions

ma_counties <- read_csv(file.path(dir_git, "src/tables/MA_counties.csv")) %>% 
  select(-X1) %>% 
  rename(NAME_2 = County)

coastal_counties <- st_read(dsn=file.path(dir_git, "spatial/shapefiles/ne_counties.shp")) %>%
  st_transform(us_alb) %>% 
  left_join(ma_counties, by = c("NAME_2")) %>% 
  rowwise() %>% 
  mutate(rgn_name = character(rgn_name),
         rgn_name = ifelse(is.na(rgn_name), NAME_1, rgn_name))


```{r zones_raster}
## we want a raster with the region names so that we can ID the regions. This is the only file I found wiht the state land, and not just water, and it doesn't have have N/S MA. ask Jamie
ne_states_ne <- ne_states %>% 
  filter(NAME %in% c("Maine", "New York", "Rhode Island", "Connecticut", "New Hampshire", "Massachusetts")) %>%
  mutate(rgn_id1 = case_when( NAME == "Maine" ~ "6", 
                              NAME == "New Hampshire" ~ "9", 
                              NAME == "Massachusetts" ~ "7",
                              NAME == "Rhode Island" ~ "11",
                              NAME == "Connecticut" ~ "5",
                              NAME == "New York" ~ "10"),
         rgn_id2 = case_when(
           State == "Massachusetts" ~ "8")) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "zone") %>% 
  filter(zone != "0") %>% 
  select(-filler, - State) %>% 
  mutate(rgn_name = case_when(
    zone == "6" ~ "Maine", 
    zone == "9" ~ "New Hampshire", 
    zone == "7" ~ "Massachusetts-North",
    zone == "8" ~ "Massachusetts-South",
    zone == "11" ~ "Rhode Island",
    zone == "5" ~ "Connecticut",
    zone == "10" ~ "New York"))
                                            
                                            
                                            NAME == "New Hampshire" ~ 9)))))
  
  mutate(zone = case_when(NAME == "Maine" ~ 6,
                          NAME == "New York" ~ 10,
                          NAME == "Rhode Island" ~ 11,
                          NAME == "Connecticut" ~ 5,
                          NAME == "New Hampshire" ~ 9,
                          NAME == "Massachusetts" ~7)) # will need to apply the same score for MA 8 later

ne_states_rast <- fasterize(ne_states_ne, r, field = "zone") %>% 
  crop(ne_ext)%>% 
  raster::resample(northeast_rast, method = "ngb") %>% ## match the resolution
  mask(buffer_rast)

plot(ne_states_rast)

plot(states)
plot(zones)

```

Extract data for each region
```{r extract_data}
# extract data for each region:
pop_stats <- zonal(ne_pop_coast, ne_states_rast, fun="mean", na.rm=TRUE, progress="text")

pop_stats_2 <- data.frame(pop_stats) %>% 
  rename(rgn_id = zone,
         avg_ppl_km2 = mean)
```

Calculate reference point
We are using the reference point of 99.99% quantile. While this number is still high (because of NYC), it is an actually population density where people have survived
```{r calc_ref_point}
vals <- ne_pop_coast %>% 
  getValues() %>%
  na.omit() 

hist(vals) ## very skewed

max <- max(vals) #38719
ref_1 <- quantile(vals, 0.9999) #26188
ref_2 <- quantile(vals, 0.999) # 16199
ref_3 <- quantile(vals, 0.99) # 3582
ref_4 <- quantile(vals, 0.5) #28.3  
```

Calculate presssure scores
```{r calc_scores}
prs_trampling <-pop_stats_2 %>% 
  mutate(ref = ref_1,
         score_av = avg_ppl_km2/ref)
prs_trampling
```









