---
title: 'Coastal Access data layer'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

#libraries
library(tidyverse)
library(sf)
```

---

#Data Source 

**Reference**: NOAA [Environmental Sensitivity Index](https://response.restoration.noaa.gov/maps-and-spatial-data/download-esi-maps-and-gis-data.html)

**Downloaded**: August, 2019

**Description**:  Spatial point data identifying biological and human use resources. 

**Native data resolution**: Point data

**Time range**: NA. There are datasets from different time periods between 2001 and 2016 but we combine all datasets as a single representative layer, therefore the time does not play a role in the final coastal access layer.

**Format**:  Geodatabase/shapefile/point file

***

# Load Data

Load the 1 mile inland buffer

```{r}
buffer <- read_sf("~/github/ne-prep/spatial/shapefiles/ohine_inland_1mile.shp")
```

# Calculate area without access

Calculate the difference between area covered by buffered points and the whole 1 mile inland buffer.
```{r}

calc_coastal_coverage <- function(state, state_abv){

  land_buffer <- buffer %>%
    filter(rgn_name == state)
  
  #get total coastal area of the 1 mile buffer
  coastal_area <- st_area(land_buffer)
  
  point_buffer <- read_sf(paste0("state_coastal_access/data/coastal_access_shapefiles/", state_abv, "_coast_access.shp")) %>%
                  st_buffer(dist = 1609) %>%  #apply a 1 mile buffer
                  st_union()
  
  #get area not covered by buffered points
  diff <- st_difference(land_buffer, point_buffer) %>%
    st_area()
  
  #calculate proportion without coverage
  wo_access <- diff/coastal_area
  
  out_df <- data.frame(state = state,
                       state_abv = toupper(state_abv),
                       coastal_area = coastal_area,
                       not_covered = diff,
                       prop_no_access = wo_access)
  
  return(out_df)
  
}

df <- data.frame(state = c("New Hampshire", "New York", "Connecticut", "Massachusetts-North", "Massachusetts-South", "Rhode Island", "Maine"),
                 state_abv = c("nh", "ny", "cy", "ma", "ma", "ri", "me"))

#i know I could use map for this but i can't figure it out right now
nh_area <- calc_coastal_coverage("New Hampshire", "nh")
ny_area <- calc_coastal_coverage("New York", "ny")
ct_area <- calc_coastal_coverage("Connecticut", "ct")
man_area <- calc_coastal_coverage("Massachusetts-North", "ma")
mas_area <- calc_coastal_coverage("Massachusetts-South", "ma")
ri_area <- calc_coastal_coverage("Rhode Island", "ri")
me_area <- calc_coastal_coverage("Maine", "me")

state_areas <- bind_rows(nh_area,ny_area,ct_area,man_area,mas_area,ri_area,me_area)

#calculate total coast without access for NE
ne_area <- sum(state_areas$coastal_area)
ne_not_covered <- sum(state_areas$not_covered)  
ne_areas <- data.frame(state = "NA",
                       state_abv = "NA",
                       coastal_area = sum(state_areas$coastal_area),
                       not_covered = sum(state_areas$not_covered)) %>%
  mutate(prop_no_access = not_covered/coastal_area)

out <- state_areas %>%
  bind_rows(ne_areas) %>%
  mutate(score = round(1-prop_no_access, digits = 2)) %>%
  left_join(rgn_data, by = c("state" = "rgn_name")) %>%
  mutate(rgn_id = ifelse(is.na(rgn_id), 12, rgn_id),
         state = ifelse(rgn_id == 12, "Northeast", state)) %>%
  select(rgn_id, rgn_name = state, score)
```

# Results

```{r}
ggplot(out, aes(x = rgn_name, y = score)) +
  geom_col() +
  theme_bw() +
  coord_flip()
```


# Save to toolbox

Need to add years and all offshore regions

```{r}

out %>%
  complete(rgn_id = 1:12) %>%
  mutate(year = 2017) %>%
  select(rgn_id, rgn_name, score, year) %>%
  write_csv("~/github/ne-scores/region/layers/tr_rao_coastal_access.csv")
```






