---
title: 'OHI-Northeast: Rhode Island Coastal Access'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

#libraries
library(tidyverse)
```

  
#Methods

## Load Data

Inland buffer

We are going to use the 1 mile inland buffer as well as the rgns to identify coastal points. 
```{r}
inland_buffer <- read_sf("~/github/ne-prep/spatial/shapefiles/ohine_inland_1km.shp") %>%
  filter(rgn_name == "Rhode Island") %>%
  select(rgn_name, rgn_id)
```

Combine inland buffer with connecticut state waters

```{r}
ri_shape <- rgns %>%
  filter(state_name == "Rhode Island") %>%
  select(rgn_name, rgn_id) %>%
  rbind(inland_buffer)
```


### Rhode Island 2016

**Socioeconomic Point Layer**

CODES
HWR - HISTORICAL WRECK
HS - HISTORICAL SITE
B - BEACH
A - AIRPORT
AS - ARCHAEOLOGICAL SITE
F3 - FACILITY


```{r}
ri_2016_soc <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/Mass_RhodeIsland_2016_GDB/MA_RI_ESI_2016.gdb"), layer = "SOCECON_POINT") %>%
  st_transform(us_alb) %>%
  st_intersection(ri_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE %in% c("HS", "B")) %>%
  select(TYPE, rgn_name, rgn_id) %>%
  mutate(year = 2016)
```

**Navigational Marine Point layer**

CODES
A2 - ACCESS POINT
BR - BOAT RAMP
LD - LOCK AND DAM
M - MARINA
F - FERRY
PT - PORT

```{r}
ri_2016_nav <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/Mass_RhodeIsland_2016_GDB/MA_RI_ESI_2016.gdb"), layer = "NAV_MARINE_POINT") %>%
  st_transform(us_alb) %>%
  st_intersection(ri_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE != "LD") %>%
  select(TYPE, rgn_name, rgn_id) %>%
  mutate(year = 2016)
```

**Managed Point layer**

```{r}
ri_2016_mgd <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/Mass_RhodeIsland_2016_GDB/MA_RI_ESI_2016.gdb"), layer = "MANAGED_POINT") %>%
  st_transform(us_alb) %>%
  st_intersection(ri_shape) 
```

There is just a single MI point in the MANAGED_POINT layer which I believe to be MILITARY...

Combine the layers

```{r}
ri_2016 <- ri_2016_nav %>%
  rbind(ri_2016_soc)
```

---

### Rhode Island 2001 points


**Socioeconomic Points**


Metadata: https://response.restoration.noaa.gov/sites/default/files/esimaps/gisdata/Mass_2001_Meta.pdf
**NOTE**: There was no associated code data for this layer 

CODES
M - MARINA/ANCHORAGE
BR - BOAT RAMP
B - BEACH
WI - WATER INTAKE
CF - COMMERCIAL FISHING
AQ - AQUACULTURE
F - FERRY  
WO - WASHOVER
HS - HISTORICAL SITE
A - AIRPORT


```{r}
ri_2001_soc <- st_read(dsn=file.path(dir_anx, "_raw_data/NOAA_ESI/RI_CT_NY_NJ_2001_Shapefiles/shape/soceconp.shp")) %>% 
  st_set_crs("+init=epsg:4269") %>%
  st_transform(us_alb) %>%
  st_intersection(ri_shape) %>%
  mutate(TYPE = as.character(TYPE)) %>%
  filter(TYPE %in% c("M", "HS", "BR", "B", "F"))
```



----

Lets look at the 2016 and 2001 points

```{r, fig.width = 8, fig.height = 8}
library(mapview)
mapview(ri_2016, col.regions = "blue") + mapview(ri_2001_soc, col.regions = "red")
```








