---
title: "OHIEC: Resilience - Fishing Pressure prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R
dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(tidyverse)
library(dplyr)
```



To be filled out!!!

NAS databast

focuses on plants and freshwater
- includes species from other countries and non-native ranges

Status: established, unknown, extirpated, collected, failed, eradicated, stocked, locally established
Min year = 1700
Max year = 2019

look at it by year, and then unique, total  
look to see if terrestrial ar eincluded

create a data set by state, graph over time number established 1970-2018
```{r load_data}
nas_raw <- read_csv(file.path(dir_anx, "_raw_data/NAS/NAS-Data-Download.csv")) 
```

```{r calculate_by_state}
nas_int <- nas_raw %>% 
  select('Scientific Name', 'Country', 'State', 'County', 'Year', 'Status') %>% 
  rename(sci_name = "Scientific Name") %>% 
  filter(Status %in% c('established', 'eradicated', 'locally established'),
         Country %in% 'United States of America',
         State %in% c("ME", "MA", "NY", "RI", "CT", "NH"),
         Year >= 2005 & Year <= 2017) %>%
  unique() %>% 
  group_by(Year, State) %>% 
  mutate(est_occur = length(which(Status == "established" | Status == "locally established")), # need to add locally established here
         era_occur = length(which(Status == "eradicated")),
         nas_score = era_occur/est_occur) %>% 
  select(-sci_name, - County, -Status, -Country) %>% 
  unique() %>% 
  ungroup()
```

```{r visualize}
nass_res_graph <- ggplot(nas_int) +
  geom_line(aes(x= Year, y=nas_score, color= State))+
  theme_classic() +
  labs(y = "year", x = "score") +
  ggtitle("NASS ")

nass_res_graph 
ggsave(file.path(dir_git, "prep/resilience/ecological/species_pollution/figs/nass_res_graph.jpg"), width=7, height=5, dpi=300)
```

Add in region name and id
```{r adding_rgns}
nas <- nas_int %>% 
 mutate(rgn_id1 = case_when(State == "ME" ~ 6,
                            State == "NH" ~ 9,
                            State == "MA" ~ 7,
                            State == "RI" ~ 11,
                            State == "CT" ~ 5,
                            State == "NY" ~ 10),
         rgn_id8 = case_when(State == "MA" ~ 8)) %>% 
  rename(rgn_name = State) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id8, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(Year, rgn_name, rgn_id, nas_score) %>% 
  mutate(rgn_name = ifelse(rgn_id==5, "Connecticut",
                            ifelse(rgn_id==6, "Maine",
                                   ifelse(rgn_id==7, "Massachusetts-North",
                                          ifelse(rgn_id==8, "Massachusetts-South",
                                                 ifelse(rgn_id==9, "New Hampshire",
                                                        ifelse(rgn_id==10, "New York",
                                                               "Rhode Island")))))))
```

save this layer
```{r_layer}
write.csv(nas, file.path(dir_git, "prep/resilience/ecological/species_pollution/data/nas.csv"))
```


Now we want to see how many species are observed in a new area in each to quantify invasive species pressures
```{r occurances}
nas_est <- nas_raw %>% 
  select('Scientific Name', 'Country', 'State', 'Year', 'Status') %>% 
  rename(sci_name = "Scientific Name") %>% 
  filter(Status %in% c('established', 'locally established', 'eradicated','extripated'),
         Country %in% 'United States of America',
         State %in% c("ME", "MA", "NY", "RI", "CT", "NH"),
         Year >= 1970 & Year <= 2018) %>% 
  select(-Country) %>% 
  unique() %>% 
  group_by(Year, State) %>% 
  mutate(est_occur = length(Status)) 
```

```{r occurances_viz}
est_gr <- ggplot(nas_est) +
  geom_line(aes(x= Year, y=est_occur, color=State))+
  theme_classic() +
  labs(y = "Number Species", x = "State") +
  ggtitle("Invasive Species")

est_gr 

ggsave("figs/established_nas.jpg", width=7, height=5, dpi=300)
```

