---
title: "OHINE: Resilience - Nonindigenous Aquatic Species USGS Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the invasive species effectiveness and compliance scores by OHINE region. 

# Data

**Nonindigenous Aquatic Species (NAS) USGS** 

**Downloaded**: September, 2019

**Description**: This dataset has occurrence and precise location data of new nonindigenous species https://nas.er.usgs.gov/

**Time range**: 1700-2019

**Format**: csv file

# About

This dataset was used as an indicator to measure enforcement and compliance of regulations aimed at preventing the introduction and spread of invasive species. The occurrence data is consolidated by state, and the eradication of species/establishment of species is calculated

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R
dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(tidyverse)
library(dplyr)
```


```{r load_data}
nas_raw <- read_csv(file.path(dir_anx, "_raw_data/NAS/NAS-Data-Download.csv")) 
```

## Calculate Scores and Graph
```{r calculate_by_state}
nas_int <- nas_raw %>% 
  select('Scientific Name', 'Country', 'State', 'County', 'Year', 'Status') %>% 
  rename(sci_name = "Scientific Name") %>% 
  filter(Status %in% c('established', 'eradicated', 'locally established'),
         Country %in% 'United States of America',
         State %in% c("ME", "MA", "NY", "RI", "CT", "NH"),
         Year >= 2005 & Year <= 2017) %>%
  unique() %>% 
  group_by(Year, State) %>% 
  mutate(est_occur = length(which(Status == "established" | Status == "locally established")), # need to add locally established here
         era_occur = length(which(Status == "eradicated")),
         nas_score = era_occur/est_occur) %>% 
  select(-sci_name, - County, -Status, -Country) %>% 
  unique() %>% 
  ungroup()
```

```{r visualize}
nas_res_graph <- ggplot(nas_int) +
  geom_line(aes(x= Year, y=nas_score, color= State))+
  theme_classic() +
  labs(y = "year", x = "score") +
  ggtitle("NASS ")

nas_res_graph 
ggsave(file.path(dir_git, "prep/resilience/ecological/species_pollution/figs/nass_res_graph.jpg"), width=7, height=5, dpi=300)
```

Add in region name and id
```{r adding_rgns}
nas <- nas_int %>% 
 mutate(rgn_id1 = case_when(State == "ME" ~ 6,
                            State == "NH" ~ 9,
                            State == "MA" ~ 7,
                            State == "RI" ~ 11,
                            State == "CT" ~ 5,
                            State == "NY" ~ 10),
         rgn_id8 = case_when(State == "MA" ~ 8)) %>% 
  rename(rgn_name = State) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id8, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  select(Year, rgn_name, rgn_id, nas_score) %>% 
  mutate(rgn_name = ifelse(rgn_id==5, "Connecticut",
                            ifelse(rgn_id==6, "Maine",
                                   ifelse(rgn_id==7, "Massachusetts-North",
                                          ifelse(rgn_id==8, "Massachusetts-South",
                                                 ifelse(rgn_id==9, "New Hampshire",
                                                        ifelse(rgn_id==10, "New York",
                                                               "Rhode Island")))))))
```

## Save to layers
save this layer
```{r_layer}
write.csv(nas, file.path(dir_git, "prep/resilience/ecological/species_pollution/data/nas.csv"))
```

## Explore data for pressures

Exploring total number of occurrences that have become established for the species pollution pressure layer
Now we want to see how many species are observed in a new area in each to quantify invasive species pressures
```{r occurances}
nas_est <- nas_raw %>% 
  select('Scientific Name', 'Country', 'State', 'Year', 'Status') %>% 
  rename(sci_name = "Scientific Name") %>% 
  filter(Status %in% c('established', 'locally established', 'eradicated','extirpated'),
         Country %in% 'United States of America',
         State %in% c("ME", "MA", "NY", "RI", "CT", "NH"),
         Year >= 1970 & Year <= 2018) %>% 
  select(-Country) %>% 
  unique() %>% 
  group_by(Year, State) %>% 
  mutate(est_occur = length(Status)) 
```

```{r occurances_viz}
est_gr <- ggplot(nas_est) +
  geom_line(aes(x= Year, y=est_occur, color=State))+
  theme_classic() +
  labs(y = "Number Species", x = "State") +
  ggtitle("Invasive Species")

est_gr 

ggsave("figs/established_nas.jpg", width=7, height=5, dpi=300)
```

