---
title: "OHINE: Resilience - Species Pollution Resilience Layer Calculations"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates species pollution resilience values by OHINE region. 

The data layers are used to quantify existence of regulations, implementation + enforcement, and effectiveness + compliance. These three parts are then used to calculate one overall water pollution resilience score for each region. 

1. Regulations: aquaculture_reg, ballast_reg
2. Implementation + Enforcement: nas
3. Effectiveness + Compliance: nas

Data Time Ranges:
aquaculture_reg: NA
ballast_reg: NA
nas: 2005-2017

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R
dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(tidyverse)
library(dplyr)
library(readxl)
```

Read in Data
```{r load_data}
aquaculture_reg <- read_xlsx(file.path(dir_anx, "_raw_data/manually_created/aquaculture_reg.xlsx"))

ballast_reg <- read_xlsx(file.path(dir_anx, "_raw_data/manually_created/ballast_reg.xlsx"))

nas_short<- read.csv("~/github/ne-prep/prep/resilience/ecological/species_pollution/data/nas.csv") %>% 
  select(-X)
```

## Calculate Scores, graph, and save layer
Creating a df to add on to the water pollution scores to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  3, "Gulf of Maine",
  2, "Georges Bank",
  4, "Mid-Atlantic Bight",
  1, "Offshore") 

miss_rgns <- miss_rgns %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) %>% 
  mutate(score = NA) %>% 
  unique()
```

Cleaning up the data to be combined and used as a layer

NAS
```{r nas_int}
nas_int <- nas_short %>% 
  rename(year = Year) %>% 
  rename(score = nas_score) %>% 
  rbind(miss_rgns) %>% 
  arrange(year, rgn_name)
```

4 entries missing in the nas data, which means that 4 times no species occurrences were logged:
RI 2008, 2011, 2012,2013. Even though Rhode Island has a score of 0 for every entry, this is because of very low invasive species occurrences  (1= lowest (and the mode) and 7=highest). So here I will give RI a score of 1 because the way that the data is represented this means that no new invasive species were introduced.

```{r ri_missing}
nas_ri_miss <- tribble(
  ~year, ~rgn_name, ~rgn_id, ~score,
  2008, "Rhode Island", 11, 1,  
  2011, "Rhode Island", 11, 1, 
  2012, "Rhode Island", 11, 1, 
  2013, "Rhode Island", 11, 1
  ) %>% 
  as.data.frame() 
```

```{r nas_final}
nas <- nas_int %>% 
  rbind(nas_ri_miss) %>% 
  rename(nas_score = score)
```

Combine all the data together and calculate a regulation, enforcement, compliance, and overall resilience score
```{r calculate_alien}
res_sp <- left_join(nas, ballast_reg, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(aquaculture_reg, by = c("year", "rgn_id", "rgn_name")) %>% 
  rowwise() %>% 
  mutate(enforcement = mean(c(ballast_reg_score, aquaculture_reg_score)),
         regulation = nas_score,
         compliance = nas_score) %>% 
  mutate(res_score = mean(c(regulation, enforcement, compliance)))
```

Save necessary information for the water pollution layer
```{r save_data}
res_species_pollution <- res_sp %>% 
  select(year, rgn_id, rgn_name, res_score) %>% 
  rename(score = res_score)

write.csv(res_species_pollution, file.path(dir_git,"prep/resilience/ecological/species_pollution/data/res_species_pollution.csv"))
```

Visualization
```{r graph_alien}
res_species_pollution_graph <- ggplot(res_species_pollution) +
  geom_line(aes(x=year, y= score, color = rgn_name))+
  ggtitle("North East Species Pollution Resilience Scores") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 

res_species_pollution_graph
ggsave("figs/res_species_pollution_graph.jpg", width=7, height=5, dpi=300)
```


Write as a resilience layer to ne-scores
```{r write_to_scores}
write.csv(res_species_pollution, file.path(dir_calc, "layers/res_alien.csv"))
```















