---
title: "Nature Serve Data"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary

We access the NatureServe API with the `natserv` R package to retrieve each species conservation status across all scales from global (IUCN) to state level (CT, NY, MA, ME, RI, NH).

# Setup

``` {r setup, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/', message = FALSE, warning = FALSE)

library(tidyverse)
#install.packages("natserv")
library(natserv)
options(NatureServeKey = Sys.getenv("NatureServeKey"))
library(taxize)

source("~/github/ne-prep/src/R/common.R")
```

We need to get scientific names for the data portal species before querying NatureServe. 
```{r get_sci_names_for_dp_species}
#ne dataportal rasters
ne_rasters <- list.files(file.path(dir_anx, "bio/portal_spp_rasters"))

#get data portal species
ne_dp_spp <- data.frame(filepath = ne_rasters) %>%
  mutate(name = sub(".tif", "", filepath),
         common = gsub("_", " ", name),
         common_name = str_trim(tolower(gsub("normalized AEA", "", common)), "both")) %>% #removing white space and the normalized aea strings
  select(common_name) %>%
  mutate(source = "dataportal")
```

The Nature Serve R package requires scientific names to query the API. We can use the `taxize` library to help.

```{r, eval = F}
#for loop to grab each data portal species name and find it's scientific name
dp_sci_names <- c()
for(i in 1:nrow(ne_dp_spp)){ #i <- 1
  sp <- ne_dp_spp[i,1]
  sci <- comm2sci(sp, db = "eol")
  
dp_sci_names <- c(dp_sci_names, sci)
}

#turn list into dataframe
dp_sci_names_df <- do.call(rbind,lapply(dp_sci_names,data.frame))
dp_sci_names_df$common <- rownames(dp_sci_names_df)

#clean up df. there are a lot of viruses?!
dp_sci_names_df_clean <- dp_sci_names_df %>%
  rename(sciname = X..i..) %>%
  mutate(common = trimws(str_extract(common, "[^.]+")), "r") %>% #remove numbers
  distinct() %>%
  filter(!str_detect(sciname, "virus")) %>%
  group_by(common) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  mutate(sci_fix = case_when( #manually doing this... taxize returns so many names
    common == "butterfish" ~ "Peprilus triacanthus",
    common == "cusk" ~ "Brosme brosme",
    common == "pollock" ~ "Pollachius virens",
    common == "spot" ~ "Leiostomus xanthurus",
    common == "tilefish" ~ "Lopholatilus chamaeleonticeps",
    common == "weakfish" ~ "Cynoscion regalis",
    common == "pigfish" ~ "Orthopristis chrysoptera",
    common == "haddock" ~ "Melanogrammus aeglefinus",
    common == "round herring" ~ "Etrumeus sadina",
    common == "goosefish" ~ "Lophius americanus",
    common == "bluefish" ~ "Pomatomus saltatrix",
    common == "horseshoe crab" ~ "Limulus polyphemus",
    common == "pinfish" ~ "Lagodon rhomboides",
    common == "spiny dogfish" ~ "Squalus acanthias",
    common == "sperm whale" ~ "Physeter macrocephalus",
    common == "sand tiger" ~ "Carcharias taurus",
    common == "sea raven" ~ "Hemitripterus americanus",
    common == "atlantic sturgeon" ~ "Acipenser oxyrinchus",
    common == "silver hake" ~ "Merluccius bilinearis",
    common == "bottlenose dolphin" ~ "Tursiops truncatus",
    common == "corys shearwater" ~ "Calonectris borealis",
    common == "herring gull" ~ "Larus smithsonianus",
    common == "atlantic herring" ~ "Clupea harengus",
    common == "bay anchovy" ~ "Anchoa mitchilli",
    common == "fin whale" ~ "Balaenoptera physalus",
    common == "fourspot flounder" ~ "Hippoglossina oblonga",
    common == "ocean pout" ~ "Zoarces americanus",
    common == "razorbill" ~ "Alca torda",
    common == "smooth dogfish" ~ "Mustelus canis",
    common == "smooth skate" ~ "Malacoraja senta",
    common == "striped bass" ~ "Morone saxatilis",
    common == "thorny skate" ~ "Amblyraja radiata",
    common == "winter skate" ~ "Leucoraja ocellata",
    common == "alewife" ~ "Alosa pseudoharengus",
    common == "american plaice" ~ "Hippoglossoides platessoides",
    common == "arctic tern" ~ "Sterna paradisaea",
    common == "atlantic spotted dolphin" ~ "Stenella frontalis",
    common == "atlantic cod" ~ "Gadus morhua",
    common == "atlantic croaker" ~ "Micropogonias undulatus",
    common == "atlantic mackerel" ~ "Scomber scombrus",
    common == "atlantic puffin" ~ "Fratercula arctica",
    common == "atlantic sharpnose shark" ~ "Rhizoprionodon terraenovae",
    common == "atlantic white sided dolphin" ~ "Lagenorhynchus acutus",
    common == "atlantic halibut" ~ "Hippoglossus hippoglossus",
    common == "atlantic wolffish" ~ "Anarhichas lupus",
    common == "banded drum" ~ "Larimus fasciatus",
    common == "black legged kittiwake" ~ "Rissa tridactyla",
    common == "black scoter" ~ "Melanitta americana",
    common == "blackbelly rosefish" ~ "Helicolenus dactylopterus",
    common == "black sea bass" ~ "Centropristis striata",
    common == "blainvilles beaked whale" ~ "Mesoplodon densirostris",
    common == "bluntnose stingray" ~ "Dasyatis say",
    common == "bonapartes gull" ~ "Chroicocephalus philadelphia",
    common == "bridled tern" ~ "Onychoprion anaethetus",
    common == "bullnose ray" ~ "Myliobatis freminvillii",
    common == "common tern" ~ "Sterna hirundo",
    common == "clearnose skate" ~ "Raja eglanteria",
    common == "common eider" ~ "Somateria mollissima",
    common == "common loon" ~ "Gavia immer",
    common == "common murre" ~ "Uria aalge",
    common == "cuviers beaked whale" ~ "Ziphius cavirostris",
    common == "double crested cormorant" ~ "Phalacrocorax auritus",
    common == "dovekie" ~ "Alle alle",
    common == "gervais beaked whale" ~ "Mesoplodon europaeus",
    common == "great shearwater" ~ "Ardenna gravis",
    common == "great skua" ~ "Stercorarius skua",
    common == "gulf stream flounder" ~ "Citharichthys arctifrons",
    common == "harbor porpoise" ~ "Phocoena phocoena",
    common == "hickory shad" ~ "Alosa mediocris",
    common == "horned grebe" ~ "Podiceps auritus",
    common == "humpback whale" ~ "Megaptera novaeangliae",
    common == "laughing gull" ~ "Leucophaeus atricilla",
    common == "leachs storm petrel" ~ "Oceanodroma leucorhoa",
    common == "least tern" ~ "Sternula antillarum",
    common == "little skate" ~ "Leucoraja erinacea",
    common == "long finned pilot whale" ~ "Globicephala melas",
    common == "long tailed duck" ~ "Clangula hyemalis",
    common == "manx shearwater" ~ "Puffinus puffinus", 
    common == "minke whale" ~ "Balaenoptera acutorostrata",
    common == "north atlantic right whale" ~ "Eubalaena glacialis",
    common == "northern fulmar" ~ "Fulmarus glacialis",
    common == "northern gannet" ~ "Morus bassanus",
    common == "northern kingfish" ~ "Menticirrhus saxatilis",
    common == "northern pipefish" ~ "Syngnathus fuscus",
    common == "red throated loon" ~ "Gavia stellata",
    common == "rissos dolphin" ~ "Grampus griseus",
    common == "roughtail stingray" ~ "Dasyatis centroura",
    common == "rosette skate" ~ "Leucoraja garmani",
    common == "royal tern" ~ "Thalasseus maximus",
    common == "sei whale" ~ "Balaenoptera borealis",
    common == "short beaked common dolphin" ~ "Delphinus delphis",
    common == "short finned pilot whale" ~ "Globicephala macrorhynchus",
    common == "sooty shearwater" ~ "Ardenna grisea",
    common == "sooty tern" ~ "Onychoprion fuscatus",
    common == "south polar skua" ~ "Stercorarius maccormicki",
    common == "spotted hake" ~ "Urophycis regia",
    common == "striped dolphin" ~ "Stenella coeruleoalba",
    common == "striped anchovy" ~ "Anchoa hepsetus",
    common == "summer flounder" ~ "Paralichthys dentatus",
    common == "surf scoter" ~ "Melanitta perspicillata",
    common == "white winged scoter" ~ "Melanitta fusca",
    common == "wilsons storm petrel" ~ "Oceanites oceanicus",
    common == "windowpane" ~ "Scophthalmus aquosus",
    common == "winter flounder" ~ "Pseudopleuronectes americanus",
    common == "yellowtail flounder" ~ "Pleuronectes ferruginea",
    common == "northern shortfin squid" ~ "Illex illecebrosus",
    common == "longfin squid" ~ "Doryteuthis pealeii",
    TRUE ~ as.character(sciname)
  )) %>%
  select(common, sciname = sci_fix) %>%
  distinct() 

write.csv(dp_sci_names_df_clean, file = "data/4_ne_dataportal_spp_common_sci.csv")
```

```{r}
#read dataportal names in
dp_sci_names_df_clean <- read_csv("data/4_ne_dataportal_spp_common_sci.csv")

#get IUCN species
ne_iucn_spp <- read.csv("data/1_iucn_spp_in_ne.csv", stringsAsFactors = F)$sciname

#combine to get unique scientific names for all of our species across data portal and iucn maps
all_sci_names <- c(ne_iucn_spp, dp_sci_names_df_clean$sciname) %>% unique()
```

```{r}

#forloop to grab species uid
full_spp_df <- data.frame(state = NA,
                          status = NA,
                          species = NA,
                          stringsAsFactors = F)

spp_status_history <- data.frame()

for (i in 1:length(all_sci_names)){
  print(i)
  
spp <- all_sci_names[i]  

  #first use ns_search to see if the species is in natureserve (tryCatch will skip to next if it's not)
  possibleError <- tryCatch(
      id <- ns_search(x = spp)$globalSpeciesUid,
      error=function(e) e)

if(inherits(possibleError, "error")) next
  
  #then grab the data from natureserve
  possibleError <- tryCatch(
      b  <- ns_data(uid = id),
      error=function(e) e)

if(inherits(possibleError, "error")) next
  
  #get IUCN status
  iucn_stat <- b[[1]]$conservationStatus$other$`IUCN Conservation Status`$statusDescription
  if(is.null(iucn_stat))
    iucn_stat <- NA
  
  # get the US national rank
  us_rank <- b[[1]]$conservationStatus$natureserve$nationalStatuses$US$rank
  if(is.null(us_rank)) next #if there is no national rank, skip to next species
  
  #create a dataframe to rbind later
  iucn_and_us_df <- data.frame(state = c("IUCN", "USA"), 
                               status = c(iucn_stat, us_rank), 
                               stringsAsFactors = F) 

  #Now grab state level information
  state <- b[[1]]$conservationStatus$natureserve$nationalStatuses$US$subnationalStatuses

  #create an empty dataframe to cycle through each state
  out <- data.frame(state = NA,
                    rank = NA)

#forloop goes through each state and finds only those in our region and gets that info. Output is a dataframe

if(length(state) > 0) {
  
  for(j in 1:length(state)){
    
    ST <- state[[j]]$subnationCode
  
    if(ST %in% c("CT", "NY", "MA", "ME", "RI", "NH")){
      state_rank <- state[[j]]$rank
      
      df <- data.frame(state = ST, rank = state_rank, stringsAsFactors = F)
    }else{
      df <- data.frame(state = NA, rank = NA, stringsAsFactors = F)
    }
    out <- rbind(out, df) %>%
      filter(!is.na(state))
  }
}else{
  df <- data.frame(state = NA, rank = NA, stringsAsFactors = F)
}

spp_df <- out %>%
  mutate(update_rank = 
           ifelse(rank %in% c("SNA", "SNR", "SU", "SNRN"), as.character(us_rank), as.character(rank))) %>%
  select(-rank) %>%
  rename(status = update_rank) %>%
  bind_rows(iucn_and_us_df) %>%
  mutate(species = spp)

full_spp_df <- rbind(spp_df, full_spp_df) 

################## 
# Also collecting the history of status reviews and changes

 # get the last year US rank was reviewed
  us_year <- substr(b[[1]]$conservationStatus$natureserve$nationalStatuses$US$statusLastReviewed, 1, 4)
  
  #get Year last reviewed and changed by natureserve
  ns_year_last_reviewed <- substr(b[[1]]$conservationStatus$natureserve$statusLastReviewed, 1, 4)
    
  ns_year_last_changed  <- substr(b[[1]]$conservationStatus$natureserve$statusLastChanged, 1, 4)
  
  hist_df <- data.frame(species                   = spp,
                        us_rank_review_yr         = ifelse(length(us_year) > 0, us_year, NA),
                        natserve_yr_last_reviewed = ifelse(length(ns_year_last_reviewed) > 0, ns_year_last_reviewed, NA),
                        natserve_yr_lst_changed   = ifelse(length(ns_year_last_changed) > 0, ns_year_last_changed, NA))
  
#spp status history df
spp_status_history <- rbind(spp_status_history, hist_df)

}

write.csv(full_spp_df, file = "data/4_natureserve_spp_status.csv")
write.csv(spp_status_history, file = "data/4_natureserve_spp_status_history.csv")
```

