---
title: "Determining why some land cells are showing up"
author: "Jamie Afflerbach"
date: "2/11/2019"
output: html_document
---

```{r setup, warning = F, message = F}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
source('~/github/ne-prep/src/R/common.R')
```

I think land cells are occurring from the IUCN maps.

```{r}
spp_rgn_raster <- raster("data/spp_status_risk.tif")

plot(spp_rgn_raster, col = cols)
```

Get extent just of long island

```{r, eval = F}
li_ext <- zoom(spp_rgn_raster, col = "red")
# class       : Extent 
# xmin        : 1653246 
# xmax        : 1950368 
# ymin        : 542017 
# ymax        : 662937 
```
```{r}
li_rast <- crop(spp_rgn_raster, li_ext)
```


Plot cellID raster

```{r}
cellid_rast <- raster("~/github/ne-prep/spatial/ocean_rasters/ne_cellids.tif") %>%
  crop(li_ext) %>%
  mask(li_rast)

plot(cellid_rast)
```

Now using the "click" tool from `raster` I can ask what those cellIds are over long island

```{r}
#i know these are all land cells over long island
li_cells <- c(500882:500886)
```

Let's filter the `iucn_cells` dataframe for species IDs in those cells

```{r}
iucn_cells <- data.table::fread(file.path(dir_anx, "bio/iucn_sid_cells.csv")) %>%
  select(-V1) %>%
  filter(cellID %in% li_cells)

li_sp <- unique(iucn_cells$SID)
```

Wow we have `r length(li_sp)` species in these cells.

```{r}
#this df tells us what regions contain each species from IUCN maps
iucn_sp <- read_csv("data/1_iucn_spp_in_ne.csv") %>%
  filter(iucn_sid %in% li_sp)
```

Let's plot some of these rasters

```{r}
iucn_rasters <- list.files(file.path(dir_anx, "bio/spp_presence_rasters"), full.names = T)

iucn_list <- stack(iucn_rasters[substr(iucn_rasters, 69, nchar(iucn_rasters)-4) %in% li_sp])
```

```{r}
plot(iucn_list[[20]], col = "black")
```

I think these all might be birds...

```{r}
library(taxize)

li_spp <- iucn_sp %>%
  mutate(common = sci2comm(sciname, db = "eol"))
```

Yes these are all bird species. I feel comfortable using land mask to remove them. But I don't think this explains some of the edge cells (e.g. in Maine). So doing the same type of thing but for coast of Maine.

```{r}
me_ext <- zoom(spp_rgn_raster, col = "red")
# class       : Extent 
# xmin        : 2151167 
# xmax        : 2182287 
# ymin        : 1070539 
# ymax        : 1097770 
```

```{r}
me_rast <- crop(spp_rgn_raster, me_ext)
```

Plot cellID raster

```{r}
cellid_rast <- raster("~/github/ne-prep/spatial/ocean_rasters/ne_cellids.tif") %>%
  crop(me_ext) %>%
  mask(me_rast)

par(mfrow = c(1,2))
plot(cellid_rast)
plot(me_rast)
```

Now using the "click" tool from `raster` I can ask what those green cellIDs are around the coast of Maine

```{r}
#i know these are all land cells over long island
me_cells <- c(91620, 91621, 82007, 102010, 98007, 98005, 98829, 94822, 94022)
```



Let's filter the `iucn_cells` dataframe for species IDs in those cells

```{r}
iucn_cells <- data.table::fread(file.path(dir_anx, "bio/iucn_sid_cells.csv")) %>%
  select(-V1) %>%
  filter(cellID %in% me_cells)

me_sp <- unique(iucn_cells$SID)
```

Wow we have `r length(me_sp)` species in these cells.

```{r}
#this df tells us what regions contain each species from IUCN maps
iucn_sp <- read_csv("data/1_iucn_spp_in_ne.csv") %>%
  filter(iucn_sid %in% me_sp)
```

Let's plot some of these rasters

```{r}
iucn_rasters <- list.files(file.path(dir_anx, "bio/spp_presence_rasters"), full.names = T)

iucn_list <- stack(iucn_rasters[substr(iucn_rasters, 69, nchar(iucn_rasters)-4) %in% me_sp])
```

```{r}
plot(iucn_list[[20]], col = "black")
```

```{r}
me_spp <- iucn_sp %>%
  mutate(common = sci2comm(sciname, db = "eol"))

View(me_spp)
```

A lot of these are birds, but there are also loggerhead and leatherback sea turtles which have high risks.

```{r}
#these are the SID for leatherback and loggerhead sea turtles respectively. I want to map them
turtles <- c(46967827, 84131194)

turtle_map <- stack(iucn_rasters[substr(iucn_rasters, 69, nchar(iucn_rasters)-4) %in% turtles])

plot(turtle_map, col = "black")
```

It's not mainly the turtles driving this 

```{r}
l <- read.csv("data/7_spp_rgns_scores.csv") %>%
  filter(sciname %in% me_spp$sciname,
         rgn_name == "Maine")

l
```



