---
title: "OHINE: Habitat Destruction Resilience - Habitat Areas Protected Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the habitat areas that are protected scores by OHINE region. This data is used to measure habitat destruction resilience regulations.

# Data

**NOAA Greater Atlantic Fisheries** 

**Downloaded**: August, 2019

**Description**: Shapefiles where fishing and habitat management measures are. https://www.greateratlantic.fisheries.noaa.gov/educational_resources/gis/data/

**Time range**: last updated 2018

**Format**: shapefiles

# About

Filtered out management areas where no dredging or trawling is allowed, year-round.
Currently waiting/working on getting this info over time, for now/if that never happens we will apply the scores to all years

1) Habitat Management Area
  - Cashes ledge and Fippennies ledge habitat management areas,
  - Ammen rock habitat management area
  - Eastern Maine HMA and Jeffrey’s Bank HMA
  - Closed Area II Habitat Closure
  - Great South Channel Habitat Management Area
2) Habitat Research Area
  - George’s Bank DHRA
3) Frank R. Lautenberg Deep-Sea Coral Protection Area
4) Groundfish Closures
  - Closed Area II 
  - Western Gulf of Maine habitat closure
5) MSB Bottom Trawling Restricted Areas
  - Oceanographer Canyon
  - Lydonia Canyon

# Methods

## Set up, libraries, load data
Used a lot of the code from the lsp.Rmd
``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/', message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

library(tidyverse)
library(sf)
library(mapview)
library(readxl)
```


```{r load_data}
#1
hab_manage <- read_sf(dsn = file.path(dir_anx, "_raw_data/NOAA_HAB/Habitat_Management_Areas/Habitat_Management_Areas.shp")) %>% 
  filter(DESCRIBE != "Exemption area") %>% 
  mutate(start_year= 2018)

#2
hab_research <- read_sf(dsn = file.path(dir_anx, "_raw_data/NOAA_HAB/Dedicated_Habitat_Research_Area/Dedicated_Habitat_Research_Area.shp")) %>% 
  filter(AREANAME == "Georges Bank Dedicated Habitat Research Area") %>% # Stellwagen allows surf clam. quahog dredge gear
  mutate(start_year= 2018)
  
#3
hab_frl <- read_sf(dsn = file.path(dir_anx, "_raw_data/NOAA_HAB/Frank_R_Lautenberg_Deep_Sea_Coral_Protection_Areas/Frank_R_Lautenberg_Deep_Sea_Coral_Protection_Areas.shp"))%>% 
  mutate(start_year= 2018)

#4
hab_msb_trawl <- read_sf(dsn = file.path(dir_anx, "_raw_data/NOAA_HAB/MSB_Bottom_Trawling_Restricted_Areas/MSB_Bottom_Trawling_Restricted_Areas.shp"))%>% 
  mutate(start_year= 2008)

#combine all together
hab_noaa <- rbind(hab_manage, hab_research, hab_frl, hab_msb_trawl) %>%
  st_transform(us_alb) %>%
  st_crop(rgns)  %>%
  st_intersection(rgns)

hab_protect <- mapview(hab_noaa) 
hab_protect
mapshot(hab_protect, url = "hab_protect.html")
```


## Calculate Scores
Here we are using a 30% protected area target 

```{r calculate_hab_protected}
hab_protect_rgn <- hab_noaa %>%
  select(rgn_id, rgn_name, AREANAME)%>%
  group_by(rgn_id, rgn_name) %>%
  summarise() %>%
  ungroup() %>%
  mutate(area = st_area(geometry)/1000000) %>% #calculate total protected area
  left_join(rgn_data) %>%
  mutate(prop_area = as.numeric(area)/area_km2, 
         type = "marine") %>%
    st_set_geometry(NULL) %>% 
  select(rgn_id, rgn_name, prop_area) %>% 
  mutate(target = 0.3,
         score = prop_area/target,
         score = if_else(score >1, 1, score)) 
```

Creating a df to add to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  9, "New Hampshire",
  7, "Massachusetts-North",
  8, "Massachusetts-South",
  11, "Rhode Island",
  5, "Connecticut",
  10, "New York") %>% 
  mutate(score = 0)
```


Create a table with only the necessary info for calculating the score
```{r calculate_hab_protected_Score}
hab_area_protected <- hab_protect_rgn %>% 
  select(rgn_id, rgn_name,score) %>% 
  rbind(miss_rgns) %>% 
  rename(hab_area_protected_score = score)
```

Save layer
```{r save_data}
write.csv(hab_area_protected, file = "data/hab_area_protected.csv")
```



