---
title: "OHIEC: Resilience - Habitat Destruction prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates habitat resilience values by OHINE region. 

The data layers are used to quantify existence of regulations, implementation + enforcement, and effectiveness + compliance. These three parts are then used to calculate one overall habitat resilience score for each region. 

1. Regulations: hab_area_protected, surfrider_beach
2. Implementation + Enforcement: ole_enforcement (from fishing pressure data folder)
3. Effectiveness + Compliance: ole_compliance (from fishing pressure data folder)

Date Ranges:
hab_area_protected: NA
surfrider_beach: NA
ole_enforcement: 2015-2018
ole_compliance: 2015-2018


``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(dplyr)
library(tidyverse)
library(readxl)
```

Read in Data
```{r}
hab_area_protected_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/habitat_destruction/data/hab_area_protected.csv") %>% 
  select(-X)

surfrider_beach_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/habitat_destruction/data/surfrider_beach.csv") %>% 
  select(-X) 

ole_enforcement_short<- read.csv("~/github/ne-prep/prep/resilience/ecological/fishing_pressure/data/ole_enforcement.csv") %>% 
  select(-X)

ole_compliance_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/fishing_pressure/data/ole_compliance.csv") %>% 
  select(-X)
```

##Data Tidying for combining and calculations
Tidying the habitat protected area data for res score calculations over time
```{r}
hab_area_protected <- hab_area_protected_short %>% 
  uncount(3, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2015, n + 2014)) %>% 
  select(-n)
```

Tidying the habitat protected area data for res score calculations over time
```{r}
surfrider_beach <- surfrider_beach_short %>% 
  uncount(3, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2015, n + 2014)) %>% 
  select(-n)
```

Tidying the OLE enforcemnet data for res score calculations by region
```{r}
ole_enforcement <- ole_enforcement_short %>% 
  select(year, ole_enforcement_score) %>% 
  filter(year <= 2017) %>% 
  uncount(11, .id = "n", .remove = F) %>%
  mutate(rgn_name = ifelse(n == 1, "Offshore", 
                       ifelse(n==2, "Georges Bank",
                               ifelse(n==3, "Gulf of Maine",
                                      ifelse(n==4, "Mid-Atlantic Bight",
                                             ifelse(n==5, "Connecticut",
                                                    ifelse(n==6, "Maine",
                                                           ifelse(n==7, "Massachusetts-North",
                                                                  ifelse(n==8, "Massachusetts-South",
                                                                         ifelse(n==9, "New Hampshire",
                                                                                ifelse(n==10, "New York",
                                                                                       "Rhode Island"))))))))))) %>%
  rename(rgn_id=n)
```

Tidying the OLE compliance data for res score calculations by region
```{r}
ole_compliance <- ole_compliance_short %>% 
  select(year, ole_compliance_score) %>% 
  filter(year <= 2017) %>% 
  uncount(11, .id = "n", .remove = F) %>%
  mutate(rgn_name = ifelse(n == 1, "Offshore", 
                       ifelse(n==2, "Georges Bank",
                               ifelse(n==3, "Gulf of Maine",
                                      ifelse(n==4, "Mid-Atlantic Bight",
                                             ifelse(n==5, "Connecticut",
                                                    ifelse(n==6, "Maine",
                                                           ifelse(n==7, "Massachusetts-North",
                                                                  ifelse(n==8, "Massachusetts-South",
                                                                         ifelse(n==9, "New Hampshire",
                                                                                ifelse(n==10, "New York",
                                                                                       "Rhode Island"))))))))))) %>%
  rename(rgn_id=n)
```

Calculate Habitat Destruction Resilience Score 
```{r}
res_hd <- left_join(hab_area_protected, surfrider_beach, by = c("year", "rgn_name", "rgn_id")) %>% 
  left_join(ole_enforcement, by = c("year", "rgn_name", "rgn_id")) %>% 
  left_join(ole_compliance, by = c("year", "rgn_name", "rgn_id")) %>% 
  rowwise() %>% 
  mutate(regulation = mean(c(hab_area_protected_score, surfrider_beach_score), na.rm=TRUE)) %>% 
 select(-surfrider_beach_score, -hab_area_protected_score) %>% 
  rename(enforcement = ole_enforcement_score) %>% 
  rename(compliance = ole_compliance_score) %>% 
  mutate(res_score = mean(c(regulation, enforcement, compliance)))
```

Visualization
```{r}
res_hd_graph <- ggplot(res_hd) +
  geom_line(aes(x=year, y= res_score, color = rgn_name))+
  ggtitle("North East Habitat Destruction Resilience Scores") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 

res_hd_graph
ggsave("figs/res_hd_graph.jpg", width=7, height=5, dpi=300)
```


Save necessary information for the hab destruction layer
```{r}
res_hab_destruction <- res_hd %>% 
  select(year, rgn_id, rgn_name, res_score) %>% 
  rename(score = res_score)

write.csv(res_hab_destruction, file.path(dir_git,"prep/resilience/ecological/water_pollution/data/res_hab_destruction.csv"))
```


Write as a resilience layer to ne-scores
```{r}
write.csv(res_hab_destruction, file.path(dir_calc, "layers/res_hab_destruction.csv"))
```

