---
title: "OHINE: Resilience - Surfrider State of the Beach Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the scores from the Surfrider State of the Beach reports by OHINE region. 

# Data

**Surfrider**

**Downloaded**: August, 2019

**Description**: Surfrider grades each state on a scale of 1-4 in 4 categories (sediment management, coastal armoring, development, sea level rise) and then adds the points for a total score (12 maximum) https://www.surfrider.org/coastal-blog/entry/2017-state-of-the-beach-report

**Time range**: 2017 (this will be continued into the future)

**Format**: pdf

# About
The Surfrider score for each region was rescaled 0-1 

# Methods

## Set up, libraries, load data
``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/', message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

library(tidyverse)
library(readxl)
```


```{r}
raw_sob <- read_xlsx(file.path(dir_anx, "_raw_data/Surfrider/state_of_the_beach_2017.xlsx"))
```

## Calculate Scores
```{r}
sob_scores <- raw_sob %>% 
  mutate(target = 12,
         score = over_all/target)
```

Creating a df to add to have the missing regions there 
```{r missing_rgns_df}
miss_rgns <- tribble(
  ~rgn_id, ~rgn_name,
  3, "Gulf of Maine",
  2, "Georges Bank",
  4, "Mid-Atlantic Bight",
  1, "Offshore") %>% 
  mutate(score= NA)
```

Create a table with only the necessary info for calculating the score
```{r}
surfrider_beach <- sob_scores %>% 
  select(state, score) %>% 
   mutate(rgn_id1 = case_when(
    state == "Maine" ~ "6", 
    state == "New Hampshire" ~ "9", 
    state == "Massachusetts" ~ "7",
    state == "Rhode Island" ~ "11",
    state == "Connecticut" ~ "5",
    state == "New York" ~ "10"),
    rgn_id2 = case_when(
    state == "Massachusetts" ~ "8"
    )) %>% 
  replace(., is.na(.), "0") %>% 
  gather(rgn_id1, rgn_id2, key= "filler", value = "rgn_id") %>% 
  filter(rgn_id != "0") %>% 
  rename(rgn_name = state) %>% 
  select(rgn_name, rgn_id, score) %>% 
  mutate(rgn_name = ifelse(rgn_name == "Massachusetts" & rgn_id== 7, "Massachusetts-North",
                           ifelse(rgn_name == "Massachusetts" & rgn_id== 8, "Massachusetts-South", 
                                  rgn_name))) %>% 
  rbind(miss_rgns) %>% 
  rename(surfrider_beach_score = score) %>% 
  mutate(rgn_id = as.numeric(rgn_id)) 
```

Save data
```{r}
write.csv(surfrider_beach, file = "data/surfrider_beach.csv")
```