---
title: "OHINE: Fishing Resilience - Fish Landed with Stock Assessments Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the percent of species with stock assessments that are adequately assessed. This data is used to measure fishing resilience regulations.

# Data

**RAM**
**Downloaded**: February 28, 2018 from ramlegacy.org

**Description**:  RAM Legacy Stock Assessment Database provides information on global fish stocks including catch, B/Bmsy and F/Fmsy among other metrics.

**Time range**: pre-1900 to 2016

**Format**:  Tabular


**NOAA Fisheries**
**Downloaded**: December 14, 2018 (emailed to us by Jeffrey Vieser at NMFS)

**Description**:  Records of Bmsy and Fmsy estimates from stock assessments conducted in the greater Northeast region

**Time range**: 2004 - 2018

**Format**:  Tabular

# About

The percentage of species with formal stocks assessments is calculated using the RAM and NMFS data. To calculate a score, we used a reference point of a stock assessment conducted at least once every 5 years. This target was provided by Jeffery Vieser, and is the same metric used in FSSI.

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(readxl)
```


Load Data
```{r load_data}
nmfs_stock_ass <- read_csv("~/github/ne-prep/prep/fis/data/nmfs_stock_assessment_data.csv")%>% select(-X1)
ram_stock_ass  <- read_csv("~/github/ne-prep/prep/fis/data/ram_stock_assessment_data.csv")%>% select(-X1)
nmfs_catch_raw <- read.csv("~/github/ne-prep/prep/fis/data/nmfs_spatial_catch_by_ohi_rgn.csv")
```

## Tidy data
First we want to use the NMFS catch data to create a data set for the species landed and tonnage.
```{r catch_tidy}
nmfs_catch <- nmfs_catch_raw %>%
  mutate(nmfs_original_species = as.character(species),
         species_low = tolower(species)) %>%
  select(nmfs_original_species, species_low, year) %>%
  distinct()
```

Now we want to clean up the species names in the NMFS catch data, including alternative names 
```{r grab_species_caught_tidy}
nmfs_catch_sp <- nmfs_catch %>%
  separate(species_low, into = c("species1", "species2", "species3"), sep = ",", remove = FALSE) %>%
  mutate(species2 = trimws(gsub("\\/.*","", species2))) %>%
  separate(species_low, into = c("alt_name1", "alt_name2", "alt_name3"), sep = "/", remove = FALSE) %>%
  mutate(nmfs_catch = str_replace_all(paste0(species3, " ", species2, " ", species1), "NA ", ""),
         nmfs_catch = trimws(gsub("\\/.*","", nmfs_catch)),
         nmfs_catch_alternate = case_when(
          !is.na(alt_name2) ~ alt_name2,
          TRUE ~ NA_character_),
         nmfs_catch_alternate_2 = case_when(
          !is.na(alt_name3) ~ alt_name3,
          TRUE ~ NA_character_)) %>%
  select(species_low, nmfs_catch, nmfs_catch_alternate, nmfs_catch_alternate_2, year) %>%
  mutate(nmfs_catch2              = ifelse(str_detect(nmfs_catch_alternate, "mixed"), 
                                         paste0(nmfs_catch, " ", nmfs_catch_alternate), NA),
         nmfs_species             = ifelse(!is.na(nmfs_catch2), tolower(nmfs_catch2), tolower(nmfs_catch)),
         nmfs_species_alternate   = trimws(ifelse(str_detect(nmfs_catch_alternate, "mixed"), NA, 
                                                tolower(nmfs_catch_alternate)), "both"),
         nmfs_species_alternate_2 = trimws(tolower(nmfs_catch_alternate_2)), "both") %>% 
  select(year, species_low, nmfs_species, nmfs_species_alternate, nmfs_species_alternate_2) %>%
  mutate(species = nmfs_species)
```

Next we use the NMFS stock assessment data to determine which species have assessments done 
```{r stock_assesments_NMFS}
nmfs_stock_years <- nmfs_stock_ass %>%
  select(year, stock) %>%
  distinct() %>%
  separate(stock, into = c("species", "location"), sep = " - ") %>%
  mutate(species = tolower(species),
         source = "NMFS",
         stockid = NA)
```

Do the same thing with the RAM stock species
```{r stock_assesments_RAM}
ram_years <- ram_stock_ass %>%
  select(year, commonname, areaname, stockid) %>%
  distinct() %>%
  mutate(species = tolower(commonname),
         source      = "RAM",
         location = as.character(areaname)) %>%
  select(-commonname, -areaname) #create a duplicate column "species" to match with noaa
```

Next we want to combine the RAM and NMFS Stock assessment data since they have location data too
```{r combine_stock_ass}
stock_matching <- nmfs_stock_years %>%
  bind_rows(ram_years)
```

Now combine catch species with stocks. Lots of data wrangling here to get these species to matchup correctly

```{r exploring_over_time}
comb <- stock_matching %>%
  left_join(nmfs_catch_sp) %>%
  ##below I manually went through and added the NMFS_original name where it has occured in the data before. this happens when the species is not caught that year, but has an assessment done
  select(species_low, species, location, source, nmfs_species, stockid, year) %>%
  left_join(nmfs_catch_sp, by = c("species" = "nmfs_species_alternate", "year")) %>%
  mutate(nmfs_species = ifelse(is.na(nmfs_species.x), nmfs_species.y, nmfs_species.x)) %>%
  select(species_low = species_low.x, species, location, source, nmfs_species, stockid, year) %>%
  left_join(nmfs_catch_sp, by = c("nmfs_species", "year")) %>%
  mutate(species_low = ifelse(is.na(species_low.x), species_low.y, species_low.x)) %>%
  select(-species.y, -species_low.y, -species_low.x) %>%
  rename(species = species.x) %>%
  mutate(nmfs_species_fix = case_when(  #manually making some fixes for species that are found in both but aren't matching
    species == "atlantic bluefin tuna" ~ "bluefin tuna",
    species == "bigeye tuna" ~ "big eye tuna",
    species == "atlantic surfclam" ~ "surf clam",
    species == "atlantic cod" ~ "cod",
    species == "red deepsea crab" ~ "red crab",
    species == "american plaice" ~ "american plaice flounder",
    species == "atlantic menhaden" ~ "menhaden",
    species == "goosefish" ~ "monkfish",
    species == "acadian redfish" ~ "redfish",
    species == "little skate" ~ "little (summer) skate",
    species == "winter skate" ~ "winter (big) skate", #we still see winter and little skate wings reported in catch...fixed below
    species == "atlantic wolffish" ~ "wolffish",
    species == "longfin inshore squid" ~ "squid", #the NMFS catch data has longfin and shortfin (loligo and illex). but stock assessment data just has longfin
    TRUE ~ NA_character_
  ),
  nmfs_species = ifelse(is.na(nmfs_species), nmfs_species_fix, nmfs_species),
  nmfs_species_alternate = ifelse(nmfs_species == 'squid', "loligo", nmfs_species_alternate)) %>%
  select(species_low, 
         stockid,
         stock_assessment_species_name = species, 
         stock_assessment_species_location = location,
         source,
         nmfs_catch_species_name = nmfs_species, 
         nmfs_catch_species_name2 = nmfs_species_alternate_2,
         year) %>%
  mutate(species_low = case_when(
    nmfs_catch_species_name == "cod" ~ "cod",
    nmfs_catch_species_name == "big eye tuna" ~ "tuna, big eye",
    nmfs_catch_species_name == "bluefin tuna" ~ "tuna, bluefin",
    nmfs_catch_species_name == "american plaice flounder" ~ "flounder, american plaice /dab",
    nmfs_catch_species_name == "surf clam" ~ "clam, surf",
    nmfs_catch_species_name == "red crab" ~ "crab, red",
    nmfs_catch_species_name == "menhaden" ~ "menhaden",
    nmfs_catch_species_name == "monkfish" ~ "monkfish / anglerfish / goosefish",
    nmfs_catch_species_name == "redfish" ~ "redfish / ocean perch",
    nmfs_catch_species_name == "little (summer) skate" ~ "skate, little (summer)",
    nmfs_catch_species_name == "winter (big) skate" ~ "skate, winter (big)",
    nmfs_catch_species_name == "wolffish" ~ "wolffish / ocean catfish",
    nmfs_catch_species_name == "squid" ~ "squid / loligo",
    TRUE ~ as.character(species_low)
  )) %>%
  left_join(nmfs_catch) %>% #add in the original catch (uppercase)
  select(year, nmfs_original_species, stock_assessment_species_name, stock_assessment_species_location, stockid, source) %>% 
  mutate(nmfs_original_species = ifelse(stock_assessment_species_name == "sea scallop", "SCALLOP, SEA",
                                 ifelse(stock_assessment_species_name == "atlantic herring", "HERRING, ATLANTIC",
                                 ifelse(stock_assessment_species_name == "spiny dogfish", "DOGFISH, SPINY",
                                 ifelse(stock_assessment_species_name == "tilefish", "TILEFISH",
                                 ifelse(stock_assessment_species_name == "atlantic mackerel", "MACKEREL, ATLANTIC",
                                 ifelse(stock_assessment_species_name == "clearnose skate", "SKATE, CLEARNOSE",
                                 ifelse(stock_assessment_species_name == "smooth skate", 	"SKATE, SMOOTH",
                                 ifelse(stock_assessment_species_name == "thorny skate", "SKATE, THORNY",
                                 ifelse(stock_assessment_species_name == "little skate", "SKATE, LITTLE (SUMMER)",
                                 ifelse(stock_assessment_species_name == "winter skate", "SKATE, WINTER (BIG)",
                                 ifelse(stock_assessment_species_name == "ocean quahog", "OCEAN QUAHOG",
                                 ifelse(stock_assessment_species_name == "atlantic bluefin tuna", "TUNA, BLUEFIN",
                                 ifelse(stock_assessment_species_name == "atlantic croaker", "CROAKER, ATLANTIC",
                                 ifelse(stock_assessment_species_name ==  "bigeye tuna", "TUNA, BIG EYE",
                                 ifelse(stock_assessment_species_name == "american lobster", "LOBSTER, AMERICAN",
                                 ifelse(stock_assessment_species_name == "atlantic menhaden", "MENHADEN",
                                 ifelse(stock_assessment_species_name == "skipjack tuna", "TUNA, SKIPJACK",
                                 ifelse(stock_assessment_species_name == "striped bass", "STRIPED BASS",
                                 ifelse(stock_assessment_species_name == "swordfish", "SWORDFISH",
                                 ifelse(stock_assessment_species_name == "weakfish", "WEAKFISH / SQUETEAGUE / GRAY SEA TROUT",
                                 ifelse(stock_assessment_species_name == "yellowfin tuna", "TUNA, YELLOWFIN",
                                 ifelse(stock_assessment_species_name == "albacore tuna", "TUNA, ALBACORE" ,
                                        nmfs_original_species)))))))))))))))))))))) )
```

Fixing the skate wings and monk fish pieces matching. Skates and Monkfish are reported in NMFS catch as full fish as well as pieces (wings, livers, heads). I'm sure there is a nice way to incorporate these species names into the data above but I'm doing it manually here...

```{r tidy_weird_names}
add_species <- data.frame(stock_assessment_species_name = c("little skate", "winter skate", "smooth skate", "goosefish", "goosefish", "goosefish", "goosefish", "goosefish", "goosefish", "tilefish", "tilefish"),
                          stock_assessment_species_location = c("Georges Bank / Southern New England", 
                                                                "Georges Bank / Southern New England",
                                                                "Gulf of Maine",
                                                                "Gulf of Maine / Northern Georges Bank",
                                                                "Gulf of Maine / Northern Georges Bank",
                                                                "Gulf of Maine / Northern Georges Bank",
                                                                "Southern Georges Bank / Mid-Atlantic",
                                                                "Southern Georges Bank / Mid-Atlantic",
                                                                "Southern Georges Bank / Mid-Atlantic",
                                                                "Mid-Atlantic Coast",
                                                                "Mid-Atlantic Coast"),
                          source = "NMFS",
                          nmfs_original_species = c("SKATE WINGS, LITTLE (SUMMER)", "SKATE WINGS, WINTER (BIG)", "SKATE, SMOOTH", 
                                                    "MONK HEADS", "MONK LIVERS", "MONK TAILS", "MONK HEADS", "MONK LIVERS", "MONK TAILS", "TILEFISH, BLUELINE", "TILEFISH, GOLDEN"),
                          stockid = NA)
```

Combine the two datasets to get the full lookup table
```{r create_full_table}
full_df <- comb %>%
  bind_rows(add_species)
```

```{r stocks_to_rgns}
#list of stocks and their rgn_ids
match_areas_to_stocks <- read.csv("~/github/ne-prep/prep/fis/data/nmfs_spatial_catch_by_ohi_rgn.csv", stringsAsFactors = F) %>%
  select(species, stock_id, stock, rgn_id, rgn_name) %>%
  filter(!is.na(stock_id)) %>%
  distinct() %>%
  mutate(stock_assessment_species_location = case_when(
    stock_id == "CODGBE" ~ "Eastern Georges Bank",
    stock_id == "CODGBW" ~ "Georges Bank",
    stock_id == "CODGMSS" ~ "Gulf of Maine",
    stock_id == "PLAGMMA" ~ "Gulf of Maine / Georges Bank",
    stock_id == "FLDSNEMA" ~ "Southern New England / Mid-Atlantic",
    stock_id == "FLGMGBSS" ~ "Gulf of Maine / Georges Bank",
    stock_id == "FLWGB" ~ "Georges Bank",
    stock_id == "FLWGMSS" ~ "Gulf of Maine",
    stock_id == "FLWSNEMA" ~ "Southern New England / Mid-Atlantic",
    stock_id == "WITGMMA" ~ "Northwestern Atlantic Coast",
    stock_id == "YELCCGM" ~ "Cape Cod / Gulf of Maine",
    stock_id == "YELGB" ~ "Georges Bank",
    stock_id == "YELSNE" ~ "Southern New England / Mid-Atlantic",
    stock_id == "HADGBE" ~ "Eastern Georges Bank",
    stock_id == "HADGBW" ~ "Georges Bank",
    stock_id == "HADGM" ~ "Gulf of Maine",
    stock_id == "HKRGMNGB" ~ "Gulf of Maine / Northern Georges Bank",
    stock_id == "HKRSGBMA" ~ "Southern Georges Bank / Mid-Atlantic",
    stock_id == "HKSGMNGB" ~ "Gulf of Maine / Northern Georges Bank",
    stock_id == "HKSSGBMA" ~ "Southern Georges Bank / Mid-Atlantic", 
    stock_id == "HKWGMMA" ~ "Gulf of Maine / Georges Bank",
    stock_id == "HALGMMA" ~ "Northwestern Atlantic Coast",
    stock_id == "OPTGMMA" ~ "Northwestern Atlantic Coast",
    stock_id == "POKGMASS" ~ "Gulf of Maine / Georges Bank",
    stock_id == "REDGMGBSS" ~ "Gulf of Maine / Georges Bank",
    stock_id == "WOLGMMA" ~ "Gulf of Maine / Georges Bank"
  ))
```

Match
```{r match_areas_to_stocks}
m <- full_df %>%
  left_join(match_areas_to_stocks, by = c("nmfs_original_species" = "species", "stock_assessment_species_location")) %>%
  mutate(stockid = ifelse(is.na(stock_id), stockid, stock_id)) %>%
  select(-stock_id)
```

There are still species with stock assessments that do not have `stockid` in the NMFS catch dataset.
```{r filter_no_ids}
filter(m, is.na(stockid)) %>% .$stock_assessment_species_name %>% unique()
```

These species do not have multiple stocks within our region, so we can use the OHI regions where they are caught to identify where they are. The one issue is Goosefish/Monkfish. There are two stocks (Southern Georges Bank / Mid-Atlantic and Gulf of Maine / Northern Georges Bank). Let's look at these two areas, which are used for other stocks, and see what OHI areas this includes.

Regions in Southern New England/Mid Atlantic include
```{r filter_southern_location}
sgb <- filter(m, stock_assessment_species_location == "Southern Georges Bank / Mid-Atlantic") %>%
  select(rgn_id, rgn_name, stock_assessment_species_location) %>%
  distinct()
sgb
```

```{r filter_gom_locations}
gom <- filter(m, stock_assessment_species_location == "Gulf of Maine / Northern Georges Bank") %>%
  select(rgn_id, rgn_name, stock_assessment_species_location) %>%
  distinct() 
gom
```

Regions 2 and 8 are in both (MA-Virginian and Georges Bank). I think we can use these distinctions to assign regions to the Goosefish/Monkfish scores.
```{r assign_monkfish_scores}
goosefish <- gom %>%
  bind_rows(sgb) %>%
  filter(!is.na(rgn_id)) %>%
  mutate(stock_assessment_species_name = "goosefish")

m2 <- m %>%
  left_join(goosefish, by = c("stock_assessment_species_name", "stock_assessment_species_location")) %>%
  mutate(rgn_id = ifelse(is.na(rgn_id.y), rgn_id.x, rgn_id.y),
         rgn_name = ifelse(is.na(rgn_name.y), rgn_name.x, rgn_name.y)) %>%
  select(-rgn_id.y, -rgn_id.x, -rgn_name.y, -rgn_name.x)
```

Now lets just get all OHI regions where the remaining stocks are caught. The easier alternative to this would be to assign all OHI regions to all of these stocks, but if this data is used outside of scores that would possibly indicate stocks existing where they do not.
```{r left_overs}
leftovers <- m2 %>%
  filter(is.na(rgn_id))

#filter catch for these leftover species
leftovers_catch <- read.csv("~/github/ne-prep/prep/fis/data/nmfs_spatial_catch_by_ohi_rgn.csv", stringsAsFactors = F) %>%
  filter(species %in% leftovers$nmfs_original_species) %>%
  select(species, rgn_id, rgn_name) %>%
  distinct()
```

```{r create_complete_data}
comp_dataset <- m2 %>%
  left_join(leftovers_catch, by = c("nmfs_original_species" = "species")) %>%
  mutate(rgn_id = ifelse(is.na(rgn_id.y), rgn_id.x, rgn_id.y),
         rgn_name = ifelse(is.na(rgn_name.y), rgn_name.x, rgn_name.y)) %>%
  select(-rgn_id.y, -rgn_id.x, -rgn_name.y, -rgn_name.x) %>% 
  filter(year >=2001 & year <= 2017)
```


We still have 7 species that do not have a region assigned to them because there is no landings data for them. So we have to go through an do them individually
atlantic salmon, rosette skate, 	barndoor skate, albacore tuna, blue marlin , sailfish, white marlin


```{r remove_no_locs}
#id the species with no region associated with it
noloc <- comp_dataset %>% 
  filter(is.na(rgn_name))

unique(noloc$stock_assessment_species_name)

#create df of just the regions
regions <- comp_dataset %>% 
  select(rgn_name, rgn_id) %>% 
  distinct() %>% 
  filter(!is.na(rgn_name))
```

atlantic salmon-Gulf of Maine
```{r fix_reg_atl_sal}
regions_salmon <- regions %>% 
  slice(rep(row_number(), 2)) %>% 
  filter(rgn_name == "Gulf of Maine" | rgn_name == "Northeast")

salmon <- comp_dataset %>% 
  filter(is.na(rgn_name),
         stock_assessment_species_name == "atlantic salmon") %>%   # assuming only caught in the GOM because don't have more info
  select(year, nmfs_original_species, stock_assessment_species_name, stock_assessment_species_location, stockid, source, stock) %>% 
  cbind(regions_salmon)
```

rosette skate-Southern New England / Mid-Atlantic
```{r fix_reg_ros_skate}
regions_ros <- regions %>% 
  slice(rep(row_number(), 10)) # 10 is the number years

ros <- comp_dataset %>% 
  filter(is.na(rgn_name),
         stock_assessment_species_name == "rosette skate") %>% 
  slice(rep(1:n(), each = 12)) %>% # 11 is the number of regions that this fish is 
  select(year, nmfs_original_species, stock_assessment_species_name, stock_assessment_species_location, stockid, source, stock) %>% 
  cbind(regions_ros)
```

blue marlin-Atlantic Ocean
```{r fix_reg_blue_mar}
regions_blue_marlin <- regions %>% 
  slice(rep(row_number(), 10)) 

blue_marlin <- comp_dataset %>% 
  filter(is.na(rgn_name),
         stock_assessment_species_name == "blue marlin" ) %>% 
  slice(rep(1:n(), each = 12)) %>% 
  select(year, nmfs_original_species, stock_assessment_species_name, stock_assessment_species_location, stockid, source, stock) %>% 
  cbind(regions_blue_marlin)
```

white marlin-Atlantic Ocean
```{r fix_reg_white_mar}
regions_white_marlin <- regions %>% 
  slice(rep(row_number(), 10))

white_marlin <- comp_dataset %>% 
  filter(is.na(rgn_name),
         stock_assessment_species_name == "white marlin" ) %>% 
  slice(rep(1:n(), each = 12)) %>% 
  select(year, nmfs_original_species, stock_assessment_species_name, stock_assessment_species_location, stockid, source, stock) %>% 
  cbind(regions_white_marlin)
```

sailfish-Western Atlantic
```{r fix_reg_sail}
regions_sail <- regions %>% 
  slice(rep(row_number(), 8))

sail <- comp_dataset %>% 
  filter(is.na(rgn_name),
         stock_assessment_species_name == "sailfish" ) %>% 
  slice(rep(1:n(), each = 12)) %>% 
  select(year, nmfs_original_species, stock_assessment_species_name, stock_assessment_species_location, stockid, source, stock) %>% 
  cbind(regions_sail)
```

combine all together!
```{r combine_missing_locs}
missing_locations <- rbind(ros, blue_marlin, white_marlin, salmon, sail)
```

```{r}
comp_dataset2 <- comp_dataset %>% 
   filter(stock_assessment_species_name != "rosette skate" & !is.na(rgn_id),
          stock_assessment_species_name != "blue marlin" & !is.na(rgn_id),
          stock_assessment_species_name != "white marlin" & !is.na(rgn_id),
          stock_assessment_species_name != "atlantic salmon" & !is.na(rgn_id),
          stock_assessment_species_name != "sailfish" & !is.na(rgn_id)) %>% 
  rbind(missing_locations)
```


Do the assessment calculatins!! and then save the data, and then use it to combine with the other assesment

```{r}
time_series_assessed <- comp_dataset2 %>% 
  mutate(assessment = 1)
```

Create df for all areas and all years to fill in for years with no assessments later
```{r all_fish}
all_fish<- comp_dataset2 %>% 
  select(nmfs_original_species, stock_assessment_species_name, stock_assessment_species_location, stockid, source, stock, rgn_name, rgn_id) %>% 
  distinct() %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>%
  select(-n) 
```


Combining
```{r all_fish_with_assessments}
assessment_frq  <- all_fish %>% 
  left_join(time_series_assessed, by = c("year", "nmfs_original_species", "stock_assessment_species_name", "stock_assessment_species_location", "stockid", "source", "stock", "rgn_name", "rgn_id")) %>% 
  mutate(assessment = ifelse(is.na(assessment), 0, assessment)) %>% 
  mutate(assessment = as.numeric(assessment)) %>% 
  group_by(stock_assessment_species_name, stock_assessment_species_location) %>% 
  mutate(assesment_sum_5yr = zoo::rollapply(assessment, 5, FUN = sum, na.rm = F, partial = T, align = "right"), 
         adeq_assessed = if_else(assesment_sum_5yr >= 1, 1, 0)) %>% 
  ungroup()

write.csv(assessment_frq, file = "data/assessment_frq.csv")
```


## Calculate scores and graph
```{r calculate_ assessed}
number_assessed <- assessment_frq %>% 
  select(nmfs_original_species,stock_assessment_species_name, rgn_id, rgn_name, year) %>% 
  group_by(year, rgn_id, rgn_name) %>% 
  count() %>% 
  rename(number_stocks=n)
  
adeq_ass <- assessment_frq %>% 
  select(nmfs_original_species,stock_assessment_species_name, rgn_id, rgn_name, year, adeq_assessed) %>% 
  group_by(year, rgn_id, rgn_name) %>% 
  dplyr::summarise(adeq_assessed = sum(adeq_assessed))
  
adeq_assessed <- left_join(number_assessed, adeq_ass, by = c("year", "rgn_id", "rgn_name") ) %>% 
  mutate(percent_adeq_ass = adeq_assessed/number_stocks)

write.csv(adeq_assessed, file = "data/adeq_assessed.csv")
```

Visualize
```{r graph}
adq_graph <- ggplot(adeq_assessed)+
  geom_line(aes(x=year, y= percent_adeq_ass, color= rgn_name))

adq_graph
```