---
title: "OHINE: Fishing Resilience - Areas Protected From Fishing Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

#Summary

This script calculates the percentage of area protected from fishing by OHINE region.This data is used to measure fishing resilience regulations.

# Data

We use the United States Geological Survey Protected Areas Database for the US. Version 2.0.

**Reference**: 
U.S. Geological Survey (USGS) Gap Analysis Project (GAP), 2018, Protected Areas Database of the United States (PAD-US): U.S. Geological Survey data release, https://doi.org/10.5066/P955KPLE.

https://www.usgs.gov/core-science-systems/science-analytics-and-synthesis/gap/science/protected-areas

**Downloaded**: June 5th, 2019

**Time range**: Protected areas list established date back to early 1900's where available

**Format**:  Spatial Geodatabase

# About
For calculating protected area score we use a target of 30% areas in the ocean closed off to fishing.

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(tidyverse)
library(dplyr)
library(sf)
```

## Tidy the data
Note: Used a lot of same code from sense of place layer
Let's look at the available layers in the USGS PADUS database
```{r look_at_gdb_layers}
st_layers(file.path(dir_anx, "_raw_data/USGS/PADUS2_0_GDB_Arc10x/PADUS2_0.gdb"))
```

For the marine layer, we are going to remove all areas that are designated as fishery/shellfish management areas as well as the Special Area Management Plans. 
```{r load_and_clean_marine_layer}
marine <- read_sf(dsn = file.path(dir_anx, "_raw_data/USGS/PADUS2_0_Shapefiles/PADUS2_0Marine.shp")) %>%
  st_transform(us_alb) %>%
  st_crop(rgns)  %>%
  st_intersection(rgns) %>%
  filter(!Loc_Ds %in% c("Fishery Management Area", "Fishery Management Areas", "Shellfish Management Area", "Conservation Area","Gear Restricted Area", "Essential Fish Habitat Conservation Area", "Special Area Management Plan", "Closure Area"), #the Conservation Areas are for Mussel Seed
         !is.na(Loc_Ds),
!Fish_Rstr %in% c("No Site Restrictions", "Restrictions Unknown"), #only keeping areas with recreational or commercial fishing is restricted or prohibited
         Constancy == "Year-round")%>% 
  select(Loc_Ds, Loc_Nm, Date_Est, rgn_name, rgn_id)
```

Map it!
```{r interactive_map, eval = F}
 #mapview(proc, col.regions = "red") + 
usgs_data <-mapview(marine, col.regions = "blue") 

## save interactive map
mapshot(usgs_data, url = "protected_areas_map.html")
```


First let's combine polygons so we don't double count areas. 
We are also accounting for the year established, some were created during our time frame 
```{r merge_overlapping}
marine_protected_by_rgn <- data.frame()

for(i in 2005:2017){
  
  print(i)
  
  s_yr <- filter(marine, Date_Est <= i | is.na(Date_Est)) %>% #all NA included as well
    group_by(rgn_id) %>%
    summarise() %>%
    ungroup() %>%
    mutate(prot_area_m = st_area(geometry)) %>% #calculate total protected area
    mutate(prot_area_km2 = as.numeric(prot_area_m)/1000000) %>% 
    left_join(rgns %>% st_set_geometry(NULL)) %>% #join to inshore buffer 
    mutate(area_km2 = as.character(area_km2)) %>% 
    mutate(area_km2 = as.numeric(area_km2)) %>% 
    mutate(prop_area = prot_area_km2/area_km2 , #calculate proportion of waters in regions 1-4 that are protected
           type = "marine",
           year = i) %>%
    st_set_geometry(NULL)
  
  marine_protected_by_rgn <- bind_rows(s_yr, marine_protected_by_rgn)
}
```

Calculate percent of marine area protected
```{r marine_protected_prop_by_rgn}
percent_water_protec <- ggplot(marine_protected_by_rgn, aes(x = year, y = prop_area, color = rgn_name)) +
  geom_line() +
  theme_classic() +
  geom_hline(yintercept=0.3)+
  labs(y = "Percentage", x = "Year") +
  ggtitle("Percentage of Offshore waters protected")

ggsave("figs/percent_offshore_water_protec.jpg", width=7, height=5, dpi=300)
```

## Calculate scores and map it

Calculate the protected area scores
```{r calculate}
protected_score <- marine_protected_by_rgn %>% 
  select(rgn_id, rgn_name, year, prop_area) %>% 
  mutate(target = 0.3) %>% 
  mutate(fish_area_score = prop_area/target,
         fish_area_score = if_else(fish_area_score > 1, 1, fish_area_score)) %>% 
  mutate(rgn_id = as.character(rgn_id))

write.csv(protected_score, file = "data/fish_area_protected.csv")
```

Visualize
```{r graph}
protected_score_graph <- ggplot(protected_score) +
  geom_line(aes(x=year, y= fish_area_score, color = rgn_id))+
  labs(title = "Protected Areas Score: Fisheries Resilience")+
  ylab("Score") +
  xlab("Year") +
  theme_classic() 

protected_score_graph
ggsave("figs/protected_areas_score_graph.jpg", width=7, height=5, dpi=300)
```













