---
title: "OHIEC: Resilience - NOAA Fisheries Observer Coverage"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(readxl)
```

###1) Total number of sea days with fisheries observer/ alloted number of sea days 

Allocated seadays from https://www.nefsc.noaa.gov/fsb/SBRM/
(total number of sea days allocated each year based on the agency budget amd the compensation rate analysis by funding source for ex:(April 2016-March 2017))

There is an 8% target coverage of ships (in report https://www.nefsc.noaa.gov/fsb/SBRM/2011/2011_SBRM_3Year_Report.pdf)

Load in the data
```{r}
fish_ob_raw <- read_xlsx(file.path(dir_anx, "_raw_data/NOAA_Fisheries_Observer/fisheries_observer.xlsx"), sheet=1)
ob_alloc_raw <- read_xlsx(file.path(dir_anx, "_raw_data/NOAA_Fisheries_Observer/fishob_seaday_alloc.xlsx"), sheet=1) %>% 
  rename(YEAR = year)
```

Calculate the enforcement score for number of fisheries observer sea days by aking the number of actual seadays/allocated sea day
Targets for allocated seadays from https://www.nefsc.noaa.gov/fsb/SBRM/
```{r}
observ_cov <- left_join(fish_ob_raw, ob_alloc_raw, by = c("YEAR")) %>% 
  mutate(score = TOTDAYS/seadays) %>% 
#  fill(score, .direction = "up") %>%  #back gap filled using the 1st year allocated sea days are avaliable
  filter(YEAR <= 2017) %>% 
  select(YEAR, score) %>% 
  rename(year = YEAR) %>% 
  rename(enforcement_score_obser= score)

write.csv(observ_cov, file = "data/observ_cov.csv")
```

Things to think about:
- more information on allocated sea days prior to 2013
- Score overestimating because allocated sea days do not include at-sea ground fish monitoring but the observer data does









