---
title: "OHINE: Fishing Resilience - NOAA Fisheries Observer Coverage Data Prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates the fisheries observer coverage scores by OHINE region. This data is used to measure climate change resilience implementation & enforcement.

# Data

**Fisheries Observer Data**
**Downloaded**: August, 2019

**Description** The data was compiled and emailed to us by Debra Duarte (debra.duarte@noaa.gov)

**Time range**: 2005-2019

**Format**:  excel spreadsheet

**Allotted Observer Seadays**
**Downloaded**: August, 2019

**Description** The data was manually complied by me (Juliette Verstaen) from reports outlining that years fisheries observer plan https://www.nefsc.noaa.gov/fsb/SBRM/

**Time range**: 2013-2019

**Format**:  excel spreadsheet

# About
For calculating protected area score we use a target of 30% areas in the ocean closed off to fishing. The total number of sea days (our target) allocated each year based on the agency budget and the compensation rate analysis by funding source for ex:(April 2016-March 2017))

There is an 8% target coverage of ships (in report https://www.nefsc.noaa.gov/fsb/SBRM/2011/2011_SBRM_3Year_Report.pdf)

Notes:
- Finding more information on allocated sea days prior to 2013 would improve the accuracy of these scores
- Score overestimating because allocated sea days do not include at-sea ground fish monitoring but the observer data does


# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(readxl)
```

Load in the data
```{r load_data}
fish_ob_raw <- read_xlsx(file.path(dir_anx, "_raw_data/NOAA_Fisheries_Observer/fisheries_observer.xlsx"), sheet=1)
ob_alloc_raw <- read_xlsx(file.path(dir_anx, "_raw_data/NOAA_Fisheries_Observer/fishob_seaday_alloc.xlsx"), sheet=1) %>% 
  rename(YEAR = year)
```

## Calculate scores
```{r calculate_scores}
observ_cov <- left_join(fish_ob_raw, ob_alloc_raw, by = c("YEAR")) %>% 
  mutate(score = TOTDAYS/seadays) %>% 
  filter(YEAR <= 2017) %>% 
  select(YEAR, score) %>% 
  rename(year = YEAR) %>% 
  rename(enforcement_score_obser= score)

write.csv(observ_cov, file = "data/observ_cov.csv")
```








