---
title: "OHIEC: Resilience - NOAA Office of Law Enforcemnet enforcement actions"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary
This script calculates the enforcement score using NOAA OLE data.

**NOAA OLE Data**
**Downloaded**: August, 2019

**Description** The data was emailed to us by Casey Brennan casey.brennan@noaa.gov and then important information was manually pulled into spread sheets by me.

**Time range**: 2015-2018

**Format**:  PDF Reports, excel spreadsheets

# About
Enforcement here is measures by taking the total number of staff, number of patrols, number of outreach events OLE partakes in and dividing it by the highest historic total.

# Methods

## Set up, libraries, load data
``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(readxl)
```

Load in Data
```{r load_data}
ole_raw <- read_xlsx(file.path(dir_anx, "_raw_data/NOAA_OLE/NE_OLE_data.xlsx"))
```

## Calculate scores
```{r calculate_scores}
ole_enforcement <- ole_raw %>% 
  filter(year >= 2015) %>% # not good enough data to use anything before 2015
  mutate(staff_score = staffing/33, #divide by the maximum of all years
         num_patrols_score = num_patrols/537,  #divide by the maximum of all years
         out_reach_num_score = out_reach_num/453) %>%  #divide by the maximum of all years
  select(year, staff_score, num_patrols_score, out_reach_num_score) %>% 
  replace(., is.na(.), 0.718) %>% #number calculated from a linear regression of first 3 data points
  rowwise() %>% 
  mutate(ole_enforcement_score=mean(c(staff_score, num_patrols_score, out_reach_num_score), na.rm=T) )

write.csv(ole_enforcement, file = "data/ole_enforcement.csv")
```
