---
title: "OHIEC: Resilience - Calculating Fishing pressure resilience"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
toc: true
---

# Summary

This script calculates fishing pressure resilience values by OHINE region. 

The data layers are used to quantify existence of regulations, implementation + enforcement, and effectiveness + compliance. These three parts are then used to calculate one overall water pollution resilience score for each region. 

1. Regulations: landed_ass.csv, fish_area_protected.csv, adeq_assessed.csv, spec_regulation.csv
2. Implementation + Enforcement: observ_cov.csv, ole_enforcement.csv
3. Effectiveness + Compliance: ole_compliance.csv

Data Time Ranges:
landed_ass: 2005-2017
fish_area_protected: NA (one score applied to all years)
adeq_assessed: 2005-2017
species_man: NA
observ_cov: 2013-2017
ole_enforcement: 2015-2018
ole_compliance: 2015-2018


``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

library(csv)
library(tidyverse)
library(dplyr)
library(Hmisc)
library(stringi)
library(stringr)
library(readxl)
library(tabulizer)
```


Read in Data
```{r}
landed_ass_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/fishing_pressure/data/landed_ass.csv") %>% select(-X) 

fish_area_protected_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/fishing_pressure/data/fish_area_protected.csv") %>% 
  select(-X) 

adeq_assessed_short<- read.csv("~/github/ne-prep/prep/resilience/ecological/fishing_pressure/data/adeq_assessed.csv") %>% 
  select(-X)

species_man_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/fishing_pressure/data/species_man.csv") %>% 
  select(-X)

observ_cov_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/fishing_pressure/data/observ_cov.csv") %>% 
  select(-X)

ole_enforcement_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/fishing_pressure/data/ole_enforcement.csv") %>% 
  select(-X)

ole_compliance_short <- read.csv("~/github/ne-prep/prep/resilience/ecological/fishing_pressure/data/ole_compliance.csv") %>% 
  select(-X)
```


Cleaning up the data to be combined and used as a layer
```{r}
landed_ass <- landed_ass_short %>% 
  select(year, rgn_id, rgn_name, per_catch_landed) %>% 
  rename(landed_ass_score = per_catch_landed)
```

```{r}
fish_area_protected <- fish_area_protected_short %>% 
  select(year, rgn_id, rgn_name, fish_area_score) %>% 
  rename(fish_area_protected_score = fish_area_score)
```

```{r}
adeq_assessed <- adeq_assessed_short %>% 
  select(year, rgn_id, rgn_name, percent_adeq_ass) %>% 
  rename(adeq_assessed_score = percent_adeq_ass)
```


```{r}
species_man <- species_man_short %>% 
  rename(species_man_score = species_regulated_score) %>% 
  uncount(13, .id = "n", .remove = F) %>%
  mutate(year = ifelse(n == 1, 2005, n + 2004)) %>% 
  select(-n) %>% 
  mutate(rgn_name = str_replace(rgn_name,"Massachusetts-Gulf of Maine", "Massachusetts-North"),
         rgn_name = str_replace(rgn_name,"Massachusetts-Virginian", "Massachusetts-South"))
```

```{r}
observ_cov <- observ_cov_short %>% 
  rename(observ_cov_score = enforcement_score_obser) %>% 
  fill(observ_cov_score, .direction = c("up")) %>% 
  uncount(11, .id = "n", .remove = F) %>%
  mutate(rgn_name = ifelse(n == 1, "Offshore", 
                       ifelse(n==2, "Georges Bank",
                               ifelse(n==3, "Gulf of Maine",
                                      ifelse(n==4, "Mid-Atlantic Bight",
                                             ifelse(n==5, "Connecticut",
                                                    ifelse(n==6, "Maine",
                                                           ifelse(n==7, "Massachusetts-North",
                                                                  ifelse(n==8, "Massachusetts-South",
                                                                         ifelse(n==9, "New Hampshire",
                                                                                ifelse(n==10, "New York",
                                                                                       "Rhode Island"))))))))))) %>%
  rename(rgn_id=n)
```


Tidying the OLE enforcemnet data for res score calculations by region
```{r}
ole_enforcement <- ole_enforcement_short %>% 
  select(year, ole_enforcement_score) %>% 
  fill(ole_enforcement_score, .direction = c("up")) %>% 
  filter(year <= 2017) %>% 
  uncount(11, .id = "n", .remove = F) %>%
  mutate(rgn_name = ifelse(n == 1, "Offshore", 
                       ifelse(n==2, "Georges Bank",
                               ifelse(n==3, "Gulf of Maine",
                                      ifelse(n==4, "Mid-Atlantic Bight",
                                             ifelse(n==5, "Connecticut",
                                                    ifelse(n==6, "Maine",
                                                           ifelse(n==7, "Massachusetts-North",
                                                                  ifelse(n==8, "Massachusetts-South",
                                                                         ifelse(n==9, "New Hampshire",
                                                                                ifelse(n==10, "New York",
                                                                                       "Rhode Island"))))))))))) %>%
  rename(rgn_id=n)
```

Tidying the OLE compliance data for res score calculations by region
```{r}
ole_compliance <- ole_compliance_short %>% 
  select(year, ole_compliance_score) %>% 
  filter(year <= 2017) %>% 
  uncount(11, .id = "n", .remove = F) %>%
  mutate(rgn_name = ifelse(n == 1, "Offshore", 
                       ifelse(n==2, "Georges Bank",
                               ifelse(n==3, "Gulf of Maine",
                                      ifelse(n==4, "Mid-Atlantic Bight",
                                             ifelse(n==5, "Connecticut",
                                                    ifelse(n==6, "Maine",
                                                           ifelse(n==7, "Massachusetts-North",
                                                                  ifelse(n==8, "Massachusetts-South",
                                                                         ifelse(n==9, "New Hampshire",
                                                                                ifelse(n==10, "New York",
                                                                                       "Rhode Island"))))))))))) %>%
  rename(rgn_id=n)
```

Combining the data and calulating one overall fisheries resilience score
```{r}
res_fs <- left_join(landed_ass, fish_area_protected, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(species_man, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(adeq_assessed, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(observ_cov, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(ole_enforcement, by = c("year", "rgn_id", "rgn_name")) %>% 
  left_join(ole_compliance, by = c("year", "rgn_id", "rgn_name")) %>% 
  fill(ole_compliance_score, .direction = c("up")) %>% 
  fill(ole_enforcement_score, .direction = c("up")) %>% 
  rowwise() %>% 
  mutate(regulation = mean(c(landed_ass_score, fish_area_protected_score,species_man_score, adeq_assessed_score)),
         enforcement = mean(c(observ_cov_score, ole_enforcement_score))) %>% 
  rename(compliance = ole_compliance_score) %>% 
  mutate(res_score = mean(c(regulation, enforcement, compliance)))
```

Visualization
```{r}
res_fs_graph <- ggplot(res_fs) +
  geom_line(aes(x=year, y= res_score, color = rgn_name))+
  ggtitle("North East Fishing Resilience Scores") +
  ylab("Score") +
  xlab("Year") +
  theme_classic() 

res_fs_graph
ggsave("figs/res_fs_graph.jpg", width=7, height=5, dpi=300)
```


Save necessary information for the fishing layer
```{r}
res_fishing <- res_fs %>% 
  select(year, rgn_id, rgn_name, res_score) %>% 
  rename(score = res_score)

write.csv(res_fishing, file.path(dir_git,"prep/resilience/ecological/fishing_pressure/data/res_fishing.csv"))
```


Write as a resilience layer to ne-scores
```{r}
write.csv(res_fishing, file.path(dir_calc, "layers/res_fishing.csv"))
```



