---
title: 'Cleaning up NOEP data'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

## Summary

This script takes individual state-level .csv data that were manually downloaded from NOEP (http://www.oceaneconomics.org/Market/ocean/oceanEcon.asp) and combines them and cleans them for layer preparation. This data is used to create the **jobs**, **wage** and **gdp** layers for use in Livelihoods & Economies as well as Tourism & Recreation.

---

## Data Source

### National Ocean Economics Program (NOEP)

**Downloaded**: [Manually downloaded by state from website](http://www.oceaneconomics.org/Market/ocean/oceanEcon.asp) on May 9, 2019.    
**Description**:  Total number of jobs and wages per sector for RI, ME, MA, CT, NY and NH counties from 2005 to 2016. The data also include number of establishments and GDP for each sector - state - year.  
**Native data resolution**: County level     
**Time range**: 2005 - 2016  
**Format**:  Tabular  
---

## Setup

```{r, message = F, warning = F}
knitr::opts_chunk$set(fig.width = 10, fig.height = 6, fig.path = 'figs/', 
                      message = FALSE, warning = FALSE)

#load common file
source('~/github/ne-prep/src/R/common.R')  ### an OHI-NE specific version of common.R

#libraries
library(tidyverse)
```

## Get data  

Load the data and fix the column classes to be numeric instead of character.
```{r}
## get a list of the files we're going to combine. These are all held on the Mazu server at NCEAS
files <- list.files(file.path(dir_anx, "_raw_data/NOEP/d2019"), full.names = T)

## read in each csv file and change columns from character to numeric while also removing commas and dollar signs ($)
out <- map_df(files, read_csv) %>%
  select(1:10) %>% #there's a weird character column of NA showing up at the end. remove here
  filter(!is.na(State)) %>% #there are some NA for state because these rows have notes
  mutate(Establishments = as.numeric(gsub(",", "", Establishments)),
         Employment     = as.numeric(gsub(",", "", Employment)),
         Wages          = gsub(",", "", Wages),
         GDP            = gsub(",", "", GDP)) %>%
  mutate(Wages          = as.numeric(gsub("[\\$,]", "", Wages))) %>% #doing a stepwise fix on gdp and wages for the $ and ,
  mutate(GDP            = as.numeric(gsub("[\\$,]", "", GDP))) 
```

Since we have two separate Massachusetts regions, we need to split the data according to how those counties border those two regions. The [`ma_ctys` table](https://github.com/OHI-Northeast/ne-prep/blob/gh-pages/src/tables/MA_counties.csv) lists what counties in Massachusetts belong to each region. For counties that are in both (Plymouth and Barnstable) we will simply divide the values by 2.


```{r split_MA}
# need to read in the MA counties since they are divided across two bioregions

ma_ctys <- read_csv("~/github/ne-prep/src/tables/MA_counties.csv") %>%
  select(-X1) %>%
  mutate(State = "Massachusetts")

clean_data <- out %>%
        left_join(ma_ctys, by = c("State", "County")) %>% #this creates duplicates for all MA counties
        mutate(rgn_name = ifelse(is.na(rgn_name), State, rgn_name)) %>%
        left_join(rgn_data, by = 'rgn_name') %>%
        mutate(rgn_id = ifelse(is.na(rgn_id.x), rgn_id.y, rgn_id.x),
               Employment     = ifelse(State == "Massachusetts" & County %in% c("Plymouth", "Barnstable"), 
                                      Employment/2, Employment),
               Establishments = ifelse(State == "Massachusetts" & County %in% c("Plymouth", "Barnstable"), 
                                       Establishments/2, Establishments),
               Wages          = ifelse(State == "Massachusetts" & County %in% c("Plymouth", "Barnstable"), 
                                       Wages/2, Wages),
               GDP            = ifelse(State == "Massachusetts" & County %in% c("Plymouth", "Barnstable"),
                                       GDP/2, GDP)) %>%
        select(-rgn_id.x, -rgn_id.y)
```

## Try to account for added seafood sector

>Note: In the recent release of the ENOW series "An additional industry, Fish and Seafood Merchant Wholesalers (NAICS 424460) was added to the Living Resources sector's Seafood Market industry for data years 2016 and beyond. The industry was not applied to earlier years." The addition of this industry significantly distorts annual comparisons and timeline analysis of the Seafood Market sector and Living Resources totals.

```{r}
maine <- clean_data %>%
  filter(State == "Maine") %>%
  group_by(County, Year, Sector) %>%
  summarize(establishments = sum(Establishments, na.rm = T),
            jobs = sum(Employment, na.rm = T),
            wages = sum(Wages, na.rm = T),
            gdp = sum(GDP, na.rm = T))

#jobs
ggplot(maine, aes(x = Year, y = gdp, color = Sector)) +
  geom_line() +
  facet_wrap(~County, scales = "free")
```
# Account for this adjustment
by subtracting the 2016 value for Living Resources from All Ocean Sectors, and instead gapfilling with 2015 data. See what this does

```{r}
#get difference in Living Resources sector from 2015 to 2016. We will use this to fix the All Ocean Sectors data for 2016
diff <- clean_data %>%
  filter(Year > 2014,
         Sector == "Living Resources") %>%
  pivot_longer(cols = c(Establishments, Wages, Employment, GDP), names_to = "layer", values_to = "value") %>%
  pivot_wider(names_from = Year, values_from = value) %>%
  mutate(diff = `2016` - `2015`) %>%
  select(County, rgn_id, layer, diff)

fix <- clean_data %>%
  filter(Year == 2016,
         Sector == "All Ocean Sectors") %>%
  pivot_longer(cols = c(Establishments, Wages, Employment, GDP), names_to = "layer", values_to = "2016_value") %>%
  left_join(diff, by = c("rgn_id", "layer", "County")) %>%
  mutate(new_2016_value = `2016_value` - diff) %>%
  mutate(new_2016_value = ifelse(is.na(new_2016_value), `2016_value`, new_2016_value)) %>%
  select(-`2016_value`, -diff) %>%
  pivot_wider(names_from = "layer", values_from = "new_2016_value")
  
#remove 2016 from clean_data and add in fix

clean_data_w_2016_fix <- clean_data %>%
  mutate(remove = ifelse(Year == 2016 & Sector == "All Ocean Sectors", 1, 0)) %>%
  filter(remove == 0) %>%
  select(-remove) %>%
  bind_rows(fix)
```


## Save

The final output if `clean_noep_data.csv` in the data folder.

```{r save}
write.csv(clean_data_w_2016_fix, "data/clean_noep_data.csv")
```
