---
title: 'OHI-Northeast: Economic access data layer'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ne-prep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup,  message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      message = FALSE, warning = FALSE)

source('~/github/ne-prep/src/R/common.R')  ### an OHINE specific version of common.R

dir_git <- '~/github/ne-prep'
dir_anx <- file.path(dir_M, 'git-annex/neprep')

#libraries
library(tidyverse)
```

#Summary
Economic access data layer is derived from gas prices compared to median wage.

***

#Data Source 
**Reference**: BLS wage data

**Downloaded**: [date downloaded or received]

**Description**:  Wage data

**Native data resolution**: County level

**Time range**: 2005-2015

**Format**:  Tabular

**Reference**: 

---
**Reference**: https://www.eia.gov/petroleum/gasdiesel/

**Downloaded**: 

**Description**: Gas price data

**Native data resolution**: County level

**Time range**: 2005-2015

**Format**:  Tabular

***
  
# Methods

## Median wage

I'm using the `blscrapeR` package and have my own API access key. We aren't using the NOEP data because we don't want to limit this to only jobs that are tied to the ocean economy.

```{r, results = 'hide'}
#devtools::install_github("keberwein/blscrapeR")
library(blscrapeR)
#read in API access key that is saved on Mazu
blsKey <- striprtf::read_rtf(file.path(dir_M,'git-annex/neprep/keys/BureauofLaborStatistics.rtf'))
#set_bls_key(blsKey, overwrite=TRUE)
```


```{r bls_avg_annual_wage_by_state}

# Grab several data sets from the BLS at once.
# NOTE on series IDs: 
# Connecticut -- Statewide Average Annual Pay All establishment sizes Total Covered Total, all industries: ENU0900050010
# Maine -- Statewide Average Annual Pay All establishment sizes Total Covered Total, all industries: ENU2300050010
# Massachusetts -- Statewide Average Annual Pay All establishment sizes Total Covered Total, all industries: ENU2500050010
# New Hampshire -- Statewide Average Annual Pay All establishment sizes Total Covered Total, all industries: ENU3300050010
# New York -- Statewide Average Annual Pay All establishment sizes Total Covered Total, all industries: ENU3600050010
# Rhode Island -- Statewide Average Annual Pay All establishment sizes Total Covered Total, all industries: ENU4400050010
bls_data <- bls_api(c("ENU0900050010", "ENU2300050010", "ENU2500050010", "ENU3300050010", "ENU3600050010", "ENU4400050010"),
              startyear = 2001, endyear = 2017, Sys.getenv("BLS_KEY")) 

avg_pay_by_state <- bls_data %>%
  mutate(state = case_when(
    seriesID == "ENU0900050010" ~ "CT",
    seriesID == "ENU2300050010" ~ "ME",
    seriesID == "ENU2500050010" ~ "MA",
    seriesID == "ENU3300050010" ~ "NH",
    seriesID == "ENU3600050010" ~ "NY",
    seriesID == "ENU4400050010" ~ "RI"
  )) %>%
  select(year, pay = value, state)

ggplot(avg_pay_by_state, aes(x = year, y = pay, color = state)) +
  geom_line() +
  labs(x = "Year",
       y = "Average annual wage") +
  theme_bw()
```


## Gas price data

The gas price data is loaded but it is provided in $/MMBtu (Millions of British thermal units), which can be converted to approximate dollars per gallon using the heat contents provided by the EIA here. Once multiplied we get dollars per barrel, and then divide by 42 since there are 42 U.S. gallons in a barrel.

```{r load_data}
raw        <- read_csv(file.path(dir_anx, "_raw_data/EIA/pr_all.csv"))
conversion <- read_csv(file.path(dir_anx, "_raw_data/EIA/MER_TA3.csv")) %>% #data to convert raw data from Btu to barrels
  filter(Description == "Motor Gasoline (Finished) Consumption Heat Content") %>%
  mutate(Year = str_sub(YYYYMM, 1, str_length(YYYYMM)-2),
         conv_factor = as.numeric(Value)) %>%
  select(Year, conv_factor)
```

Clean data and multiply by Btu conversion to get dollars per gallon.
```{r clean_gas_data}
clean_gas_data <- raw %>%
  filter(State %in% c("NY", "CT", "RI", "MA", "NH", "ME"), #filter for just our states
         MSN == "MGACD") %>% #this is the code for gasoline prices for the transportation sector
  gather(key = "Year", value = "dollar_per_MMBtu", -Data_Status, -State, -MSN) %>%
  left_join(conversion) %>%
  mutate(Year = as.numeric(Year),
    dollar_per_gallon = dollar_per_MMBtu*conv_factor/42) #multiply $/btu by conversion to get $/barrel, then divide by 42 to get dollar per gallon since there are 42 gallons in a barrel

```

# Visualize

Cost of gas in US Dollars per Gallon for Northeast states.

```{r dollar_per_gallon}
ggplot(clean_gas_data, aes(x = Year, y = dollar_per_gallon, color = State, group = State)) +
  geom_line() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
```

## Get ratio

Compare wage to gas price each year.

From OHI West Coast: the current ratio of diesel gas price (Gc) to median income IC relative to a reference ratio of gas price (GT) to median income (IT) from 5 years prior, 

```{r ratio_gas_to_salary}

gas_to_income <- clean_gas_data %>%
  select(year = Year, state = State, dollar_per_gallon) %>%
  left_join(avg_pay_by_state) %>%
  filter(!is.na(pay)) %>%
  mutate(ratio = dollar_per_gallon/pay)

write_csv(gas_to_income, "data/gas_to_income_ratio.csv")

ggplot(gas_to_income, aes(x = year, y = ratio, color = state)) +
  geom_line() +
  theme_bw() +
  labs(y = "Ratio of $/gallon of gas to average annual pay")

```

## Calculate year over year change and score

The target for this goal is to keep the ratio constant year over year. In this case, if it doubles then they get a 0. If the change is negative, indicating a drop in the ratio, they get 100.

```{r calculate_yearly_change}
change <- gas_to_income %>%
  arrange(year) %>%
  group_by(state) %>%
  mutate(last_years_ratio = lag(ratio, n = 1)) %>%
  ungroup() %>%
  mutate(change = ratio-last_years_ratio,
         score  = ifelse(change < 0, 1, 1-(change/last_years_ratio)))
  
ggplot(change, aes(x = year, y = score, color = state)) +
  geom_line() +
  theme_bw()
```


## Save data

Saving data to the toolbox and to the data folder

```{r save_wage_data}
write_csv(avg_pay_by_state, "data/avg_pay_by_state.csv")
```


```{r save_gas_layer}
rgn_data %>%
  left_join(clean_gas_data, by = c("state_abv" = "State")) %>%
  group_by(rgn_id) %>%
  complete(Year = 1970:2017) %>%
  filter(Year > 2004) %>%
  select(rgn_id, year = Year, rgn_name, dollar_per_gallon) %>%
  write_csv("data/gas_prices_by_rgn.csv")
```

Toolbox layer includes all regions (even offshore) and the ratio of gas to salary

```{r}
out <- change %>%
  left_join(rgns, by = c("state" = "state_abv")) %>%
  select(rgn_id, year, score) %>%
  complete(rgn_id = 1:11, year) %>%
  filter(year > 2004)

write_csv(out, "~/github/ne-scores/region/layers/rao_econ_access.csv")
```

